### 2013-08-21 paper.tex
Added the paper

\documentclass[aoas, preprint]{imsart}
\RequirePackage{natbib}

%\long\def\authornote#1{%
%        \leavevmode\unskip\raisebox{-3.5pt}{\rlap{$\scriptstyle\diamond$}}%
%        \marginpar{\raggedright\hbadness=10000
%        \def\baselinestretch{0.8}\tiny
%        \it #1\par}}
%\newcommand{\ville}[1]{\authornote{NOTE TO SELF: #1}}

%\usepackage{amsthm,amsmath,natbib}
%\RequirePackage[colorlinks,citecolor=blue,urlcolor=blue]{hyperref}

% provide arXiv number if available:
%\arxiv{arXiv:0000.0000}

% put your definitions there:
\startlocaldefs
\endlocaldefs

\newcommand{\argmin}{\operatornamewithlimits{arg\ min}}
\newcommand{\argmax}{\operatornamewithlimits{arg\ max}}
\newcommand\ville[1]{{\color{red} #1 }}
%    \mymarginpar{\raggedright\hbadness=10000\tiny\it #1\par}}

%\usepackage[section]{placeins}
\usepackage{amsmath} 
\usepackage{times}
\usepackage{amsmath,amsthm,amssymb}
\usepackage{fancyhdr}
\usepackage{moreverb}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{url}
\usepackage{multirow} 
\usepackage[boxed, section]{algorithm}
%\usepackage{algorithm}
\usepackage{algorithmic}
%\usepackage{cite}
\usepackage{multirow} 
\usepackage{rotating}
\usepackage{geometry}
\usepackage{fix-cm}
\usepackage{subfigure}
\usepackage{bbm}
\usepackage{color} 
\usepackage{natbib}
\usepackage{xcolor,xparse}
\usepackage{xparse}
%\usepackage{setspace}
\usepackage{booktabs}
%\usepackage{subfig}
%\usepackage{caption, subcaption}

%\renewcommand{\baselinestretch}{1.2}
%\setlength{\topmargin}{-0.3in}
%\setlength{\textwidth}{6in}
%\setlength{\textheight}{8.5in}
%\setlength{\oddsidemargin}{0.25in}
%\setlength{\evensidemargin}{0.25in}
%\raggedbottom

%\doublespacing

\allowdisplaybreaks

% Math Macros.  It would be better to use the AMS LaTeX package,
% including the Bbb fonts, but I'm showing how to get by with the most
% primitive version of LaTeX.  I follow the naming convention to begin
% user-defined macro and variable names with the prefix "my" to make it
% easier to distiguish user-defined macros from LaTeX commands.
%
\newcommand{\myN}{\hbox{N\hspace*{-.9em}I\hspace*{.4em}}}
\newcommand{\myZ}{\hbox{Z}^+}
\newcommand{\R}{\mathbb{R}}
\renewcommand{\P}{\mathbb{P}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\logit}{\text{logit}}
\NewDocumentCommand{\framecolorbox}{oommm}
 {% #1 = width (optional)
  % #2 = inner alignment (optional)
  % #3 = frame color
  % #4 = background color
  % #5 = text
  \IfValueTF{#1}
   {%
    \IfValueTF{#2}
     {\fcolorbox{#3}{#4}{\makebox[#1][#2]{#5}}}
     {\fcolorbox{#3}{#4}{\makebox[#1]{#5}}}%
   }
   {\fcolorbox{#3}{#4}{#5}}%
 }

\newcommand{\myfunction}[3]
{${#1} : {#2} \rightarrow {#3}$ }

\newcommand{\myzrfunction}[1]
{\myfunction{#1}{{\myZ}}{{\myR}}}

\newcommand{\mysection}[1]
{\noindent {\bf {#1}}}

%%%%%% Begin document with header and title %%%%%%%%%

\begin{document}
\begin{frontmatter}
% "Title of the paper"
\title{Probability Aggregation: Dynamic Hierarchical Modeling of Expert Beliefs in Geopolitical Events}
\runtitle{Hierarchical Model for Expert Beliefs}

% indicate corresponding author with \corref{}
% \author{\fnms{John} \snm{Smith}\corref{}\ead[label=e1]{smith@foo.com}\thanksref{t1}}
% \thankstext{t1}{Thanks to somebody} 
% \address{line 1\\ line 2\\ printead{e1}}
% \affiliation{Some University}

%\begin{aug}
%  \author{\fnms{First}  \snm{Author}\corref{}\thanksref{t2}\ead[label=e1]{first@somewhere.com}},
%  \author{\fnms{Second} \snm{Author}\ead[label=e2]{second@somewhere.com}}
%  \and
%  \author{\fnms{Third}  \snm{Author}%
%  \ead[label=e3]{third@somewhere.com}%
%  \ead[label=u1,url]{http://www.foo.com}}
%
%  \thankstext{t2}{Footnote to the first author with the `thankstext' command.}
%
%  \runauthor{F. Author et al.}
%
%  \affiliation{Some University and Another University}
%
%  \address{Address of the First and Second authors,\\ 
%          \printead{e1,e2}}
%
%  \address{Address of the Third author,\\
%          \printead{e3,u1}}
%
%\end{aug}

\begin{aug}
\author{\fnms{Ville A.} \snm{Satop\"a\"a}\corref{}\ead[label=e1]{satopaa@wharton.upenn.edu}},
\author{\fnms{Shane T.} \snm{Jensen}\ead[label=e3]{stjensen@wharton.upenn.edu}}
\and
\author{\fnms{Lyle H.} \snm{Ungar}\ead[label=e2]{ungar@cis.upenn.edu}}

\runauthor{Satop\"a\"a et al.}

 \affiliation{Department of Statistics\\ The Wharton School of the University of Pennsylvania   \printead{e1,e3}}

\affiliation{Department of Computer and Information Science\\ University of Pennsylvania, \printead{e2}}
 \address{Philadelphia, PA 19104- 6340, USA\\  \printead{e1}}

\end{aug}

\begin{abstract}
Most research on subjective probability aggregation concentrates on procedures that consult the expert advice only once before making the final aggregation. This paper makes an advance towards unexplored areas of probability aggregation by considering a dynamic context with a time period during which the experts are allowed to update their beliefs. The analysis is performed on a real-world dataset that includes over 2,300 experts making probability forecasts on a number of 166 different international political events. The model is able to take  into account the forecaster's level of self-reported expertise and produce aggregate probabilities that are sharp and well-calibrated both in- and out-of-sample. Using the aggregate probabilities as the no-bias reference-line, the groups of experts with different levels of self-reported expertise are found to be under-confident, with the level of under-confidence decreasing as a function of self-reported expertise.
\end{abstract}

%\begin{keyword}[class=MSC]
%\kwd[Primary ]{}
%\kwd{}
%\kwd[; secondary ]{}
%\end{keyword}

\begin{keyword}
\kwd{Probability Aggregation}
\kwd{Dynamic Linear Model}
\kwd{Hierarchical Modeling}
\kwd{Expert Forecast}
\kwd{Subjective Probability}
\kwd{Bias Estimation}
\kwd{Calibration}
\kwd{Time Series}
\end{keyword}
\end{frontmatter}

%
%\title{Dynamic Hierarchical Model for Belief Streams}
%\author{by\\
%\\
%Ville A. Satop\"a\"a\\
%Shane T. Jensen and Lyle H. Ungar, Faculty Advisors\\
%\\
% \small Department of Statistics\\
% \small The Wharton School of the University of Pennsylvania\\
% \small Philadelphia, PA 19104- 6340, USA\\ [-0.25in]} \date{}
%\maketitle
%
%\pagestyle{myheadings}
%\markboth{Ville A. Satop\"a\"a}{Calibrated Crowd Belief}
%\thispagestyle{empty}

\section{Introduction}
\label{intro}
Individual experts can differ radically from one another in their abilities to assess probabilities of future events. The quality of their assessments is typically evaluated with an important attribute called \textit{calibration}. An expert is said to be well-calibrated if 
the proportion of events that actually occurred is, say, 60\% for all those events for which he assessed a probability of 0.60. 
%the long-run frequency of occurrence of the events agrees with his probability forecasts. This means that, for instance, among all those events for which an expert assesses a probability of  0.60, the proportion of events that actually occur is 60\%.  
Even though several experiments have been conducted to show that experts are generally poorly calibrated (see, e.g., \citet{cooke1991experts, shlyakhter1994quantifying}), some personal differences among the experts have been found to associate with calibration. In particular, \citet{wright1994coherence} argue that a higher self-rated expertise on the subject matter implies better calibration.

%Therefore self-reported expertise should be taken in to account when modeling probability forecasts given by a diverse group of experts. 
Unfortunately, calibration by itself does not guarantee usefulness of a probability assessment. For instance, if the empirical frequency of  the last 10 years of a relatively stationary process is 45\%, the expert can assess a constant probability of 0.45 and turn out well-calibrated. His probability, however, is not very useful in decision making. In general, a probability assessments that is near 0.0 or 1.0, is considered more informative, and an expert who makes such assessments is called \textit{sharp}.
%called \textit{smooth} or \textit{not sharp}. If, on other hand, most of his probabilities are near 0.0 or 1.0, he is deemed \textit{sharp}. 
%This notion leads to a second important attribute called \textit{sharpness}, which is a statement about the distribution of the assessed probabilities. If most of the probabilities are near 0.0 and 1.0, the assessor is said to be sharp.
In the forecasting literature sharpness is often combined with calibration into a single goal of maximizing sharpness subject to calibration (see, e.g., \citet{raftery2005using, Ranjan08}). An expert that meets this goal is considered to give perfect probability forecasts (\citet{gneiting2004verifying}). Even though expecting perfect forecasts from an expert is not reasonable in practice, this criterion has led to a wide range of novel and insightful observations in probability forecasting.

One such observation is related to the aggregation of multiple probabilities. There is strong empirical evidence that bringing together the strengths of different experts by combining their probability forecasts into a single consensus, known as the \textit{crowd belief},  results in better predictive performance. Being motivated by the long list of applications of probability forecasts, including medical diagnosis (\citet{wilson1998prediction, pepe2003statistical}, political and socio-economic foresight (\citet{tetlock2005expert}), and meteorology (\citet{sanders1963subjective, vislocky1995improved, baars2005performance}), researchers have proposed many different approaches to combining probability forecasts (see, e.g., \citet{Ranjan08, satopaa} for some recent studies, and \citet{Genest, Wallsten97evaluatingand, clemen2007aggregating, primo2009calibration} for a comprehensive overview). The general focus, however, has been on developing one-time aggregation procedures that consult the experts' advice only once before the future event resolves.

Therefore many complex areas of probability aggregation still remain rather unexplored. 
%Therefore this paper makes an advance towards rather unexplored areas of probability aggregation by modeling forecasting 
%This paper introduces an aggregation procedure that allows multiple probability assessments from a single expert before the event resolves. 
%The context considered in this paper, on the other hand, allows the experts to update their probability assessments over a period of time before the future event resolves.
%
%considering a dynamic context with a time period during which the experts are allowed to update their beliefs. 
For instance, consider an investor aiming to assess whether a stock index will finish trading at or above some threshold on a given date. In order to maximize his overall predictive accuracy, he may consult a group of experts every day and adjust his estimate of the aggregate probability accordingly. Since the experts are allowed to update their probability assessments on a daily basis, the aggregation should be performed by taking into account the correlation and time-series nature of their advice. Analyzing this kind of future events under daily updating of the forecasts is the main focus of this paper. The core research goal is to construct an interpretable time-series model that incorporates self-reported expertise in order to estimate a sharp and well-calibrated version of the crowd belief. The model is then used for
\begin{itemize}
\item the analysis of group-level under- and overconfidence across different levels of self-reported expertise,
%\item confidence bands for the crowd belief,
\item accurate probability forecasts, and
\item many question-specific quantities that have easy interpretations and can be used to gain novel insight in the social sciences. 
\end{itemize}

The paper begins with a description of our probability forecasting data that were collected by asking over 2,300 experts to give probability forecasts and to self-assess their level of expertise on a number of 166 geopolitical binary events. After summarizing the dataset, the paper introduces a dynamic hierarchical model for capturing the crowd belief. The model is estimated in a two-step procedure: First, constrained parameter estimates are obtained via Gibbs sampling. These estimates are then transformed in to their unconstrained equivalents by a calibrating step. An extension of this model to polychotomous outcomes is briefly discussed before the model evaluation sections. The first evaluation section uses synthetic data to study the accuracy of the two-step procedure to capture parameter values. The second and much longer evaluation section uses the model on our probability forecasting data. The main goal is to illustrate and also to study the specific research objectives itemized above. The paper concludes with a discussion on future research directions and model limitations.

%Two commonly desired attributes of an aggregator are \textit{calibration} and \textit{sharpness}:

%Sharpness, on the other hand, is a statement about the spread of the forecasts. The aggregator is said to be sharp if the resulting forecasts are near 0 and 1. These two attributes are typically combined into a single goal of finding an aggregator that maximizes sharpness subject to calibration (see, e.g., \citet{gneiting2004verifying, raftery2005using}).

%
%
%
%
%
%\noindent
%\\
%Element 1: MAIN THEME: The broad theme is subjective probability aggregation. Use a catchy initial phrase: "Individuals can differ radically from one another in the precision to which they are able to  evaluate the likelihood of a future event happening." This is makes a statement about human cognitive process and does not include any odd terms. \\
%Element 2: WHY CARE: Explain the importance of probability aggregation (in a static context) for decision making. Explain why time-series advice is more realistic. (1 REABSTC-LogN) Lack of previous literature on time-series, subjective probability. (2 REASON)  \\
%Element 3: SUMMARY OF LITERATURE: See the guidelines. We want to to site work done on static case. There is also some work done on repeated forecasts, i.e. see raftery's paper.\\
%Element 4: MAIN CONTRIBUTION: "Our main contribution arises from our abilities ... " Advertise the data agin.\\
%Element 5: WE NEED A CLEAR INDICATION OF THE FOLLOWING:\\
%(a) The core research question: The main goal of this study is to find an interpretable model for detecting a a calibrated crowd belief.\\
%(b) More specific research objectives: With such a model we can make predictions, answer questions about bias-vs-expertise, and evaluate question difficulty. It may be a good idea to use bullet points here for this. \\
%(c) Context: Geopolitcal events. Describe the data we have briefly.\\
%(d) Units of analysis: Group of experts with different self-reported expertise; different questions in the dataset.\\
%Element 6: STRUCTURE OF THE PAPER\\
%%
%\ville{Much of the earlier paper can be included: define calibration and sharpness, normality around the calibrated crowd belief}
%
%\ville{Not studied in time-series - to our knowledge: REQUIRES LITERATURE REVIEW}
%
%\ville{Define calibration}
%
%\ville{Time series allows many more predictions and hence study of calibration -- this is not possible in static context where we would have only 166 predictions. Note that this method can still be applied to any given time point.}
%
%\ville{what are we hoping to learn from the data (applications) and why}
%
%\ville{Structure of the paper: theory, results (insight + precition)}
%
\section{Geopolitical Forecasting Data}
\label{data}
The data collection began with a recruitment of 2,365 forecasters ranging from graduate students to forecasting and political science faculty and practitioners. The recruiting was made from professional societies, research centers, alumni associations, science bloggers, and word of mouth. Requirements included at least a Bachelor's degree and completion of psychological and political tests that took roughly two hours. These measures assessed cognitive styles, cognitive abilities, personality traits, political attitudes, and real-world knowledge. The experts were asked to give probability forecasts (to the second decimal point) and to self-assess their level of expertise (measured on a 1-to-5 scale with 1 = Not At All Expert and 5 = Extremely Expert) on a number of 166 geopolitical binary events taking place between September 29, 2011 and May 8, 2013. Each question was active for a period of time during which the participating experts were allowed to update their forecasts as frequently as they liked without being penalized. The forecasters knew that their probability estimates would be assessed for accuracy using Brier scores\footnote{The Brier score is the squared distance between the probability forecast and the event indicator that equals 1.0 or 0.0 depending on whether the event happened or not, respectively. See \citet{Brier} for the original introduction.}. This incentivized them to report their true beliefs instead of attempting to game the system (\citet{winkler1968good}). In addition to receiving \$150 for meeting minimum participation requirements that did not depend on prediction accuracy, the forecasters also received status rewards for their performance via leader-boards displaying Brier scores for the top 20 forecasters. Since a typical expert participated only in a small subset of the 166 questions, the forecasters are considered indistinguishable conditional on the self-reported level of expertise.
Tables \ref{DataStats} and \ref{ExpertiseTable} provide more relevant summary statistics on the data. Notice that the distribution of the self-reported expertise is skewed to the right
%: some questions involved very few experts who associated themselves with the highest level of expertise. Notice also 
and that some questions remained active longer than others.  For more details on the data and its collection see \citet{ungar2012good}.

To illustrate the nature of the data with some concrete examples, Figures \ref{Example1} and \ref{Example2} show scatterplots of the probability forecasts given for (a) \textit{Will the expansion of the European bailout fund be ratified by all 17 Eurozone nations before 1 November 2011?}, and (b) \textit{Will the Nikkei 225 index finish trading at or above 9,500 on 30 September 2011?}. The points have been jittered slightly to make overlaps visible.
%Overall, Figures \ref{Example1} and \ref{Example2} contain 872 forecasts made by 710 experts and 1,860 forecasts made by 1,376 experts, respectively. 
The darkness of the points are positively associated with the self-reported expertise levels. Given that the European bailout fund was ratified before 1 November 2011 and that the Nikkei 225 index finished trading at around 8,700 on 30 September 2011, the general trend of the probability forecasts converge to the correct answers. There is, however, much disagreement among the experts. This disagreement persists even near the closing dates of the questions. Unfortunately, the data is not only noisy but also quite sparse. As mentioned before, the forecasters were allowed to update their forecast as long as the questions remained active. Updating, however, was done on a very infrequent basis. An expert gave, on average, only 0.0159 forecasts per day. A typical expert, therefore, made a forecast about every 62 days. As a result the average response rate was around 13.5 forecasts per day from a large group of experts.

\vspace{-1em}
\begin{figure}[h!]
\centering
\hspace*{0em} 	\includegraphics[width=  \textwidth]{Figures/LegendExamplePlot} % requires the graphicx package
\vspace{-4.5em}

\hspace{-0.5em}
\subfigure[\textit{Will the expansion of the European bailout fund be ratified by all 17 Eurozone nations before 1 November 2011?}]{
\hspace*{-1em}  \includegraphics[width= 0.49\textwidth]{Figures/ExamplePlot1}
%\hspace*{-1em}  \includegraphics[width= 1.5in]{CalibrationPLATTSPH10}
%\caption{$\beta = 1$, $\alpha = 0$}
\label{Example1}
} \hspace{0.5em}
\subfigure[\textit{Will the Nikkei 225 index finish trading at or above 9,500 on 30 September 2011?}]{
\hspace*{-1em}  \includegraphics[width= 0.49\textwidth]{Figures/ExamplePlot2}
%\caption{\textit{Sample-Then-Calibrate} with logarithmic score}
\label{Example2}
}

\caption[Optional caption for list of figures]{Scatterplots of the probability forecasts given for two questions in our dataset. The shadings represents the self-reported expertise of the forecaster who provided the probability forecast.}
\label{ExamplePlots}
\end{figure}

\begin{table}
\begin{center}
\begin{tabular}{c c c c c c c c }
\hline
Statistic & Min. & $Q_1$ & Median & Mean & $Q_3$ & Max.\\ \hline
\# of Days a Problem is Active  & 4.00 &  35.25 &  72.00 & 106.30 & 145.20 & 418.00 \\
\# of Experts per Problem & 212.0 &   543.2 &  693.5 &  783.7 &  983.2 & 1690.0\\ 
\# Forecasts given by each Expert on a Problem & 1.0 &  1.0 &   1.0  &  1.8 &  2.0 & 131.0 \\
\# Problems participated by an Expert & 1.0 &  14.0 &   36.0  &  55.0 &  90.0 & 166.0 \\ \hline
\end{tabular}
\caption{Five-number summaries of the real-world data used in this paper: \textit{\# of Days Active} represents the number of days that the experts were given to update their forecast before the correct answer was revealed; \textit{\# of Experts per Problem } denotes the number of experts participating in a question; \textit{\# Forecasts given by each Expert on a Problem} gives the number of forecasts for a question made by an active expert; \textit{\# Problems participated by an Expert} represents the number of different questions a given expert participated in.}
\label{DataStats}
\end{center}
\end{table}

\begin{table}
\begin{center}
\begin{tabular}{c c c c c c c c }
\hline
Expertise Level & 1 & 2 & 3 & 4 & 5\\
Frequency (\%) & 25.3 & 30.7 & 33.6 &  8.2 &  2.1 \\\hline
\end{tabular}
\caption{Frequencies of the expertise levels across all the 166 questions in the dataset.}
\label{ExpertiseTable}
\end{center}
\end{table}

%
% For instance, two of the questions were
%\begin{quote}\textit{
%Will the expansion of the European bailout fund be ratified by all 17 Eurozone nations before 1 November 2011? %1021
%}\end{quote}
%and
%\begin{quote}\textit{
%Will the Nikkei 225 index finish trading at or above 9,500 on 30 September 2011? %1011
%}
%\end{quote}

\section{Model}
\label{model}
Let $p_{i,t,k} \in (0,1)$ be the probability forecast given by the $i$th expert at time $t$ for the $k$th question, where $i = i, \dots, I_k$, $t = 1, \dots, T_k$, and $k = 1, \dots, K$. Denote the logit-probabilities with $Y_{i,t,k} = \logit(p_{i,t,k}) = \log \left( \frac{p_{i,t,k}}{1-p_{i,t,k}}\right) \in \R$ and collect the logit-probability forecasts given for question $k$ at time $t$ into a vector $\boldsymbol{Y}_{t,k} = [Y_{1,t,k}\ Y_{2,t,k}\ \dots\ Y_{I_{k},t,k}]^T$. Partition the experts into $J$ groups based on some individual feature, such as self-reported expertise, with each group sharing a common multiplicative bias term $b_{j} \in \R$ for $j = 1, \dots, J$. Collect these bias terms into a bias vector $\boldsymbol{b} = [b_{1}\ b_{2}\ \dots\ b_{J}]^T$ and let $\boldsymbol{M}_k$ be a $I_k \times J$ matrix denoting  the group-memberships of the forecasters in question $k$. That is, if the $i$th expert participating in the $k$th question belongs to the $j$th group, then the $i$th row of $\boldsymbol{M}_k$ is the $j$th standard basis vector, $\boldsymbol{e}_j$. The bias vector $\boldsymbol{b}$ does not include a subindex because it is considered shared among all the $K$ questions. To secure model identifiability, it is sufficient to share only one of the elements of $\boldsymbol{b}$ among the questions. This element defines a baseline under which it is possible to estimate the remaining $J-1$ bias terms separately within each of the questions. In this paper, however, the entire vector $\boldsymbol{b}$ is shared because some of the questions in our real-world data set involve very few experts with the highest level of self-reported expertise. Under this notation, the model for the $k$th question can be expressed as 
\begin{eqnarray}
\boldsymbol{Y}_{t, k} &=&  \boldsymbol{M}_k \boldsymbol{b} X_{t, k} + \boldsymbol{v}_{t, k} \label{observedpr} \\
X_{t, k} &=& \gamma_k X_{t-1, k} + w_{t, k} \label{hiddenpr}\\
X_{0,k} &\sim& \mathcal{N}(\mu_0, \sigma^2_0) \nonumber
\end{eqnarray}
where Equation (\ref{observedpr}) denotes the observed process, Equation (\ref{hiddenpr}) shows the hidden process that is driven by the constant $\gamma_k \in \R$, the error terms are independent and identically distributed normal random variables with mean zero
\begin{eqnarray*}
\boldsymbol{v}_{t, k} | \sigma^2_k &\stackrel{i.i.d.}{\sim}& \mathcal{N}_{I_k}(\boldsymbol{0}, \sigma^2_k \boldsymbol{I}_{I_k})\\
w_{t, k} | \tau^2_k &\stackrel{i.i.d.}{\sim}& \mathcal{N}(0, \tau^2_k),
\end{eqnarray*}
and $(\mu_0, \sigma_0^2) \in (\R, \R^+)$ are hyper-parameters chosen a priori. The hidden state $X_{t,k}$ represents the sharp and calibrated logit-probability for the $k$th event given the information available up to and including time $t$. To make this more specific, let $Z_k \in \{0, 1\}$ indicate whether the event associated with the $k$th question happened $\left(Z_k = 1\right)$ or did not happen $\left(Z_k = 0\right)$. If $\{\mathcal{F}_{t,k}\}_{t=1}^{T_k}$ is a filtration representing the information available up to and including a given time point, then $\E[Z_k | \mathcal{F}_{t,k}] = \P(Z_k = 1| \mathcal{F}_{t,k}) = \logit^{-1}(X_{t,k})$. These probabilities are made estimable by assuming that the observed process is associated with the filtration $\{\mathcal{F}_{t,k}\}_{t=1}^{T_k}$ and hence can be used as a proxy for the amount of information available at any given time point. Therefore it is not unreasonable to interpreted the hidden process as the crowd belief across time. 
%Each modeling task should begin with observations about the nature of the question. With a good understanding of the question, the modeler can proceed to construct a model that has an interpretable and intuitive connection to the real world.

To give intuitive interpretations of the other model parameters, notice that the error term in the observed process has mean zero. Therefore the forecasters are assumed to be, on average, a multiplicative constant $\boldsymbol{b}$ away from the calibrated crowd belief. An individual element of $\boldsymbol{b}$ can be interpreted as a group-specific \textit{systematic bias} that labels the group either as over-confident $\left( b_j \in (1, \infty)\right)$ or as under-confident $\left(b_j \in (0,1)\right)$. The bias terms are analyzed and discussed more in Section \ref{ExpertBias}. Any other deviation from the calibrated crowd belief is considered \textit{random noise}. This noise is measured in terms of $\sigma^2_k$ and can be assumed to be caused by momentary over-optimism (or pessimism), false beliefs, or other misconceptions.

 Similarly to the observed process, the hidden process  has a systematic and a random component. The \textit{random fluctuations} are measured in terms of $\tau^2_k$ and can be assumed to represent changes or shocks to the underlying circumstances that ultimately decide the outcome of the event. The \textit{systematic component} $\gamma_k$, on the other hand, allows the model to incorporate a constant signal stream that drifts the hidden process to infinity (when $\gamma_k \in (1, \infty)$) or zero (when $\gamma_k \in (0, 1)$). If the uncertainty in the question diminishes as the current time point $t$ approaches $T_k$, the hidden process drifts to infinity. Alternatively, the hidden process can drift to zero in which case any available information about the target event does not improve predictive accuracy. Since each of the $K$ questions in our dataset was resolved within a pre-specified timeframe, $\gamma_k$ is expected to fall within the interval $(1, \infty)$ for all $k = 1, \dots, K$.

%
%\textit{Random noise, $\sigma^2_k$:} The variability of each observed value $Y_{i,t,k}$ around the biased hidden state $b_j \tilde{X}_{t,k}$. This can be thought of as being caused by momentary over-optimism (or pessimism), false beliefs, or other misconceptions. 
%Note that only the most informed expert making an objective estimation based solely on facts can be expected to give $X_t$ as his prediction. Unfortunately, human experts hardly make such objective estimates. Therefore this noise component is an important part of our model. 
%
%\textit{Random signal, $\tau^2_k$:} This variability is caused by changes in the amount of true information that is available to the experts. It allows the model to incorporate changes or shocks to the underlying circumstances that ultimately decide the outcome of the event. 
%
%\textit{Systematic noise, $b_{j}$:} This is a bias in the observed values given by the $j$th most expert group and can be interpreted as groupwise over- or under-confidence that pulls the observed values away from the hidden state values. 
%
%\textit{Systematic signal, $\gamma_k$:} Since the event associated with the $k$th question in our dataset realizes by time $T_k$, the uncertainty in the question reduces as the current time point $t$ approaches the final time point $T_k$. This reduction in uncertainty can be viewed as a constant signal stream that pushes the hidden process towards either extreme as more time passes by.

%These interpretations have been summarized in Table \ref{signalnoise}.
%\begin{table}
%\centering
%\begin{tabular}{| c | c |  c | }\hline
% Model Factors & Signal & Noise \\ 
%Random & $\tau^2_k$ & $\sigma^2_k$ \\ 
%Systematic & $\gamma_k$ & $b_{j,k}$ \\ \hline
%\end{tabular}
%\caption{The signal/noise structure of the dynamic linear model}
%\label{signalnoise}
%\end{table}

\section{Model Estimation}
\label{identifiability}
%Unfortunately, the model as described by Equations (\ref{observedpr}) and (\ref{hiddenpr}) is not identifiable. To see this, let $a$ be any positive constant. Then $(\boldsymbol{b} a, X_{t,k}/a, a \tau) \neq (\boldsymbol{b}, X_{t,k}, \tau)$ yield the same likelihood. As a result, the parameter values tend to drift during the estimation process.  Several solutions are as follows.
%
%\begin{enumerate}
%\item[a.]  Place informative prior distributions on some of the non-identifiable parameters. These prior distributions, however, must be highly informative in order to provide identifiability. Since studying the expertise biases is one of the research objectives of this paper, placing a highly restrictive prior distribution on $\boldsymbol{b}$ is not satisfactory. Having strong apriori knowledge of $\tau_k$, on the other hand, seems difficult to justify.  
%\item[b.] The parameters could be constrained to reduce the number of degrees of freedom in the model. A popular choice is to fix one of the bias terms to some constant. For instance, the third expertise group could be chosen as the baseline group by fixing $b_3 = 1$. This, however, assumes that the third expertise group is unbiased. Such an assumption is hardly realistic and would hinder our analysis of under- and overconfidence among the expertise groups. 
%\item[c.] Reduce the degrees of freedom by introducing more information to the modeling process. 
% \end{enumerate}
% 
%The next two subsections introduce two hierarchical approaches that combine solutions b. and c. to restore identifiability without sacrificing the interpretability of the model. The first approach, called  \textit{Sample-Then-Calibrate}, is flexible and results into an efficient sampling algorithm. The second approach, named \textit{Sample-Only}, on the other hand, produces uncertainty estimates for all the unknown parameters in a coherent, fully Bayesian manner.

%

The most challenging part of our model estimation is to incorporate the event indicators $\{Z_k\}_{k=1}^K$ such that the hidden process is well-calibrated and the interpretability of our model is not compromised. This section introduces a two-step procedure, called \textit{Sample-Then-Calibrate} or simply \textit{STC}, that achieves this goal in a flexible, yet efficient manner by first estimating  the model parameters under a constraint (\textit{Sampling Step}) and then performing a one-dimension optimization procedure to transform the constrained estimates into their unconstrained counterparts (\textit{Calibration Step}).

%The second approach, named \textit{Sample-Only} (SO), on the other hand, produces uncertainty estimates for all the unknown parameters in a fully Bayesian manner.

\subsection{Sampling Step}
\label{sampling_step}
Since $\left(a \boldsymbol{b}, X_{t,k}/a, a^2 \tau_k^2\right) \neq \left(\boldsymbol{b}, X_{t,k}, \tau_k^2\right)$ for any $a > 0$ yield the same  likelihood for $\boldsymbol{Y}_{t,k}$, the model as described by Equations (\ref{observedpr}) and (\ref{hiddenpr}) is not identifiable and the parameter estimates tend to drift during the sampling process. A well-known solution is to choose one of the elements of $\boldsymbol{b}$, say $b_3$, as the reference point and fix $b_3 = 1$.  Denote the constrained version of the model by
\begin{eqnarray*}
\boldsymbol{Y}_{t, k} &=&  \boldsymbol{M}_k \boldsymbol{b} (1) X_{t, k}(1)+ \boldsymbol{v}_{t, k}\\
X_{t, k}(1) &=& \gamma_k(1) X_{t-1, k}(1) + w_{t, k}\\
\boldsymbol{v}_{t, k} | \sigma^2_k(1) &\stackrel{i.i.d.}{\sim}& \mathcal{N}_{I_k}(\boldsymbol{0}, \sigma^2_k(1) \boldsymbol{I}_{I_k})\\
w_{t, k} | \tau^2_k(1) &\stackrel{i.i.d.}{\sim}& \mathcal{N}\left(0, \tau^2_k(1)\right),
\end{eqnarray*}
where the trailing input parameter emphasizes the constraint $b_{3} = 1$. Since this version is identifiable, estimates of the model parameters can be obtained. Denote the estimates by placing a hat on the parameter symbol. For instance, $\hat{\boldsymbol{b}}(1)$ and $\hat{X}_{t, k}(1)$ represent the estimates of $\boldsymbol{b}(1)$ and $X_{t, k}(1)$, respectively. These estimates are found via a Gibbs sampler that only makes use of standard distributions. This procedure is standard practice in Bayesian statistics and is therefore not described in great detail here. See Appendix \ref{appendix} for the technical details of our sampler, and, e.g., \citet{gelman2003bayesian} for a discussion on the general principles of Gibbs sampling.

%See Appendix \ref{appendix} for the technical details.

\subsection{Calibration Step}
\label{calibration_step}
%First choose one of the $J$ groups as the baseline group. Denote this group by $j^*$ and fix its multiplicative bias term $b_{j^*} = 1$.  Denote the constrained version of the observed process by
%\begin{eqnarray*}
%\boldsymbol{Y}_{t, k} &=&  \boldsymbol{M}_k \boldsymbol{b} (1) X_{t, k}(1)+ \boldsymbol{v}_{t, k},
%\end{eqnarray*}
%where the input parameter emphasizes the constraint $b_{j^*} = 1$. Since this constrained model is identifiable, estimates of $\boldsymbol{b}(1)$ and $X_{t, k}(1)$ can be obtained. Denote these estimates by $\hat{\boldsymbol{b}}(1)$ and $\hat{X}_{t, k}(1)$, respectively. 
Since the model parameters can be estimated under a constrained version of $b_3$, the next step is to determine how $b_3$ should be constrained such that both the sharpness and calibration of the estimated hidden process are maximized. This section introduces an efficient procedure that finds the optimal constraint without requiring any additional runs of the sampling step. First, assume that parameter estimates  $\hat{\boldsymbol{b}}(1)$ and $\hat{X}_{t, k}(1)$ have already been obtained via the constrained sampling step described in Section \ref{sampling_step}. Since for any $\beta \in \R / \{0\}$,
\begin{eqnarray*}
\boldsymbol{Y}_{t, k} &=&  \boldsymbol{M}_k \boldsymbol{b} (1) X_{t, k}(1)+ \boldsymbol{v}_{t, k}\\
 &=&   \boldsymbol{M}_k \left(\boldsymbol{b} (1) \beta \right) \left(X_{t, k}(1)/\beta\right)  + \boldsymbol{v}_{t, k}\\
&=&   \boldsymbol{M}_k \boldsymbol{b} (\beta) X_{t, k}(\beta) + \boldsymbol{v}_{t, k},
\end{eqnarray*}
the parameter values under $b_{3} = \beta$ can be obtained from $\boldsymbol{b} (\beta) = \boldsymbol{b} (1) \beta $ and $X_{t, k}(\beta) = X_{t, k}(1)/\beta$. 
%Furthermore, an estimate of $X_{t, k}(\beta)$ is given by $\hat{X}_{t, k}(1)/\beta$. 
This means that $X_{t,k} = X_{t,k}(1)/\beta$ when $\beta$ is equal to the true value of $b_3$. Since the hidden process $X_{t, k}$ is assumed to be sharp and well-calibrated, $b_3$ can be estimated by the value of $\beta$ that simultaneously maximizes the sharpness and calibration of $\hat{X}_{t, k}(1) / \beta$. A natural criterion for this maximization is provided by the class of \textit{proper scoring rules} that combine sharpness and calibration (\citet{gneiting2008rejoinder, buja2005loss}). Due to the possibility of \textit{complete separation} in any one question (see, e.g., \citet{gelman2008weakly}), the maximization must be performed over multiple questions. Therefore,
\begin{eqnarray}
%\hat{\beta}_{OSE} &=& \argmax_{\beta \in \R / \{0\}} \sum_{k=1}^K \frac{1}{T_k}  \sum_{t=1}^{T_k} \int_{\R} S(Z_k,  X_{k,t}(\beta)) d\P(X_{k,t} | \boldsymbol{Y}) \nonumber\\
%&\approx& \argmax_{\beta \in \R / \{0\}} \sum_{k=1}^K \frac{1}{T_k} \sum_{t=1}^{T_k}   \frac{1}{G} \left( \sum_{g=1}^{G}  S\left(Z_k,  \hat{X}_{k,t}^{(g)}(\beta) \right) \right) \label{OSE}
 \hat{\beta}&=&  \argmax_{\beta \in \R / \{0\}} \sum_{k=1}^K \sum_{t=1}^{T_k}  S\left(Z_k, \hat{X}_{k,t}(1) / \beta \right) \label{OSE}
\end{eqnarray}
where $S$ is a strictly proper scoring rule such as 
%that is a natural choice for binary-classification questions (\citet{buja2005loss}) and can be conveniently adapted to the question at hand (\citet{Gneiting04strictlyproper}). 
the negative Brier score (\citet{Brier})
\begin{eqnarray*}
S_{BRI}(Z, X) &=& -(Z - \logit^{-1}(X))^2
\end{eqnarray*}
or the logarithmic score (\citet{good1952rational})
\begin{eqnarray*}
S_{LOG}(Z, X) &=& Z \log\left(\logit^{-1}(X)\right) + (1-Z) \log\left(1-\logit^{-1}(X)\right)
\end{eqnarray*}
%The proper scoring rule can be conveniently adapted to the question at hand (\citet{Gneiting04strictlyproper}). 
%Notice that Equation (\ref{OSE}) under $S_{LOG}$ is equivalent to finding the maximum likelihood estimator of the (inverse) slope coefficient for a (centered) logistic regression model with $Z_k$ as the response and $\hat{X}_{k,t}$ as the explanatory variable.
%Since the expected utility under a strictly proper scoring rule for an observation $Z \sim p$ is uniquely maximized by forecasting exactly $p$ (\citet{Gneiting04strictlyproper}), the estimated log-odds $\hat{X}_{t, k}(1) / \hat{\beta}_{OSE}$ can be expected to be well-calibrated. 
Since it is not clear which rule should be used for predicting geopolitical events, such as the ones in our dataset, the \textit{Sample-Then-Calibrate} procedure is evaluated  separately under both rules in Sections \ref{syntheticData} and \ref{realData}. Once $\hat{\beta}$ has been computed, estimates of the unconstrained model parameters are given by
\begin{eqnarray}
 \hat{X}_{t,k}&=& \hat{X}_{k,t}(1) / \hat{\beta} \nonumber\\
 \hat{\boldsymbol{b}}_{t,k}&=& \hat{\boldsymbol{b}}(1) \hat{\beta} \nonumber\\
   \hat{\tau}_{k}^2&=& \hat{\tau}_{k}^2(1) \hat{\beta}^2\nonumber\\
  \hat{\sigma}_{k}^2&=& \hat{\sigma}_{k}^2(1)\nonumber\\
  \hat{\gamma}_{k}&=& \hat{\gamma}_{k}(1)\nonumber
\end{eqnarray}
Notice that the parameters $\sigma^2_k$ and $\gamma_k$ are not affected by the constraint. Therefore their constrained and unconstrained estimates are the same.

%\subsection{Sample-Only}
%In order to clarify the upcoming notation, denote the calibrated logit-probabilities and  the event indicators across all $K$ questions with $\boldsymbol{X}$ and $\boldsymbol{Z}$, respectively. A Bayesian version of Equation (\ref{OSE}) under $S_{LOG}$ can be derived by specifying a prior distribution $p(\beta | \boldsymbol{X})$ and a likelihood function $p(\boldsymbol{Z} | \beta, \boldsymbol{X})$. The posterior distribution of $\beta$ conditional on $\boldsymbol{X}$ is then given by $p(\beta | \boldsymbol{X}, \boldsymbol{Z}) \propto p( \boldsymbol{Z} | \beta, \boldsymbol{X}) p(\beta | \boldsymbol{X})$. Recall that Equation (\ref{OSE}) under $S_{LOG}$ is equivalent to fitting a logistic regression model with $Z_k$ as the response and $X_{k,t}$ as the explanatory variable. To give a Bayesian version of this logistic regression, the likelihood is chosen to be proportional to
%\begin{eqnarray}
%p( \boldsymbol{Z} | \beta, \boldsymbol{X}) &\propto& \prod_{k=1}^K \prod_{t=1}^{T_k}  \logit^{-1} \left(X_{k,t}/\beta  \right)^{Z_k} \left( 1-  \logit^{-1} \left( X_{k,t}/\beta  \right) \right)^{1-Z_k} \label{OSE2}
%\end{eqnarray}
%Similarly to  \citet{gelman2003bayesian} the prior is chosen to be locally uniform and independent, $p(1/\beta) \propto 1$. Posterior estimates of $\beta$ can be sampled from Equation (\ref{OSE2}) using generic sampling algorithms such as the Metropolis algorithm (\citet{metropolis1953equation}) or slice sampling (\citet{neal2003slice}).

%Notice that with this procedure it is not necessary to estimate $\boldsymbol{b}$ under a constraint. Instead we could let $p(Z_k | X_{t,k}) \propto \logit^{-1} \left(X_{k,t}  \right)^{Z_k} \left( 1-  \logit^{-1} \left( X_{k,t}  \right) \right)^{1-Z_k}$ independently across all time points and questions similarly to Equation (\ref{OSE2}), and estimate the elements of the bias vector $\boldsymbol{b}$ directly. Even though this would make $\beta$ unnecessary, the resulting procedure, which is equivalent to the version with $\beta$, does not make a clear connection to the \textit{Sample-Then-Calibrate} approach.

%Overall, this approach can be viewed as a Bayesian equivalent of Equation (\ref{OSE}) under the logarithmic score, $S_{LOG}$.
%This assumes an independent $Z_k$ for each time point in the $k$th question. This is a coarse approximation since these indicators are, in fact, perfectly correlated. It does, however, reinforce good calibration and is therefore of practical use.

\subsection{Discussion}
If the class labels in the data are balanced with respect to the time points, the calibration step with the logarithmic scoring rule is approximately equivalent to \textit{Platt calibration}, which has been shown to yield good calibration under various modeling scenarios (see, e.g., \citet{platt1999probabilistic, niculescu2005obtaining}). To see this, recall that the Platt calibrated logit-probabilities are given by $\hat{A} + \hat{B}\hat{X}_{t,k}(1)$, where 
\begin{eqnarray}
\left(\hat{A}, \hat{B} \right) =  \argmax_{A, B \in \R} \sum_{k=1}^K \sum_{t=1}^{T_k}  S_{LOG}(Z_k, A + B\hat{X}_{t,k}(1) ) \label{platt}
\end{eqnarray}
This is equivalent to fitting a logistic regression model with $Z_k$ as the response and $\hat{X}_{t,k}(1)$ as the explanatory variable.
%model with $Z_k$ as the response and $\hat{X}_{t,k}(1)$ as the explanatory variable. 
To understand the behavior of the coefficients $A$ and $B$, express the logistic regression as linear regression
\begin{eqnarray*}
\logit(\P(Z_k = 1 | \hat{X}_{t,k})) = A + B\hat{X}_{t,k} + e_{t,k}
\end{eqnarray*}
with $e_{t,k} \stackrel{i.i.d.}{\sim} \mathcal{N}(0,\sigma^2)$. If the data are balanced with respect to the time points, then exactly half of the summands in Equation (\ref{platt}) have $Z_k = 1$ and the average response logit-probability $\logit(\P(Z_k = 1 | \hat{X}_{t,k})) $ is close to zero. Since the values of $\hat{X}_{t,k}$ are estimated logit-probabilities of the same $K$ events across different time points, their overall average is also around zero. Therefore both the response and explanatory variables are approximately centered. This means that the intercept term $A$ is near zero reducing Platt calibration to Equation (\ref{OSE}) under the logarithmic scoring rule. If the data is not balanced, Platt calibration can be easily incorporated into our model via an additional intercept parameter. This, however, reduces the interpretability of our model. Fortunately, compromising interpretability is rarely necessary because it is often possible to use the data in a well-balanced form. One procedure to attain this is described in the beginning of Section \ref{realData}.

%In terms of the prior distributions, we assume that $\alpha$ and $\beta$ are a priori independent with 
%\begin{eqnarray*}
%\beta &\sim& \text{Cauchy}(1, 1/2.5)\\
%\alpha &\sim& \text{Cauchy}(0, 10),
%\end{eqnarray*}
%which are essentially the weakly informative default priors introduced in \citet{gelman2008weakly}. The only difference is in the centering of the prior $p(\beta)$. \citet{gelman2008weakly} recommend centering this prior at 0. Since a priori we have no reason to assume under- or overconfidence of the $\tilde{X}_{t,k}$ process, the prior $p(\beta)$ should be centered at 1 instead. 
%
%Unlike the \textit{Sample-Then-Calibrate} approach, the \textit{Sample-Only} approach models the uncertainty in estimating $(\beta, \alpha)$. The \textit{Sample-Only} approach also results into an algorithm that produces all estimates from a single procedure; hence, making it appear more coherent.  The \textit{Sample-Then-Calibrate} approach, on other hand, enjoys the flexibility to choose a score function, $S$, that suits the application at hand (see \citet{Gneiting04strictlyproper}). Furthermore, the  \textit{Sample-Then-Calibrate} approach results into a slightly more efficient sampling procedure.

%Even though both approaches assume a hierarchical model with a common $\boldsymbol{b}$ across the questions,

\section{Extension: Polychotomous Outcomes}
If the future event can take upon $M > 2$ possible outcomes, the hidden state $X_{t,k}$ must be extended to a vector of size $M-1$. One of the outcomes, e.g., the $M$th one, is chosen as the base-case.  This gives us a total of $M-1$ observed processes and one hidden process
\begin{eqnarray*}
\boldsymbol{Y}_{1,t,k} &=&  \boldsymbol{M}_k\boldsymbol{b}_{1}X_{1,t,k} + \boldsymbol{v}_{1,t,k}\\
\boldsymbol{Y}_{2,t,k} &=&  \boldsymbol{M}_k\boldsymbol{b}_{2}X_{2,t,k} + \boldsymbol{v}_{2,t,k}\\
&\vdots&\\
\boldsymbol{Y}_{M-1,t,k} &=&  \boldsymbol{M}_k\boldsymbol{b}_{M-1}X_{M-1,t,k} + \boldsymbol{v}_{M-1,t, k}\\
\boldsymbol{X}_{t,k} &=& \boldsymbol{\gamma}^T_k \boldsymbol{I}_{M-1} \boldsymbol{X}_{t-1,k} + \boldsymbol{w}_{t,k}
\end{eqnarray*}
where $\boldsymbol{X}_{t,k}$ is a $(M-1) \times 1$ matrix of calibrated logit-probabilities. Notice that each process has a separate bias vector. The error terms are independent and identically distributed multivariate normal random variables 
\begin{eqnarray*}
\boldsymbol{v}_{m, t, k} | \sigma^2_{j,k} &\stackrel{i.i.d.}{\sim}& \mathcal{N}_{I_k} \left(\boldsymbol{0}, \sigma^2_{j,k} \boldsymbol{I}_{I_k} \right)\\
\boldsymbol{w}_{t,k} | \tau_k^2 &\stackrel{i.i.d.}{\sim}& \mathcal{N}_{M-1} \left(\boldsymbol{0}, \tau_k^2 \boldsymbol{I}_{M-1} \right)
\end{eqnarray*}
On the left-hand side of observed process
\begin{eqnarray*}
\boldsymbol{Y}_{m,t,k} &=&  \left[ \begin{matrix} Y_{m,1,t,k} & Y_{m,2,t,k} & \dots & Y_{m,I_k,t,k} \end{matrix} \right]^T\\
&=&  \left[ \begin{matrix} \log \left( \frac{p_{m,1,t,k}}{p_{M,1,t,k}} \right) &  \log \left( \frac{p_{m,2,t,k}}{p_{M,2,t,k}} \right) & \dots &  \log \left( \frac{p_{m,I_k,t}}{p_{M,I_k,t,k}} \right) \end{matrix} \right]^T
\end{eqnarray*}
is a $I_k \times 1$ matrix of the expert logit-probabilities for the $m$th outcome of the $k$th question at time $t$, and on the left-hand side of the hidden process
\begin{eqnarray*}
\boldsymbol{X}_{t,k} &=&  \left[ \begin{matrix}X_{1,t,k} &X_{2,t,k} & \dots &X_{M-1,t,k} \end{matrix} \right]^T
%&=&  \left[ \begin{matrix} X_{1,t,k} + \alpha_1 & X_{2,t,k} + \alpha_2  & \dots & X_{M-1,t,k} + \alpha_{M-1}  \end{matrix} \right]^T\\
%&=&  \left[ \begin{matrix} \log \left( p_{1,t,k}/p_{M,t,k} \right) & \log \left( p_{2,t,k}/p_{M,t,k} \right) & \dots & \log \left( p_{M-1,t,k}/p_{M,t,k} \right) \end{matrix} \right]^T
\end{eqnarray*}
is the matrix of sharp and calibrated logit-probabilities at time $t$. Choosing one of the outcomes as the base-case, ensures that the probabilities will sum to one at any given time point. Since this multinomial extension is equivalent to having $M-1$ independent binary-outcome models (see Section \ref{model}), the estimation can be done separately for each outcome. Therefore, even though this paper focuses on modeling binary events, all the properties and discussion generalize trivially to the multi-outcome case.

%\subsection*{Incomplete Data or Missing Observations}
%If the observation $\boldsymbol{Y}_{t,k}$ is completely missing at time $t$, the update step is skipped and the state estimates are sampled from
%\begin{align*}
%\mathcal{N}(\gamma_k \hat{X}_{t-1|t-1}, \gamma^2 P_{t-1|t-1} + \tau^2)
%\end{align*}
%If, on the other hand, only some subset of the experts report their forecast at time $t$, then the observation vector $\boldsymbol{Y}_t$ has a smaller dimension. Fortunately, the formulae derived in the previous subsection remain valid even if the dimension of the observation vector is allowed to change; see, e.g., Theorem 2.1 in \textit{Kalman filter with outliers and missing observations} by Cipra and Romera.

\section{Synthetic Data Results \ville{TODO}}
\label{syntheticData}

The synthetic data is not generated directly from the model for several reasons: (i) showing good performance on a dataset directly generated from the model assumptions is hardly any news, and (ii) generating data from the dynamic model description does not produce calibrated hidden states. The latter is important for our interpretation of the hidden states as the calibrated crowd belief. In order to achieve this, let $\{ Z_{t,k} \}_{t=1}^{T_k}$ denote the values of a standard Brownian motion at time points $t = 1, \dots, T_k, T_k $. Then,
\begin{eqnarray*}
Z_k &=& \mathbbm{1}\left( Z_{T_k, k}  > 0\right)\\
 X_{t,k} &=& \logit \left[ \Phi \left(\frac{Z_{t, k}}{\sqrt{T_k- t}} \right) \right]
\end{eqnarray*}
Once the hidden process has been generated, the expert logit forecasts are produced by adding white noise to the given sequences of $X_{t,k}$.

The simulation is performed over a four-dimensional grid of values: The number of questions, $K$, varies between $10, 25, 50, 75$, and $100$, the values of $\beta$ change from $0.5$ to $1.5$ with steps of size 0.25, the values of $\alpha$ range from $-0.5$ to $0.5$ with steps of size 0.25, and finally the values of $\sigma^2$ vary from $0.25$ to $1.5$ with steps of size 0.25. At each grid point the simulation runs a total of two times; each time measuring how well the approaches estimate the hidden process in terms of quadratic loss. The average loss from the two runs is used as the performance measure at that grid point. Since the final results are summarized by choosing one of the grid variables (e.g. $\alpha$) and then averaging over the remaining three variables (e.g. $\beta$, $K$, and $\sigma^2$), each average value in the final results represents an average of a $2 \times 5^3 = 250$ values.

\begin{figure}[h!]
\centering
\hspace*{2em} 	\includegraphics[width=  \textwidth]{Figures/LegendSynthetic} % requires the graphicx package
\vspace{-3em}

\subfigure[]{
\hspace*{-1em}  \includegraphics[width= 0.33\textwidth]{Figures/SyntheticSigma2}
%\hspace*{-1em}  \includegraphics[width= 1.5in]{CalibrationPLATTSPH10}
%\caption{$\beta = 1$, $\alpha = 0$}
\label{SyntheticSigma2}
}
\subfigure[]{
\hspace*{-1em} \includegraphics[width=  0.33\textwidth]{Figures/SyntheticBeta}
%\caption{\textit{Sample-Then-Calibrate} with Brier score}
\label{SyntheticBeta}
}
\subfigure[]{
\hspace*{-1em} \includegraphics[width=  0.33\textwidth]{Figures/SyntheticK}
%\caption{\textit{Sample-Only} }
\label{SyntheticK}
}

\vspace{-2em}
\subfigure[]{
\hspace*{-1em}  \includegraphics[width= 0.33\textwidth]{Figures/SyntheticSigma2bias}
%\hspace*{-1em}  \includegraphics[width= 1.5in]{CalibrationPLATTSPH10}
%\caption{$\beta = 1$, $\alpha = 0$}
\label{SyntheticSigma2}
}
\subfigure[]{
\hspace*{-1em} \includegraphics[width=  0.33\textwidth]{Figures/SyntheticBetabias}
%\caption{\textit{Sample-Then-Calibrate} with Brier score}
\label{SyntheticBeta}
}
\subfigure[]{
\hspace*{-1em} \includegraphics[width=  0.33\textwidth]{Figures/SyntheticKbias}
%\caption{\textit{Sample-Only} }
\label{SyntheticK}
}

\caption[Optional caption for list of figures]{Comparing the approaches on synthetic data. (a) Performance under different values of $\sigma^2$. (b) Performance under different values of $\alpha$. (c) Performance under different values of $\beta$. (d) Performance under different numbers of questions, $K$.}
\label{Synthetic}
\end{figure}

Figure \ref{Synthetic} summarizes the results by choosing one grid variable at a time and then averaging over the remaining three variables. The \textit{Sample-Then-Calibrate} with log loss outperforms the other two approaches almost uniformly across the grid.

\section{Geopolitical Data Results}
\label{realData}
\noindent
This section discusses results related to the real-world data described in Section \ref{data}. The goal is to provide application specific insight by discussing the specific research objectives itemized in Section \ref{intro} and also to evaluate the \textit{Sample-Then-Calibrate}-procedure in terms of predictive power and calibration. Before presenting the results, however, two practical matters, which need to be taken into account when aggregating real-world  probability forecasting data, are discussed.

The first matter regards human experts making probability forecasts of 0.0 or 1.0 even when they are not completely sure of the outcome of the future event. For instance, all the 166 questions in our dataset contained both a zero and a one. Transforming such forecasts into the logit-space yields infinities that are highly influential and cause problems in model estimation. To avoid this, \citet{Ariely00theeffects} suggest changing $p = 0.0$ and $1.0$ to $p = 0.02$ and $0.98$, respectively. This is similar to winsorising which sets the extreme probabilities to a specified percentile of the data (see, e.g., \citet{hastings1947low} for more details on winsorising).  \citet{Geo}, on the other hand, consider only probabilities that fall within a constrained interval, say $[0.001, 0.999]$, and discard the rest. Since this implies ignoring a portion of the data, we decided to adopt the first approach by changing $p = 0.0$ and $1.0$ to $p = 0.01$ and $0.99$, respectively.

%Given that the experts were allowed to report their probability forecasts to the second decimal place, this compromise was considered rather natural.

The second matter is related to the distribution of the class labels in the dataset. If the set of events that occurred is much larger than the set of events that did not occur (or \textit{vice versa}), the dataset is called \textit{imbalanced}. On such data the model can end up over-focusing on the larger class, and as a result, give very accurate forecast performance over the larger class at the cost of performing poorly over the smaller class (see, e.g., \citet{chen2009learning, wallace2012class}). Fortunately, it is often possible to detect a well-balanced form of the dataset. The first step is to find a partition, $S_0$ and $S_1$, of the question indices, $\{1, 2, \dots, K\}$, such that the equality $\sum_{k \in S_0} T_k  = \sum_{k \in S_1} T_k$ is as closely approximated as possible. This is equivalent to an NP-hard problem known in computer science as the \textit{Partition Problem}:  determine whether a given set of positive integers can be partitioned into two sets such that the sums of the two sets equal to each other (see, e.g., \citet{karmarkar1982differencing, hayes2002easiest}). A simple solution is to use a greedy algorithm, which iterates through the values of $T_k$ in descending order, assigning each $T_k$ to the subset that currently has the smaller sum (see, e.g. \citet{kellerer2004knapsack, gent1996phase} for more details on the Partition Problem). After finding a well-balanced partition, the next step is to assign the class labels such that the labels for the questions in $S_x$ are equal to $x$ with $x = 0$ or $1$. Recall from section \ref{calibration_step} that $Z_k$ represents the event indicator for the event associated with the $k$th question. To define a balanced set of indicators, for all $k \in S_x$, let
\begin{align*}
 \tilde{Z}_k &= x\\
%\{\tilde{p}_{i,t,k} | i = 1, \dots, I_k, t = 1, \dots, T_k\} &=  \begin{cases} 
%\{1-p_{i,t,k} | i = 1, \dots, I_k, t = 1, \dots, T_k\} & \text{if } Z_k = 1-x\\
%\{p_{i,t,k} | i = 1, \dots, I_k, t = 1, \dots, T_k\} & \text{if } Z_k = x 
\tilde{p}_{i,t,k} &=  \begin{cases} 
1-p_{i,t,k} & \text{if } Z_k = 1-x\\
p_{i,t,k}& \text{if } Z_k = x 
\end{cases}
\end{align*}
with $i = 1, \dots, I_k$, and $t = 1, \dots, T_k$. The set $\left\{\left(\tilde{Z}_k, \{\tilde{p}_{i,t,k} | i = 1, \dots, I_k, t = 1, \dots, T_k \}\right) \right\}_{k=1}^K$ is a balanced version of the data. 
%Notice that this procedure extends trivially to multi-outcome (polychotomous) events because many of the multinomial models, such as the multinomial logistic regression (see, e.g., \citet{agresti2002categorical}) or the multi-outcome extension of our model, can be expressed as a set of independent binary-response models to which the above balancing procedure can be applied directly. 
This procedure was used to balance our real-world dataset both in terms of events and time points. The final output splits the events exactly in half ($|S_0| = |S_1| = 83$) such that number of time points in the first and second halves are 8,737 and 8,738, respectively.

%\subsection{Incoherent and Imbalanced Data}
%\label{practicalmatters}
%The first issue that needs to be addressed arises from human experts making probability forecasts of 0 or 1 even when they are not completely sure of the outcome. For instance, each of the forecast pools associated with the 166 questions in our data contained both a zero and a one. Transforming such forecasts into the logit-space yields infinities which are highly influential and hence cause questions in model estimation. To avoid this, \citet{Ariely00theeffects} suggest changing $p = 0$ and $1$ to $p = 0.02$ and $0.98$, respectively. This is similar to winsorising which sets the extreme probabilities to a specified percentile of the data (see, e.g., \citet{hastings1947low} for more details on winsorising).  \citet{Geo}, on the other hand, consider only probabilities that fall within a constrained interval, say $[0.001, 0.999]$, and discard the rest. Since this implies ignoring a portion of the data, we have adopted the first approach by changing $p = 0$ and $1$ to $p = 0.01$ and $0.99$, respectively. Given that the experts were allowed to report their probability forecasts with a precision of 0.01, this practical compromise was considered a natural choice.

% The second issue is related to the distribution of the class labels in the dataset. If the set of events that occurred is much larger than the set of events that did not occur (or \textit{vice versa}), the dataset is called \textit{imbalanced}. On such data the model can end up over-focusing on the larger class and, as a results, give very accurate forecast performance over the larger class but poor forecast performance over the  smaller class (see, e.g., \citet{chen2009learning, wallace2012class}).
% %Imbalanced data can also cause Platt calibration to overfit (see \citet{mitchell2007multi}). 
%Furthermore, as discussed in Section \ref{Sample-Then-Calibrate-Ident}, balancing the data allows the model to imitate Platt Calibration without losing interpretability. 
%
%To describe the process of balancing the data, recall that our goal is to construct a model that is well-calibrated at any time point across all questions. The class labels in the data must therefore be balanced with respect to the total number of time points. The first step is to find a partition, $S_0$ and $S_1$, of the question indices, $\{1, 2, \dots, K\}$, such that the equality $\sum_{k \in S_0} T_k  = \sum_{k \in S_1} T_k$
%is as closely approximated as possible. This is equivalent to an NP-hard question known in computer science as the \textit{Partition Problem}:  determine whether a given set of positive integers can be partitioned into two sets such that the sums of the two sets equal to each other (see, e.g., \citet{karmarkar1982differencing, hayes2002easiest}). A simple option is to use a greedy algorithm, which iterates through the values of $T_k$ in an ascending order, assigning each $T_k$ to the subset that currently has the smaller sum (see, e.g. \citet{kellerer2004knapsack, gent1996phase} for more details on the Partition Problem). After finding such a partition, the final step is to assign the class labels such that the labels for the questions in $S_x$ are equal to $x$ with $x = 0$ or $1$. In other words, let $\tilde{Z}_k = x$ and
%\begin{align*}
%\{\tilde{p}_{i,t,k | i = 1, \dots, I_k, t = 1, \dots, T_k}\} &=  \begin{cases} 
%\{1-p_{i,t,k} | i = 1, \dots, I_k, t = 1, \dots, T_k\} & \text{if } Z_k = 1-x\\
%\{p_{i,t,k} | i = 1, \dots, I_k, t = 1, \dots, T_k\} & \text{if } Z_k = x 
%\end{cases}
%\end{align*}
%for $k \in S_x$. The final set $\left\{\left(\tilde{Z}_k, \{\tilde{p}_{i,t,k | i = 1, \dots, I_k, t = 1, \dots, T_k} \}\right) \right\}_{k=1}^K$ is a balanced version of the data. This procedure simultaneously balances our dataset both in terms of events and time points. The final output splits the events exactly in half ($|S_0| = |S_1| = 83$) such that number of time points in the first and second halves are 8,737 and 8,738, respectively.

% Both of these approaches, however, are rather \textit{ad-hoc} and unsatisfactory: while the first one requires the user to pick a specified percentile or range within the unit interval, the second one ignores a portion of the data. 
% 
% Therefore we put in our efforts to propose a simple, yet slightly more principled approach: First, fit a cubic smoothing spline to a set of equidistant probabilities in the open interval $(0,1)$ and their corresponding logit probabilities. Second, use this spline to extrapolate the extreme logit probability $\logit(1.0)$. Third, assign the negative of this extrapolated logit probability to the other extreme logit probability, $\logit(0.0)$. The initial choice of the probability grid matters: generally finer grids lead to larger extrapolated values for $\logit(1.0)$. Fortunately, the dataset at hand often offers a natural choice for the grid. For instance, in our case the experts reported their probability forecasts with an accuracy of 0.01. Therefore  $0.50, 0.51, \dots, 0.99$ gives us the initial probability grid. Under this choice, the extrapolation yields $\widehat{\logit(1.0)} \approx 5.509$. This is approximately equivalent to changing $p = 0.0$ and $1.0$ to $p = 0.004$ and $0.996$, respectively. 
% 
% Overall, this approach seems more reasonable than the earlier described solutions by \citet{Ariely00theeffects} and  \citet{Geo} because (a) it does not discard any data and (b) it uses the application-specific scale to construct a smooth, strictly increasing, yet finite transformation from probability scale to logit scale.

%\subsection{Unbalanced Data}
%Since in our case the explanatory variables are probabilities associated with dichotomous events, the data can be balanced by choosing between $\{p_i\}_{i=1}^{I_k}$ and $\{1-p_i\}_{i=1}^{I_k}$ for the first $K/2$ events depending on the whether the events occurred or not, respectively. That is, if the event occurred (or did not occur), choose $\{p_i\}_{i=1}^{I_k}$ (or $\{1-p_i\}_{i=1}^{I_k}$, respectively). Set $Z_k$ for $k = 1, \dots, K/2$ equal to 1. For the remaining $K/2$ questions, on the other hand, choose $\{1-p_i\}_{i=1}^{I_k}$ (or $\{p_i\}_{i=1}^{I_k}$) if the event happened (or did not happen, respectively) and then set $Z_k$ for $k = K/2+1, \dots, K$ equal to 0.
% event $k = 1, \dots, K$ such that
%\begin{eqnarray}
%|\{k : \{p_i\}_{i=1}^{I_k} \text{, } Z_k = 1 \}| + |\{k : \{1-p_i\}_{i=1}^{I_k}  \text{, } Z_k = 0\}| &=& K/2 \label{balanceddata}
%\end{eqnarray}
%and then defining a new set of event indicators denoted $\{Z_k^b\}_{k=1}^K$, where $Z_k^b = 1$ if  $k$th event is in either set described in Equation (\ref{balanceddata}) and $Z_k^b = 0$ otherwise. 
%

%To make this more specific, let $S_1$ be an index set for events that both occurred and for which we choose $\{p_i\}_{i=1}^{I_k}$. Correspondingly, let $S_2$ be an index set for events that both did not occur and for which we choose $\{1-p_i\}_{i=1}^{I_k}$. The dataset is balanced if the sets $S_1$ and $S_2$ satisfy
%\begin{eqnarray*}
%\sum_{k \in S_1 \cup S_2} T_k  &=& \frac{\sum_{k=1}^K T_k}{2}
%%|\{(k, t) : \{p_i\}_{i=1}^{I_k} \text{, } Z_k = 1\text{, } t \in \{1, \dots, T_k\} \}|\\
%% + |\{(k,t) : \{1-p_i\}_{i=1}^{I_k}  \text{, } Z_k = 0\text{, } t \in \{1, \dots, T_k\} \}| &=& \frac{\sum_{k=1}^K T_k}{2}
%\end{eqnarray*}
%with $Z_k = 1$ if $k \in S_1 \cup S_2$ and $Z_k = 0$ otherwise. This i

\subsection{Out-of-Sample Forecasting}
\label{forecasting}
This section evaluates the out-of-sample predictive performance of \textit{Sample-Then-Calibrate} against several other probability aggregation procedures. All the models are allowed to utilize a training set before making predictions on an independent testing set. In order to clarify some of the upcoming notation, let $S_{train}$ and $S_{test}$ be index sets that partition the data into training and testing sets of sizes $|S_{train}| = N_{train}$ and $|S_{test}| = 166 - N_{test}$, respectively. This means that the $k$th question is in the training set if and only if $k \in S_{train}$. The competing models are as follows.

\begin{enumerate}
\item \textit{Simple Dynamic Linear Model (SDLM)}. This procedure fixes $\boldsymbol{b} = \boldsymbol{1}$ and $\beta = 1$ in the hierarchical model described in Section \ref{model}. The model then reduces to
\begin{eqnarray*}
\boldsymbol{Y}_{t, k} &=&  X_{t, k} + \boldsymbol{v}_{t, k}  \\
X_{t, k} &=& \gamma_k X_{t-1, k} + w_{t, k},
\end{eqnarray*}
where $X_{t,k}$ is the logit-probability used for prediction. Since this model is not hierarchical, estimates of the hidden process can be obtained directly for the questions in the testing set without fitting the model first on the training set. In order to make predictions, the sampler is run for 500 iterations of which the first 200 are used for burn-in. The remaining sample of 300 iterations is thinned by discarding every other observation, leaving a final predictive sample of 150 observations.

\item \textit{The Sample-Then-Calibrate procedure both under the Brier (STC-Bri) and the logarithmic score (STC-Log)}. The model is first fit on the training set by running the sampling step for 3,000 iterations of which the first 500 iterations are used for burn-in. After thinning by only keeping every fifth observation, the calibration step is performed for each of the remaining 500 observations. The out-of-sample prediction is done by running the sampling step for 500 iterations with each consecutive iteration reading in and conditioning on the next value of $\beta$ and $\boldsymbol{b}$ that were found during training period. The first 200 observations are used for burn-in. The remaining sample of 300 iterations is thinned by discarding every other observation, leaving a final predictive sample of 150 observations.

\item \textit{A fully Bayesian version of STC-Log (BSTC-Log)}. Denote the (constrained) calibrated logit-probabilities and  the event indicators across all $K$ questions with $\boldsymbol{X}(1)$ and $\boldsymbol{Z}$, respectively.  The posterior distribution of $\beta$ conditional on $\boldsymbol{X}(1)$ is then given by $p(\beta | \boldsymbol{X}(1), \boldsymbol{Z}) \propto p( \boldsymbol{Z} | \beta, \boldsymbol{X}(1)) p(\beta | \boldsymbol{X}(1))$. Recall that the calibration step under $S_{LOG}$ is equivalent to fitting a logistic regression model with $Z_k$ as the response and $\hat{X}_{k,t}(1)$ as the explanatory variable. Therefore the likelihood for the Bayesian version is
\begin{eqnarray}
p( \boldsymbol{Z} | \beta, \boldsymbol{X}(1)) &\propto& \prod_{k=1}^K \prod_{t=1}^{T_k}  \logit^{-1} \left(X_{k,t}(1)/\beta  \right)^{Z_k} \left( 1-  \logit^{-1} \left( X_{k,t}(1)/\beta  \right) \right)^{1-Z_k} \label{OSE2}
\end{eqnarray}
Similarly to  \citet{gelman2003bayesian} the prior is chosen to be locally uniform, $p(1/\beta) \propto 1$. Posterior estimates of $\beta$ can be sampled from Equation (\ref{OSE2}) using generic sampling algorithms such as the Metropolis algorithm (\citet{metropolis1953equation}) or slice sampling (\citet{neal2003slice}). Since the sampling procedure conditions on the event indicators, the full conditional distribution of the hidden states is a non-standard form. Therefore the Metropolis algorithm is also used for sampling the hidden states. Predictions are made with the same choices of thinning and burn-in as described under \textit{Sample-Then-Calibrate}.

\item Due to the lack of previous literature on dynamic aggregation of expert probability forecasts,  the main competitors are exponentially weighted versions of procedures that have been proposed for static probability aggregation:

\begin{enumerate}

\item \textit{Exponentially Weighted Moving Average (EWMA)}. If 
\begin{eqnarray*}
\bar{p}_{t,k}  = \frac{1}{N_{t,k}} \sum_{i=1}^{N_{t,k}} p_{i,t,k},
\end{eqnarray*}
is the average probability forecast given at time $t$ for the $k$th question, then the EWMA forecasts for the $k$th problem are obtained recursively from
\begin{eqnarray*}
\hat{p}_{t,k}(\alpha) =
\begin{cases}
 \bar{p}_{1,k} & \text{ for } t  = 1 \\
\alpha \bar{p}_{t,k}  + (1-\alpha) \hat{p}_{t-1,k}(\alpha) & \text{ for } t > 1
\end{cases}
%\hat{p}_1 &=& \bar{p}_1 \\	
%\hat{p}_t &=& \alpha \bar{p}_t  + (1-\alpha) \hat{p}_{t-1}, \text{ for } t > 1
\end{eqnarray*}
where the input parameter $\alpha$ is learned from the training set by 
\begin{eqnarray*}
\hat{\alpha} &=& \argmin_{\alpha \in [0,1]} \sum_{k \in S_{train}} \sum_{t=1}^{T_k} \left(Z_k - \hat{p}_{t,k}(\alpha) \right)^2
\end{eqnarray*}

%
%\item Many time-series models have been developed for univariate time series. Our data, however, is not univariate nor directly multivariate. The response vector changes in size across time. Therefore in order to use the standard time series models, we turn the data into a univariate time series by considering the average forecast at each time point. 
%
%Autoregressive integrated moving average, ARIMA($p,d,q$), applied to the average forecasts given at each time point, $t$. The parameters, $p,d$, and $q$, are learnt form the training set by
%\begin{eqnarray*}
%(p,d,q) &=& \argmin_{p,d,q \in \mathbb{Z}^+} \sum_{k=1}^K \sum_{t=1}^{T_k} \left(Z_k - \hat{p}_{t,k}\right)^2,
%\end{eqnarray*}
%where $\hat{p}_{t,k}$ is the ARIMA prediction for the $k$th question at time $t$. 
%
%\item  \textit{Exponentially Weighted Moving Logit Aggregator} (EWMLA). This is a moving version of the aggregator, $\hat{p}_G(a)$, introduced in \cite{satopaa}. If $\boldsymbol{p}_{t,k}$ is a vector collecting all the probability forecasts made for the $k$th question at time $t$, then the EWMLA forecasts are found recursively as
%\begin{eqnarray*}
%\hat{p}_{t,k}(\alpha, \nu) =
%\begin{cases}
%G_{t,k}(\nu ) & \text{ for } t  = 1 \\
%\alpha G_{t,k}(\nu )  + (1-\alpha) \hat{p}_{t-1,k}(\alpha, \nu)  & \text{ for } t > 1
%\end{cases}
%\end{eqnarray*}
%where
%\begin{eqnarray*}
%G_{t,k}(\nu )  &=& \frac{\left[ \prod\limits_{i=1}^{N_{t,k}} \left( \frac{p_{i, t, k}}{1-p_{i, t, k}} \right)^{1/N_{t,k}} \right]^\nu}{1+\left[ \prod\limits_{i=1}^{N_{t,k}} \left( \frac{p_{i, t, k}}{1-p_{i, t, k}} \right)^{1/N_{t,k}} \right]^\nu}
%\end{eqnarray*}
%The tuning parameters, $(\alpha, \nu) $, are learned from the training set by
%\begin{eqnarray*}
%(\hat{\alpha}, \hat{\nu}) &=& \argmin_{\nu \in \R, \alpha \in [0,1]} \sum_{k \in S_{train}} \sum_{t=1}^{T_k} \left(Z_k - \hat{p}_{t,k}(\alpha, \nu)\right)^2
%\end{eqnarray*}

\item  \textit{Exponentially Weighted Moving Logit Aggregator (EWMLA)}. This is a moving version of the aggregator $\hat{p}_G(\boldsymbol{b})$ that was introduced in \cite{satopaa}. If $\boldsymbol{p}_{t,k}$ is a vector collecting all the probability forecasts made for the $k$th question at time $t$, then the EWMLA forecasts are found recursively from
\begin{eqnarray*}
\hat{p}_{t,k}(\alpha, \boldsymbol{b}) =
\begin{cases}
G_{t,k}(\boldsymbol{b} ) & \text{ for } t  = 1 \\
\alpha G_{t,k}(\boldsymbol{b} )  + (1-\alpha) \hat{p}_{t-1,k}(\alpha, \boldsymbol{b})  & \text{ for } t > 1
\end{cases}
\end{eqnarray*}
where
\begin{eqnarray*}
G_{t,k}(\nu )  &=& \left( \prod\limits_{i=1}^{N_{t,k}} \left( \frac{p_{i, t, k}}{1-p_{i, t, k}} \right)^{ \frac{\boldsymbol{e}_{i,k}'\boldsymbol{b}}{N_{t,k}}} \right) \Bigg/ \left(1+ \prod\limits_{i=1}^{N_{t,k}} \left( \frac{p_{i, t, k}}{1-p_{i, t, k}} \right)^{\frac{\boldsymbol{e}_{i,k}' \boldsymbol{b}}{N_{t,k}}} \right)
\end{eqnarray*}
The vector $\boldsymbol{b}$ collects the bias terms of the different expertise groups. Therefore it is equivalent to the bias vector found under \textit{Sample-Then-Calibrate}. The term $\boldsymbol{e}_{i,k}$ is a vector of length 5 indicating which level of self-reported expertise the $i$th forecaster in the $k$th question belongs to. For instance, if $\boldsymbol{e}_{i,k} = [0, 1, 0, 0, 0]$, then the expert identifies himself with the expertise level two. The tuning parameters $(\alpha, \boldsymbol{b})$ are learned from the training set by
\begin{eqnarray*}
(\hat{\alpha}, \hat{ \boldsymbol{b}}) &=& \argmin_{ \boldsymbol{b} \in \R^5, \alpha \in [0,1]} \sum_{k \in S_{train}} \sum_{t=1}^{T_k} \left(Z_k - \hat{p}_{t,k}(\alpha,  \boldsymbol{b})\right)^2
\end{eqnarray*}

\item  \textit{Exponentially Weighted Moving Beta-transformed Aggregator (EWMBA)}. The static version of the Beta-transformed (linear) aggregator was introduced in \cite{Ranjan08}. A dynamic version can be obtained by replacing $G_{t,k}(\nu )$ in the EWMLA description with
\begin{eqnarray*}
H_{\nu, \tau} \left( \bar{p}_{t,k}\right),
\end{eqnarray*}
where $H_{\nu, \tau}$ is the cumulative distribution function of the Beta distribution and $\bar{p}_{t,k}$ is the average probability forecast defined under EWMA. The tuning parameters $(\alpha, \nu, \tau)$ are learned from the training set by
\begin{eqnarray*}
(\hat{\alpha}, \hat{\nu}, \hat{\tau}) &=& \argmin_{\nu, \tau > 0\ \alpha \in [0,1]} \sum_{k \in S_{train}} \sum_{t=1}^{T_k} \left(Z_k - \hat{p}_{t,k}(\alpha, \nu, \tau)\right)^2
\end{eqnarray*}

\end{enumerate}
\end{enumerate}

The competing models are evaluated via a 10-fold cross-validation\footnote{A 5-fold cross-validation was also performed. The results were, however, very similar to the 10-fold cross-validation and hence not presented in the paper.} that first partitions the 166 questions into 10 sets. The partition is chosen such that each of the 10 sets has approximately the same number of questions (16 or 17 questions per set in our case) and the same number of time points (between 1760 and 1764 time points per set in our case). The evaluation then iterates 10 times, each time using one of the 10 sets as the testing set and the remaining 9 sets as the training set. Each question is therefore used nine times for training and exactly once for testing. The testing proceeds sequentially one testing question at a time as follows: First, for a question with a time horizon of $T_k$, make a prediction based on the first two days. Compute the Brier score for the aggregate forecast of the second day. Next, make a prediction based on the first three days and  compute the Brier score for the most recent day, namely, the third day. Repeat this process until the prediction is made on all of the $T_k-1$ days. This leads to $T_k-1$ Brier scores per testing question and a total of  17,475 Brier scores across the entire dataset.

Table \ref{prediction} summarizes different ways to aggregate these scores: The first option, denoted by \textit{Scores by Day}, weighs the questions by the number of days the question remained open. This is performed by computing the average of the 17,475 scores. The second option, denoted by \textit{Scores by Problem}, gives each question an equal weight regardless how long the question remained open. This is done by first averaging the scores within a question, and then averaging the average scores across all the questions. Both scores can be further broken down into subcategories by considering the length of the questions. The questions are divided into \textit{Short} questions (30 days or fewer), \textit{Medium} questions (between 31 and 59 days), and \textit{Long} Problems (60 days or more). The number of questions in these subcategories were 36, 32 and 98, respectively. The bolded scores indicate the lowest score in each column. The values in the parenthesis quantify the variability in the scores: Under \textit{Scores by Day} the values give the standard errors of all the scores. Under \textit{Scores by Problem}, on other hand, the values represent the standard errors of the average scores across the questions.

% Requires the booktabs if the memoir class is not being used
\begin{table}[htbp]
   \centering
      \begin{tabular}{l llll} % Column formatting, @{} suppresses leading/trailing space
 \multicolumn{5}{ c }{Scores by Day}\\
Model & \multicolumn{1}{ c }{All Problems} & \multicolumn{1}{ c }{Short} & \multicolumn{1}{ c }{Medium} & \multicolumn{1}{ c }{Long}\\ \hline
SDLM & 0.100 (0.156) & 0.066 (0.116) & 0.098 (0.154) & 0.102 (0.157)\\ 
BSTC & 0.097 (0.213) & \textbf{0.053} (0.147) & 0.100 (0.215) & 0.098 (0.215)\\ 
STC-Bri & 0.096 (0.190) & 0.056 (0.134) & 0.097 (0.190) & 0.098 (0.192)\\ 
STC-Log & \textbf{0.096} (0.191) & 0.056 (0.134) & \textbf{0.096} (0.189) & \textbf{0.098} (0.193)\\ 
EWMBA & 0.102 (0.203) & 0.060 (0.124) & 0.110 (0.201) & 0.103 (0.206)\\ 
EWMLA & 0.102 (0.199) & 0.061 (0.130) & 0.111 (0.214) & 0.103 (0.200)\\ 
EWMA & 0.111 (0.142) & 0.089 (0.100) & 0.111 (0.136) & 0.112 (0.144)\\ 
&&&&\\
\multicolumn{5}{ c }{Scores by Problem}\\
Model & \multicolumn{1}{ c }{All Problems} & \multicolumn{1}{ c }{Short} & \multicolumn{1}{ c }{Medium} & \multicolumn{1}{ c }{Long}\\ \hline
SDLM & 0.089 (0.116) & 0.064 (0.085) & 0.106 (0.141) & 0.092 (0.117)\\ 
BSTC & 0.083 (0.160) & \textbf{0.052} (0.103) & 0.110 (0.198) & 0.085 (0.162)\\ 
STC-Bri & 0.083 (0.142) & 0.055 (0.096) & 0.106 (0.174) & 0.085 (0.144)\\ 
STC-Log & \textbf{0.082} (0.142) & 0.055 (0.096) & \textbf{0.105} (0.174) & \textbf{0.085} (0.144)\\ 
EWMBA & 0.090 (0.156) & 0.063 (0.101) & 0.118 (0.186) & 0.091 (0.161)\\ 
EWMLA & 0.090 (0.159) & 0.064 (0.109) & 0.120 (0.200) & 0.090 (0.159)\\ 
EWMA & 0.104 (0.105) & 0.092 (0.081) & 0.119 (0.125) & 0.103 (0.107)\\ 
   \end{tabular}
   \caption{Brier Scores based on 10-fold cross-validation. \textit{Scores by Day} weighs a question by the number of days the question remained open. \textit{Scores by Problem} gives each question an equal weight regardless how long the question remained open. The bolded values indicate the lowest scores in each column. The values in the parenthesis represent standard errors in the scores. }
   \label{prediction}
\end{table}

Overall, STC-Bri and STC-Log achieve the lowest average scores across all columns except \textit{Short} where they are slightly outperformed by BSTC-Log.  BSTC-Log, however, turns out to be overconfident (see Section \ref{calibration} with a discussion on calibration and sharpness). This means that BSTC-Log underestimates the uncertainty in the events and outputs probability forecasts that are typically too near 0.0 or 1.0 to be calibrated. As a result, the forecasts are either very close to the correct answer or very far from it.  As can be seen in Table \ref{prediction}, this results into high variance in performance. The short questions, however, involve very little uncertainty and were generally the easiest to forecast. On such easy questions, overconfidence can pay off frequently enough to compensate for the few potential large scores that arise from an overconfident and incorrect forecast. In contrast to the BSTC-Log-model, SDLM lacks sharpness and is highly under-confident. This means that SDLM makes conservative forecasts that result into low average scores and low performance variability. The \textit{Sample-Then-Calibrate}-model can therefore be viewed as a well-performing compromise that avoids over-confidence without being too conservative.

 Of the exponentially weighted moving aggregators, EWMBA performs the best and EWMA the worst. Given that the static versions of EWMLA and EWMBA depend on similar transformations, the results suggest that EWMLA is unable to make efficient use of the self-rated expertise information. Both EWMLA and EWMBA, however, suffer from over-confidence and are therefore unable to outperform the \textit{Sample-Then-Calibrate}-model.

The dataset contained a few long questions with an abrupt change of the crowd belief from one answer option to another. Since these abrupt changes happened generally towards the end of each question, during most days the Brier scores were really high. Given that these problems were typically very long, they gained a higher weight under \textit{Scores by Day} than under \textit{Scores by Problem}. This explains why the \textit{Scores by Day} are higher than the \textit{Scores by Problem} under the columns \textit{All Problems} and \textit{Long}.

%% Requires the booktabs if the memoir class is not being used
%\begin{table}[htbp]
%   \centering
%   
%   %\topcaption{Table captions are better up top} % requires the topcapt package
%   \begin{tabular}{l llllllll} % Column formatting, @{} suppresses leading/trailing space
%\hline
%& \multicolumn{2}{ c }{All Problems} & \multicolumn{2}{ c }{Short} & \multicolumn{2}{ c }{Medium} & \multicolumn{2}{ c }{Long}\\
%Method & Days &Problems & Days &Problems & Days &Problems & Days &Problems\\ \hline
%SDLM & 0.101 (0.024) & 0.066 (0.013) & 0.098 (0.024) & 0.102 (0.025)\\ 
%BSTC-Log & 0.097 (0.045) & 0.053 (0.022) & 0.100 (0.046) & 0.099 (0.046)\\ 
%STC-Bri & 0.096 (0.036) & 0.056 (0.018) & 0.097 (0.036) & 0.098 (0.037)\\ 
%STC-Log & 0.096 (0.036) & 0.056 (0.018) & 0.096 (0.036) & 0.098 (0.037)\\ 
%EWMBA & 0.102 (0.041) & 0.060 (0.015) & 0.110 (0.040) & 0.103 (0.042)\\ 
%EWMLA & 0.102 (0.040) & 0.061 (0.017) & 0.111 (0.046) & 0.103 (0.040)\\ 
%EWMA & 0.111 (0.020) & 0.089 (0.010) & 0.111 (0.019) & 0.112 (0.021)\\ 
%   \end{tabular}
%   \caption{Mean Brier Scores based on 10-fold cross-validation. The bolded values indicate the lowest scores in each column.}
%   \label{prediction}
%\end{table}

\subsection{In- and Out-of-Sample Sharpness and Calibration}
\label{calibration}
A calibration plot is a simple tool for assessing the sharpness and calibration of a model. The idea is to plot the probability forecasts against the observed empirical frequencies. Therefore any deviation from the diagonal line suggest poor calibration. A model is considered under-confident (or over-confidence) if the points form an S-shaped (or \reflectbox{S}-shaped) trend. In order to assess sharpness of the model, it is common practice to place a histogram of the given forecasts within the plot.

\begin{figure}[h!]
\subfigure[\scriptsize SDLM In-Sample]{
\hspace*{-1.1em}
  \includegraphics[width= 0.28\textwidth]{Figures/CalibrationNONE10}
%\hspace*{-1em}  \includegraphics[width= 1.5in]{CalibrationPLATTSPH10}
%\caption{$\beta = 1$, $\alpha = 0$}
\label{CalibrationNONEa}
}
\subfigure[\scriptsize STC-Log In-Sample]{
\hspace*{-3em}  \includegraphics[width= 0.28\textwidth]{Figures/CalibrationLOG10}
%\caption{\textit{Sample-Then-Calibrate} with logarithmic score}
\label{CalibrationPLATTLOG10}
}
\subfigure[\scriptsize STC-Bri In-Sample]{
\hspace*{-3em} \includegraphics[width= 0.28\textwidth]{Figures/CalibrationBRI10}
%\caption{\textit{Sample-Then-Calibrate} with Brier score}
\label{CalibrationPLATTBRI10}
}
\subfigure[\scriptsize BSTC-Log In-Sample]{
\hspace*{-3em} \includegraphics[width= 0.28\textwidth]{Figures/CalibrationFULL10}
%\caption{\textit{Sample-Only} }
\label{CalibrationPLATTFULL10}
}

\vspace{-1.5em}
\subfigure[\scriptsize SDLM Out-of-Sample]{
\hspace*{-1.1em}  \includegraphics[width= 0.28\textwidth]{Figures/CalibrationNONE}
%\hspace*{-1em}  \includegraphics[width= 1.5in]{CalibrationPLATTSPH10}
%\caption{$\beta = 1$, $\alpha = 0$}
\label{CalibrationNONEb}
}
\subfigure[\scriptsize STC-Log Out-of-Sample]{
\hspace*{-3em}  \includegraphics[width= 0.28\textwidth]{Figures/CalibrationSTO-Log}
%\caption{\textit{Sample-Then-Calibrate} with logarithmic score}
\label{CalibrationSTC-Log}
}
\subfigure[\scriptsize STC-Bri Out-of-Sample]{
\hspace*{-3em} \includegraphics[width= 0.28\textwidth]{Figures/CalibrationSTO-Bri}
%\caption{\textit{Sample-Then-Calibrate} with Brier score}
\label{CalibrationSTC-Bri}
}
\subfigure[\scriptsize BSTC-Log Out-of-Sample]{
\hspace*{-3em} \includegraphics[width= 0.28\textwidth]{Figures/CalibrationSO}
%\caption{\textit{Sample-Only} }
\label{CalibrationBSTC-Log}
}

\caption[Optional caption for list of figures]{The top and bottom rows show in- and out-of-sample calibration and sharpness, respectively. The models are Simple Dynamic Linear Model (SDLM), the \textit{Sample-Then-Calibrate} approach that optimizes over the logarithmic score (STC-Log), the \textit{Sample-Then-Calibrate} approach that optimizes over the Brier score (STC-Bri), and the fully-Bayesian version of the \textit{Sample-Then-Calibrate} approach that optimizes over the logarithmic score (BSTC-Log).
}
\label{Calibration-Out}
\end{figure}

The top and bottom rows of Figure \ref{Calibration-Out} present calibration plots for SDLM, STC-Log, STC-Bri, and BSTC-Log under in- and out-of-sample probability estimation, respectively. Each case is of interest in its own right:  Good in-sample calibration is crucial for model interpretability. In particular, if the estimated crowd belief is well-calibrated, then the elements of the bias vector $\boldsymbol{b}$ can be used to study the amount of under- or over-confidence in the different expertise groups. Good out-of-sample calibration and sharpness, on other hand, are necessary properties in predicting future events with high accuracy.  To guide our assessment, the dashed bands around the diagonal connect the point-wise, Bonferroni-corrected (\citet{bonferroni}) 95\% lower and upper critical values under the null hypothesis of calibration. These have been computed by running the bootstrap technique described in \citet{brocker2007increasing} for 10,000 iterations. The in-sample results were obtained by running the models were for 10,200 iterations leading to a final posterior sample of 1,000 observations after thinning and using the first 200 iterations for burn-in. The out-of-sample results were computed based on the 10-fold cross-validation discussed in Section \ref{forecasting}.

Overall, the \textit{Sample-Then-Calibrate}-procedure is sharp and well-calibrated both in- and out-of-sample with only a few points barely falling outside the \textit{point-wise} critical values. Since the calibration does not change drastically from the top to the bottom row, the \textit{Sample-Then-Calibrate}-procedure can be considered to present robustness against over-fitting. This is, however, not the case with BSTC-Log that is well-calibrated in-sample but presents over-confidence out-of-sample. Figures \ref{CalibrationNONEa} and \ref{CalibrationNONEb} serve as baselines by showing the reliability plots for the SDLM-model. Since this model does not perform any explicit calibration, it is not surprising to see most points outside the critical values. The pattern in the deviations suggests drastic under-confidence, and the inset histogram reveals drastic lack of sharpness.

% Sharpness refers to the distribution of the forecast probabilities. The closer the forecasts are to zero and one, the sharper the model is said to be. This is an important feature in forecasting, where the main objective typically is to maximize sharpness with respect to calibration (see, e.g., \cite{Ranjan08,raftery2005using}).

% Overall, both in-sample calibration and sharpness increase from (a) to (d) with \textit{BSTC-Log} being the sharpest and most calibrated.

\subsection{Group-Level Expertise Bias}
\label{ExpertBias}
%In Section \ref{calibration} it was briefly mentioned that the experts as a group are under-confident relative to the calibrated crowd belief.
%This section analyzes the biases across different levels of self-reported expertise. 
%Figure \ref{Biases} is similar in nature to the plots seen in the previous section. This time, however, instead of comparing the predicted probabilities against the empirical occurrence rate, we are comparing the expected expert predictions against the calibrated probabilities under different modeling approaches; that is,  $\logit^{-1}(b_j (X +\alpha))$ against $\logit^{-1}(X)$ for $j = 1, \dots, 5$ and $X \in \R$. Furthermore, we separate the results by the self-reported expertise levels. 
%Recall from Section \ref{data} that we asked the experts to self-assess their expertise level on a 1-to-5 scale (1 = Not At All Expert to 5 = Extremely Expert). 
%Recall from Section \ref{model} that the vector $\boldsymbol{b}$ collects the bias terms specific to each self-reported expertise group. 
Recall from Section \ref{data} that the experts were asked to self-assess their expertise level (on a 1-to-5 scale with 1 = Not At All Expert to 5 = Extremely Expert) on  all of the 166 questions in our real-world dataset. Based on these self-assessments the experts were then divided into 5 expertise groups, and each of the expertise groups was assigned a separate multiplicative bias term. This section uses the \textit{Sample-Then-Calibrate}-procedure to explore the posterior distribution of these group-level bias terms. Figure \ref{Biases} presents the posterior distributions of the bias terms with side-by-side box plots. Since the distributions fall completely below the \textit{no-bias} reference-line at 1.0, the expertise groups are deemed under-confident.

The under-confidence, however, decreases as the level of the self-reported expertise increases. In order to understand this result, it is helpful to view the process of making a subjective probability assessment as Bayesian updating: A completely ignorant expert aiming to minimize a reasonable loss function, such as the Brier score, has no reason to give anything but 0.5 as his probability forecast. However, as soon as the forecaster gains some knowledge about the event, he produces an updated forecast that is a compromise between his initial forecast and the new information acquired. The updated forecast is therefore conservative and necessarily too close to 0.5 as long as the forecaster remains only partially informed about the event. If most forecasters fall somewhere on this spectrum between ignorance and full information, their average forecast tends to fall strictly between 0.5 and the most-informed probability forecast (see \citet{Baron} for more details). Since expertise is to a large extent determined by subject-matter knowledge, the level of under-confidence can be expected to decrease as a function of the group's level of self-reported expertise.

Overall, finding under-confidence is a rather surprising result given that many previous studies have been conducted to show that experts are likely to be over-confident. For instance,  \citet{lichtenstein1977calibration, morgan1992uncertainty, bier2004implications} summarize the results from numerous calibration studies, and conclude that experts are systematically over-conÞdent about their probability assessments. Our result, however, is a statement about groups of experts and hence does not invalidate the possibility of the individual experts being overconfident. To make conclusions at the individual-level based on the group-level bias terms would result into a \textit{ecological inference fallacy} (see, e.g., \citet{lubinski1996seeing}).

%This follows from the well-known Simpson's Paradox (\citet{simpson1951interpretation}) that  explains why the trend of the individuals can be reversed after they are aggregated into a single group.

%The well-informed experts, in the contrary, place themselves well among the two groups with the levels of self-reported expertise.

\begin{figure}[!ht]
%\centering
%\hspace*{-42em} 	\includegraphics[width=  0.8\textwidth]{Figures/Figures/LegendBias} % requires the graphicx package
%\vspace{-2.7em}
%
%\subfigure[]{
%  \includegraphics[width = 0.32 \textwidth]{Figures/BiasesPLATTLOG}
%%\caption{\textit{Sample-Then-Calibrate} with logarithmic score}
%\label{BiasesPLATTLOG}
%}
%\subfigure[]{
%\hspace*{-1em} \includegraphics[width = 0.32 \textwidth]{Figures/BiasesPLATTBRI}
%%\caption{\textit{Sample-Then-Calibrate} with Brier score}
%\label{BiasesPLATTBRI}
%}

%\subfigure[]{
%\hspace*{-1em} \includegraphics[width = 0.5 \textwidth]{Figures/BiasesFULL}
%%\caption{\textit{Sample-Only} }
%\label{BiasesFULL}
%}
\vspace*{-1em} 
\includegraphics{Figures/BiasesBoxplots}
%\caption{\textit{Sample-Then-Calibrate} with Brier score}
\label{BiasesLOGBoxplots}

%\subfigure[STC-Log]{
%\hspace*{-1em} \includegraphics[width = 0.5 \textwidth]{Figures/BiasesLOGBoxplots}
%%\caption{\textit{Sample-Then-Calibrate} with Brier score}
%\label{BiasesLOGBoxplots}
%}
%\subfigure[STC-Bri]{
%\hspace*{-1em} \includegraphics[width = 0.5 \textwidth]{Figures/BiasesBRIBoxplots}
%%\caption{\textit{Sample-Then-Calibrate} with Brier score}
%\label{BiasesBRIBoxplots}
%}
%\subfigure[BSTC-Log]{
%\hspace*{-1em} \includegraphics[width = 0.32 \textwidth]{Figures/BiasesFULLBoxplots}
%%\caption{\textit{Sample-Then-Calibrate} with Brier score}
%\label{BiasesFULLBoxplots}
%}

\caption[Optional caption for list of figures]{Comparing the bias-levels across self-reported expertise under different approaches. 
Posterior distributions of $b_j$ for $j = 1, \dots, 5$ under the \textit{Sample-Then-Calibrate} approach that optimizes over the logarithmic score and the \textit{Sample-Then-Calibrate} approach that optimizes over the Brier score.
}
\label{Biases}
\end{figure}

%Under the \textit{Sample-Only} approach, the posterior mean of $\alpha$ is around 0.155, which is very small in the logit scale. Therefore 
%\begin{eqnarray*}
%\E[\boldsymbol{Y}_{t, k}] &=&  \boldsymbol{M}_k \boldsymbol{b} (X_{t, k} +\alpha) \approx \boldsymbol{M}_k \boldsymbol{b} X_{t, k}
%\end{eqnarray*}
%and t
More specific group-level inference can be performed by computing probabilities based the estimated posterior distributions. For instance, the posterior probability that the most expert group is the least under-confident is approximately equal to $1.0$, and 
%More specifically, $\P(b_5 > b_j \text{ for } j = 1, \dots, 4) \approx 1.0$.
the posterior probability of a strictly decreasing level of under-confidence is approximately 0.87. 
 %In other words, $\P(b_5 > b_4 > \dots > b_1) \approx 0.871$.  
 The latter probability is driven down by the inseparability of the two groups associated with the lowest levels of self-reported expertise. The fact that these groups are very similar suggests that the forecasters are poor at assessing how little they know about a subject that is strange to them. If these groups are combined into a single group, the posterior probability of a strictly decreasing level of under-confidence is approximately 1.0.

\subsection{Question Difficulty and Other Measures}
\label{QuestionDifficulty}
One advantage of our model arises from its ability to produce estimates of interpretable question-specific parameters $\gamma_k$, $\sigma^2_k$, and $\tau^2_k$. These quantities can be combined in many interesting ways to answer questions about different groups of experts or the questions themselves. For instance, being able to assess the difficulty of a question could lead to more principled ways of aggregating performance measures across questions or to novel insight on the kind of questions that are found difficult by experts (see, e.g., a discussion on the \textit{Hard-Easy Effect} in \cite{Wilson94cognitivefactors}). To illustrate, recall that higher values of $\sigma^2_k$ suggest greater disagreement among the participating experts. Since experts are more likely to disagree over a difficult question than an easy one, it is reasonable to assume that $\sigma^2_k$ has a positive relationship with question difficulty. An alternative measure is given by $\tau_k$ that quantifies the volatility of the underlying circumstances that ultimately decide the outcome of the event. Therefore a high value of $\tau_k$ can cause the outcome of the event appear unstable and difficult to predict.

%This constant drives the hidden process and is therefore directly indicative of the rate at which the calibrated crowd belief approaches 0.0 or 1.0.  If this rate is very slow, the day-to-day accumulation of available and relevant information does not

%is, how early on does the question become solvable.

%Since $\sigma^2_k$ has a positive and $\gamma_k$ a negative relationship with question difficulty, these two quantities could be then combined via multiplication, $\sigma^2_k / \gamma_k$, into a single measure of question difficulty.

As a final illustration of our model, consider the two questions used for illustrative purposes in Section \ref{data}. Figure \ref{ExamplePlotsFinal} is a copy of Figure \ref{ExamplePlots} with the addition of a solid line surrounded by a dashed band. The solid line represents the posterior mean of the calibrated crowd belief as estimated by STC-Log. The dashed lines connect the point-wise 95\% posterior intervals across different time points. Since $\hat{\sigma}_k^2 = 2.43$  and $\hat{\sigma}_k^2 = 1.77$ for the questions depicted in Figures \ref{Example1b} and \ref{Example2b}, respectively, the first question provokes more disagreement among the experts than the second one. Intuitively this makes sense because the target event in Figure \ref{Example1b}  is determined by several conditions that may change radically from one day to the next while the target event in Figure \ref{Example2b} is determined by a relatively steady stock market index. Therefore it is not surprising to find that the first question has $\hat{\tau}_k^2 = 1.38$ while the second one has  $\hat{\tau}_k^2 = 0.198$. We conclude that the first question can be considered inherently more difficult than the second one.

\begin{figure}[h!]
\centering
\hspace*{0em} 	\includegraphics[width=  \textwidth]{Figures/LegendExamplePlot} % requires the graphicx package
\vspace{-4.5em}

\hspace{-0.5em}
\subfigure[\textit{Will the expansion of the European bailout fund be ratified by all 17 Eurozone nations before 1 November 2011?}]{
\hspace*{-1em}  \includegraphics[width= 0.49\textwidth]{Figures/ExamplePlot1b}
%\hspace*{-1em}  \includegraphics[width= 1.5in]{CalibrationPLATTSPH10}
%\caption{$\beta = 1$, $\alpha = 0$}
\label{Example1b}
} \hspace{0.5em}
\subfigure[\textit{Will the Nikkei 225 index finish trading at or above 9,500 on 30 September 2011?}]{
\hspace*{-1em}  \includegraphics[width= 0.49\textwidth]{Figures/ExamplePlot2b}
%\caption{\textit{Sample-Then-Calibrate} with logarithmic score}
\label{Example2b}
}

\caption[Optional caption for list of figures]{Scatterplots of the probability forecasts given for two questions in our dataset. The shadings represents the self-reported expertise of the forecaster who provided the probability forecast. The solid line gives the posterior mean of the calibrated crowd belief as estimated by STC-Log. The surrounding dashed lines connect the point-wise 95\% posterior intervals.}
\label{ExamplePlotsFinal}
\end{figure}

%
%\begin{enumerate}
%\item Expert Noise $ \sigma^2$.
%\item Distance from 0.5 and how quickly the process approaches the extreme probabilities 0.0 and 1.0. This is determined largely by the value of $\gamma$. 
%\item Rank-correlation
%\end{enumerate}

\section{Discussion}
This paper introduced an interpretable time-series model that incorporates self-reported expertise and captures a sharp and well-calibrated crowd belief across time. The model is estimated in two-steps: The first step samples constrained versions of the model parameters via a Gibbs sampler that only makes use of standard distributions. The second step involves a one-dimensional optimization procedure that transforms the constrained parameter values into their unconstrained counterparts. Due to the rather unorthodox nature of our dataset, our work expands the literature towards less-studied areas of  probability aggregation.

\subsection{Summary of Findings}
The model was applied to an unusually large dataset on probability forecasting done by human experts. The estimated crowd belief was found to be sharp and well-calibrated under both in- and out-of-sample probability estimation. This has direct implications  on predictive power and model interpretability. Firstly, the model was shown to outperform other probability aggregators in terms of forecasting ability. Secondly, the crowd belief was used as the no-bias reference point to study the distributions of the bias terms for different expertise-groups. All the groups were found to be under-confident. The under-confidence, however, decreased as the level of self-reported expertise increased. This result is about groups of experts and hence does not conflict with the well-known result of the individuals being over-confident (see, e.g.,  \citet{lichtenstein1977calibration, morgan1992uncertainty, bier2004implications}). Besides studying group-level bias, the model can be used to generate estimates of many problem-specific quantities. These quantities have clear interpretations and can be combined in many interesting ways to answer a range of hypotheses about the experts and the questions.

\subsection{Future Directions and Limitations}
 The model can be extended in various ways to meet a wide range of research objectives. For instance, the degree to which the group-level under-confidence decreases towards the resolution of the target event, could be studied by allowing the bias vector to depend on time. Furthermore, the model is not limited to the study of group-level bias at different levels of self-reported expertise. The experts could equally well be partitioned based on other features, such as gender, education, or age, to produce a range of novel insight on the level of confidence among different groups of expert forecasters.

Other future directions could aim to remove some of the obvious limitations of our model. For instance, recall that the random components are assumed to follow a normal distribution. This is a strong assumption that may not always be justified. Logit-probabilities, however, have been modeled with a normal distribution before (see, e.g., \citet{Erev1994}). Furthermore, the normal distribution is a rather standard assumption in psychological models (see, e.g.,  signal-detection theory  in \citet{tanner1954decision}). A second limitation resides in the assumption that both the observed and hidden processes are expected to grow linearly. This assumption could be relaxed, for instance, by adding higher order terms to the model. A more complex model, however, is likely to sacrifice interpretability. Given that our model is able to detect very intricate patterns in the crowd belief  (see Figure \ref{ExamplePlotsFinal}), compromising interpretability for the sake of facilitating non-linear growth is hardly necessary.

%Despite these and potentially other unmentioned limitations, the model yields very good predictive performance and calibration on real data.

%
%Ideally the model could be further developed into a non-hierarchical version that does not share any information among the forecasting questions. Such a model could be used to study group-level bias across different questions and to make aggregate probability forecasts without a separate training set. Due to \textit{perfect separation}, however, some sharing among the questions is necessary unless the modeler has access to prior information on the bias terms, the hidden states, or their variability. Under strong enough prior information the estimation procedure can be made regular enough to be performed separately for each question. This, however, was considered to be outside the scope of this paper and was therefore not discussed in detail. 
%

%
%This leads to several research objectives including
%\begin{itemize}
%\item the analysis of under- and overconfidence across different levels of self-reported expertise,
%\item confidence bands for the crowd belief,
%\item accurate probability forecasts, and
%\item many question-specific quantities that have easy interpretations and can be used to gain novel insight in the social sciences. 
%\end{itemize}
%

%Allow observed variance to vary across expert groups. Allow more terms (e.g. the bias and variance terms) to vary over time. Model the bias based on expert characteristic. Problem difficulty.

\appendix
\section{Technical Details of the Sampling Step}
\label{appendix}
%The implementation of the hierarchical procedure introduced in Section \ref{model} consist of two steps: \ref{GibbsSampler} Run the Gibbs sampler that to obtain estimates of the constrained parameters. \ref{Optimization} Perform a one-dimensional maximization to transform the constrained parameter estimates into their unconstrained counterparts.

%The main difference arises in the way the hidden states and the calibration parameter, $\beta$, are sampled.

%\subsection{Technical Details: \textit{Sample-Then-Calibrate}}
%\label{Sample-Then-Calibrate}
%This subsection describes the details of \textit{Sample-Then-Calibrate} approach one block of model parameters at a time.
%\subsection{Sampling Step}
%\label{GibbsSampler}
The Gibbs sampler iteratively samples all the unknown parameters from their full-conditionals one block of parameters at a time. Since this is performed under the constraint $b_3 = 1$ to ensure model identifiability, the constrained parameter estimates should be denoted with a trailing $(1)$ to maintain consistency with earlier notation. For instance, the constrained estimate of $\gamma_k$ should be denoted by $\hat{\gamma}_k(1)$ while the unconstrained estimate is denoted by $\hat{\gamma}_k$. For the sake of clarity, however, the constraint suffix is omitted in this section. Nonetheless, it is important to keep in mind that all the estimates in this section are constrained.

\begin{center}
\framecolorbox[\textwidth]{gray}{gray!15}{Sample $X_{t,k}$}
\end{center}
The hidden states are sampled via the \textit{Forward-Filtering-Backward-Sampling} (FFBS) algorithm that first predicts the hidden states using a Kalman Filter and then performs a backward sampling procedure that treats these predicted states as additional observations (see, e.g., \cite{carter1994gibbs, migon2005dynamic} for details on FFBS). More specifically, the first part, namely the Kalman Filter, is deterministic and consists of a predict and an update step. Given all the other parameters except the hidden states, the predict step for the $k$th question is
\begin{align*}
X_{t|t-1,k} &= \gamma_k X_{t-1|t-1,k} \\
P_{t|t-1, k} &= \gamma_k^2 P_{t-1|t-1, k} + \tau_k^2,
\end{align*}
where the initial values, $X_{0|0,k}$ and $P_{0|0, k}$, are equal to $0$ and $1$, respectively. The update step is 
\begin{align*}
e_{t,k} &= Y_{i,t,k} - b_{i,k} X_{t | t-1, k} \\
S_{t,k} &=  \sigma_k^2 + b_{i,k}^2 P_{t|t-1, k}\\
K_{t,k} &=  P_{t|t-1, k} b_{i,k} S_{t,k}^{-1} \\
X_{t|t, k} &= X_{t|t-1, k} + K_{t,k} e_{t,k} \\
P_{t|t,k} &= (1 - K_{t,k} b_{i,k}) P_{t|t-1,k},
\end{align*}
where $b_{i,k}$ is the corresponding bias term for the $i$th expert in the $k$th question. The update step is repeated sequentially for each observation $Y_{i,t,k}$ given at time $t$. For each such repetition of the update step, the previous posterior values, $X_{t|t, k}$ and $P_{t|t, k}$, should be considered as the new prior values, $X_{t|t-1, k}$ and $P_{t|t-1, k}$. After running the Kalman Filter up to the final time point at $t = T_k$, the final hidden state is sampled from $X_{T_k,k} \sim \mathcal{N}(X_{T_k|T_k, k}, P_{T_k|T_k, k})$. The remaining states are obtained via the backward sampling that is performed in reverse from
\begin{align*}
X_{t-1, k} &\sim  \mathcal{N} \left(V\left( \frac{\gamma_kX_{t,k}}{\tau_k^2}  + \frac{X_{t|t,k}}{P_{t|t,k} } \right),  V\right),
\end{align*}
where
\begin{align*}
V &= \left( \frac{\gamma_k^2}{\tau_k^2} + \frac{1}{P_{t|t,k}}\right)^{-1}
\end{align*}
This can be viewed as backward updating that considers the Kalman Filter estimates as additional observations at each given time point. If the observation $\boldsymbol{Y}_{t,k}$ is completely missing at time $t$, the update step is skipped and the state estimates are sampled from
\begin{align*}
\mathcal{N}\left(\gamma_k X_{t-1|t-1,k}, \gamma_k^2 P_{t-1|t-1,k} + \tau_k^2\right)
\end{align*}

\begin{center}
\framecolorbox[\textwidth]{gray}{gray!15}{Sample $\boldsymbol{b}$ and $\sigma^2_k$}
\end{center}
First, vectorize all the response vectors $\boldsymbol{Y}_{t,k}$ into a single vector denoted $\boldsymbol{Y}_k = \left[\boldsymbol{Y}_{1,k}^T, \dots, \boldsymbol{Y}_{T_k,k}^T\right]^T$. Since each $\boldsymbol{Y}_{t,k}$ is matched with $X_{t,k}$ via the time index $t$, we can form a $|\boldsymbol{Y}_k| \times J$ design-matrix by letting $\boldsymbol{X}_k = \left[ (\boldsymbol{M}_kX_{1,k})^T, \dots, (\boldsymbol{M}_kX_{T_k,k})^T \right]^T$. Given that the goal is to borrow strength across questions by assuming a common bias vector $\boldsymbol{b}$, the parameter values must be estimated in parallel for each question such that the matrices $\boldsymbol{X}_k$ can be further concatenated into $\boldsymbol{X} = [\boldsymbol{X}_1^T, \dots, \boldsymbol{X}_K^T]^T$ during every iteration. Similarly, $\boldsymbol{Y}_k$ must be further vectorized into a vector $\boldsymbol{Y} = [\boldsymbol{Y}_1^T, \dots, \boldsymbol{Y}_K^T]^T$. The question-specific variance terms are taken into account by letting $\boldsymbol{\Sigma} = \text{diag}(\sigma^2_1 \boldsymbol{1}_{1 \times T_1}, \dots, \sigma^2_K \boldsymbol{1}_{1 \times T_K})$.  After adopting the non-informative prior $p(\boldsymbol{b}, \sigma_k^2 | \boldsymbol{X}_k) \propto \sigma_k^{-2}$ for each $k = 1, \dots, K$, the bias vector are sampled from
\begin{eqnarray}
\boldsymbol{b} | \dots &\sim& \mathcal{N}_J \left( (\boldsymbol{X}^T \boldsymbol{\Sigma}^{-1} \boldsymbol{X})^{-1} \boldsymbol{X}^T \boldsymbol{\Sigma}^{-1} \boldsymbol{Y}, (\boldsymbol{X}^T \boldsymbol{\Sigma}^{-1} \boldsymbol{X})^{-1} \right) \label{BiasVector}
\end{eqnarray}
Since the covariance matrix in Equation (\ref{BiasVector}) is diagonal, the constraint is enforced at this point by letting $b_3 = 1$. The variance parameter is then sampled from
\begin{eqnarray*}
\sigma^2_k | \dots &\sim& \text{Inv-}\chi^2\left(|\boldsymbol{Y}_k| - J, \frac{1}{|\boldsymbol{Y}_k| - J} (\boldsymbol{Y}_k - \boldsymbol{X}_k \boldsymbol{b})^T (\boldsymbol{Y}_k - \boldsymbol{X}_k \boldsymbol{b}) \right)
\end{eqnarray*}
Since the experts are not required to give a new prediction at every time unit, the design matrices must be trimmed accordingly such that their dimensions match up with the dimensions of the observed matrices.

%
%Adopting the non-informative prior $p(\boldsymbol{b}, \sigma_k^2 | \boldsymbol{X}_k) \propto \sigma_k^{-2}$, the parameters $\boldsymbol{b}_k$ and $\sigma^2_k$ can be sampled via a standard Bayesian regression setup as follows.
%\begin{eqnarray*}
%\boldsymbol{b}_k | \dots &\sim& \mathcal{MVN}_J \left( (\boldsymbol{X}_k^T \boldsymbol{X}_k)^{-1} \boldsymbol{X}_k^T \boldsymbol{Y}_k, (\boldsymbol{X}_k^T \boldsymbol{X}_k)^{-1} \sigma^2_k\right) \\
%\sigma^2_k | \dots &\sim& \text{Inv-}\chi^2\left(|\boldsymbol{Y}_k| - J, \frac{1}{|\boldsymbol{Y}_k| - J} (\boldsymbol{Y}_k - \boldsymbol{X}_k \boldsymbol{b}_k)^T (\boldsymbol{Y}_k - \boldsymbol{X}_k \boldsymbol{b}_k) \right),
%\end{eqnarray*}
%where the final distribution is scaled inverse-$\chi^2$ (see, e.g., \citet{gelman2003bayesian}). For identifiability reasons, one of the elements of $\boldsymbol{b}_k$ should be fixed to some pre-specified constant, say, 1. The remaining elements can be estimated separately for each question. 
%

\begin{center}
\framecolorbox[\textwidth]{gray}{gray!15}{Sample $\gamma_k$ and $\tau^2_k$}
\end{center}
Estimating the parameters related to the hidden process are estimated via a regression setup. More specifically, after adopting the non-informative prior $p(\gamma_k, \tau_k^2 | \boldsymbol{X}_k) \propto \tau_k^{-2}$, the parameter values are sampled from
\begin{eqnarray*}
\gamma_k | \dots  &\sim&  \mathcal{N} \left( \frac{\sum_{t=2}^{T_k} X_{t,k}X_{t-1,k}}{\sum_{t=1}^{T_k-1} X_{t,k}^2} , \frac{\tau_k^2}{\sum_{t=1}^{T_k-1} X_{t,k}^2} \right)\\
\tau_k^2 | \dots &\sim& \text{Inv-}\chi^2 \left(T_k-1, \frac{1}{T_k-1} \sum_{t=2}^{T_k} \left(X_{t,k} - \gamma_k X_{t-1,k} \right)^2 \right),
\end{eqnarray*}
where the final distribution is a scaled inverse-$\chi^2$ (see, e.g., \citet{gelman2003bayesian}).

%\begin{center}
%\framecolorbox[\textwidth]{Figures/gray}{gray!15}{Estimate $\beta$}
%\end{center}
%\subsection{Calibration Step}
%\label{Optimization}
%
%The calibration parameter $\beta$ is estimated via a one-dimensional maximization procedure
%\begin{eqnarray*}
% \hat{\beta}&=&  \argmax_{\beta \in \R / \{0\}} \sum_{k=1}^K \sum_{t=1}^{T_k}  S\left(Z_k, \hat{X}_{k,t}(1) / \beta \right),\\
%\end{eqnarray*}
%where $S$ is a proper scoring rule. The unconstrained parameter estimates are then given by
%\begin{eqnarray*}
% \hat{X}_{t,k}&=& \hat{X}_{k,t}(1) / \hat{\beta} \\
% \hat{\boldsymbol{b}}_{t,k}&=& \hat{\boldsymbol{b}}(1) \hat{\beta}\\
%  \hat{\tau}_{k}^2&=& \hat{\tau}_{k}^2(1) \hat{\beta}^2\\
%  \hat{\sigma}_{k}^2&=& \hat{\sigma}_{k}^2(1)\\
%  \hat{\gamma}_{k}&=& \hat{\gamma}_{k}(1)
%\end{eqnarray*}

%\subsection{Technical Details: \textit{Sample-Only}}
%
%This subsection describes the details of \textit{Sample-Only} approach one block of model parameters at a time.
%\begin{center}
%\framecolorbox[\textwidth]{Figures/gray}{gray!15}{Sample $X_{t,k}$}
%\end{center}
%Since the \textit{Sample-Only} approach conditions on $\{Z_k\}_{k=1}^K$, the FFBS algorithm cannot be used for sampling the hidden states. Instead, he hidden states are sampled in an increasing order one at a time from
%\begin{eqnarray*}
%X_{t,k} | \dots &\propto&  \logit^{-1} \left( X_{k,t}/\beta \right)^{Z_k} \left( 1-  \logit^{-1} \left( X_{k,t}/\beta \right) \right)^{1-Z_k}\\
%&& \times \exp\left( -\frac{1}{2\sigma_k^{2}} \left( \boldsymbol{Y}_{t,k} - \boldsymbol{M}_k \boldsymbol{b}X_{t,k}\right)^T \left( \boldsymbol{Y}_{t,k} - \boldsymbol{M}_k \boldsymbol{b}X_{t,k} \right)\right)\\
%&& \times \exp\left(-\frac{1}{2\tau^{2}_k}\left(X_{t,k} - \gamma_k X_{t-1,k} \right)^2 \right),
%\end{eqnarray*}
%where the initial state is sampled from $X_{0,k} \sim \mathcal{N}(0,1)$. The sampling is done via the Metropolis algorithm (\citet{metropolis1953equation}) with a Normal proposal centered at the current value and variance tuned to give an acceptance ratio around 50\%.
%
%
%% Since the final forecasts $\{\boldsymbol{Y}_{T_k, k}\}_{k=1}^K$ contain all the information within $\{Z_k\}_{k=1}^K$, we have that 
%%\begin{eqnarray*}
%%p(\boldsymbol{X}_{k} | \boldsymbol{Y}_{1:T_k, k},  Z_k, \dots) &=& p(\boldsymbol{X}_{k} | \boldsymbol{Y}_{1:T_k, k}, \dots)
%%\end{eqnarray*}
%%Therefore the FFBS-algorithm can be used to sample the hidden states also in the \textit{Sample-Only} approach.
%
%
%
%\begin{center}
%\framecolorbox[\textwidth]{Figures/gray}{gray!15}{Sample $\left(\boldsymbol{b}, \sigma^2\right)$ and $\left(\gamma, \tau^2\right)$}
%\end{center}
%The same as for the \text{Sample-Then-Calibrate} approach. See Section \ref{Sample-Then-Calibrate}.
%
%\begin{center}
%\framecolorbox[\textwidth]{Figures/gray}{gray!15}{Sample $\beta$}
%\end{center}
%This is equivalent to logistic regression. By adopting a locally uniform prior $p(1/\beta) \propto 1$, the calibration parameter can be sampled from
%\begin{eqnarray*}
%\beta | \dots &\propto& \prod_{k=1}^K \prod_{t=1}^{T_k}  \logit^{-1} \left( X_{k,t}/\beta \right)^{Z_k} \left( 1-  \logit^{-1} \left( X_{k,t}/\beta \right) \right)^{1-Z_k}
%\end{eqnarray*}
%This is performed via the Metropolis algorithm (\citet{metropolis1953equation})  with a Normal proposal centered at the current value and variance tuned to give an acceptance ratio around 50\%.

% \bibliographystyle{plainnat}
\bibliographystyle{imsart-nameyear}
\bibliography{biblio}		% expects file "myrefs.bib"

\end{document}

### 2013-08-21 paper.tex
Edit
Added Barb and Phil. Also some cleanup.

@@
 \author{\fnms{Ville A.} \snm{Satop\"a\"a}\corref{}\ead[label=e1]{satopaa@wharton.upenn.edu}},
-\author{\fnms{Shane T.} \snm{Jensen}\ead[label=e3]{stjensen@wharton.upenn.edu}}
+\author{\fnms{Shane T.} \snm{Jensen}\ead[label=e3]{stjensen@wharton.upenn.edu}},
+\author{\fnms{Lyle H.} \snm{Ungar}\ead[label=e2]{ungar@cis.upenn.edu}},
+\author{\fnms{Barbara A.} \snm{Mellers}\ead[label=e4]{mellers@wharton.upenn.edu}},
 \and
-\author{\fnms{Lyle H.} \snm{Ungar}\ead[label=e2]{ungar@cis.upenn.edu}}
+\author{\fnms{Phil E.} \snm{Tetlock}\ead[label=e5]{tetlock@wharton.upenn.edu}}
 
@@
 
- \affiliation{Department of Statistics\\ The Wharton School of the University of Pennsylvania   \printead{e1,e3}}
-
-\affiliation{Department of Computer and Information Science\\ University of Pennsylvania, \printead{e2}}
+ \affiliation{Department of Statistics, The Wharton School of the University of Pennsylvania\\   \printead{e1,e3}}
+ \affiliation{Department of Computer and Information Science, University of Pennsylvania\\ \printead{e2}}
+ \affiliation{Department of Psychology, University of Pennsylvania\\ \printead{e4,e5}}
  \address{Philadelphia, PA 19104- 6340, USA\\  \printead{e1}}
@@
 the proportion of events that actually occurred is, say, 60\% for all those events for which he assessed a probability of 0.60.
-%the long-run frequency of occurrence of the events agrees with his probability forecasts.
-This means that, for instance, among all those events for which an expert assesses a probability of  0.60, the proportion of events that actually occur is 60\%.
 Even though several experiments have been conducted to show that experts are generally poorly calibrated (see, e.g., \citet{cooke1991experts, shlyakhter1994quantifying}), some personal differences among the experts have been found to associate with calibration.
@@
 
-%Therefore self-reported expertise should be taken in to account when modeling probability forecasts given by a diverse group of experts. 
 Unfortunately, calibration by itself does not guarantee usefulness of a probability assessment.
@@
 In general, a probability assessments that is near 0.0 or 1.0, is considered more informative, and an expert who makes such assessments is called \textit{sharp}.
-%called \textit{smooth} or \textit{not sharp}. If, on other hand, most of his probabilities are near 0.0 or 1.0, he is deemed \textit{sharp}. 
-%This notion leads to a second important attribute called \textit{sharpness}, which is a statement about the distribution of the assessed probabilities.
-If most of the probabilities are near 0.0 and 1.0, the assessor is said to be sharp.
 In the forecasting literature sharpness is often combined with calibration into a single goal of maximizing sharpness subject to calibration (see, e.g., \citet{raftery2005using, Ranjan08}).
@@
 Therefore many complex areas of probability aggregation still remain rather unexplored.
-%Therefore this paper makes an advance towards rather unexplored areas of probability aggregation by modeling forecasting 
-%This paper introduces an aggregation procedure that allows multiple probability assessments from a single expert before the event resolves. 
-%The context considered in this paper, on the other hand, allows the experts to update their probability assessments over a period of time before the future event resolves.
-%
-%considering a dynamic context with a time period during which the experts are allowed to update their beliefs. 
 For instance, consider an investor aiming to assess whether a stock index will finish trading at or above some threshold on a given date.
@@
 
-%Two commonly desired attributes of an aggregator are \textit{calibration} and \textit{sharpness}:
-
-%Sharpness, on the other hand, is a statement about the spread of the forecasts.
-The aggregator is said to be sharp if the resulting forecasts are near 0 and 1.
-These two attributes are typically combined into a single goal of finding an aggregator that maximizes sharpness subject to calibration (see, e.g., \citet{gneiting2004verifying, raftery2005using}).
-
-%
-%
-%
-%
-%
-%\noindent
-%\\
-%Element 1: MAIN THEME: The broad theme is subjective probability aggregation.
-Use a catchy initial phrase: "Individuals can differ radically from one another in the precision to which they are able to  evaluate the likelihood of a future event happening." This is makes a statement about human cognitive process and does not include any odd terms.
-\\
-%Element 2: WHY CARE: Explain the importance of probability aggregation (in a static context) for decision making.
-Explain why time-series advice is more realistic.
-(1 REABSTC-LogN) Lack of previous literature on time-series, subjective probability.
-(2 REASON)  \\
-%Element 3: SUMMARY OF LITERATURE: See the guidelines. We want to to site work done on static case. There is also some work done on repeated forecasts, i.e. see raftery's paper.\\
-%Element 4: MAIN CONTRIBUTION: "Our main contribution arises from our abilities ... " Advertise the data agin.\\
-%Element 5: WE NEED A CLEAR INDICATION OF THE FOLLOWING:\\
-%(a) The core research question: The main goal of this study is to find an interpretable model for detecting a a calibrated crowd belief.\\
-%(b) More specific research objectives: With such a model we can make predictions, answer questions about bias-vs-expertise, and evaluate question difficulty.
-It may be a good idea to use bullet points here for this.
-\\
-%(c) Context: Geopolitcal events. Describe the data we have briefly.\\
-%(d) Units of analysis: Group of experts with different self-reported expertise; different questions in the dataset.\\
-%Element 6: STRUCTURE OF THE PAPER\\
-%%
-%\ville{Much of the earlier paper can be included: define calibration and sharpness, normality around the calibrated crowd belief}
-%
-%\ville{Not studied in time-series - to our knowledge: REQUIRES LITERATURE REVIEW}
-%
-%\ville{Define calibration}
-%
-%\ville{Time series allows many more predictions and hence study of calibration -- this is not possible in static context where we would have only 166 predictions.
-Note that this method can still be applied to any given time point.}
-%
-%\ville{what are we hoping to learn from the data (applications) and why}
-%
-%\ville{Structure of the paper: theory, results (insight + precition)}
-%
 \section{Geopolitical Forecasting Data}
@@
 Since a typical expert participated only in a small subset of the 166 questions, the forecasters are considered indistinguishable conditional on the self-reported level of expertise.
-Tables \ref{DataStats} and \ref{ExpertiseTable} provide more relevant summary statistics on the data. Notice that the distribution of the self-reported expertise is skewed to the right
-%: some questions involved very few experts who associated themselves with the highest level of expertise. Notice also 
-and that some questions remained active longer than others.  For more details on the data and its collection see \citet{ungar2012good}.
+Tables \ref{DataStats} and \ref{ExpertiseTable} provide more relevant summary statistics on the data.
+Notice that the distribution of the self-reported expertise is skewed to the right and that some questions remained active longer than others.
+For more details on the data and its collection see \citet{ungar2012good}.
 
@@
 The points have been jittered slightly to make overlaps visible.
-%Overall, Figures \ref{Example1} and \ref{Example2} contain 872 forecasts made by 710 experts and 1,860 forecasts made by 1,376 experts, respectively. 
 The darkness of the points are positively associated with the self-reported expertise levels.
@@
 \hspace*{-1em}  \includegraphics[width= 0.49\textwidth]{Figures/ExamplePlot1}
-%\hspace*{-1em}  \includegraphics[width= 1.5in]{CalibrationPLATTSPH10}
-%\caption{$\beta = 1$, $\alpha = 0$}
 \label{Example1}
@@
 \hspace*{-1em}  \includegraphics[width= 0.49\textwidth]{Figures/ExamplePlot2}
-%\caption{\textit{Sample-Then-Calibrate} with logarithmic score}
 \label{Example2}
@@
 \end{table}
-
-%
-% For instance, two of the questions were
-%\begin{quote}\textit{
-%Will the expansion of the European bailout fund be ratified by all 17 Eurozone nations before 1 November 2011? %1021
-%}\end{quote}
-%and
-%\begin{quote}\textit{
-%Will the Nikkei 225 index finish trading at or above 9,500 on 30 September 2011? %1011
-%}
-%\end{quote}
 
@@
 Therefore it is not unreasonable to interpreted the hidden process as the crowd belief across time.
-%Each modeling task should begin with observations about the nature of the question.
-With a good understanding of the question, the modeler can proceed to construct a model that has an interpretable and intuitive connection to the real world.
 
@@
 
-%
-%\textit{Random noise, $\sigma^2_k$:} The variability of each observed value $Y_{i,t,k}$ around the biased hidden state $b_j \tilde{X}_{t,k}$.
-This can be thought of as being caused by momentary over-optimism (or pessimism), false beliefs, or other misconceptions.
-%Note that only the most informed expert making an objective estimation based solely on facts can be expected to give $X_t$ as his prediction.
-Unfortunately, human experts hardly make such objective estimates.
-Therefore this noise component is an important part of our model.
-%
-%\textit{Random signal, $\tau^2_k$:} This variability is caused by changes in the amount of true information that is available to the experts.
-It allows the model to incorporate changes or shocks to the underlying circumstances that ultimately decide the outcome of the event.
-%
-%\textit{Systematic noise, $b_{j}$:} This is a bias in the observed values given by the $j$th most expert group and can be interpreted as groupwise over- or under-confidence that pulls the observed values away from the hidden state values.
-%
-%\textit{Systematic signal, $\gamma_k$:} Since the event associated with the $k$th question in our dataset realizes by time $T_k$, the uncertainty in the question reduces as the current time point $t$ approaches the final time point $T_k$.
-This reduction in uncertainty can be viewed as a constant signal stream that pushes the hidden process towards either extreme as more time passes by.
-
-%These interpretations have been summarized in Table \ref{signalnoise}.
-%\begin{table}
-%\centering
-%\begin{tabular}{| c | c |  c | }\hline
-% Model Factors & Signal & Noise \\ 
-%Random & $\tau^2_k$ & $\sigma^2_k$ \\ 
-%Systematic & $\gamma_k$ & $b_{j,k}$ \\ \hline
-%\end{tabular}
-%\caption{The signal/noise structure of the dynamic linear model}
-%\label{signalnoise}
-%\end{table}
-
 \section{Model Estimation}
 \label{identifiability}
-%Unfortunately, the model as described by Equations (\ref{observedpr}) and (\ref{hiddenpr}) is not identifiable.
-To see this, let $a$ be any positive constant.
-Then $(\boldsymbol{b} a, X_{t,k}/a, a \tau) \neq (\boldsymbol{b}, X_{t,k}, \tau)$ yield the same likelihood.
-As a result, the parameter values tend to drift during the estimation process.
-Several solutions are as follows.
-%
-%\begin{enumerate}
-%\item[a.]  Place informative prior distributions on some of the non-identifiable parameters.
-These prior distributions, however, must be highly informative in order to provide identifiability.
-Since studying the expertise biases is one of the research objectives of this paper, placing a highly restrictive prior distribution on $\boldsymbol{b}$ is not satisfactory.
-Having strong apriori knowledge of $\tau_k$, on the other hand, seems difficult to justify.
-%\item[b.] The parameters could be constrained to reduce the number of degrees of freedom in the model.
-A popular choice is to fix one of the bias terms to some constant.
-For instance, the third expertise group could be chosen as the baseline group by fixing $b_3 = 1$.
-This, however, assumes that the third expertise group is unbiased.
-Such an assumption is hardly realistic and would hinder our analysis of under- and overconfidence among the expertise groups.
-%\item[c.] Reduce the degrees of freedom by introducing more information to the modeling process. 
-% \end{enumerate}
-% 
-%The next two subsections introduce two hierarchical approaches that combine solutions b.
-and c.
-to restore identifiability without sacrificing the interpretability of the model.
-The first approach, called  \textit{Sample-Then-Calibrate}, is flexible and results into an efficient sampling algorithm.
-The second approach, named \textit{Sample-Only}, on the other hand, produces uncertainty estimates for all the unknown parameters in a coherent, fully Bayesian manner.
-
-%
 
@@
 This section introduces a two-step procedure, called \textit{Sample-Then-Calibrate} or simply \textit{STC}, that achieves this goal in a flexible, yet efficient manner by first estimating  the model parameters under a constraint (\textit{Sampling Step}) and then performing a one-dimension optimization procedure to transform the constrained estimates into their unconstrained counterparts (\textit{Calibration Step}).
-
-%The second approach, named \textit{Sample-Only} (SO), on the other hand, produces uncertainty estimates for all the unknown parameters in a fully Bayesian manner.
 
@@
 
-%See Appendix \ref{appendix} for the technical details.
-
 \subsection{Calibration Step}
 \label{calibration_step}
-%First choose one of the $J$ groups as the baseline group. Denote this group by $j^*$ and fix its multiplicative bias term $b_{j^*} = 1$.  Denote the constrained version of the observed process by
-%\begin{eqnarray*}
-%\boldsymbol{Y}_{t, k} &=&  \boldsymbol{M}_k \boldsymbol{b} (1) X_{t, k}(1)+ \boldsymbol{v}_{t, k},
-%\end{eqnarray*}
-%where the input parameter emphasizes the constraint $b_{j^*} = 1$.
-Since this constrained model is identifiable, estimates of $\boldsymbol{b}(1)$ and $X_{t, k}(1)$ can be obtained.
-Denote these estimates by $\hat{\boldsymbol{b}}(1)$ and $\hat{X}_{t, k}(1)$, respectively.
 Since the model parameters can be estimated under a constrained version of $b_3$, the next step is to determine how $b_3$ should be constrained such that both the sharpness and calibration of the estimated hidden process are maximized.
@@
 \begin{eqnarray}
-%\hat{\beta}_{OSE} &=& \argmax_{\beta \in \R / \{0\}} \sum_{k=1}^K \frac{1}{T_k}  \sum_{t=1}^{T_k} \int_{\R} S(Z_k,  X_{k,t}(\beta)) d\P(X_{k,t} | \boldsymbol{Y}) \nonumber\\
-%&\approx& \argmax_{\beta \in \R / \{0\}} \sum_{k=1}^K \frac{1}{T_k} \sum_{t=1}^{T_k}   \frac{1}{G} \left( \sum_{g=1}^{G}  S\left(Z_k,  \hat{X}_{k,t}^{(g)}(\beta) \right) \right) \label{OSE}
  \hat{\beta}&=&  \argmax_{\beta \in \R / \{0\}} \sum_{k=1}^K \sum_{t=1}^{T_k}  S\left(Z_k, \hat{X}_{k,t}(1) / \beta \right) \label{OSE}
 \end{eqnarray}
-where $S$ is a strictly proper scoring rule such as 
-%that is a natural choice for binary-classification questions (\citet{buja2005loss}) and can be conveniently adapted to the question at hand (\citet{Gneiting04strictlyproper}). 
-the negative Brier score (\citet{Brier})
+where $S$ is a strictly proper scoring rule such as the negative Brier score (\citet{Brier})
 \begin{eqnarray*}
@@
 \end{eqnarray*}
-%The proper scoring rule can be conveniently adapted to the question at hand (\citet{Gneiting04strictlyproper}). 
-%Notice that Equation (\ref{OSE}) under $S_{LOG}$ is equivalent to finding the maximum likelihood estimator of the (inverse) slope coefficient for a (centered) logistic regression model with $Z_k$ as the response and $\hat{X}_{k,t}$ as the explanatory variable.
-%Since the expected utility under a strictly proper scoring rule for an observation $Z \sim p$ is uniquely maximized by forecasting exactly $p$ (\citet{Gneiting04strictlyproper}), the estimated log-odds $\hat{X}_{t, k}(1) / \hat{\beta}_{OSE}$ can be expected to be well-calibrated.
 Since it is not clear which rule should be used for predicting geopolitical events, such as the ones in our dataset, the \textit{Sample-Then-Calibrate} procedure is evaluated  separately under both rules in Sections \ref{syntheticData} and \ref{realData}.
@@
 
-%\subsection{Sample-Only}
-%In order to clarify the upcoming notation, denote the calibrated logit-probabilities and  the event indicators across all $K$ questions with $\boldsymbol{X}$ and $\boldsymbol{Z}$, respectively.
-A Bayesian version of Equation (\ref{OSE}) under $S_{LOG}$ can be derived by specifying a prior distribution $p(\beta | \boldsymbol{X})$ and a likelihood function $p(\boldsymbol{Z} | \beta, \boldsymbol{X})$.
-The posterior distribution of $\beta$ conditional on $\boldsymbol{X}$ is then given by $p(\beta | \boldsymbol{X}, \boldsymbol{Z}) \propto p( \boldsymbol{Z} | \beta, \boldsymbol{X}) p(\beta | \boldsymbol{X})$.
-Recall that Equation (\ref{OSE}) under $S_{LOG}$ is equivalent to fitting a logistic regression model with $Z_k$ as the response and $X_{k,t}$ as the explanatory variable.
-To give a Bayesian version of this logistic regression, the likelihood is chosen to be proportional to
-%\begin{eqnarray}
-%p( \boldsymbol{Z} | \beta, \boldsymbol{X}) &\propto& \prod_{k=1}^K \prod_{t=1}^{T_k}  \logit^{-1} \left(X_{k,t}/\beta  \right)^{Z_k} \left( 1-  \logit^{-1} \left( X_{k,t}/\beta  \right) \right)^{1-Z_k} \label{OSE2}
-%\end{eqnarray}
-%Similarly to  \citet{gelman2003bayesian} the prior is chosen to be locally uniform and independent, $p(1/\beta) \propto 1$.
-Posterior estimates of $\beta$ can be sampled from Equation (\ref{OSE2}) using generic sampling algorithms such as the Metropolis algorithm (\citet{metropolis1953equation}) or slice sampling (\citet{neal2003slice}).
-
-%Notice that with this procedure it is not necessary to estimate $\boldsymbol{b}$ under a constraint.
-Instead we could let $p(Z_k | X_{t,k}) \propto \logit^{-1} \left(X_{k,t}  \right)^{Z_k} \left( 1-  \logit^{-1} \left( X_{k,t}  \right) \right)^{1-Z_k}$ independently across all time points and questions similarly to Equation (\ref{OSE2}), and estimate the elements of the bias vector $\boldsymbol{b}$ directly.
-Even though this would make $\beta$ unnecessary, the resulting procedure, which is equivalent to the version with $\beta$, does not make a clear connection to the \textit{Sample-Then-Calibrate} approach.
-
-%Overall, this approach can be viewed as a Bayesian equivalent of Equation (\ref{OSE}) under the logarithmic score, $S_{LOG}$.
-%This assumes an independent $Z_k$ for each time point in the $k$th question.
-This is a coarse approximation since these indicators are, in fact, perfectly correlated.
-It does, however, reinforce good calibration and is therefore of practical use.
-
 \subsection{Discussion}
@@
 
-%In terms of the prior distributions, we assume that $\alpha$ and $\beta$ are a priori independent with 
-%\begin{eqnarray*}
-%\beta &\sim& \text{Cauchy}(1, 1/2.5)\\
-%\alpha &\sim& \text{Cauchy}(0, 10),
-%\end{eqnarray*}
-%which are essentially the weakly informative default priors introduced in \citet{gelman2008weakly}.
-The only difference is in the centering of the prior $p(\beta)$.
-\citet{gelman2008weakly} recommend centering this prior at 0.
-Since a priori we have no reason to assume under- or overconfidence of the $\tilde{X}_{t,k}$ process, the prior $p(\beta)$ should be centered at 1 instead.
-%
-%Unlike the \textit{Sample-Then-Calibrate} approach, the \textit{Sample-Only} approach models the uncertainty in estimating $(\beta, \alpha)$.
-The \textit{Sample-Only} approach also results into an algorithm that produces all estimates from a single procedure; hence, making it appear more coherent.
-The \textit{Sample-Then-Calibrate} approach, on other hand, enjoys the flexibility to choose a score function, $S$, that suits the application at hand (see \citet{Gneiting04strictlyproper}).
-Furthermore, the  \textit{Sample-Then-Calibrate} approach results into a slightly more efficient sampling procedure.
-
-%Even though both approaches assume a hierarchical model with a common $\boldsymbol{b}$ across the questions,
-
 \section{Extension: Polychotomous Outcomes}
@@
 \boldsymbol{X}_{t,k} &=&  \left[ \begin{matrix}X_{1,t,k} &X_{2,t,k} & \dots &X_{M-1,t,k} \end{matrix} \right]^T
-%&=&  \left[ \begin{matrix} X_{1,t,k} + \alpha_1 & X_{2,t,k} + \alpha_2  & \dots & X_{M-1,t,k} + \alpha_{M-1}  \end{matrix} \right]^T\\
-%&=&  \left[ \begin{matrix} \log \left( p_{1,t,k}/p_{M,t,k} \right) & \log \left( p_{2,t,k}/p_{M,t,k} \right) & \dots & \log \left( p_{M-1,t,k}/p_{M,t,k} \right) \end{matrix} \right]^T
 \end{eqnarray*}
@@
 Therefore, even though this paper focuses on modeling binary events, all the properties and discussion generalize trivially to the multi-outcome case.
-
-%\subsection*{Incomplete Data or Missing Observations}
-%If the observation $\boldsymbol{Y}_{t,k}$ is completely missing at time $t$, the update step is skipped and the state estimates are sampled from
-%\begin{align*}
-%\mathcal{N}(\gamma_k \hat{X}_{t-1|t-1}, \gamma^2 P_{t-1|t-1} + \tau^2)
-%\end{align*}
-%If, on the other hand, only some subset of the experts report their forecast at time $t$, then the observation vector $\boldsymbol{Y}_t$ has a smaller dimension.
-Fortunately, the formulae derived in the previous subsection remain valid even if the dimension of the observation vector is allowed to change; see, e.g., Theorem 2.1 in \textit{Kalman filter with outliers and missing observations} by Cipra and Romera.
 
@@
 \hspace*{-1em}  \includegraphics[width= 0.33\textwidth]{Figures/SyntheticSigma2}
-%\hspace*{-1em}  \includegraphics[width= 1.5in]{CalibrationPLATTSPH10}
-%\caption{$\beta = 1$, $\alpha = 0$}
 \label{SyntheticSigma2}
@@
 \hspace*{-1em} \includegraphics[width=  0.33\textwidth]{Figures/SyntheticBeta}
-%\caption{\textit{Sample-Then-Calibrate} with Brier score}
 \label{SyntheticBeta}
@@
 \hspace*{-1em} \includegraphics[width=  0.33\textwidth]{Figures/SyntheticK}
-%\caption{\textit{Sample-Only} }
 \label{SyntheticK}
@@
 \hspace*{-1em}  \includegraphics[width= 0.33\textwidth]{Figures/SyntheticSigma2bias}
-%\hspace*{-1em}  \includegraphics[width= 1.5in]{CalibrationPLATTSPH10}
-%\caption{$\beta = 1$, $\alpha = 0$}
 \label{SyntheticSigma2}
@@
 \hspace*{-1em} \includegraphics[width=  0.33\textwidth]{Figures/SyntheticBetabias}
-%\caption{\textit{Sample-Then-Calibrate} with Brier score}
 \label{SyntheticBeta}
@@
 \hspace*{-1em} \includegraphics[width=  0.33\textwidth]{Figures/SyntheticKbias}
-%\caption{\textit{Sample-Only} }
 \label{SyntheticK}
@@
 Since this implies ignoring a portion of the data, we decided to adopt the first approach by changing $p = 0.0$ and $1.0$ to $p = 0.01$ and $0.99$, respectively.
-
-%Given that the experts were allowed to report their probability forecasts to the second decimal place, this compromise was considered rather natural.
 
@@
  \tilde{Z}_k &= x\\
-%\{\tilde{p}_{i,t,k} | i = 1, \dots, I_k, t = 1, \dots, T_k\} &=  \begin{cases} 
-%\{1-p_{i,t,k} | i = 1, \dots, I_k, t = 1, \dots, T_k\} & \text{if } Z_k = 1-x\\
-%\{p_{i,t,k} | i = 1, \dots, I_k, t = 1, \dots, T_k\} & \text{if } Z_k = x 
 \tilde{p}_{i,t,k} &=  \begin{cases} 
@@
 The set $\left\{\left(\tilde{Z}_k, \{\tilde{p}_{i,t,k} | i = 1, \dots, I_k, t = 1, \dots, T_k \}\right) \right\}_{k=1}^K$ is a balanced version of the data.
-%Notice that this procedure extends trivially to multi-outcome (polychotomous) events because many of the multinomial models, such as the multinomial logistic regression (see, e.g., \citet{agresti2002categorical}) or the multi-outcome extension of our model, can be expressed as a set of independent binary-response models to which the above balancing procedure can be applied directly.
 This procedure was used to balance our real-world dataset both in terms of events and time points.
 The final output splits the events exactly in half ($|S_0| = |S_1| = 83$) such that number of time points in the first and second halves are 8,737 and 8,738, respectively.
-
-%\subsection{Incoherent and Imbalanced Data}
-%\label{practicalmatters}
-%The first issue that needs to be addressed arises from human experts making probability forecasts of 0 or 1 even when they are not completely sure of the outcome.
-For instance, each of the forecast pools associated with the 166 questions in our data contained both a zero and a one.
-Transforming such forecasts into the logit-space yields infinities which are highly influential and hence cause questions in model estimation.
-To avoid this, \citet{Ariely00theeffects} suggest changing $p = 0$ and $1$ to $p = 0.02$ and $0.98$, respectively.
-This is similar to winsorising which sets the extreme probabilities to a specified percentile of the data (see, e.g., \citet{hastings1947low} for more details on winsorising).
-\citet{Geo}, on the other hand, consider only probabilities that fall within a constrained interval, say $[0.001, 0.999]$, and discard the rest.
-Since this implies ignoring a portion of the data, we have adopted the first approach by changing $p = 0$ and $1$ to $p = 0.01$ and $0.99$, respectively.
-Given that the experts were allowed to report their probability forecasts with a precision of 0.01, this practical compromise was considered a natural choice.
-
-% The second issue is related to the distribution of the class labels in the dataset.
-If the set of events that occurred is much larger than the set of events that did not occur (or \textit{vice versa}), the dataset is called \textit{imbalanced}.
-On such data the model can end up over-focusing on the larger class and, as a results, give very accurate forecast performance over the larger class but poor forecast performance over the  smaller class (see, e.g., \citet{chen2009learning, wallace2012class}).
-% %Imbalanced data can also cause Platt calibration to overfit (see \citet{mitchell2007multi}). 
-%Furthermore, as discussed in Section \ref{Sample-Then-Calibrate-Ident}, balancing the data allows the model to imitate Platt Calibration without losing interpretability. 
-%
-%To describe the process of balancing the data, recall that our goal is to construct a model that is well-calibrated at any time point across all questions.
-The class labels in the data must therefore be balanced with respect to the total number of time points.
-The first step is to find a partition, $S_0$ and $S_1$, of the question indices, $\{1, 2, \dots, K\}$, such that the equality $\sum_{k \in S_0} T_k  = \sum_{k \in S_1} T_k$
-%is as closely approximated as possible.
-This is equivalent to an NP-hard question known in computer science as the \textit{Partition Problem}:  determine whether a given set of positive integers can be partitioned into two sets such that the sums of the two sets equal to each other (see, e.g., \citet{karmarkar1982differencing, hayes2002easiest}).
-A simple option is to use a greedy algorithm, which iterates through the values of $T_k$ in an ascending order, assigning each $T_k$ to the subset that currently has the smaller sum (see, e.g.
-\citet{kellerer2004knapsack, gent1996phase} for more details on the Partition Problem).
-After finding such a partition, the final step is to assign the class labels such that the labels for the questions in $S_x$ are equal to $x$ with $x = 0$ or $1$.
-In other words, let $\tilde{Z}_k = x$ and
-%\begin{align*}
-%\{\tilde{p}_{i,t,k | i = 1, \dots, I_k, t = 1, \dots, T_k}\} &=  \begin{cases} 
-%\{1-p_{i,t,k} | i = 1, \dots, I_k, t = 1, \dots, T_k\} & \text{if } Z_k = 1-x\\
-%\{p_{i,t,k} | i = 1, \dots, I_k, t = 1, \dots, T_k\} & \text{if } Z_k = x 
-%\end{cases}
-%\end{align*}
-%for $k \in S_x$.
-The final set $\left\{\left(\tilde{Z}_k, \{\tilde{p}_{i,t,k | i = 1, \dots, I_k, t = 1, \dots, T_k} \}\right) \right\}_{k=1}^K$ is a balanced version of the data.
-This procedure simultaneously balances our dataset both in terms of events and time points.
-The final output splits the events exactly in half ($|S_0| = |S_1| = 83$) such that number of time points in the first and second halves are 8,737 and 8,738, respectively.
-
-% Both of these approaches, however, are rather \textit{ad-hoc} and unsatisfactory: while the first one requires the user to pick a specified percentile or range within the unit interval, the second one ignores a portion of the data.
-% 
-% Therefore we put in our efforts to propose a simple, yet slightly more principled approach: First, fit a cubic smoothing spline to a set of equidistant probabilities in the open interval $(0,1)$ and their corresponding logit probabilities.
-Second, use this spline to extrapolate the extreme logit probability $\logit(1.0)$.
-Third, assign the negative of this extrapolated logit probability to the other extreme logit probability, $\logit(0.0)$.
-The initial choice of the probability grid matters: generally finer grids lead to larger extrapolated values for $\logit(1.0)$.
-Fortunately, the dataset at hand often offers a natural choice for the grid.
-For instance, in our case the experts reported their probability forecasts with an accuracy of 0.01.
-Therefore  $0.50, 0.51, \dots, 0.99$ gives us the initial probability grid.
-Under this choice, the extrapolation yields $\widehat{\logit(1.0)} \approx 5.509$.
-This is approximately equivalent to changing $p = 0.0$ and $1.0$ to $p = 0.004$ and $0.996$, respectively.
-% 
-% Overall, this approach seems more reasonable than the earlier described solutions by \citet{Ariely00theeffects} and  \citet{Geo} because (a) it does not discard any data and (b) it uses the application-specific scale to construct a smooth, strictly increasing, yet finite transformation from probability scale to logit scale.
-
-%\subsection{Unbalanced Data}
-%Since in our case the explanatory variables are probabilities associated with dichotomous events, the data can be balanced by choosing between $\{p_i\}_{i=1}^{I_k}$ and $\{1-p_i\}_{i=1}^{I_k}$ for the first $K/2$ events depending on the whether the events occurred or not, respectively.
-That is, if the event occurred (or did not occur), choose $\{p_i\}_{i=1}^{I_k}$ (or $\{1-p_i\}_{i=1}^{I_k}$, respectively).
-Set $Z_k$ for $k = 1, \dots, K/2$ equal to 1.
-For the remaining $K/2$ questions, on the other hand, choose $\{1-p_i\}_{i=1}^{I_k}$ (or $\{p_i\}_{i=1}^{I_k}$) if the event happened (or did not happen, respectively) and then set $Z_k$ for $k = K/2+1, \dots, K$ equal to 0.
-% event $k = 1, \dots, K$ such that
-%\begin{eqnarray}
-%|\{k : \{p_i\}_{i=1}^{I_k} \text{, } Z_k = 1 \}| + |\{k : \{1-p_i\}_{i=1}^{I_k}  \text{, } Z_k = 0\}| &=& K/2 \label{balanceddata}
-%\end{eqnarray}
-%and then defining a new set of event indicators denoted $\{Z_k^b\}_{k=1}^K$, where $Z_k^b = 1$ if  $k$th event is in either set described in Equation (\ref{balanceddata}) and $Z_k^b = 0$ otherwise. 
-%
-
-%To make this more specific, let $S_1$ be an index set for events that both occurred and for which we choose $\{p_i\}_{i=1}^{I_k}$.
-Correspondingly, let $S_2$ be an index set for events that both did not occur and for which we choose $\{1-p_i\}_{i=1}^{I_k}$.
-The dataset is balanced if the sets $S_1$ and $S_2$ satisfy
-%\begin{eqnarray*}
-%\sum_{k \in S_1 \cup S_2} T_k  &=& \frac{\sum_{k=1}^K T_k}{2}
-%%|\{(k, t) : \{p_i\}_{i=1}^{I_k} \text{, } Z_k = 1\text{, } t \in \{1, \dots, T_k\} \}|\\
-%% + |\{(k,t) : \{1-p_i\}_{i=1}^{I_k}  \text{, } Z_k = 0\text{, } t \in \{1, \dots, T_k\} \}| &=& \frac{\sum_{k=1}^K T_k}{2}
-%\end{eqnarray*}
-%with $Z_k = 1$ if $k \in S_1 \cup S_2$ and $Z_k = 0$ otherwise. This i
 
@@
 \end{cases}
-%\hat{p}_1 &=& \bar{p}_1 \\	
-%\hat{p}_t &=& \alpha \bar{p}_t  + (1-\alpha) \hat{p}_{t-1}, \text{ for } t > 1
 \end{eqnarray*}
@@
 \end{eqnarray*}
-
-%
-%\item Many time-series models have been developed for univariate time series.
-Our data, however, is not univariate nor directly multivariate.
-The response vector changes in size across time.
-Therefore in order to use the standard time series models, we turn the data into a univariate time series by considering the average forecast at each time point.
-%
-%Autoregressive integrated moving average, ARIMA($p,d,q$), applied to the average forecasts given at each time point, $t$. The parameters, $p,d$, and $q$, are learnt form the training set by
-%\begin{eqnarray*}
-%(p,d,q) &=& \argmin_{p,d,q \in \mathbb{Z}^+} \sum_{k=1}^K \sum_{t=1}^{T_k} \left(Z_k - \hat{p}_{t,k}\right)^2,
-%\end{eqnarray*}
-%where $\hat{p}_{t,k}$ is the ARIMA prediction for the $k$th question at time $t$. 
-%
-%\item  \textit{Exponentially Weighted Moving Logit Aggregator} (EWMLA).
-This is a moving version of the aggregator, $\hat{p}_G(a)$, introduced in \cite{satopaa}.
-If $\boldsymbol{p}_{t,k}$ is a vector collecting all the probability forecasts made for the $k$th question at time $t$, then the EWMLA forecasts are found recursively as
-%\begin{eqnarray*}
-%\hat{p}_{t,k}(\alpha, \nu) =
-%\begin{cases}
-%G_{t,k}(\nu ) & \text{ for } t  = 1 \\
-%\alpha G_{t,k}(\nu )  + (1-\alpha) \hat{p}_{t-1,k}(\alpha, \nu)  & \text{ for } t > 1
-%\end{cases}
-%\end{eqnarray*}
-%where
-%\begin{eqnarray*}
-%G_{t,k}(\nu )  &=& \frac{\left[ \prod\limits_{i=1}^{N_{t,k}} \left( \frac{p_{i, t, k}}{1-p_{i, t, k}} \right)^{1/N_{t,k}} \right]^\nu}{1+\left[ \prod\limits_{i=1}^{N_{t,k}} \left( \frac{p_{i, t, k}}{1-p_{i, t, k}} \right)^{1/N_{t,k}} \right]^\nu}
-%\end{eqnarray*}
-%The tuning parameters, $(\alpha, \nu) $, are learned from the training set by
-%\begin{eqnarray*}
-%(\hat{\alpha}, \hat{\nu}) &=& \argmin_{\nu \in \R, \alpha \in [0,1]} \sum_{k \in S_{train}} \sum_{t=1}^{T_k} \left(Z_k - \hat{p}_{t,k}(\alpha, \nu)\right)^2
-%\end{eqnarray*}
 
@@
 
-% Requires the booktabs if the memoir class is not being used
 \begin{table}[htbp]
@@
 
-%% Requires the booktabs if the memoir class is not being used
-%\begin{table}[htbp]
-%   \centering
-%   
-%   %\topcaption{Table captions are better up top} % requires the topcapt package
-%   \begin{tabular}{l llllllll} % Column formatting, @{} suppresses leading/trailing space
-%\hline
-%& \multicolumn{2}{ c }{All Problems} & \multicolumn{2}{ c }{Short} & \multicolumn{2}{ c }{Medium} & \multicolumn{2}{ c }{Long}\\
-%Method & Days &Problems & Days &Problems & Days &Problems & Days &Problems\\ \hline
-%SDLM & 0.101 (0.024) & 0.066 (0.013) & 0.098 (0.024) & 0.102 (0.025)\\ 
-%BSTC-Log & 0.097 (0.045) & 0.053 (0.022) & 0.100 (0.046) & 0.099 (0.046)\\ 
-%STC-Bri & 0.096 (0.036) & 0.056 (0.018) & 0.097 (0.036) & 0.098 (0.037)\\ 
-%STC-Log & 0.096 (0.036) & 0.056 (0.018) & 0.096 (0.036) & 0.098 (0.037)\\ 
-%EWMBA & 0.102 (0.041) & 0.060 (0.015) & 0.110 (0.040) & 0.103 (0.042)\\ 
-%EWMLA & 0.102 (0.040) & 0.061 (0.017) & 0.111 (0.046) & 0.103 (0.040)\\ 
-%EWMA & 0.111 (0.020) & 0.089 (0.010) & 0.111 (0.019) & 0.112 (0.021)\\ 
-%   \end{tabular}
-%   \caption{Mean Brier Scores based on 10-fold cross-validation. The bolded values indicate the lowest scores in each column.}
-%   \label{prediction}
-%\end{table}
-
 \subsection{In- and Out-of-Sample Sharpness and Calibration}
@@
   \includegraphics[width= 0.28\textwidth]{Figures/CalibrationNONE10}
-%\hspace*{-1em}  \includegraphics[width= 1.5in]{CalibrationPLATTSPH10}
-%\caption{$\beta = 1$, $\alpha = 0$}
 \label{CalibrationNONEa}
@@
 \hspace*{-3em}  \includegraphics[width= 0.28\textwidth]{Figures/CalibrationLOG10}
-%\caption{\textit{Sample-Then-Calibrate} with logarithmic score}
 \label{CalibrationPLATTLOG10}
@@
 \hspace*{-3em} \includegraphics[width= 0.28\textwidth]{Figures/CalibrationBRI10}
-%\caption{\textit{Sample-Then-Calibrate} with Brier score}
 \label{CalibrationPLATTBRI10}
@@
 \hspace*{-3em} \includegraphics[width= 0.28\textwidth]{Figures/CalibrationFULL10}
-%\caption{\textit{Sample-Only} }
 \label{CalibrationPLATTFULL10}
@@
 \hspace*{-1.1em}  \includegraphics[width= 0.28\textwidth]{Figures/CalibrationNONE}
-%\hspace*{-1em}  \includegraphics[width= 1.5in]{CalibrationPLATTSPH10}
-%\caption{$\beta = 1$, $\alpha = 0$}
 \label{CalibrationNONEb}
@@
 \hspace*{-3em}  \includegraphics[width= 0.28\textwidth]{Figures/CalibrationSTO-Log}
-%\caption{\textit{Sample-Then-Calibrate} with logarithmic score}
 \label{CalibrationSTC-Log}
@@
 \hspace*{-3em} \includegraphics[width= 0.28\textwidth]{Figures/CalibrationSTO-Bri}
-%\caption{\textit{Sample-Then-Calibrate} with Brier score}
 \label{CalibrationSTC-Bri}
@@
 \hspace*{-3em} \includegraphics[width= 0.28\textwidth]{Figures/CalibrationSO}
-%\caption{\textit{Sample-Only} }
 \label{CalibrationBSTC-Log}
@@
 
-% Sharpness refers to the distribution of the forecast probabilities.
-The closer the forecasts are to zero and one, the sharper the model is said to be.
-This is an important feature in forecasting, where the main objective typically is to maximize sharpness with respect to calibration (see, e.g., \cite{Ranjan08,raftery2005using}).
-
-% Overall, both in-sample calibration and sharpness increase from (a) to (d) with \textit{BSTC-Log} being the sharpest and most calibrated.
-
 \subsection{Group-Level Expertise Bias}
 \label{ExpertBias}
-%In Section \ref{calibration} it was briefly mentioned that the experts as a group are under-confident relative to the calibrated crowd belief.
-%This section analyzes the biases across different levels of self-reported expertise. 
-%Figure \ref{Biases} is similar in nature to the plots seen in the previous section.
-This time, however, instead of comparing the predicted probabilities against the empirical occurrence rate, we are comparing the expected expert predictions against the calibrated probabilities under different modeling approaches; that is,  $\logit^{-1}(b_j (X +\alpha))$ against $\logit^{-1}(X)$ for $j = 1, \dots, 5$ and $X \in \R$.
-Furthermore, we separate the results by the self-reported expertise levels.
-%Recall from Section \ref{data} that we asked the experts to self-assess their expertise level on a 1-to-5 scale (1 = Not At All Expert to 5 = Extremely Expert). 
-%Recall from Section \ref{model} that the vector $\boldsymbol{b}$ collects the bias terms specific to each self-reported expertise group. 
 Recall from Section \ref{data} that the experts were asked to self-assess their expertise level (on a 1-to-5 scale with 1 = Not At All Expert to 5 = Extremely Expert) on  all of the 166 questions in our real-world dataset.
@@
 Overall, finding under-confidence is a rather surprising result given that many previous studies have been conducted to show that experts are likely to be over-confident.
-For instance,  \citet{lichtenstein1977calibration, morgan1992uncertainty, bier2004implications} summarize the results from numerous calibration studies, and conclude that experts are systematically over-conÞdent about their probability assessments.
+For instance,  \citet{lichtenstein1977calibration, morgan1992uncertainty, bier2004implications} summarize the results from numerous calibration studies, and conclude that experts are systematically over-conﬁdent about their probability assessments.
 Our result, however, is a statement about groups of experts and hence does not invalidate the possibility of the individual experts being overconfident.
@@
 
-%This follows from the well-known Simpson's Paradox (\citet{simpson1951interpretation}) that  explains why the trend of the individuals can be reversed after they are aggregated into a single group.
-
-%The well-informed experts, in the contrary, place themselves well among the two groups with the levels of self-reported expertise.
-
 \begin{figure}[!ht]
-%\centering
-%\hspace*{-42em} 	\includegraphics[width=  0.8\textwidth]{Figures/Figures/LegendBias} % requires the graphicx package
-%\vspace{-2.7em}
-%
-%\subfigure[]{
-%  \includegraphics[width = 0.32 \textwidth]{Figures/BiasesPLATTLOG}
-%%\caption{\textit{Sample-Then-Calibrate} with logarithmic score}
-%\label{BiasesPLATTLOG}
-%}
-%\subfigure[]{
-%\hspace*{-1em} \includegraphics[width = 0.32 \textwidth]{Figures/BiasesPLATTBRI}
-%%\caption{\textit{Sample-Then-Calibrate} with Brier score}
-%\label{BiasesPLATTBRI}
-%}
-
-%\subfigure[]{
-%\hspace*{-1em} \includegraphics[width = 0.5 \textwidth]{Figures/BiasesFULL}
-%%\caption{\textit{Sample-Only} }
-%\label{BiasesFULL}
-%}
 \vspace*{-1em} 
 \includegraphics{Figures/BiasesBoxplots}
-%\caption{\textit{Sample-Then-Calibrate} with Brier score}
 \label{BiasesLOGBoxplots}
-
-%\subfigure[STC-Log]{
-%\hspace*{-1em} \includegraphics[width = 0.5 \textwidth]{Figures/BiasesLOGBoxplots}
-%%\caption{\textit{Sample-Then-Calibrate} with Brier score}
-%\label{BiasesLOGBoxplots}
-%}
-%\subfigure[STC-Bri]{
-%\hspace*{-1em} \includegraphics[width = 0.5 \textwidth]{Figures/BiasesBRIBoxplots}
-%%\caption{\textit{Sample-Then-Calibrate} with Brier score}
-%\label{BiasesBRIBoxplots}
-%}
-%\subfigure[BSTC-Log]{
-%\hspace*{-1em} \includegraphics[width = 0.32 \textwidth]{Figures/BiasesFULLBoxplots}
-%%\caption{\textit{Sample-Then-Calibrate} with Brier score}
-%\label{BiasesFULLBoxplots}
-%}
 
@@
 
-%Under the \textit{Sample-Only} approach, the posterior mean of $\alpha$ is around 0.155, which is very small in the logit scale. Therefore 
-%\begin{eqnarray*}
-%\E[\boldsymbol{Y}_{t, k}] &=&  \boldsymbol{M}_k \boldsymbol{b} (X_{t, k} +\alpha) \approx \boldsymbol{M}_k \boldsymbol{b} X_{t, k}
-%\end{eqnarray*}
-%and t
 More specific group-level inference can be performed by computing probabilities based the estimated posterior distributions.
 For instance, the posterior probability that the most expert group is the least under-confident is approximately equal to $1.0$, and
-%More specifically, $\P(b_5 > b_j \text{ for } j = 1, \dots, 4) \approx 1.0$.
 the posterior probability of a strictly decreasing level of under-confidence is approximately 0.87.
- %In other words, $\P(b_5 > b_4 > \dots > b_1) \approx 0.871$.  
 The latter probability is driven down by the inseparability of the two groups associated with the lowest levels of self-reported expertise.
@@
 
-%This constant drives the hidden process and is therefore directly indicative of the rate at which the calibrated crowd belief approaches 0.0 or 1.0.
-If this rate is very slow, the day-to-day accumulation of available and relevant information does not
-
-%is, how early on does the question become solvable.
-
-%Since $\sigma^2_k$ has a positive and $\gamma_k$ a negative relationship with question difficulty, these two quantities could be then combined via multiplication, $\sigma^2_k / \gamma_k$, into a single measure of question difficulty.
-
 As a final illustration of our model, consider the two questions used for illustrative purposes in Section \ref{data}.
@@
 \hspace*{-1em}  \includegraphics[width= 0.49\textwidth]{Figures/ExamplePlot1b}
-%\hspace*{-1em}  \includegraphics[width= 1.5in]{CalibrationPLATTSPH10}
-%\caption{$\beta = 1$, $\alpha = 0$}
 \label{Example1b}
@@
 \hspace*{-1em}  \includegraphics[width= 0.49\textwidth]{Figures/ExamplePlot2b}
-%\caption{\textit{Sample-Then-Calibrate} with logarithmic score}
 \label{Example2b}
@@
 \end{figure}
-
-%
-%\begin{enumerate}
-%\item Expert Noise $ \sigma^2$.
-%\item Distance from 0.5 and how quickly the process approaches the extreme probabilities 0.0 and 1.0. This is determined largely by the value of $\gamma$. 
-%\item Rank-correlation
-%\end{enumerate}
 
@@
 
-%Despite these and potentially other unmentioned limitations, the model yields very good predictive performance and calibration on real data.
-
-%
-%Ideally the model could be further developed into a non-hierarchical version that does not share any information among the forecasting questions.
-Such a model could be used to study group-level bias across different questions and to make aggregate probability forecasts without a separate training set.
-Due to \textit{perfect separation}, however, some sharing among the questions is necessary unless the modeler has access to prior information on the bias terms, the hidden states, or their variability.
-Under strong enough prior information the estimation procedure can be made regular enough to be performed separately for each question.
-This, however, was considered to be outside the scope of this paper and was therefore not discussed in detail.
-%
-
-%
-%This leads to several research objectives including
-%\begin{itemize}
-%\item the analysis of under- and overconfidence across different levels of self-reported expertise,
-%\item confidence bands for the crowd belief,
-%\item accurate probability forecasts, and
-%\item many question-specific quantities that have easy interpretations and can be used to gain novel insight in the social sciences. 
-%\end{itemize}
-%
-
-%Allow observed variance to vary across expert groups. Allow more terms (e.g. the bias and variance terms) to vary over time. Model the bias based on expert characteristic. Problem difficulty.
-
 \appendix
@@
 \label{appendix}
-%The implementation of the hierarchical procedure introduced in Section \ref{model} consist of two steps: \ref{GibbsSampler} Run the Gibbs sampler that to obtain estimates of the constrained parameters.
-\ref{Optimization} Perform a one-dimensional maximization to transform the constrained parameter estimates into their unconstrained counterparts.
-
-%The main difference arises in the way the hidden states and the calibration parameter, $\beta$, are sampled.
-
-%\subsection{Technical Details: \textit{Sample-Then-Calibrate}}
-%\label{Sample-Then-Calibrate}
-%This subsection describes the details of \textit{Sample-Then-Calibrate} approach one block of model parameters at a time.
-%\subsection{Sampling Step}
-%\label{GibbsSampler}
+
 The Gibbs sampler iteratively samples all the unknown parameters from their full-conditionals one block of parameters at a time.
@@
 
-%
-%Adopting the non-informative prior $p(\boldsymbol{b}, \sigma_k^2 | \boldsymbol{X}_k) \propto \sigma_k^{-2}$, the parameters $\boldsymbol{b}_k$ and $\sigma^2_k$ can be sampled via a standard Bayesian regression setup as follows.
-%\begin{eqnarray*}
-%\boldsymbol{b}_k | \dots &\sim& \mathcal{MVN}_J \left( (\boldsymbol{X}_k^T \boldsymbol{X}_k)^{-1} \boldsymbol{X}_k^T \boldsymbol{Y}_k, (\boldsymbol{X}_k^T \boldsymbol{X}_k)^{-1} \sigma^2_k\right) \\
-%\sigma^2_k | \dots &\sim& \text{Inv-}\chi^2\left(|\boldsymbol{Y}_k| - J, \frac{1}{|\boldsymbol{Y}_k| - J} (\boldsymbol{Y}_k - \boldsymbol{X}_k \boldsymbol{b}_k)^T (\boldsymbol{Y}_k - \boldsymbol{X}_k \boldsymbol{b}_k) \right),
-%\end{eqnarray*}
-%where the final distribution is scaled inverse-$\chi^2$ (see, e.g., \citet{gelman2003bayesian}).
-For identifiability reasons, one of the elements of $\boldsymbol{b}_k$ should be fixed to some pre-specified constant, say, 1.
-The remaining elements can be estimated separately for each question.
-%
-
 \begin{center}
@@
 
-%\begin{center}
-%\framecolorbox[\textwidth]{Figures/gray}{gray!15}{Estimate $\beta$}
-%\end{center}
-%\subsection{Calibration Step}
-%\label{Optimization}
-%
-%The calibration parameter $\beta$ is estimated via a one-dimensional maximization procedure
-%\begin{eqnarray*}
-% \hat{\beta}&=&  \argmax_{\beta \in \R / \{0\}} \sum_{k=1}^K \sum_{t=1}^{T_k}  S\left(Z_k, \hat{X}_{k,t}(1) / \beta \right),\\
-%\end{eqnarray*}
-%where $S$ is a proper scoring rule. The unconstrained parameter estimates are then given by
-%\begin{eqnarray*}
-% \hat{X}_{t,k}&=& \hat{X}_{k,t}(1) / \hat{\beta} \\
-% \hat{\boldsymbol{b}}_{t,k}&=& \hat{\boldsymbol{b}}(1) \hat{\beta}\\
-%  \hat{\tau}_{k}^2&=& \hat{\tau}_{k}^2(1) \hat{\beta}^2\\
-%  \hat{\sigma}_{k}^2&=& \hat{\sigma}_{k}^2(1)\\
-%  \hat{\gamma}_{k}&=& \hat{\gamma}_{k}(1)
-%\end{eqnarray*}
-
-%\subsection{Technical Details: \textit{Sample-Only}}
-%
-%This subsection describes the details of \textit{Sample-Only} approach one block of model parameters at a time.
-%\begin{center}
-%\framecolorbox[\textwidth]{Figures/gray}{gray!15}{Sample $X_{t,k}$}
-%\end{center}
-%Since the \textit{Sample-Only} approach conditions on $\{Z_k\}_{k=1}^K$, the FFBS algorithm cannot be used for sampling the hidden states.
-Instead, he hidden states are sampled in an increasing order one at a time from
-%\begin{eqnarray*}
-%X_{t,k} | \dots &\propto&  \logit^{-1} \left( X_{k,t}/\beta \right)^{Z_k} \left( 1-  \logit^{-1} \left( X_{k,t}/\beta \right) \right)^{1-Z_k}\\
-%&& \times \exp\left( -\frac{1}{2\sigma_k^{2}} \left( \boldsymbol{Y}_{t,k} - \boldsymbol{M}_k \boldsymbol{b}X_{t,k}\right)^T \left( \boldsymbol{Y}_{t,k} - \boldsymbol{M}_k \boldsymbol{b}X_{t,k} \right)\right)\\
-%&& \times \exp\left(-\frac{1}{2\tau^{2}_k}\left(X_{t,k} - \gamma_k X_{t-1,k} \right)^2 \right),
-%\end{eqnarray*}
-%where the initial state is sampled from $X_{0,k} \sim \mathcal{N}(0,1)$.
-The sampling is done via the Metropolis algorithm (\citet{metropolis1953equation}) with a Normal proposal centered at the current value and variance tuned to give an acceptance ratio around 50\%.
-%
-%
-%% Since the final forecasts $\{\boldsymbol{Y}_{T_k, k}\}_{k=1}^K$ contain all the information within $\{Z_k\}_{k=1}^K$, we have that 
-%%\begin{eqnarray*}
-%%p(\boldsymbol{X}_{k} | \boldsymbol{Y}_{1:T_k, k},  Z_k, \dots) &=& p(\boldsymbol{X}_{k} | \boldsymbol{Y}_{1:T_k, k}, \dots)
-%%\end{eqnarray*}
-%%Therefore the FFBS-algorithm can be used to sample the hidden states also in the \textit{Sample-Only} approach.
-%
-%
-%
-%\begin{center}
-%\framecolorbox[\textwidth]{Figures/gray}{gray!15}{Sample $\left(\boldsymbol{b}, \sigma^2\right)$ and $\left(\gamma, \tau^2\right)$}
-%\end{center}
-%The same as for the \text{Sample-Then-Calibrate} approach. See Section \ref{Sample-Then-Calibrate}.
-%
-%\begin{center}
-%\framecolorbox[\textwidth]{Figures/gray}{gray!15}{Sample $\beta$}
-%\end{center}
-%This is equivalent to logistic regression. By adopting a locally uniform prior $p(1/\beta) \propto 1$, the calibration parameter can be sampled from
-%\begin{eqnarray*}
-%\beta | \dots &\propto& \prod_{k=1}^K \prod_{t=1}^{T_k}  \logit^{-1} \left( X_{k,t}/\beta \right)^{Z_k} \left( 1-  \logit^{-1} \left( X_{k,t}/\beta \right) \right)^{1-Z_k}
-%\end{eqnarray*}
-%This is performed via the Metropolis algorithm (\citet{metropolis1953equation})  with a Normal proposal centered at the current value and variance tuned to give an acceptance ratio around 50\%.
-
-% \bibliographystyle{plainnat}
 \bibliographystyle{imsart-nameyear}

### 2013-08-21 paper.tex
Edited Abstract

@@
 \begin{abstract}
-Most research on subjective probability aggregation concentrates on procedures that consult the expert advice only once before making the final aggregation.
-This paper makes an advance towards unexplored areas of probability aggregation by considering a dynamic context with a time period during which the experts are allowed to update their beliefs.
-The analysis is performed on a real-world dataset that includes over 2,300 experts making probability forecasts on a number of 166 different international political events.
-The model is able to take  into account the forecaster's level of self-reported expertise and produce aggregate probabilities that are sharp and well-calibrated both in- and out-of-sample.
-Using the aggregate probabilities as the no-bias reference-line, the groups of experts with different levels of self-reported expertise are found to be under-confident, with the level of under-confidence decreasing as a function of self-reported expertise.
+Most subjective probability aggregation procedures use a single
+probability judgement from each expert, even though it is common for
+experts studying real problems to update their probability estimates
+over time.  This paper makes an advance towards unexplored areas of
+probability aggregation by considering a dynamic context in which
+experts are allowed to update their beliefs at random intervals.  We
+demonstrate our method on a real-world dataset that includes over
+2,300 experts making multiple probability forecasts on different
+subsets of 166 different international political events. We present a
+dynamic model that takes into account the forecaster’s level of
+self-reported expertise and produces aggregate probabilities that are
+sharp and well-calibrated both in- and out-of-sample. Using the
+aggregate probabilities as the no-bias reference-line, the lower the
+self-reported expertise in a group, the more under-confident it is
+found to be.
 \end{abstract}

### 2013-08-23 paper.tex
Edits

@@
 \begin{abstract}
-Most subjective probability aggregation procedures use a single
-probability judgement from each expert, even though it is common for
-experts studying real problems to update their probability estimates
-over time.  This paper makes an advance towards unexplored areas of
-probability aggregation by considering a dynamic context in which
-experts are allowed to update their beliefs at random intervals.  We
-demonstrate our method on a real-world dataset that includes over
-2,300 experts making multiple probability forecasts on different
-subsets of 166 different international political events. We present a
-dynamic model that takes into account the forecaster’s level of
-self-reported expertise and produces aggregate probabilities that are
-sharp and well-calibrated both in- and out-of-sample. Using the
-aggregate probabilities as the no-bias reference-line, the lower the
-self-reported expertise in a group, the more under-confident it is
-found to be.
+Most subjective probability aggregation procedures use a single probability judgement from each expert, even though it is common for experts studying real problems to update their probability estimates over time.
+This paper makes an advance towards unexplored areas of probability aggregation by considering a dynamic context in which experts are allowed to update their beliefs at random intervals.
+The updates are assumed to occur very infrequently, resulting into a highly sparse dataset that cannot be modeled by standard time-series procedures.
+In response to the lack of appropriate methodology, this paper presents a dynamic model that takes into account the forecaster’s level of self-reported expertise and produces aggregate probabilities that are sharp and well-calibrated both in- and out-of-sample.
+The model is demonstrated on a real-world dataset that includes over 2,300 experts making multiple probability forecasts on different subsets of 166 different international political events.
 \end{abstract}
@@
 Individual experts can differ radically from one another in their abilities to assess probabilities of future events.
-The quality of their assessments is typically evaluated with an important attribute called \textit{calibration}.
-An expert is said to be well-calibrated if
-the proportion of events that actually occurred is, say, 60\% for all those events for which he assessed a probability of 0.60.
-Even though several experiments have been conducted to show that experts are generally poorly calibrated (see, e.g., \citet{cooke1991experts, shlyakhter1994quantifying}), some personal differences among the experts have been found to associate with calibration.
-In particular, \citet{wright1994coherence} argue that a higher self-rated expertise on the subject matter implies better calibration.
-
-Unfortunately, calibration by itself does not guarantee usefulness of a probability assessment.
-For instance, if the empirical frequency of  the last 10 years of a relatively stationary process is 45\%, the expert can assess a constant probability of 0.45 and turn out well-calibrated.
-His probability, however, is not very useful in decision making.
-In general, a probability assessments that is near 0.0 or 1.0, is considered more informative, and an expert who makes such assessments is called \textit{sharp}.
-In the forecasting literature sharpness is often combined with calibration into a single goal of maximizing sharpness subject to calibration (see, e.g., \citet{raftery2005using, Ranjan08}).
-An expert that meets this goal is considered to give perfect probability forecasts (\citet{gneiting2004verifying}).
-Even though expecting perfect forecasts from an expert is not reasonable in practice, this criterion has led to a wide range of novel and insightful observations in probability forecasting.
+The quality of a probability assessment is typically evaluated in regards to an important attribute called \textit{calibration}.
+An expert is said to be well-calibrated if his probabilities agree with the long-term frequency of the events.
+This means that, for instance, the proportion of events that actually occurred is 60\% for all those events for which he assessed a probability of 0.60.
+Even though several experiments have been conducted to show that experts are generally poorly calibrated (see, e.g., \citet{cooke1991experts, shlyakhter1994quantifying}), the level of expertise has been found to be associated with calibration.
+In particular, \citet{wright1994coherence} argue that a higher self-reported expertise on the subject matter implies better calibration.
+
+Unfortunately, calibration by itself does not guarantee usefulness.
+To see this, consider a relatively stationary process, such as rain in some geographic region, whose empirical frequency of occurrence in the last 10 years is 45\%.
+In this setting, an expert can assess a constant probability of 0.45 and be well-calibrated.
+This probability, however, is very near 0.5 and hence too uninformative to be useful in decision making.
+Compare this to the advice of  another expert whose typical probability assessment is near 0.0 or 1.0; that is, he is a \textit{sharp} assessor.
+If the expert is both sharp and well-calibrated, then he is able to forecast the behavior of the process both with high certainty and accuracy.
+From the decision-maker's perspective, such advice is highly useful as it involves very little risk.
+Therefore the goal in probability estimation should be to maximize sharpness subject to calibration (see, e.g., \citet{raftery2005using, Ranjan08}).
+This is a well-known goal that has led to a wide range of novel and insightful observations in probability forecasting literature.
 
@@
 Being motivated by the long list of applications of probability forecasts, including medical diagnosis (\citet{wilson1998prediction, pepe2003statistical}, political and socio-economic foresight (\citet{tetlock2005expert}), and meteorology (\citet{sanders1963subjective, vislocky1995improved, baars2005performance}), researchers have proposed many different approaches to combining probability forecasts (see, e.g., \citet{Ranjan08, satopaa} for some recent studies, and \citet{Genest, Wallsten97evaluatingand, clemen2007aggregating, primo2009calibration} for a comprehensive overview).
-The general focus, however, has been on developing one-time aggregation procedures that consult the experts' advice only once before the future event resolves.
-
-Therefore many complex areas of probability aggregation still remain rather unexplored.
-For instance, consider an investor aiming to assess whether a stock index will finish trading at or above some threshold on a given date.
-In order to maximize his overall predictive accuracy, he may consult a group of experts every day and adjust his estimate of the aggregate probability accordingly.
-Since the experts are allowed to update their probability assessments on a daily basis, the aggregation should be performed by taking into account the correlation and time-series nature of their advice.
-Analyzing this kind of future events under daily updating of the forecasts is the main focus of this paper.
-The core research goal is to construct an interpretable time-series model that incorporates self-reported expertise in order to estimate a sharp and well-calibrated version of the crowd belief.
+The general focus, however, has been on developing one-time aggregation procedures that consult the expert's advice only once before the event resolves.
+
+Consequently, many areas of probability aggregation still remain rather unexplored.
+For instance, consider an investor aiming to assess whether a stock index will finish trading above some threshold on a given date.
+In order to maximize his overall predictive accuracy, he may consult a group of experts repeatedly over a period of time and adjust his estimate of the aggregate probability accordingly.
+Since the experts are allowed to update their probability assessments, the aggregation should be performed by taking into account the temporal correlation in their advice.
+Many standard time-series procedures could be used to perform this aggregation as long as the experts update their advice consistently on a daily basis.
+
+%If the consulted experts diligently make probability assessments on a daily basis, many previously-studied time-series procedures could be used to aggregate their assessments.
+
+This paper adds another layer of complexity by assuming a heterogeneous set of experts, most of whom only make one or two probability assessments over the hundred or so days before the event resolves.
+The problem at hand is therefore strikingly different from many time-series estimation problems, where one has an observation at every time point -- or almost every time point.
+Therefore standard time-series procedures like ARIMA (see, e.g., \citet{mills1991time}) are not applicable without ignoring the heterogeneity among the experts.
+In response to the lack of procedures to model sparse and heterogeneous probability forecasting data, this paper introduces an interpretable time-series model that incorporates self-reported expertise and captures a sharp and well-calibrated estimate of the crowd belief.
 The model is then used for
@@
 
-The paper begins with a description of our probability forecasting data that were collected by asking over 2,300 experts to give probability forecasts and to self-assess their level of expertise on a number of 166 geopolitical binary events.
+The paper begins with a description of our probability forecasting data that were collected by asking over 2,300 experts to give probability forecasts and to self-assess their level of expertise on a subset of 166 geopolitical binary events.
 After summarizing the dataset, the paper introduces a dynamic hierarchical model for capturing the crowd belief.
-The model is estimated in a two-step procedure: First, constrained parameter estimates are obtained via Gibbs sampling.
-These estimates are then transformed in to their unconstrained equivalents by a calibrating step.
-An extension of this model to polychotomous outcomes is briefly discussed before the model evaluation sections.
-The first evaluation section uses synthetic data to study the accuracy of the two-step procedure to capture parameter values.
-The second and much longer evaluation section uses the model on our probability forecasting data.
-The main goal is to illustrate and also to study the specific research objectives itemized above.
+The model is estimated in a two-step procedure.
+First, constrained parameter estimates are obtained via Gibbs sampling (see \citet{geman1984stochastic} for the original introduction of Gibbs sampling).
+These estimates are then transformed to their unconstrained equivalents by a calibration step.
+An extension of this model to polychotomous outcomes is briefly discussed before model evaluation.
+The first evaluation section uses synthetic data to study the accuracy to which the two-step procedure is able to estimate parameter values.
+The second evaluation section discusses different aspects of the model by applying it to our real-world probability forecasting data.
 The paper concludes with a discussion on future research directions and model limitations.
@@
 \label{data}
-The data collection began with a recruitment of 2,365 forecasters ranging from graduate students to forecasting and political science faculty and practitioners.
+
+\begin{table}
+\begin{center}
+\begin{tabular}{c c c c c c c c }
+\hline
+Statistic & Min. & $Q_1$ & Median & Mean & $Q_3$ & Max.\\ \hline
+\# of Days a Problem is Active  & 4 &  35.6 &  72.0 & 106.3 & 145.20 & 418 \\
+\# of Experts per Problem & 212 &   543.2 &  693.5 &  783.7 &  983.2 & 1690\\ 
+\# Forecasts given by each Expert on a Problem & 1 &  1.0 &   1.0  &  1.8 &  2.0 & 131 \\
+\# Problems participated by an Expert & 1 &  14.0 &   36.0  &  55.0 &  90.0 & 166 \\ \hline
+\end{tabular}
+\caption{Five-number summaries of our real-world data.
+%: \textit{\# of Days Active} represents the number of days that the experts were given to update their forecast before the correct answer was revealed; \textit{\# of Experts per Problem } denotes the number of experts participating in a question; \textit{\# Forecasts given by each Expert on a Problem} gives the number of forecasts for a question made by an active expert; \textit{\# Problems participated by an Expert} represents the number of different questions a given expert participated in.
+}
+\label{DataStats}
+\end{center}
+\end{table}
+
+\begin{table}
+\begin{center}
+\begin{tabular}{c c c c c c c c }
+\hline
+Expertise Level & 1 & 2 & 3 & 4 & 5\\
+Frequency (\%) & 25.3 & 30.7 & 33.6 &  8.2 &  2.1 \\\hline
+\end{tabular}
+\caption{Frequencies of the self-reported expertise (1 = Not At All Expert and 5 = Extremely Expert) levels across all the 166 questions in our real-world data.}
+\label{ExpertiseTable}
+\end{center}
+\end{table}
+
+The data collection began with a recruitment of 2,365 forecasters ranging from graduate students to political science faculty and practitioners.
 The recruiting was made from professional societies, research centers, alumni associations, science bloggers, and word of mouth.
@@
 
+\vspace{-1em}
+\begin{figure}[h!]
+\centering
+\hspace*{0em} 	\includegraphics[width=  \textwidth]{Figures/LegendExamplePlot} % requires the graphicx package
+\vspace{-4.5em}
+
+\hspace{-0.5em}
+\subfigure[\textit{Will the expansion of the European bailout fund be ratified by all 17 Eurozone nations before 1 November 2011?}]{
+\hspace*{-1em}  \includegraphics[width= 0.49\textwidth]{Figures/ExamplePlot1}
+\label{Example1}
+} \hspace{0.5em}
+\subfigure[\textit{Will the Nikkei 225 index finish trading at or above 9,500 on 30 September 2011?}]{
+\hspace*{-1em}  \includegraphics[width= 0.49\textwidth]{Figures/ExamplePlot2}
+\label{Example2}
+}
+
+\caption[Optional caption for list of figures]{Scatterplots of the probability forecasts given for two questions in our dataset.
+The shadings represents the self-reported expertise of the forecaster who provided the probability forecast.}
+\label{ExamplePlots}
+\end{figure}
+
 To illustrate the nature of the data with some concrete examples, Figures \ref{Example1} and \ref{Example2} show scatterplots of the probability forecasts given for (a) \textit{Will the expansion of the European bailout fund be ratified by all 17 Eurozone nations before 1 November 2011?}, and (b) \textit{Will the Nikkei 225 index finish trading at or above 9,500 on 30 September 2011?}.
@@
 As mentioned before, the forecasters were allowed to update their forecast as long as the questions remained active.
-Updating, however, was done on a very infrequent basis.
-An expert gave, on average, only 0.0159 forecasts per day.
-A typical expert, therefore, made a forecast about every 62 days.
+Updating was done on a very infrequent basis.
+An expert gave, on average, only 0.0159 forecasts per day, and so a typical expert made a forecast about every 62 days.
 As a result the average response rate was around 13.5 forecasts per day from a large group of experts.
-
-\vspace{-1em}
-\begin{figure}[h!]
-\centering
-\hspace*{0em} 	\includegraphics[width=  \textwidth]{Figures/LegendExamplePlot} % requires the graphicx package
-\vspace{-4.5em}
-
-\hspace{-0.5em}
-\subfigure[\textit{Will the expansion of the European bailout fund be ratified by all 17 Eurozone nations before 1 November 2011?}]{
-\hspace*{-1em}  \includegraphics[width= 0.49\textwidth]{Figures/ExamplePlot1}
-\label{Example1}
-} \hspace{0.5em}
-\subfigure[\textit{Will the Nikkei 225 index finish trading at or above 9,500 on 30 September 2011?}]{
-\hspace*{-1em}  \includegraphics[width= 0.49\textwidth]{Figures/ExamplePlot2}
-\label{Example2}
-}
-
-\caption[Optional caption for list of figures]{Scatterplots of the probability forecasts given for two questions in our dataset.
-The shadings represents the self-reported expertise of the forecaster who provided the probability forecast.}
-\label{ExamplePlots}
-\end{figure}
-
-\begin{table}
-\begin{center}
-\begin{tabular}{c c c c c c c c }
-\hline
-Statistic & Min. & $Q_1$ & Median & Mean & $Q_3$ & Max.\\ \hline
-\# of Days a Problem is Active  & 4.00 &  35.25 &  72.00 & 106.30 & 145.20 & 418.00 \\
-\# of Experts per Problem & 212.0 &   543.2 &  693.5 &  783.7 &  983.2 & 1690.0\\ 
-\# Forecasts given by each Expert on a Problem & 1.0 &  1.0 &   1.0  &  1.8 &  2.0 & 131.0 \\
-\# Problems participated by an Expert & 1.0 &  14.0 &   36.0  &  55.0 &  90.0 & 166.0 \\ \hline
-\end{tabular}
-\caption{Five-number summaries of the real-world data used in this paper: \textit{\# of Days Active} represents the number of days that the experts were given to update their forecast before the correct answer was revealed; \textit{\# of Experts per Problem } denotes the number of experts participating in a question; \textit{\# Forecasts given by each Expert on a Problem} gives the number of forecasts for a question made by an active expert; \textit{\# Problems participated by an Expert} represents the number of different questions a given expert participated in.}
-\label{DataStats}
-\end{center}
-\end{table}
-
-\begin{table}
-\begin{center}
-\begin{tabular}{c c c c c c c c }
-\hline
-Expertise Level & 1 & 2 & 3 & 4 & 5\\
-Frequency (\%) & 25.3 & 30.7 & 33.6 &  8.2 &  2.1 \\\hline
-\end{tabular}
-\caption{Frequencies of the expertise levels across all the 166 questions in the dataset.}
-\label{ExpertiseTable}
-\end{center}
-\end{table}
 
@@
 Let $p_{i,t,k} \in (0,1)$ be the probability forecast given by the $i$th expert at time $t$ for the $k$th question, where $i = i, \dots, I_k$, $t = 1, \dots, T_k$, and $k = 1, \dots, K$.
-Denote the logit-probabilities with $Y_{i,t,k} = \logit(p_{i,t,k}) = \log \left( \frac{p_{i,t,k}}{1-p_{i,t,k}}\right) \in \R$ and collect the logit-probability forecasts given for question $k$ at time $t$ into a vector $\boldsymbol{Y}_{t,k} = [Y_{1,t,k}\ Y_{2,t,k}\ \dots\ Y_{I_{k},t,k}]^T$.
+Denote the logit-probabilities with
+\begin{eqnarray*}
+Y_{i,t,k} = \logit(p_{i,t,k}) = \log \left( \frac{p_{i,t,k}}{1-p_{i,t,k}}\right) \in \R
+\end{eqnarray*}
+and collect the logit-probability forecasts given for question $k$ at time $t$ into a vector $\boldsymbol{Y}_{t,k} = [Y_{1,t,k}\ Y_{2,t,k}\ \dots\ Y_{I_{k},t,k}]^T$.
 Partition the experts into $J$ groups based on some individual feature, such as self-reported expertise, with each group sharing a common multiplicative bias term $b_{j} \in \R$ for $j = 1, \dots, J$.
@@
 \end{eqnarray}
-where Equation (\ref{observedpr}) denotes the observed process, Equation (\ref{hiddenpr}) shows the hidden process that is driven by the constant $\gamma_k \in \R$, the error terms are independent and identically distributed normal random variables with mean zero
+where Equation (\ref{observedpr}) denotes the observed process, Equation (\ref{hiddenpr}) shows the hidden process that is driven by the constant $\gamma_k \in \R$.
+The error terms are independent and identically distributed normal random variables with mean zero
 \begin{eqnarray*}
@@
 \end{eqnarray*}
-and $(\mu_0, \sigma_0^2) \in (\R, \R^+)$ are hyper-parameters chosen a priori.
+and $(\mu_0, \sigma_0^2) \in (\R, \R^+)$ are hyper-parameters chosen \textit{a priori}.
 The hidden state $X_{t,k}$ represents the sharp and calibrated logit-probability for the $k$th event given the information available up to and including time $t$.
@@
 An individual element of $\boldsymbol{b}$ can be interpreted as a group-specific \textit{systematic bias} that labels the group either as over-confident $\left( b_j \in (1, \infty)\right)$ or as under-confident $\left(b_j \in (0,1)\right)$.
-The bias terms are analyzed and discussed more in Section \ref{ExpertBias}.
+Unfortunately, due to the high sparsity of our data, estimating a bias term and analyzing the bias separately for each expert is not possible.
+See Section \ref{ExpertBias} for an analysis and discussion on the bias terms.
 Any other deviation from the calibrated crowd belief is considered \textit{random noise}.
@@
 The most challenging part of our model estimation is to incorporate the event indicators $\{Z_k\}_{k=1}^K$ such that the hidden process is well-calibrated and the interpretability of our model is not compromised.
-This section introduces a two-step procedure, called \textit{Sample-Then-Calibrate} or simply \textit{STC}, that achieves this goal in a flexible, yet efficient manner by first estimating  the model parameters under a constraint (\textit{Sampling Step}) and then performing a one-dimension optimization procedure to transform the constrained estimates into their unconstrained counterparts (\textit{Calibration Step}).
+This section introduces a two-step procedure, called \textit{Sample-Then-Calibrate} or simply \textit{STC}, that achieves this goal in a flexible and efficient manner by first estimating  the model parameters under a constraint (\textit{Sampling Step}) and then performing a one-dimension optimization procedure to transform the constrained estimates into their unconstrained counterparts (\textit{Calibration Step}).
 
@@
 These estimates are found via a Gibbs sampler that only makes use of standard distributions.
-This procedure is standard practice in Bayesian statistics and is therefore not described in great detail here.
 See Appendix \ref{appendix} for the technical details of our sampler, and, e.g., \citet{gelman2003bayesian} for a discussion on the general principles of Gibbs sampling.
@@
 Due to the possibility of \textit{complete separation} in any one question (see, e.g., \citet{gelman2008weakly}), the maximization must be performed over multiple questions.
-Therefore,
+Therefore we estimate
 \begin{eqnarray}
@@
 \end{eqnarray}
-where $S$ is a strictly proper scoring rule such as the negative Brier score (\citet{Brier})
+where $Z_k \in \{0, 1\}$ indicate whether the event associated with the $k$th question happened $\left(Z_k = 1\right)$ or did not happen $\left(Z_k = 0\right)$.
+The function $S$ is a strictly proper scoring rule such as the negative Brier score (\citet{Brier})
 \begin{eqnarray*}
@@
 Since this multinomial extension is equivalent to having $M-1$ independent binary-outcome models (see Section \ref{model}), the estimation can be done separately for each outcome.
-Therefore, even though this paper focuses on modeling binary events, all the properties and discussion generalize trivially to the multi-outcome case.
+Therefore, even though this paper focuses on modeling binary events, all the properties and discussion generalize to the multi-outcome case.
 
@@
 The goal is to provide application specific insight by discussing the specific research objectives itemized in Section \ref{intro} and also to evaluate the \textit{Sample-Then-Calibrate}-procedure in terms of predictive power and calibration.
-Before presenting the results, however, two practical matters, which need to be taken into account when aggregating real-world  probability forecasting data, are discussed.
+Before presenting the results two practical matters are discussed.
+
+%These matters must be taken into account when aggregating real-world  probability forecasting data, are discussed.
 
@@
 SDLM & 0.100 (0.156) & 0.066 (0.116) & 0.098 (0.154) & 0.102 (0.157)\\ 
-BSTC & 0.097 (0.213) & \textbf{0.053} (0.147) & 0.100 (0.215) & 0.098 (0.215)\\ 
+BSTC-Log & 0.097 (0.213) & \textbf{0.053} (0.147) & 0.100 (0.215) & 0.098 (0.215)\\ 
 STC-Bri & 0.096 (0.190) & 0.056 (0.134) & 0.097 (0.190) & 0.098 (0.192)\\ 
@@
 SDLM & 0.089 (0.116) & 0.064 (0.085) & 0.106 (0.141) & 0.092 (0.117)\\ 
-BSTC & 0.083 (0.160) & \textbf{0.052} (0.103) & 0.110 (0.198) & 0.085 (0.162)\\ 
+BSTC-Log & 0.083 (0.160) & \textbf{0.052} (0.103) & 0.110 (0.198) & 0.085 (0.162)\\ 
 STC-Bri & 0.083 (0.142) & 0.055 (0.096) & 0.106 (0.174) & 0.085 (0.144)\\ 
@@
 As a result, the forecasts are either very close to the correct answer or very far from it.
-As can be seen in Table \ref{prediction}, this results into high variance in performance.
+As can be seen in Table \ref{prediction}, this results into high variance in performance of BSTC-Log.
 The short questions, however, involve very little uncertainty and were generally the easiest to forecast.
@@
 However, as soon as the forecaster gains some knowledge about the event, he produces an updated forecast that is a compromise between his initial forecast and the new information acquired.
-The updated forecast is therefore conservative and necessarily too close to 0.5 as long as the forecaster remains only partially informed about the event.
+The updated forecast is therefore conservative and too close to 0.5 as long as the forecaster remains only partially informed about the event.
 If most forecasters fall somewhere on this spectrum between ignorance and full information, their average forecast tends to fall strictly between 0.5 and the most-informed probability forecast (see \citet{Baron} for more details).
@@
 For instance, the posterior probability that the most expert group is the least under-confident is approximately equal to $1.0$, and
-the posterior probability of a strictly decreasing level of under-confidence is approximately 0.87.
+the posterior probability of a strictly decreasing level of under-confidence by expertise ranking is approximately 0.87.
 The latter probability is driven down by the inseparability of the two groups associated with the lowest levels of self-reported expertise.
@@
 \section{Discussion}
-This paper introduced an interpretable time-series model that incorporates self-reported expertise and captures a sharp and well-calibrated crowd belief across time.
+This paper began with an introduction of a rather unorthodox but nonetheless realistic probability forecasting setting, in which forecasts are made very infrequently by a heterogeneous group of experts over a period of time.
+Unfortunately, the resulting data is too sparse to be modeled efficiently with standard time-series methods.
+In response to this lack of previous methodology, our work
+%expands the literature towards less-studied areas of  probability aggregation by
+introduces an interpretable time-series model that incorporates self-reported expertise and captures a sharp and well-calibrated crowd belief across time.
 The model is estimated in two-steps: The first step samples constrained versions of the model parameters via a Gibbs sampler that only makes use of standard distributions.
 The second step involves a one-dimensional optimization procedure that transforms the constrained parameter values into their unconstrained counterparts.
-Due to the rather unorthodox nature of our dataset, our work expands the literature towards less-studied areas of  probability aggregation.
+
+%Due to the unconventional nature of our dataset, our work expands the literature towards less-studied areas of  probability aggregation.
 
@@
 This result is about groups of experts and hence does not conflict with the well-known result of the individuals being over-confident (see, e.g.,  \citet{lichtenstein1977calibration, morgan1992uncertainty, bier2004implications}).
-Besides studying group-level bias, the model can be used to generate estimates of many problem-specific quantities.
+Besides making predictions or studying group-level bias, the model can be used to generate estimates of many problem-specific quantities.
 These quantities have clear interpretations and can be combined in many interesting ways to answer a range of hypotheses about the experts and the questions.
@@
 \subsection{Future Directions and Limitations}
- The model can be extended in various ways to meet a wide range of research objectives.
+ The model can be extended in different ways to meet a wide range of research objectives.
 For instance, the degree to which the group-level under-confidence decreases towards the resolution of the target event, could be studied by allowing the bias vector to depend on time.
-Furthermore, the model is not limited to the study of group-level bias at different levels of self-reported expertise.
-The experts could equally well be partitioned based on other features, such as gender, education, or age, to produce a range of novel insight on the level of confidence among different groups of expert forecasters.
+The model is also by no means limited to the study of group-level bias at different levels of self-reported expertise.
+The experts could equally well be partitioned based on other features, such as gender, education, or specialty, to produce a range of novel insight on the level of confidence among different groups of expert forecasters.
+Similarly, the model could be applied to questions grouped by types, such as economic, domestic, or international, to learn about expert forecasting under a wide variety of problem attributes.
 
@@
 
-The Gibbs sampler iteratively samples all the unknown parameters from their full-conditionals one block of parameters at a time.
+The Gibbs sampler (\citet{geman1984stochastic}) iteratively samples all the unknown parameters from their full-conditional posterior distributions one block of parameters at a time.
 Since this is performed under the constraint $b_3 = 1$ to ensure model identifiability, the constrained parameter estimates should be denoted with a trailing $(1)$ to maintain consistency with earlier notation.
@@
 \begin{eqnarray*}
-\sigma^2_k | \dots &\sim& \text{Inv-}\chi^2\left(|\boldsymbol{Y}_k| - J, \frac{1}{|\boldsymbol{Y}_k| - J} (\boldsymbol{Y}_k - \boldsymbol{X}_k \boldsymbol{b})^T (\boldsymbol{Y}_k - \boldsymbol{X}_k \boldsymbol{b}) \right)
-\end{eqnarray*}
+\sigma^2_k | \dots &\sim& \text{Inv-}\chi^2\left(|\boldsymbol{Y}_k| - J, \frac{1}{|\boldsymbol{Y}_k| - J} (\boldsymbol{Y}_k - \boldsymbol{X}_k \boldsymbol{b})^T (\boldsymbol{Y}_k - \boldsymbol{X}_k \boldsymbol{b}) \right),
+\end{eqnarray*}
+where the distribution is a scaled inverse-$\chi^2$ (see, e.g., \citet{gelman2003bayesian}).
 Since the experts are not required to give a new prediction at every time unit, the design matrices must be trimmed accordingly such that their dimensions match up with the dimensions of the observed matrices.

### 2013-08-23 paper.tex
Edits

@@
 The quality of a probability assessment is typically evaluated in regards to an important attribute called \textit{calibration}.
-An expert is said to be well-calibrated if his probabilities agree with the long-term frequency of the events.
+An expert is said to be well-calibrated if his probabilities match with the long-term frequencies of the events.
 This means that, for instance, the proportion of events that actually occurred is 60\% for all those events for which he assessed a probability of 0.60.
-Even though several experiments have been conducted to show that experts are generally poorly calibrated (see, e.g., \citet{cooke1991experts, shlyakhter1994quantifying}), the level of expertise has been found to be associated with calibration.
-In particular, \citet{wright1994coherence} argue that a higher self-reported expertise on the subject matter implies better calibration.
+Even though several experiments have been conducted to show that experts are generally poorly calibrated (see, e.g., \citet{cooke1991experts, shlyakhter1994quantifying}), expertise has been found to be associated with calibration.
+In particular, \citet{wright1994coherence} argue that a higher level of self-reported expertise on the subject matter implies better calibration.
 
@@
 To see this, consider a relatively stationary process, such as rain in some geographic region, whose empirical frequency of occurrence in the last 10 years is 45\%.
-In this setting, an expert can assess a constant probability of 0.45 and be well-calibrated.
-This probability, however, is very near 0.5 and hence too uninformative to be useful in decision making.
-Compare this to the advice of  another expert whose typical probability assessment is near 0.0 or 1.0; that is, he is a \textit{sharp} assessor.
-If the expert is both sharp and well-calibrated, then he is able to forecast the behavior of the process both with high certainty and accuracy.
-From the decision-maker's perspective, such advice is highly useful as it involves very little risk.
-Therefore the goal in probability estimation should be to maximize sharpness subject to calibration (see, e.g., \citet{raftery2005using, Ranjan08}).
-This is a well-known goal that has led to a wide range of novel and insightful observations in probability forecasting literature.
+In this setting an expert can assess a constant probability of 0.45 and be well-calibrated.
+This probability, however, is close to 0.5 and hence too uninformative to be very useful in decision making.
+For the sake of comparison, consider another expert whose typical probability assessment is near 0.0 or 1.0.
+In the forecasting literature such an expert is called \textit{sharp}.
+If the expert is not only sharp but also well-calibrated, then he is able to forecast the behavior of the process both with high certainty and accuracy.
+From the decision-maker's perspective, his advice is highly useful as it involves very little risk and much information.
+For these reasons, a common goal in probability estimation is to maximize sharpness subject to calibration (see, e.g., \citet{raftery2005using, Ranjan08}).
+This is a well-defined goal that has led to a wide range of novel and insightful observations in probability forecasting literature.
 
@@
 \label{ExpertBias}
-Recall from Section \ref{data} that the experts were asked to self-assess their expertise level (on a 1-to-5 scale with 1 = Not At All Expert to 5 = Extremely Expert) on  all of the 166 questions in our real-world dataset.
-Based on these self-assessments the experts were then divided into 5 expertise groups, and each of the expertise groups was assigned a separate multiplicative bias term.
-This section uses the \textit{Sample-Then-Calibrate}-procedure to explore the posterior distribution of these group-level bias terms.
+Recall from Section \ref{data} that the experts were asked to self-assess their level of expertise (on a 1-to-5 scale with 1 = Not At All Expert to 5 = Extremely Expert) on any questions that they participated in.
+The self-reported expertise then divides the experts into 5  groups, with each group assigned a separate multiplicative bias term.
+This section uses the \textit{Sample-Then-Calibrate}-procedure to explore the posterior distribution of these multiplicative bias terms.
 Figure \ref{Biases} presents the posterior distributions of the bias terms with side-by-side box plots.
-Since the distributions fall completely below the \textit{no-bias} reference-line at 1.0, the expertise groups are deemed under-confident.
-
-The under-confidence, however, decreases as the level of the self-reported expertise increases.
-In order to understand this result, it is helpful to view the process of making a subjective probability assessment as Bayesian updating: A completely ignorant expert aiming to minimize a reasonable loss function, such as the Brier score, has no reason to give anything but 0.5 as his probability forecast.
-However, as soon as the forecaster gains some knowledge about the event, he produces an updated forecast that is a compromise between his initial forecast and the new information acquired.
-The updated forecast is therefore conservative and too close to 0.5 as long as the forecaster remains only partially informed about the event.
-If most forecasters fall somewhere on this spectrum between ignorance and full information, their average forecast tends to fall strictly between 0.5 and the most-informed probability forecast (see \citet{Baron} for more details).
+Since the distributions fall completely below the \textit{no-bias} reference-line at 1.0, all the groups are deemed under-confident.
+
+The under-confidence, however, decreases as the level of  expertise increases.
+For instance, the posterior probability that the most expert group is the least under-confident is approximately equal to $1.0$, and the posterior probability of a strictly decreasing level of under-confidence is approximately 0.87.
+The latter probability is driven down by the inseparability of the two groups associated with the lowest levels of self-reported expertise.
+The fact that these groups are very similar suggests that the forecasters are poor at assessing how little they know about a subject that is strange to them.
+If these groups are combined into a single group, the posterior probability of a strictly decreasing level of under-confidence is approximately 1.0.
+
+The decreasing trend in under-confidence can be reasoned by viewing the process of making a subjective probability  as Bayesian updating: At first a completely ignorant expert aiming to minimize a reasonable loss function, such as the Brier score, has no reason to give anything but 0.5 as his probability forecast.
+However, as soon as the expert gains some knowledge about the event, he produces an updated forecast that is a compromise between his initial forecast and the new information acquired.
+The updated forecast is therefore conservative and too close to 0.5 as long as the expert remains only partially informed about the event.
+If most experts fall somewhere on this spectrum between ignorance and full information, their average forecast tends to fall strictly between 0.5 and the most-informed probability forecast (see \citet{Baron} for more details).
 Since expertise is to a large extent determined by subject-matter knowledge, the level of under-confidence can be expected to decrease as a function of the group's level of self-reported expertise.
 
-Overall, finding under-confidence is a rather surprising result given that many previous studies have been conducted to show that experts are likely to be over-confident.
+Finding under-confidence in all the groups is a rather surprising result given that many previous studies have been conducted to show that experts are likely to be over-confident.
 For instance,  \citet{lichtenstein1977calibration, morgan1992uncertainty, bier2004implications} summarize the results from numerous calibration studies, and conclude that experts are systematically over-conﬁdent about their probability assessments.
 Our result, however, is a statement about groups of experts and hence does not invalidate the possibility of the individual experts being overconfident.
-To make conclusions at the individual-level based on the group-level bias terms would result into a \textit{ecological inference fallacy} (see, e.g., \citet{lubinski1996seeing}).
+To make conclusions at the individual-level based on the group-level bias terms would be considered an \textit{ecological inference fallacy} (see, e.g., \citet{lubinski1996seeing}).
 
@@
 
-More specific group-level inference can be performed by computing probabilities based the estimated posterior distributions.
+%More specific group-level inference can be performed by computing probabilities based the estimated posterior distributions.
 For instance, the posterior probability that the most expert group is the least under-confident is approximately equal to $1.0$, and
-the posterior probability of a strictly decreasing level of under-confidence by expertise ranking is approximately 0.87.
+%the posterior probability of a strictly decreasing level of under-confidence by expertise ranking is approximately 0.87.
 The latter probability is driven down by the inseparability of the two groups associated with the lowest levels of self-reported expertise.
@@
 This paper began with an introduction of a rather unorthodox but nonetheless realistic probability forecasting setting, in which forecasts are made very infrequently by a heterogeneous group of experts over a period of time.
-Unfortunately, the resulting data is too sparse to be modeled efficiently with standard time-series methods.
-In response to this lack of previous methodology, our work
+The resulting data is too sparse to be modeled efficiently with standard time-series methods.
+In response to this lack of appropriate methodology, our work
 %expands the literature towards less-studied areas of  probability aggregation by

### 2013-08-24 paper.tex
Edits

@@
 Individual experts can differ radically from one another in their abilities to assess probabilities of future events.
-The quality of a probability assessment is typically evaluated in regards to an important attribute called \textit{calibration}.
-An expert is said to be well-calibrated if his probabilities match with the long-term frequencies of the events.
-This means that, for instance, the proportion of events that actually occurred is 60\% for all those events for which he assessed a probability of 0.60.
-Even though several experiments have been conducted to show that experts are generally poorly calibrated (see, e.g., \citet{cooke1991experts, shlyakhter1994quantifying}), expertise has been found to be associated with calibration.
-In particular, \citet{wright1994coherence} argue that a higher level of self-reported expertise on the subject matter implies better calibration.
-
-Unfortunately, calibration by itself does not guarantee usefulness.
-To see this, consider a relatively stationary process, such as rain in some geographic region, whose empirical frequency of occurrence in the last 10 years is 45\%.
+The quality of a probability assessment is typically evaluated in regards to \textit{calibration} that measures the degree to which the frequency of occurrence of the events agree with the given probabilities.
+This means that, for instance, the proportion of events that actually occurred is 60\% for all those events for which a well-calibrated expert assessed a probability of 0.60.
+Unfortunately, several experiments have been conducted to show that experts are generally poorly calibrated (see, e.g., \citet{cooke1991experts, shlyakhter1994quantifying}).
+\citet{wright1994coherence}, however, argue that a higher level of self-reported expertise on the subject-matter can be associated with better calibration.
+
+Calibration by itself, however, is not sufficient for a useful probability assessment.
+To see this, consider a relatively stationary process, such rain on different days in some geographic region, whose empirical frequency of occurrence in the last 10 years is 45\%.
 In this setting an expert can assess a constant probability of 0.45 and be well-calibrated.
-This probability, however, is close to 0.5 and hence too uninformative to be very useful in decision making.
-For the sake of comparison, consider another expert whose typical probability assessment is near 0.0 or 1.0.
-In the forecasting literature such an expert is called \textit{sharp}.
-If the expert is not only sharp but also well-calibrated, then he is able to forecast the behavior of the process both with high certainty and accuracy.
-From the decision-maker's perspective, his advice is highly useful as it involves very little risk and much information.
-For these reasons, a common goal in probability estimation is to maximize sharpness subject to calibration (see, e.g., \citet{raftery2005using, Ranjan08}).
-This is a well-defined goal that has led to a wide range of novel and insightful observations in probability forecasting literature.
+This assessment, however, can be made without any subject-matter expertise or special training.
+For this reason the long-term frequency is often considered as the baseline probability -- a naive assessment that provides the decision-maker very little extra information.
+A consulting expert should therefore aim to make probability assessments that are as far from the baseline as possible.
+The extent to which his probabilities differ from the baseline is measured with an important attribute called \textit{sharpness} (\citet{gneiting2008rejoinder, winkler2008comments}).
+If the expert is both sharp and well-calibrated, he is able to forecast the behavior of the process with high certainty and accuracy.
+From the decision-maker's perspective, his advice is highly useful as it involves much information and very little risk.
+Therefore useful probability estimation should aim to maximize sharpness subject to calibration (see, e.g., \citet{raftery2005using, murphy1987general}).
+This is a well-defined goal that has led to a wide range of novel and insightful observations in probability forecasting.
 
@@
 
-%If the consulted experts diligently make probability assessments on a daily basis, many previously-studied time-series procedures could be used to aggregate their assessments.
-
 This paper adds another layer of complexity by assuming a heterogeneous set of experts, most of whom only make one or two probability assessments over the hundred or so days before the event resolves.
+This means that the decision-maker faces a different group of experts  every day, with only a few experts returning later on for a second round of advice.
 The problem at hand is therefore strikingly different from many time-series estimation problems, where one has an observation at every time point -- or almost every time point.
-Therefore standard time-series procedures like ARIMA (see, e.g., \citet{mills1991time}) are not applicable without ignoring the heterogeneity among the experts.
-In response to the lack of procedures to model sparse and heterogeneous probability forecasting data, this paper introduces an interpretable time-series model that incorporates self-reported expertise and captures a sharp and well-calibrated estimate of the crowd belief.
+As a result, standard time-series procedures like ARIMA (see, e.g., \citet{mills1991time}) are not directly applicable to our setting.
+In response to this lack of appropriate procedures, we introduce an interpretable time-series model that incorporates self-reported expertise and captures a sharp and well-calibrated estimate of the crowd belief.
 The model is then used for
@@
 \section{Discussion}
-This paper began with an introduction of a rather unorthodox but nonetheless realistic probability forecasting setting, in which forecasts are made very infrequently by a heterogeneous group of experts over a period of time.
+This paper began with an introduction of a rather unorthodox but nonetheless realistic time-series setting in which probability forecasts are made very infrequently by a heterogeneous group of experts.
 The resulting data is too sparse to be modeled efficiently with standard time-series methods.
 In response to this lack of appropriate methodology, our work
-%expands the literature towards less-studied areas of  probability aggregation by
 introduces an interpretable time-series model that incorporates self-reported expertise and captures a sharp and well-calibrated crowd belief across time.
-The model is estimated in two-steps: The first step samples constrained versions of the model parameters via a Gibbs sampler that only makes use of standard distributions.
-The second step involves a one-dimensional optimization procedure that transforms the constrained parameter values into their unconstrained counterparts.
+The model is estimated in two-steps: The first step, known as the \textit{sampling step}, samples constrained versions of the model parameters via a Gibbs sampler that only makes use of standard distributions.
+The second step, known as the \textit{calibration step}, involves a one-dimensional optimization procedure that transforms the constrained parameter values into their unconstrained counterparts.
 
@@
 Besides making predictions or studying group-level bias, the model can be used to generate estimates of many problem-specific quantities.
-These quantities have clear interpretations and can be combined in many interesting ways to answer a range of hypotheses about the experts and the questions.
+These quantities have clear interpretations and can be combined in many interesting ways to answer a range of hypotheses about different types of questions and expert behavior.
 
 \subsection{Future Directions and Limitations}
- The model can be extended in different ways to meet a wide range of research objectives.
-For instance, the degree to which the group-level under-confidence decreases towards the resolution of the target event, could be studied by allowing the bias vector to depend on time.
-The model is also by no means limited to the study of group-level bias at different levels of self-reported expertise.
+This paper developed a time-series model that is parsimonious subject to addressing the main challenges that arise from sparse probability forecasting data.
+The model can be extended in various ways using standard techniques.
+For instance, most of the model parameters were assumed constant over time.
+It is intuitively reasonable, however, to assume that the behavior of the model parameters is different a hundred days before as compared to a day before the event resolves.
+The evolution of the parameters over time could be studied by allowing the parameters to depend on the time index.
+
+Similarly, the model parameters can be allowed to depend on a variety of expert characteristic.
+In this paper the multiplicative bias was assumed to depend on the level of self-reported expertise.
+Similar dependencies, however, could be constructed based on any expert characteristic, such as sex, education, specialty, or even a combination of characteristics, to produce a range of novel insight on the level of confidence across different groups of experts.
+%The model is also by no means limited to the study of group-level bias at different levels of self-reported expertise.
 The experts could equally well be partitioned based on other features, such as gender, education, or specialty, to produce a range of novel insight on the level of confidence among different groups of expert forecasters.
-Similarly, the model could be applied to questions grouped by types, such as economic, domestic, or international, to learn about expert forecasting under a wide variety of problem attributes.
+Furthermore, instead of grouping the experts, the questions could be grouped by types, such as economic, domestic, or international.
+These groups could be analyzed separately to learn about forecasting behavior under a wide variety of problem attributes.
+
+%The model could be applied to questions grouped by types, such as economic, domestic, or international, to learn about expert forecasting under a wide variety of problem attributes.
 

### 2013-08-24 paper.tex
Edits

@@
 As a result, standard time-series procedures like ARIMA (see, e.g., \citet{mills1991time}) are not directly applicable to our setting.
-In response to this lack of appropriate procedures, we introduce an interpretable time-series model that incorporates self-reported expertise and captures a sharp and well-calibrated estimate of the crowd belief.
-The model is then used for
+
+In response to the lack of appropriate modeling procedures, this paper introduces a time-series model that incorporates self-reported expertise and captures a sharp and well-calibrated estimate of the crowd belief.
+The model is highly interpretable and can be used for
 \begin{itemize}
@@
 The model is estimated in a two-step procedure.
-First, constrained parameter estimates are obtained via Gibbs sampling (see \citet{geman1984stochastic} for the original introduction of Gibbs sampling).
-These estimates are then transformed to their unconstrained equivalents by a calibration step.
+First, a sampling step produces constrained parameter estimates via Gibbs sampling (see \citet{geman1984stochastic} for the original introduction of Gibbs sampling).
+Second, a calibration step transforms these estimates to their unconstrained equivalents via a one-dimension optimization procedure.
 An extension of this model to polychotomous outcomes is briefly discussed before model evaluation.
@@
 \begin{figure}[h!]
+\vspace{-1em}
 \centering
@@
 The resulting data is too sparse to be modeled efficiently with standard time-series methods.
-In response to this lack of appropriate methodology, our work
-introduces an interpretable time-series model that incorporates self-reported expertise and captures a sharp and well-calibrated crowd belief across time.
+In response, our work introduces an interpretable time-series model that incorporates self-reported expertise and captures a sharp and well-calibrated crowd belief across time.
 The model is estimated in two-steps: The first step, known as the \textit{sampling step}, samples constrained versions of the model parameters via a Gibbs sampler that only makes use of standard distributions.
-The second step, known as the \textit{calibration step}, involves a one-dimensional optimization procedure that transforms the constrained parameter values into their unconstrained counterparts.
-
-%Due to the unconventional nature of our dataset, our work expands the literature towards less-studied areas of  probability aggregation.
+The second step, known as the \textit{calibration step}, involves a one-dimensional optimization procedure that transforms the constrained parameter values to their unconstrained counterparts.
 
 \subsection{Summary of Findings}
-The model was applied to an unusually large dataset on probability forecasting done by human experts.
+The model was applied to an unusually large dataset on probability forecasting.
 The estimated crowd belief was found to be sharp and well-calibrated under both in- and out-of-sample probability estimation.
@@
 Firstly, the model was shown to outperform other probability aggregators in terms of forecasting ability.
-Secondly, the crowd belief was used as the no-bias reference point to study the distributions of the bias terms for different expertise-groups.
+In particular, the model was deemed a well-balanced compromise that avoids over-fitting without begin overly conservative.
+Secondly, the crowd belief was used as the no-bias reference point to study the bias among groups of experts with different levels of self-reported expertise.
 All the groups were found to be under-confident.
@@
 
-\subsection{Future Directions and Limitations}
-This paper developed a time-series model that is parsimonious subject to addressing the main challenges that arise from sparse probability forecasting data.
-The model can be extended in various ways using standard techniques.
-For instance, most of the model parameters were assumed constant over time.
-It is intuitively reasonable, however, to assume that the behavior of the model parameters is different a hundred days before as compared to a day before the event resolves.
-The evolution of the parameters over time could be studied by allowing the parameters to depend on the time index.
-
-Similarly, the model parameters can be allowed to depend on a variety of expert characteristic.
-In this paper the multiplicative bias was assumed to depend on the level of self-reported expertise.
-Similar dependencies, however, could be constructed based on any expert characteristic, such as sex, education, specialty, or even a combination of characteristics, to produce a range of novel insight on the level of confidence across different groups of experts.
-%The model is also by no means limited to the study of group-level bias at different levels of self-reported expertise.
-The experts could equally well be partitioned based on other features, such as gender, education, or specialty, to produce a range of novel insight on the level of confidence among different groups of expert forecasters.
-Furthermore, instead of grouping the experts, the questions could be grouped by types, such as economic, domestic, or international.
-These groups could be analyzed separately to learn about forecasting behavior under a wide variety of problem attributes.
-
-%The model could be applied to questions grouped by types, such as economic, domestic, or international, to learn about expert forecasting under a wide variety of problem attributes.
+\subsection{Limitations and Directions for Future Research}
+This paper developed a time-series procedure that addresses the main challenges in modeling sparse probability forecasting data.
+The introduced model was kept parsimonious and can be viewed as a basic template that is amenable to many future extensions.
+For instance, most of the model parameters were assumed to remain constant over time.
+It is intuitively reasonable, however, that the behavior of the model parameters is different earlier as opposed to later on in the question.
+For instance, the level of disagreement, $\sigma^2_k$, among the experts can be expected to decrease towards the final time point when the the question finally resolves.
+This hypothesis could be investigated by letting $\sigma^2_{t,k}$ depend on the previous term $\sigma^2_{t-1,k}$ similarly to the hidden process, $X_{t,k}$.
+
+If personal information on the experts were inquired during the data collection process, it is possible to study the forecasting behavior of different kinds of experts.
+For instance, in this paper the multiplicative bias was modeled as a function of self-reported expertise.
+The model, however, is by no means restricted to the study of self-reported expertise.
+Parameter dependencies could be constructed based on many other expert characteristic, such as sex, education or specialty, to produce a range of novel insight on the forecasting behavior across different groups of experts.
+Furthermore, expert characteristics could be interacted with question types, such as economic, domestic or international, to learn about the kind of experts who generally perform well on certain types of questions.
+This would be of interest to the decision-maker who could use the information as a basis for consulting only a high-performing subset of the available experts.
 
@@
 This is a strong assumption that may not always be justified.
-Logit-probabilities, however, have been modeled with a normal distribution before (see, e.g., \citet{Erev1994}).
+Logit-probabilities, however, have been modeled with the normal distribution before (see, e.g., \citet{Erev1994}).
 Furthermore, the normal distribution is a rather standard assumption in psychological models (see, e.g.,  signal-detection theory  in \citet{tanner1954decision}).

### 2013-08-24 paper.tex
Edits

@@
 The goal is to provide application specific insight by discussing the specific research objectives itemized in Section \ref{intro} and also to evaluate the \textit{Sample-Then-Calibrate}-procedure in terms of predictive power and calibration.
-Before presenting the results two practical matters are discussed.
-
-%These matters must be taken into account when aggregating real-world  probability forecasting data, are discussed.
-
-The first matter regards human experts making probability forecasts of 0.0 or 1.0 even when they are not completely sure of the outcome of the future event.
+Before presenting the results, however, we discuss two practical matters that must be taken into account when aggregating real-world  probability forecasts.
+
+\subsection{Incoherent and Imbalanced Data}
+The first matter regards human experts making probability forecasts of 0.0 or 1.0 even if they are not completely sure of the outcome of the event.
 For instance, all the 166 questions in our dataset contained both a zero and a one.
@@
 To avoid this, \citet{Ariely00theeffects} suggest changing $p = 0.0$ and $1.0$ to $p = 0.02$ and $0.98$, respectively.
-This is similar to winsorising which sets the extreme probabilities to a specified percentile of the data (see, e.g., \citet{hastings1947low} for more details on winsorising).
+This is similar to \textit{winsorising} that sets the extreme probabilities to a specified percentile of the data (see, e.g., \citet{hastings1947low} for more details on winsorising).
 \citet{Geo}, on the other hand, consider only probabilities that fall within a constrained interval, say $[0.001, 0.999]$, and discard the rest.
-Since this implies ignoring a portion of the data, we decided to adopt the first approach by changing $p = 0.0$ and $1.0$ to $p = 0.01$ and $0.99$, respectively.
-
-The second matter is related to the distribution of the class labels in the dataset.
-If the set of events that occurred is much larger than the set of events that did not occur (or \textit{vice versa}), the dataset is called \textit{imbalanced}.
-On such data the model can end up over-focusing on the larger class, and as a result, give very accurate forecast performance over the larger class at the cost of performing poorly over the smaller class (see, e.g., \citet{chen2009learning, wallace2012class}).
-Fortunately, it is often possible to detect a well-balanced form of the dataset.
+Since this implies ignoring a portion of the data, we have decided to adopt the first approach by changing $p = 0.0$ and $1.0$ to $p = 0.01$ and $0.99$, respectively.
+
+The second matter is related to the distribution of the class labels in the data.
+If the set of occurred events is much larger than the set of events that did not occur (or \textit{vice versa}), the dataset is called \textit{imbalanced}.
+On such data the modeling procedure can end up over-focusing on the larger class, and as a result, give very accurate forecast performance over the larger class at the cost of performing poorly over the smaller class (see, e.g., \citet{chen2009learning, wallace2012class}).
+Fortunately, it is often possible to use a well-balanced version of the data.
 The first step is to find a partition, $S_0$ and $S_1$, of the question indices, $\{1, 2, \dots, K\}$, such that the equality $\sum_{k \in S_0} T_k  = \sum_{k \in S_1} T_k$ is as closely approximated as possible.
@@
 Recall from section \ref{calibration_step} that $Z_k$ represents the event indicator for the event associated with the $k$th question.
-To define a balanced set of indicators, for all $k \in S_x$, let
+To define a balanced set of indicators, $ \tilde{Z}_k$ for all $k \in S_x$, let
 \begin{align*}
@@
 with $i = 1, \dots, I_k$, and $t = 1, \dots, T_k$.
-The set $\left\{\left(\tilde{Z}_k, \{\tilde{p}_{i,t,k} | i = 1, \dots, I_k, t = 1, \dots, T_k \}\right) \right\}_{k=1}^K$ is a balanced version of the data.
+The resulting set $$\left\{\left(\tilde{Z}_k, \{\tilde{p}_{i,t,k} | i = 1, \dots, I_k, t = 1, \dots, T_k \}\right) \right\}_{k=1}^K$$ is a balanced version of the data.
 This procedure was used to balance our real-world dataset both in terms of events and time points.
@@
 This section evaluates the out-of-sample predictive performance of \textit{Sample-Then-Calibrate} against several other probability aggregation procedures.
-All the models are allowed to utilize a training set before making predictions on an independent testing set.
+The models are allowed to utilize a training set before making predictions on an independent testing set.
 In order to clarify some of the upcoming notation, let $S_{train}$ and $S_{test}$ be index sets that partition the data into training and testing sets of sizes $|S_{train}| = N_{train}$ and $|S_{test}| = 166 - N_{test}$, respectively.
@@
 \begin{enumerate}
-\item \textit{Simple Dynamic Linear Model (SDLM)}.
-This procedure fixes $\boldsymbol{b} = \boldsymbol{1}$ and $\beta = 1$ in the hierarchical model described in Section \ref{model}.
-The model then reduces to
+\item \textit{Simple Dynamic Linear Model (SDLM)}. This is equivalent to the dynamic model from Section \ref{model} but with $\boldsymbol{b} = \boldsymbol{1}$ and $\beta = 1$. Thus,
 \begin{eqnarray*}
@@
 In order to make predictions, the sampler is run for 500 iterations of which the first 200 are used for burn-in.
-The remaining sample of 300 iterations is thinned by discarding every other observation, leaving a final predictive sample of 150 observations.
+The remaining 300 iterations is thinned by discarding every other observation, leaving a final predictive sample of 150 observations.
 
@@
 The model is first fit on the training set by running the sampling step for 3,000 iterations of which the first 500 iterations are used for burn-in.
-After thinning by only keeping every fifth observation, the calibration step is performed for each of the remaining 500 observations.
-The out-of-sample prediction is done by running the sampling step for 500 iterations with each consecutive iteration reading in and conditioning on the next value of $\beta$ and $\boldsymbol{b}$ that were found during training period.
-The first 200 observations are used for burn-in.
-The remaining sample of 300 iterations is thinned by discarding every other observation, leaving a final predictive sample of 150 observations.
+After thinning by only keeping every fifth observation, the calibration step is performed for the remaining 500 observations.
+The out-of-sample prediction is done by running the sampling step for 500 iterations with each consecutive iteration reading in and conditioning on the next value of $\beta$ and $\boldsymbol{b}$ that were found during the training period.
+The first 200 iterations are used for burn-in.
+The remaining 300 iterations are thinned by discarding every other observation, leaving a final predictive sample of 150 observations.
 
@@
 \end{eqnarray}
-Similarly to  \citet{gelman2003bayesian} the prior is chosen to be locally uniform, $p(1/\beta) \propto 1$.
+Similarly to  \citet{gelman2003bayesian}, the prior is chosen to be locally uniform, $p(1/\beta) \propto 1$.
 Posterior estimates of $\beta$ can be sampled from Equation (\ref{OSE2}) using generic sampling algorithms such as the Metropolis algorithm (\citet{metropolis1953equation}) or slice sampling (\citet{neal2003slice}).
-Since the sampling procedure conditions on the event indicators, the full conditional distribution of the hidden states is a non-standard form.
+Since the sampling procedure conditions on the event indicators, the full conditional distribution of the hidden states is not in a standard form.
 Therefore the Metropolis algorithm is also used for sampling the hidden states.
@@
 \item  \textit{Exponentially Weighted Moving Beta-transformed Aggregator (EWMBA)}.
-The static version of the Beta-transformed (linear) aggregator was introduced in \cite{Ranjan08}.
+The static version of the Beta-transformed  aggregator was introduced in \cite{Ranjan08}.
 A dynamic version can be obtained by replacing $G_{t,k}(\nu )$ in the EWMLA description with
@@
 
-Table \ref{prediction} summarizes different ways to aggregate these scores: The first option, denoted by \textit{Scores by Day}, weighs the questions by the number of days the question remained open.
+Table \ref{prediction} summarizes different ways to aggregate these scores.
+The first option, denoted by \textit{Scores by Day}, weighs each question by the number of days the question remained open.
 This is performed by computing the average of the 17,475 scores.
@@
 Both scores can be further broken down into subcategories by considering the length of the questions.
-The questions are divided into \textit{Short} questions (30 days or fewer), \textit{Medium} questions (between 31 and 59 days), and \textit{Long} Problems (60 days or more).
+The final three columns of Table \ref{prediction} divide the questions into \textit{Short} questions (30 days or fewer), \textit{Medium} questions (between 31 and 59 days), and \textit{Long} Problems (60 days or more).
 The number of questions in these subcategories were 36, 32 and 98, respectively.
 The bolded scores indicate the lowest score in each column.
-The values in the parenthesis quantify the variability in the scores: Under \textit{Scores by Day} the values give the standard errors of all the scores.
-Under \textit{Scores by Problem}, on other hand, the values represent the standard errors of the average scores across the questions.
+The values in the parenthesis quantify the variability in the scores.
+Under \textit{Scores by Day} the values give the standard errors of all the scores.
+Under \textit{Scores by Problem}, on other hand, the values represent the standard errors of the average scores of the different questions.
 
@@
  \multicolumn{5}{ c }{Scores by Day}\\
-Model & \multicolumn{1}{ c }{All Problems} & \multicolumn{1}{ c }{Short} & \multicolumn{1}{ c }{Medium} & \multicolumn{1}{ c }{Long}\\ \hline
+Model & \multicolumn{1}{ c }{All} & \multicolumn{1}{ c }{Short} & \multicolumn{1}{ c }{Medium} & \multicolumn{1}{ c }{Long}\\ \hline
 SDLM & 0.100 (0.156) & 0.066 (0.116) & 0.098 (0.154) & 0.102 (0.157)\\ 
@@
 \multicolumn{5}{ c }{Scores by Problem}\\
-Model & \multicolumn{1}{ c }{All Problems} & \multicolumn{1}{ c }{Short} & \multicolumn{1}{ c }{Medium} & \multicolumn{1}{ c }{Long}\\ \hline
+Model & \multicolumn{1}{ c }{All} & \multicolumn{1}{ c }{Short} & \multicolumn{1}{ c }{Medium} & \multicolumn{1}{ c }{Long}\\ \hline
 SDLM & 0.089 (0.116) & 0.064 (0.085) & 0.106 (0.141) & 0.092 (0.117)\\ 
@@
 Overall, STC-Bri and STC-Log achieve the lowest average scores across all columns except \textit{Short} where they are slightly outperformed by BSTC-Log.
-BSTC-Log, however, turns out to be overconfident (see Section \ref{calibration} with a discussion on calibration and sharpness).
-This means that BSTC-Log underestimates the uncertainty in the events and outputs probability forecasts that are typically too near 0.0 or 1.0 to be calibrated.
-As a result, the forecasts are either very close to the correct answer or very far from it.
-As can be seen in Table \ref{prediction}, this results into high variance in performance of BSTC-Log.
+BSTC-Log, however, turns out to be overconfident (see Section \ref{calibration}).
+This means that BSTC-Log underestimates the uncertainty in the events and outputs probability forecasts that are typically too near 0.0 or 1.0.
+As a result, the aggregate forecasts are either very close to the correct answer or very far from it.
+As can be seen in Table \ref{prediction}, this results into highly variable forecasting performance.
 The short questions, however, involve very little uncertainty and were generally the easiest to forecast.
-On such easy questions, overconfidence can pay off frequently enough to compensate for the few potential large scores that arise from an overconfident and incorrect forecast.
-In contrast to the BSTC-Log-model, SDLM lacks sharpness and is highly under-confident.
-This means that SDLM makes conservative forecasts that result into low average scores and low performance variability.
-The \textit{Sample-Then-Calibrate}-model can therefore be viewed as a well-performing compromise that avoids over-confidence without being too conservative.
-
- Of the exponentially weighted moving aggregators, EWMBA performs the best and EWMA the worst.
-Given that the static versions of EWMLA and EWMBA depend on similar transformations, the results suggest that EWMLA is unable to make efficient use of the self-rated expertise information.
-Both EWMLA and EWMBA, however, suffer from over-confidence and are therefore unable to outperform the \textit{Sample-Then-Calibrate}-model.
-
-The dataset contained a few long questions with an abrupt change of the crowd belief from one answer option to another.
+On such easy questions, overconfidence can pay off frequently enough to compensate for a few large scores arising from the overconfident and incorrect forecasts.
+
+In comparison to BSTC-Log, the baseline SDLM-model lacks sharpness and is highly under-confident (see Section \ref{calibration}).
+This behavior is expected as the experts are under-confident at the group-level (see Section \ref{ExpertBias}) and the SDLM-procedure does not use the training set to explicitly calibrate its forecasts.
+Instead, it merely smooths the forecasts given by the experts.
+The resulting aggregate forecasts are therefore necessarily conservative and result into high average scores with low variability.
+
+ Similarly behavior is exhibited by EWMA that performs the worst of all the competing models.
+The other two exponentially weighted aggregators, namely EWMLA and EWMBA, on other hand, make efficient use of the training set and present good forecasting performance in most columns of Table \ref{prediction}.
+Recall that EWMBA uses the cumulative distribution function of the Beta distribution to transform the average forecast probability.
+This transformation depends on two parameters and is more flexible than the transformation used by EWMLA.
+On other hand, only EWLMA is given access to the self-reported expertise information.
+Based on the results, however, it is unable to use this information to outperform EWMBA.
+Overall, these two aggregators perform very similarly.
+They both suffer from over-confidence and are hence unable to outperform the \textit{Sample-Then-Calibrate}-procedure.
+
+%The dataset contained a few long questions with an abrupt change of the crowd belief from one answer option to another.
 Since these abrupt changes happened generally towards the end of each question, during most days the Brier scores were really high.
@@
 The idea is to plot the probability forecasts against the observed empirical frequencies.
-Therefore any deviation from the diagonal line suggest poor calibration.
-A model is considered under-confident (or over-confidence) if the points form an S-shaped (or \reflectbox{S}-shaped) trend.
-In order to assess sharpness of the model, it is common practice to place a histogram of the given forecasts within the plot.
+Any deviation from the diagonal line therefore suggests poor calibration.
+A model is considered under-confident (or over-confidence) if the points follow an S-shaped (or \reflectbox{S}-shaped) trend.
+In order to assess sharpness of the model, it is common practice to place a histogram of the given forecasts in the corner of the plot.
 
@@
 The top and bottom rows of Figure \ref{Calibration-Out} present calibration plots for SDLM, STC-Log, STC-Bri, and BSTC-Log under in- and out-of-sample probability estimation, respectively.
-Each case is of interest in its own right:  Good in-sample calibration is crucial for model interpretability.
+Each setting is of interest in its own right:  Good in-sample calibration is crucial for model interpretability.
 In particular, if the estimated crowd belief is well-calibrated, then the elements of the bias vector $\boldsymbol{b}$ can be used to study the amount of under- or over-confidence in the different expertise groups.
@@
 These have been computed by running the bootstrap technique described in \citet{brocker2007increasing} for 10,000 iterations.
-The in-sample results were obtained by running the models were for 10,200 iterations leading to a final posterior sample of 1,000 observations after thinning and using the first 200 iterations for burn-in.
-The out-of-sample results were computed based on the 10-fold cross-validation discussed in Section \ref{forecasting}.
+The in-sample predictions were obtained by running the models for 10,200 iterations, leading to a final posterior sample of 1,000 observations after thinning and using the first 200 iterations for burn-in.
+The out-of-sample predictions were given by the 10-fold cross-validation discussed in Section \ref{forecasting}.
 
@@
 Since this model does not perform any explicit calibration, it is not surprising to see most points outside the critical values.
-The pattern in the deviations suggests drastic under-confidence, and the inset histogram reveals drastic lack of sharpness.
+The pattern in the deviations suggests drastic under-confidence.
+Furthermore, the inset histogram reveals drastic lack of sharpness.
+Therefore the \textit{Sample-Then-Calibrate}-model can be viewed as a well-performing compromise between SDLM and BSTC-Log that avoids over-confidence without being too conservative.
 

### 2013-08-26 paper.tex
Edits

@@
 Individual experts can differ radically from one another in their abilities to assess probabilities of future events.
-The quality of a probability assessment is typically evaluated in regards to \textit{calibration} that measures the degree to which the frequency of occurrence of the events agree with the given probabilities.
-This means that, for instance, the proportion of events that actually occurred is 60\% for all those events for which a well-calibrated expert assessed a probability of 0.60.
-Unfortunately, several experiments have been conducted to show that experts are generally poorly calibrated (see, e.g., \citet{cooke1991experts, shlyakhter1994quantifying}).
-\citet{wright1994coherence}, however, argue that a higher level of self-reported expertise on the subject-matter can be associated with better calibration.
+The quality of a probability assessment is typically evaluated in regards to \textit{calibration}, which is an attribute measuring the degree to which the frequency of occurrence of the events agree with the given probabilities.
+This means that, for instance, the proportion of occurred events is 60\% for all those events for which a well-calibrated expert assessed a probability of 0.60.
+Even though several experiments have been conducted to show that experts are generally poorly calibrated (see, e.g., \citet{cooke1991experts, shlyakhter1994quantifying}), relative differences can occur between different types of experts.
+In particular, \citet{wright1994coherence} argue that a higher level of self-reported expertise on the subject-matter can be associated with better calibration.
 
@@
 In order to assess sharpness of the model, it is common practice to place a histogram of the given forecasts in the corner of the plot.
+Since the data were balanced, any deviation from the the baseline probability at 0.5 suggests improved sharpness.
 

### 2013-08-26 paper.tex
Edits (Second Draft)

@@
 The updates are assumed to occur very infrequently, resulting into a highly sparse dataset that cannot be modeled by standard time-series procedures.
-In response to the lack of appropriate methodology, this paper presents a dynamic model that takes into account the forecaster’s level of self-reported expertise and produces aggregate probabilities that are sharp and well-calibrated both in- and out-of-sample.
+In response to the lack of appropriate methodology, this paper presents a dynamic model that takes into account the forecaster's level of self-reported expertise and produces aggregate probabilities that are sharp and well-calibrated both in- and out-of-sample.
 The model is demonstrated on a real-world dataset that includes over 2,300 experts making multiple probability forecasts on different subsets of 166 different international political events.
@@
 As a result, standard time-series procedures like ARIMA (see, e.g., \citet{mills1991time}) are not directly applicable to our setting.
-
 In response to the lack of appropriate modeling procedures, this paper introduces a time-series model that incorporates self-reported expertise and captures a sharp and well-calibrated estimate of the crowd belief.
@@
 
-The data collection began with a recruitment of 2,365 forecasters ranging from graduate students to political science faculty and practitioners.
+The data collection began with recruitment of 2,365 forecasters ranging from graduate students to political science faculty and practitioners.
 The recruiting was made from professional societies, research centers, alumni associations, science bloggers, and word of mouth.
@@
 These measures assessed cognitive styles, cognitive abilities, personality traits, political attitudes, and real-world knowledge.
-The experts were asked to give probability forecasts (to the second decimal point) and to self-assess their level of expertise (measured on a 1-to-5 scale with 1 = Not At All Expert and 5 = Extremely Expert) on a number of 166 geopolitical binary events taking place between September 29, 2011 and May 8, 2013.
+The experts were asked to give probability forecasts (to the second decimal point) and to self-assess their level of expertise (on a 1-to-5 scale with 1 = Not At All Expert and 5 = Extremely Expert) on a number of 166 geopolitical binary events taking place between September 29, 2011 and May 8, 2013.
 Each question was active for a period of time during which the participating experts were allowed to update their forecasts as frequently as they liked without being penalized.
@@
 In addition to receiving \$150 for meeting minimum participation requirements that did not depend on prediction accuracy, the forecasters also received status rewards for their performance via leader-boards displaying Brier scores for the top 20 forecasters.
-Since a typical expert participated only in a small subset of the 166 questions, the forecasters are considered indistinguishable conditional on the self-reported level of expertise.
+Since a typical expert participated only in a small subset of the 166 questions, the forecasters are considered indistinguishable conditional on the level of self-reported expertise.
 Tables \ref{DataStats} and \ref{ExpertiseTable} provide more relevant summary statistics on the data.
@@
 The points have been jittered slightly to make overlaps visible.
-The darkness of the points are positively associated with the self-reported expertise levels.
+The darkness of the points is positively associated with the self-reported expertise.
 Given that the European bailout fund was ratified before 1 November 2011 and that the Nikkei 225 index finished trading at around 8,700 on 30 September 2011, the general trend of the probability forecasts converge to the correct answers.
@@
 Unfortunately, the data is not only noisy but also quite sparse.
-As mentioned before, the forecasters were allowed to update their forecast as long as the questions remained active.
-Updating was done on a very infrequent basis.
-An expert gave, on average, only 0.0159 forecasts per day, and so a typical expert made a forecast about every 62 days.
+As mentioned before, the forecasters were allowed to update their forecast as long as the questions remained open.
+Updating, however, was done on a very infrequent basis.
+An expert gave, on average, only 0.016 forecasts per day, and so a typical expert made a forecast about every 62 days.
 As a result the average response rate was around 13.5 forecasts per day from a large group of experts.
@@
 The resulting data is too sparse to be modeled efficiently with standard time-series methods.
-In response, our work introduces an interpretable time-series model that incorporates self-reported expertise and captures a sharp and well-calibrated crowd belief across time.
-The model is estimated in two-steps: The first step, known as the \textit{sampling step}, samples constrained versions of the model parameters via a Gibbs sampler that only makes use of standard distributions.
+In response to the lack of appropriate modeling procedures, our work introduces an interpretable time-series model that incorporates self-reported expertise to capture a sharp and well-calibrated estimate of the crowd belief.
+The model is estimated in two-steps: The first step, known as the \textit{sampling step}, samples constrained versions of the model parameters via a Gibbs sampler that only makes use of standard distributions (see Appendix \ref{appendix} for technical details).
 The second step, known as the \textit{calibration step}, involves a one-dimensional optimization procedure that transforms the constrained parameter values to their unconstrained counterparts.
@@
 For instance, the level of disagreement, $\sigma^2_k$, among the experts can be expected to decrease towards the final time point when the the question finally resolves.
-This hypothesis could be investigated by letting $\sigma^2_{t,k}$ depend on the previous term $\sigma^2_{t-1,k}$ similarly to the hidden process, $X_{t,k}$.
+This hypothesis could be investigated by letting $\sigma^2_{t,k}$ depend on the previous term $\sigma^2_{t-1,k}$.
 
@@
 The model, however, is by no means restricted to the study of self-reported expertise.
-Parameter dependencies could be constructed based on many other expert characteristic, such as sex, education or specialty, to produce a range of novel insight on the forecasting behavior across different groups of experts.
-Furthermore, expert characteristics could be interacted with question types, such as economic, domestic or international, to learn about the kind of experts who generally perform well on certain types of questions.
+Parameter dependencies could be constructed based on many other expert characteristic, such as gender, education, or specialty, to produce a range of novel insight on the forecasting behavior across different groups of experts.
+Furthermore, expert characteristics could be interacted with question types, such as economic, domestic, or international, to learn about the kind of experts who generally perform well on certain types of questions.
 This would be of interest to the decision-maker who could use the information as a basis for consulting only a high-performing subset of the available experts.

### 2013-08-26 paper.tex
Edits

@@
 \section{Discussion}
-This paper began with an introduction of a rather unorthodox but nonetheless realistic time-series setting in which probability forecasts are made very infrequently by a heterogeneous group of experts.
-The resulting data is too sparse to be modeled efficiently with standard time-series methods.
-In response to the lack of appropriate modeling procedures, our work introduces an interpretable time-series model that incorporates self-reported expertise to capture a sharp and well-calibrated estimate of the crowd belief.
-The model is estimated in two-steps: The first step, known as the \textit{sampling step}, samples constrained versions of the model parameters via a Gibbs sampler that only makes use of standard distributions (see Appendix \ref{appendix} for technical details).
-The second step, known as the \textit{calibration step}, involves a one-dimensional optimization procedure that transforms the constrained parameter values to their unconstrained counterparts.
+This paper began with an introduction of a rather unorthodox but nonetheless realistic time-series setting where probability forecasts are made very infrequently by a heterogeneous group of experts.
+The resulting data is too sparse to be modeled well with standard time-series methods.
+In response to this lack of appropriate modeling procedures, our work introduces an interpretable time-series model that incorporates self-reported expertise to capture a sharp and well-calibrated estimate of the crowd belief.
+The model estimation is performed in two steps.
+The first step, known as the \textit{sampling step}, samples constrained versions of the model parameters via a Gibbs sampler.
+The sampling is done from standard distributions with fast convergence to the target distribution (see Appendix \ref{appendix} for technical details).
+The second step, known as the \textit{calibration step}, uses a one-dimensional optimization procedure to transforms the constrained parameter values to their unconstrained counterparts.
+With minimal sacrifices in model interpretability and efficiency, this procedure is able to extend forecasting literature towards rather unexplored areas of probability aggregation.
 
 \subsection{Summary of Findings}
-The model was applied to an unusually large dataset on probability forecasting.
-The estimated crowd belief was found to be sharp and well-calibrated under both in- and out-of-sample probability estimation.
+The model was applied to an unusually large dataset on expert probability forecasts.
+The estimated crowd belief was found to be sharp and well-calibrated under both in- and out-of-sample settings.
 This has direct implications  on predictive power and model interpretability.
@@
 This result is about groups of experts and hence does not conflict with the well-known result of the individuals being over-confident (see, e.g.,  \citet{lichtenstein1977calibration, morgan1992uncertainty, bier2004implications}).
-Besides making predictions or studying group-level bias, the model can be used to generate estimates of many problem-specific quantities.
-These quantities have clear interpretations and can be combined in many interesting ways to answer a range of hypotheses about different types of questions and expert behavior.
+Besides making predictions or studying group-level bias, the model can be used to generate estimates of many problem-specific parameters.
+These quantities have clear interpretations and can be combined in many interesting ways to explore a range of hypotheses about different types of questions and expert behavior.
 
 \subsection{Limitations and Directions for Future Research}
-This paper developed a time-series procedure that addresses the main challenges in modeling sparse probability forecasting data.
-The introduced model was kept parsimonious and can be viewed as a basic template that is amenable to many future extensions.
-For instance, most of the model parameters were assumed to remain constant over time.
-It is intuitively reasonable, however, that the behavior of the model parameters is different earlier as opposed to later on in the question.
-For instance, the level of disagreement, $\sigma^2_k$, among the experts can be expected to decrease towards the final time point when the the question finally resolves.
-This hypothesis could be investigated by letting $\sigma^2_{t,k}$ depend on the previous term $\sigma^2_{t-1,k}$.
-
-If personal information on the experts were inquired during the data collection process, it is possible to study the forecasting behavior of different kinds of experts.
-For instance, in this paper the multiplicative bias was modeled as a function of self-reported expertise.
-The model, however, is by no means restricted to the study of self-reported expertise.
-Parameter dependencies could be constructed based on many other expert characteristic, such as gender, education, or specialty, to produce a range of novel insight on the forecasting behavior across different groups of experts.
+Our model pertains model parsimony while addressing the main challenges that arise from modeling sparse probability forecasting data.
+Therefore it can be viewed as a basis for many future extensions.
+To illustrate some potential directions, recall that most of the model parameters are assumed to remain constant over time.
+It is intuitively reasonable, however, that these model parameters behave differently during different time intervals of the question.
+For instance, the level of disagreement (represented by $\sigma^2_k$ in our model) among the experts can be expected to decrease towards the final time point when the the question finally resolves.
+This hypothesis could be explored by letting $\sigma^2_{t,k}$ evolve dynamically as a function of the previous term $\sigma^2_{t-1,k}$ and random noise.
+
+If personal information on the experts were inquired during the data collection process, it may be of interest to study the forecasting behavior of different kinds of experts.
+For instance, this paper modeled the bias separately within each expertise group.
+This, however, is by no means restricted to the study of bias or its relation to self-reported expertise.
+Different parameter dependencies could be constructed based on many other expert characteristic, such as gender, education, or specialty, to produce a range of novel insight on the forecasting behavior across different groups of experts.
 Furthermore, expert characteristics could be interacted with question types, such as economic, domestic, or international, to learn about the kind of experts who generally perform well on certain types of questions.

### 2013-08-29 paper.tex
Edits

@@
 % "Title of the paper"
-\title{Probability Aggregation: Dynamic Hierarchical Modeling of Expert Beliefs in Geopolitical Events}
+\title{Probability Aggregation: Dynamic Hierarchical Modeling of Sparse Expert Beliefs}
 \runtitle{Hierarchical Model for Expert Beliefs}
@@
 
-\vspace{-1em}
 \begin{figure}[h!]
+\vspace{-2em}
 \centering
-\hspace*{0em} 	\includegraphics[width=  \textwidth]{Figures/LegendExamplePlot} % requires the graphicx package
-\vspace{-4.5em}
+\hspace*{0em} 	\includegraphics{Figures/LegendExamplePlot} % requires the graphicx package
+\vspace{-3.5em}
 
@@
 
-\section{Synthetic Data Results \ville{TODO}}
+\section{Synthetic Data Results}
 \label{syntheticData}
 
-The synthetic data is not generated directly from the model for several reasons: (i) showing good performance on a dataset directly generated from the model assumptions is hardly any news, and (ii) generating data from the dynamic model description does not produce calibrated hidden states.
-The latter is important for our interpretation of the hidden states as the calibrated crowd belief.
-In order to achieve this, let $\{ Z_{t,k} \}_{t=1}^{T_k}$ denote the values of a standard Brownian motion at time points $t = 1, \dots, T_k, T_k $.
+The synthetic data is not generated directly from the model for two reasons: (i) showing good performance on a dataset directly generated from the model assumptions is hardly any news, and (ii) generating data from the dynamic model description does not produce well-calibrated hidden states.
+The latter is important for our interpretation of the hidden states as the sharp and well-calibrated crowd belief.
+In order to achieve this, let $\{ Z_{t,k} \}_{t=1}^{T_k}$ denote the values of a standard Brownian motion at time points $t = 1, \dots, T_k $.
 Then,
@@
 \end{eqnarray*}
-Once the hidden process has been generated, the expert logit forecasts are produced by adding white noise to the given sequences of $X_{t,k}$.
-
-The simulation is performed over a four-dimensional grid of values: The number of questions, $K$, varies between $10, 25, 50, 75$, and $100$, the values of $\beta$ change from $0.5$ to $1.5$ with steps of size 0.25, the values of $\alpha$ range from $-0.5$ to $0.5$ with steps of size 0.25, and finally the values of $\sigma^2$ vary from $0.25$ to $1.5$ with steps of size 0.25.
-At each grid point the simulation runs a total of two times; each time measuring how well the approaches estimate the hidden process in terms of quadratic loss.
-The average loss from the two runs is used as the performance measure at that grid point.
-Since the final results are summarized by choosing one of the grid variables (e.g.
-$\alpha$) and then averaging over the remaining three variables (e.g.
-$\beta$, $K$, and $\sigma^2$), each average value in the final results represents an average of a $2 \times 5^3 = 250$ values.
+Once the hidden process has been generated, the expert logit-probabilities are produced from Equation (\ref{observedpr}) by applying bias and white noise to the given sequences of $X_{t,k}$.
+Following the real-world data setting in Section \ref{realData}, extreme logit-probabilities are truncated to avoid overly influential observations.
+More specifically, any $p_{i,t,k} = \logit^{-1}(Y_{i,t,k})$ less than 0.01 (or greater than 0.99) is truncated to 0.01 (or to 0.99).
+
+The simulation is then performed over a three-dimensional grid of values: The number of questions, $K$, varies between $20$ and $80$ with steps of 15, the bias vector changes as $$\boldsymbol{b} = [0.50, 0.75, \dots, 1.50]^T \beta,$$ where $\beta$ takes values in $1/2, 3/4, 1, 4/3$, and $2/1$ symmetrically around 1 in the log-scale, and finally the values of $\sigma^2$ vary from $0.5$ to $2.0$ with steps of size 0.5.
+These grid values were considered realistic based on our experience with the real-world probability forecasting data.
+At each grid point the simulation runs a total of 40 times; each time measuring how well the approaches estimate the hidden process and bias vector.
+The average loss from the 40 runs is used as the performance measure at that grid point.
+
+Figure \ref{Synthetic} summarizes the results in a series of plots.
+The first row shows each grid variable's marginal effect on the accuracy to which the \textit{Sample-Then-Calibrate} under the logarithmic loss (STC-Log) and the Brier score (STC-Bri) are able to estimate the hidden process.
+The second row is similar in nature but measures the accuracy of estimating the bias vector instead.
+The plots were computed by choosing one grid variable (e.g.
+$\beta$)  at a time and averaging over the remaining two variables  (e.g.
+$K$, and $\sigma^2$).
+Therefore each value in the plots represents an average of a $40 \times 5^2 = 1,000$ values.
+
+Based on the first row of plots, STC-Log and STC-Bri estimate the hidden process to a similar accuracy.
+Even though some relative differences are visible in Figure \ref{SyntheticK}, the loss values are very small in absolute terms; hence suggesting good accuracy to capture the hidden process.
+The main differences arise in estimating the bias vector.
+Based on the second row of plots, the procedures capture the bias vector rather precisely as long as the true values of $\beta$ and $\sigma^2$ are not very small.
+Surprisingly, even though STC-Log is more accurate at estimating the bias vector across all three grid dimensions, it is unable to outperform STC-Bri in estimating the hidden process.
+
+\ville{Talk about strictly decreasing scores in beta}
 
@@
 \centering
-\hspace*{2em} 	\includegraphics[width=  \textwidth]{Figures/LegendSynthetic} % requires the graphicx package
-\vspace{-3em}
+ \hspace{-5em}	\includegraphics{Figures/LegendSynthetic} % requires the graphicx package
 
 \subfigure[]{
-\hspace*{-1em}  \includegraphics[width= 0.33\textwidth]{Figures/SyntheticSigma2}
+\hspace{-1em} \includegraphics{Figures/SyntheticSigma2}
 \label{SyntheticSigma2}
@@
 \subfigure[]{
-\hspace*{-1em} \includegraphics[width=  0.33\textwidth]{Figures/SyntheticBeta}
+\hspace{-1.5em} \includegraphics{Figures/SyntheticBeta}
 \label{SyntheticBeta}
@@
 \subfigure[]{
-\hspace*{-1em} \includegraphics[width=  0.33\textwidth]{Figures/SyntheticK}
+\hspace{-1.5em} \includegraphics{Figures/SyntheticK}
 \label{SyntheticK}
@@
 
-\vspace{-2em}
 \subfigure[]{
-\hspace*{-1em}  \includegraphics[width= 0.33\textwidth]{Figures/SyntheticSigma2bias}
-\label{SyntheticSigma2}
+\hspace{-1em}  \includegraphics{Figures/SyntheticSigma2bias}
+\label{SyntheticSigma2b}
 }
 \subfigure[]{
-\hspace*{-1em} \includegraphics[width=  0.33\textwidth]{Figures/SyntheticBetabias}
-\label{SyntheticBeta}
+\hspace{-1.5em}  \includegraphics{Figures/SyntheticBetabias}
+\label{SyntheticBetab}
 }
 \subfigure[]{
-\hspace*{-1em} \includegraphics[width=  0.33\textwidth]{Figures/SyntheticKbias}
-\label{SyntheticK}
-}
-
-\caption[Optional caption for list of figures]{Comparing the approaches on synthetic data.
+\hspace{-1.5em}   \includegraphics{Figures/SyntheticKbias}
+\label{SyntheticKb}
+}
+
+\caption[Optional caption for list of figures]{Evaluating the \textit{Sample-Then-Calibrate}-model under synthetic data.
 (a) Performance under different values of $\sigma^2$.
-(b) Performance under different values of $\alpha$.
-(c) Performance under different values of $\beta$.
-(d) Performance under different numbers of questions, $K$.}
+(b) Performance under different values of $\beta$.
+(c) Performance under different numbers of questions, $K$.}
 \label{Synthetic}
 \end{figure}
-
-Figure \ref{Synthetic} summarizes the results by choosing one grid variable at a time and then averaging over the remaining three variables.
-The \textit{Sample-Then-Calibrate} with log loss outperforms the other two approaches almost uniformly across the grid.
 
@@
 \begin{figure}[h!]
+\newgeometry{left=5cm, right = 1cm, bottom=0.1cm}
+
 \subfigure[\scriptsize SDLM In-Sample]{
-\hspace*{-1.1em}
-  \includegraphics[width= 0.28\textwidth]{Figures/CalibrationNONE10}
+\hspace*{-3em}  \includegraphics{Figures/CalibrationNONE10}
 \label{CalibrationNONEa}
@@
 \subfigure[\scriptsize STC-Log In-Sample]{
-\hspace*{-3em}  \includegraphics[width= 0.28\textwidth]{Figures/CalibrationLOG10}
+\hspace*{-1.15em} \includegraphics{Figures/CalibrationLOG10}
 \label{CalibrationPLATTLOG10}
@@
 \subfigure[\scriptsize STC-Bri In-Sample]{
-\hspace*{-3em} \includegraphics[width= 0.28\textwidth]{Figures/CalibrationBRI10}
+\hspace*{-1.15em} \includegraphics{Figures/CalibrationBRI10}
 \label{CalibrationPLATTBRI10}
@@
 \subfigure[\scriptsize BSTC-Log In-Sample]{
-\hspace*{-3em} \includegraphics[width= 0.28\textwidth]{Figures/CalibrationFULL10}
-\label{CalibrationPLATTFULL10}
-}
-
-\vspace{-1.5em}
-\subfigure[\scriptsize SDLM Out-of-Sample]{
-\hspace*{-1.1em}  \includegraphics[width= 0.28\textwidth]{Figures/CalibrationNONE}
+\hspace*{-1.15em} \includegraphics{Figures/CalibrationFULL10}
+\hspace*{-2em} \label{CalibrationPLATTFULL10}
+}
+
+\subfigure[ \scriptsize SDLM Out-of-Sample]{
+\hspace*{-3em} \includegraphics{Figures/CalibrationNONE}
 \label{CalibrationNONEb}
 }
-\subfigure[\scriptsize STC-Log Out-of-Sample]{
-\hspace*{-3em}  \includegraphics[width= 0.28\textwidth]{Figures/CalibrationSTO-Log}
+\subfigure[ \scriptsize STC-Log Out-of-Sample]{
+\hspace*{-1.15em} \includegraphics{Figures/CalibrationSTO-Log}
 \label{CalibrationSTC-Log}
 }
-\subfigure[\scriptsize STC-Bri Out-of-Sample]{
-\hspace*{-3em} \includegraphics[width= 0.28\textwidth]{Figures/CalibrationSTO-Bri}
+\subfigure[ \scriptsize STC-Bri Out-of-Sample]{
+\hspace*{-1.15em} \includegraphics{Figures/CalibrationSTO-Bri}
 \label{CalibrationSTC-Bri}
@@
 \subfigure[\scriptsize  BSTC-Log Out-of-Sample]{
-\hspace*{-3em} \includegraphics[width= 0.28\textwidth]{Figures/CalibrationSO}
-\label{CalibrationBSTC-Log}
-}
-
+\hspace*{-1.15em} \includegraphics{Figures/CalibrationSO}
+\hspace*{-2em}\label{CalibrationBSTC-Log}
+}
+\restoregeometry
 \caption[Optional caption for list of figures]{The top and bottom rows show in- and out-of-sample calibration and sharpness, respectively.
@@
 \label{BiasesLOGBoxplots}
+\vspace*{-2em}
 
@@
 \centering
-\hspace*{0em} 	\includegraphics[width=  \textwidth]{Figures/LegendExamplePlot} % requires the graphicx package
-\vspace{-4.5em}
+\hspace*{0em} 	\includegraphics{Figures/LegendExamplePlot} % requires the graphicx package
+\vspace{-3.5em}
 
@@
 \subfigure[\textit{Will the expansion of the European bailout fund be ratified by all 17 Eurozone nations before 1 November 2011?}]{
-\hspace*{-1em}  \includegraphics[width= 0.49\textwidth]{Figures/ExamplePlot1b}
+\hspace*{-1em}  \includegraphics{Figures/ExamplePlot1b}
 \label{Example1b}
@@
 \subfigure[\textit{Will the Nikkei 225 index finish trading at or above 9,500 on 30 September 2011?}]{
-\hspace*{-1em}  \includegraphics[width= 0.49\textwidth]{Figures/ExamplePlot2b}
+\hspace*{-1em}  \includegraphics{Figures/ExamplePlot2b}
 \label{Example2b}

### 2013-08-31 paper.tex
Edits

@@
 The synthetic data is not generated directly from the model for two reasons: (i) showing good performance on a dataset directly generated from the model assumptions is hardly any news, and (ii) generating data from the dynamic model description does not produce well-calibrated hidden states.
-The latter is important for our interpretation of the hidden states as the sharp and well-calibrated crowd belief.
-In order to achieve this, let $\{ Z_{t,k} \}_{t=1}^{T_k}$ denote the values of a standard Brownian motion at time points $t = 1, \dots, T_k $.
-Then,
+The latter is important for our interpretation of the hidden states as a sharp and well-calibrated versions of the crowd belief.
+To generate a well-calibrated hidden process, generate a path of the standard Brownian motion until time $T_k$ for question $k$, and let $Z_{t,k}$ denote the value of the path at time $t$.
+Then, for $t = 1, \dots, T_k$,
 \begin{eqnarray*}
@@
 \end{eqnarray*}
-Once the hidden process has been generated, the expert logit-probabilities are produced from Equation (\ref{observedpr}) by applying bias and white noise to the given sequences of $X_{t,k}$.
-Following the real-world data setting in Section \ref{realData}, extreme logit-probabilities are truncated to avoid overly influential observations.
-More specifically, any $p_{i,t,k} = \logit^{-1}(Y_{i,t,k})$ less than 0.01 (or greater than 0.99) is truncated to 0.01 (or to 0.99).
-
-The simulation is then performed over a three-dimensional grid of values: The number of questions, $K$, varies between $20$ and $80$ with steps of 15, the bias vector changes as $$\boldsymbol{b} = [0.50, 0.75, \dots, 1.50]^T \beta,$$ where $\beta$ takes values in $1/2, 3/4, 1, 4/3$, and $2/1$ symmetrically around 1 in the log-scale, and finally the values of $\sigma^2$ vary from $0.5$ to $2.0$ with steps of size 0.5.
-These grid values were considered realistic based on our experience with the real-world probability forecasting data.
-At each grid point the simulation runs a total of 40 times; each time measuring how well the approaches estimate the hidden process and bias vector.
-The average loss from the 40 runs is used as the performance measure at that grid point.
+gives a sequence of $T_k$ calibrated logit-probabilities for the event $Z_k = 1$.
+
+We used this procedure to generate the hidden processes for $K$ questions with $T_k = 100$.
+Each question involved 50 experts allocated evenly among the five expertise groups.
+Their reported logit-probabilities, $Y_{t,k}$, were generated from Equation (\ref{observedpr}) by applying bias and noise to the hidden process.
+Similarly to the real-world data setting discussed in Section \ref{realData}, extreme logit-probabilities are truncated to avoid overly influential observations; that is, any $p_{i,t,k} = \logit^{-1}(Y_{i,t,k})$ less than 0.01 (or greater than 0.99) is truncated to 0.01 (or to 0.99).
+The simulation was then performed over a three-dimensional grid of parameter values:
+\begin{eqnarray*}
+\beta &\in& \{1/2, 3/4, 1, 4/3, 2/1\}\\
+\sigma^2 &\in& \{0.5, 1.0, 1.5, 2.0, 2.5\}\\
+K &\in& \{20, 35, 50, 65, 80\},
+\end{eqnarray*}
+where $\beta$ generates the bias vector by  $\boldsymbol{b} = [0.50, 0.75, \dots, 1.50]^T \beta$. 
+The simulation runs a total of 40 times at every grid point. Each run records the accuracy to which the approaches are able to estimate the hidden process and bias vector.
+
+%The number of questions, $K$, varies between $20$ and $80$ with increments of 15 problems, the bias vector changes as $$\boldsymbol{b} = [0.50, 0.75, \dots, 1.50]^T \beta,$$ where $\beta$ takes values in $1/2, 3/4, 1, 4/3$, and $2/1$ symmetrically around 1 in the log-scale, and finally the values of $\sigma^2$ vary from $0.5$ to $2.0$ with increments of $0.5$.
 
 Figure \ref{Synthetic} summarizes the results in a series of plots.
-The first row shows each grid variable's marginal effect on the accuracy to which the \textit{Sample-Then-Calibrate} under the logarithmic loss (STC-Log) and the Brier score (STC-Bri) are able to estimate the hidden process.
+The first row shows the marginal effects of the three grid variables on the accuracy to which the \textit{Sample-Then-Calibrate} under the logarithmic loss (STC-Log) and the Brier score (STC-Bri) are able to estimate the hidden process.
+The accuracy is measured in terms of the average quadratic loss in the probability space, $\sum_{k=1}^K \sum_{t=1}^{T_k} (\hat{p}_{t,k} - \logit^{-1}(X_{t,k}))^2 / \sum_{k=1}^K T_k$.
 The second row is similar in nature but measures the accuracy of estimating the bias vector instead.
+The accuracy is measured in terms of the average quadratic loss, $|| \boldsymbol{b} - \hat{\boldsymbol{b}}||^2/5$.
 The plots were computed by choosing one grid variable (e.g.
@@
 Based on the first row of plots, STC-Log and STC-Bri estimate the hidden process to a similar accuracy.
-Even though some relative differences are visible in Figure \ref{SyntheticK}, the loss values are very small in absolute terms; hence suggesting good accuracy to capture the hidden process.
-The main differences arise in estimating the bias vector.
-Based on the second row of plots, the procedures capture the bias vector rather precisely as long as the true values of $\beta$ and $\sigma^2$ are not very small.
-Surprisingly, even though STC-Log is more accurate at estimating the bias vector across all three grid dimensions, it is unable to outperform STC-Bri in estimating the hidden process.
-
-\ville{Talk about strictly decreasing scores in beta}
+Even though some relative differences are visible, the loss values are very small in absolute terms; hence suggesting good overall accuracy to capture the hidden process.
+While increasing the number of questions improves the accuracy to a small extent, estimating the hidden process becomes more difficult as either $\sigma^2$ or $\beta$ gets larger.
+The latter seems more unintuitive and can be understood better by observing that larger values of $\beta$ lead to more extreme expert logit-probabilities.
+As a result, more logit-probabilities are modified by truncation, leading to poorer estimation accuracy.
+
+Based on the second row of plots, STC-Log and STC-Bri  capture the bias vector rather precisely as long as the true values of $\beta$ and $\sigma^2$ are not very large.
+This behavior can be considered as being caused  to a large extent by the truncation of the extreme expert logit-probabilities.
+Surprisingly, even though STC-Log is more accurate in estimating the bias vector, it is unable to outperform STC-Bri in estimating the hidden process.
 
 \begin{figure}[h!]
+\vspace{-5em}
 \centering
- \hspace{-5em}	\includegraphics{Figures/LegendSynthetic} % requires the graphicx package
-
-\subfigure[]{
+ \hspace{-30em}	\includegraphics{Figures/LegendSynthetic} % requires the graphicx package
+\vspace{-0.5em}
+
+\subfigure{
 \hspace{-1em} \includegraphics{Figures/SyntheticSigma2}
@@
 }
-\subfigure[]{
+\subfigure{
 \hspace{-1.5em} \includegraphics{Figures/SyntheticBeta}
@@
 }
-\subfigure[]{
+\subfigure{
 \hspace{-1.5em} \includegraphics{Figures/SyntheticK}
@@
 
-\subfigure[]{
+\subfigure{
 \hspace{-1em}  \includegraphics{Figures/SyntheticSigma2bias}
@@
 }
-\subfigure[]{
+\subfigure{
 \hspace{-1.5em}  \includegraphics{Figures/SyntheticBetabias}
@@
 }
-\subfigure[]{
+\subfigure{
 \hspace{-1.5em}   \includegraphics{Figures/SyntheticKbias}
@@
 
-\caption[Optional caption for list of figures]{Evaluating the \textit{Sample-Then-Calibrate}-model under synthetic data.
-(a) Performance under different values of $\sigma^2$.
-(b) Performance under different values of $\beta$.
-(c) Performance under different numbers of questions, $K$.}
+\caption[Optional caption for list of figures]{Comparing \textit{Sample-Then-Calibrate} that optimizes over the logarithmic score (STC-Log) and \textit{Sample-Then-Calibrate} that optimizes over the Brier score (STC-Bri) under synthetic data.
+The top row presents the accuracy to capture the hidden process.
+The bottom row presents the accuracy to capture the bias vector.
+}
 \label{Synthetic}

### 2013-08-31 paper.tex
Edits

@@
 
-The data collection began with recruitment of 2,365 forecasters ranging from graduate students to political science faculty and practitioners.
+The data collection began with a recruitment of 2,365 forecasters ranging from graduate students to political science faculty and practitioners.
 The recruiting was made from professional societies, research centers, alumni associations, science bloggers, and word of mouth.
@@
 The darkness of the points is positively associated with the self-reported expertise.
-Given that the European bailout fund was ratified before 1 November 2011 and that the Nikkei 225 index finished trading at around 8,700 on 30 September 2011, the general trend of the probability forecasts converge to the correct answers.
+Given that the European bailout fund was ratified before 1 November 2011 and that the Nikkei 225 index finished trading at around 8,700 on 30 September 2011, the general trends of the probability forecasts converge to the correct answers.
 There is, however, much disagreement among the experts.
@@
 \label{model}
-Let $p_{i,t,k} \in (0,1)$ be the probability forecast given by the $i$th expert at time $t$ for the $k$th question, where $i = i, \dots, I_k$, $t = 1, \dots, T_k$, and $k = 1, \dots, K$.
+Let $p_{i,t,k} \in (0,1)$ be the probability forecast given by the $i$th expert at time $t$ for the $k$th question, where $i = 1, \dots, I_k$, $t = 1, \dots, T_k$, and $k = 1, \dots, K$.
 Denote the logit-probabilities with
@@
 and $(\mu_0, \sigma_0^2) \in (\R, \R^+)$ are hyper-parameters chosen \textit{a priori}.
-The hidden state $X_{t,k}$ represents the sharp and calibrated logit-probability for the $k$th event given the information available up to and including time $t$.
+The hidden state $X_{t,k}$ represents the sharp and well-calibrated logit-probability for the $k$th event given the information available up to and including time $t$.
 To make this more specific, let $Z_k \in \{0, 1\}$ indicate whether the event associated with the $k$th question happened $\left(Z_k = 1\right)$ or did not happen $\left(Z_k = 0\right)$.
@@
 These probabilities are made estimable by assuming that the observed process is associated with the filtration $\{\mathcal{F}_{t,k}\}_{t=1}^{T_k}$ and hence can be used as a proxy for the amount of information available at any given time point.
-Therefore it is not unreasonable to interpreted the hidden process as the crowd belief across time.
+%Therefore it is not unreasonable to interpreted the hidden process as the crowd belief across time.
 
@@
 For instance, $\hat{\boldsymbol{b}}(1)$ and $\hat{X}_{t, k}(1)$ represent the estimates of $\boldsymbol{b}(1)$ and $X_{t, k}(1)$, respectively.
-These estimates are found via a Gibbs sampler that only makes use of standard distributions.
+These estimates are found via Gibbs sampling that only makes use of standard distributions.
 See Appendix \ref{appendix} for the technical details of our sampler, and, e.g., \citet{gelman2003bayesian} for a discussion on the general principles of Gibbs sampling.
@@
 The synthetic data is not generated directly from the model for two reasons: (i) showing good performance on a dataset directly generated from the model assumptions is hardly any news, and (ii) generating data from the dynamic model description does not produce well-calibrated hidden states.
-The latter is important for our interpretation of the hidden states as a sharp and well-calibrated versions of the crowd belief.
-To generate a well-calibrated hidden process, generate a path of the standard Brownian motion until time $T_k$ for question $k$, and let $Z_{t,k}$ denote the value of the path at time $t$.
-Then, for $t = 1, \dots, T_k$,
+The latter is important for our interpretation of the hidden process as a sharp and well-calibrated version of the crowd belief.
+To ensure that the hidden process for question $k$ is well-calibrated, generate a path of the standard Brownian motion until time $T_k$.
+If $Z_{t,k}$ denotes the value of the path at time $t$, then
 \begin{eqnarray*}
@@
 
-We used this procedure to generate the hidden processes for $K$ questions with $T_k = 100$.
-Each question involved 50 experts allocated evenly among the five expertise groups.
-Their reported logit-probabilities, $Y_{t,k}$, were generated from Equation (\ref{observedpr}) by applying bias and noise to the hidden process.
+We use this procedure to generate the hidden processes for $K$ questions with $T_k = 100$ for $k = 1, \dots, K$.
+Each question involves 50 experts allocated evenly among the five expertise groups.
+The expert logit-probabilities $Y_{t,k}$ are generated from Equation (\ref{observedpr}) by applying bias and noise to the hidden process $X_{t,k}$.
 Similarly to the real-world data setting discussed in Section \ref{realData}, extreme logit-probabilities are truncated to avoid overly influential observations; that is, any $p_{i,t,k} = \logit^{-1}(Y_{i,t,k})$ less than 0.01 (or greater than 0.99) is truncated to 0.01 (or to 0.99).
-The simulation was then performed over a three-dimensional grid of parameter values:
+
+The simulation is then performed over a three-dimensional grid of parameter values:
 \begin{eqnarray*}
@@
 where $\beta$ generates the bias vector by  $\boldsymbol{b} = [0.50, 0.75, \dots, 1.50]^T \beta$. 
-The simulation runs a total of 40 times at every grid point. Each run records the accuracy to which the approaches are able to estimate the hidden process and bias vector.
-
-%The number of questions, $K$, varies between $20$ and $80$ with increments of 15 problems, the bias vector changes as $$\boldsymbol{b} = [0.50, 0.75, \dots, 1.50]^T \beta,$$ where $\beta$ takes values in $1/2, 3/4, 1, 4/3$, and $2/1$ symmetrically around 1 in the log-scale, and finally the values of $\sigma^2$ vary from $0.5$ to $2.0$ with increments of $0.5$.
+The simulation runs a total of 40 times at every grid point.
+Each run records the accuracy to which the approaches are able to estimate the hidden process and bias vector.
+The first is measured in terms of the average quadratic loss in the probability space, $\sum_{k=1}^K \sum_{t=1}^{T_k} (\hat{p}_{t,k} - \logit^{-1}(X_{t,k}))^2 / \sum_{k=1}^K T_k$, while the latter is measured in terms of the average quadratic loss, $|| \boldsymbol{b} - \hat{\boldsymbol{b}}||^2/5$.
+
+\begin{figure}[h!]
+\vspace{-5em}
+\centering
+ \hspace{-28em}	\includegraphics{Figures/LegendSynthetic} % requires the graphicx package
+\vspace{-0.5em}
+
+\subfigure{
+\hspace{-1em} \includegraphics{Figures/SyntheticSigma2}
+\label{SyntheticSigma2}
+}
+\subfigure{
+\hspace{-1.5em} \includegraphics{Figures/SyntheticBeta}
+\label{SyntheticBeta}
+}
+\subfigure{
+\hspace{-1.5em} \includegraphics{Figures/SyntheticK}
+\label{SyntheticK}
+}
+
+\vspace{-1em}
+
+\subfigure{
+\hspace{-1em}  \includegraphics{Figures/SyntheticSigma2bias}
+\label{SyntheticSigma2b}
+}
+\subfigure{
+\hspace{-1.5em}  \includegraphics{Figures/SyntheticBetabias}
+\label{SyntheticBetab}
+}
+\subfigure{
+\hspace{-1.5em}   \includegraphics{Figures/SyntheticKbias}
+\label{SyntheticKb}
+}
+
+\caption[Optional caption for list of figures]{Comparing \textit{Sample-Then-Calibrate} that optimizes over the logarithmic score (STC-Log) and \textit{Sample-Then-Calibrate} that optimizes over the Brier score (STC-Bri) under synthetic data.
+The top row presents the accuracy to capture the hidden process.
+The bottom row presents the accuracy to capture the bias vector.
+}
+\label{Synthetic}
+\end{figure}
 
@@
 The first row shows the marginal effects of the three grid variables on the accuracy to which the \textit{Sample-Then-Calibrate} under the logarithmic loss (STC-Log) and the Brier score (STC-Bri) are able to estimate the hidden process.
-The accuracy is measured in terms of the average quadratic loss in the probability space, $\sum_{k=1}^K \sum_{t=1}^{T_k} (\hat{p}_{t,k} - \logit^{-1}(X_{t,k}))^2 / \sum_{k=1}^K T_k$.
 The second row is similar in nature but measures the accuracy of estimating the bias vector instead.
-The accuracy is measured in terms of the average quadratic loss, $|| \boldsymbol{b} - \hat{\boldsymbol{b}}||^2/5$.
 The plots were computed by choosing one grid variable (e.g.
@@
 Even though some relative differences are visible, the loss values are very small in absolute terms; hence suggesting good overall accuracy to capture the hidden process.
-While increasing the number of questions improves the accuracy to a small extent, estimating the hidden process becomes more difficult as either $\sigma^2$ or $\beta$ gets larger.
-The latter seems more unintuitive and can be understood better by observing that larger values of $\beta$ lead to more extreme expert logit-probabilities.
-As a result, more logit-probabilities are modified by truncation, leading to poorer estimation accuracy.
+Estimating the hidden process, however, becomes more difficult as either $\sigma^2$ or $\beta$ gets larger.
+The latter seems more unintuitive and can be understood better by observing that larger values of $\beta$ lead to more extreme expert logit-probabilities and hence to more frequent truncation.
+This causes the estimation accuracy to degrade because  a high level of truncation results to a drastic deviation from the model assumptions.
 
 Based on the second row of plots, STC-Log and STC-Bri  capture the bias vector rather precisely as long as the true values of $\beta$ and $\sigma^2$ are not very large.
-This behavior can be considered as being caused  to a large extent by the truncation of the extreme expert logit-probabilities.
+This behavior can be again considered as being caused  to a large extent by the truncation of the extreme expert logit-probabilities.
 Surprisingly, even though STC-Log is more accurate in estimating the bias vector, it is unable to outperform STC-Bri in estimating the hidden process.
-
-\begin{figure}[h!]
-\vspace{-5em}
-\centering
- \hspace{-30em}	\includegraphics{Figures/LegendSynthetic} % requires the graphicx package
-\vspace{-0.5em}
-
-\subfigure{
-\hspace{-1em} \includegraphics{Figures/SyntheticSigma2}
-\label{SyntheticSigma2}
-}
-\subfigure{
-\hspace{-1.5em} \includegraphics{Figures/SyntheticBeta}
-\label{SyntheticBeta}
-}
-\subfigure{
-\hspace{-1.5em} \includegraphics{Figures/SyntheticK}
-\label{SyntheticK}
-}
-
-\subfigure{
-\hspace{-1em}  \includegraphics{Figures/SyntheticSigma2bias}
-\label{SyntheticSigma2b}
-}
-\subfigure{
-\hspace{-1.5em}  \includegraphics{Figures/SyntheticBetabias}
-\label{SyntheticBetab}
-}
-\subfigure{
-\hspace{-1.5em}   \includegraphics{Figures/SyntheticKbias}
-\label{SyntheticKb}
-}
-
-\caption[Optional caption for list of figures]{Comparing \textit{Sample-Then-Calibrate} that optimizes over the logarithmic score (STC-Log) and \textit{Sample-Then-Calibrate} that optimizes over the Brier score (STC-Bri) under synthetic data.
-The top row presents the accuracy to capture the hidden process.
-The bottom row presents the accuracy to capture the bias vector.
-}
-\label{Synthetic}
-\end{figure}
 

### 2013-08-31 paper.tex
Edits

@@
 \title{Probability Aggregation: Dynamic Hierarchical Modeling of Sparse Expert Beliefs}
-\runtitle{Hierarchical Model for Expert Beliefs}
+\runtitle{Dynamic Hierarchical Modeling of Sparse Expert Beliefs}
 
@@
 Each question involves 50 experts allocated evenly among the five expertise groups.
-The expert logit-probabilities $Y_{t,k}$ are generated from Equation (\ref{observedpr}) by applying bias and noise to the hidden process $X_{t,k}$.
-Similarly to the real-world data setting discussed in Section \ref{realData}, extreme logit-probabilities are truncated to avoid overly influential observations; that is, any $p_{i,t,k} = \logit^{-1}(Y_{i,t,k})$ less than 0.01 (or greater than 0.99) is truncated to 0.01 (or to 0.99).
-
-The simulation is then performed over a three-dimensional grid of parameter values:
+The expert logit-probabilities $Y_{i,t,k}$ are generated from Equation (\ref{observedpr}) by applying bias and noise to the hidden process $X_{t,k}$.
+Any extreme logit-probabilities are truncated to avoid overly influential observations; that is, any $p_{i,t,k} = \logit^{-1}(Y_{i,t,k})$ less than 0.01 (or greater than 0.99) is truncated to 0.01 (or to 0.99).
+This is akin to the real-world data setting, where extreme probabilities must be truncated (see Section \ref{realData}).
+Therefore, to understand the effect that truncation has on the real-world data results, it should be first analyzed and understood under the controlled environment of synthetic data.
+
+Throughout the evaluation data is generated under different choices of parameter values. More specifically, the evaluation procedure iterates over a three-dimensional grid of parameter values
 \begin{eqnarray*}
@@
 \end{eqnarray*}
-where $\beta$ generates the bias vector by  $\boldsymbol{b} = [0.50, 0.75, \dots, 1.50]^T \beta$. 
-The simulation runs a total of 40 times at every grid point.
-Each run records the accuracy to which the approaches are able to estimate the hidden process and bias vector.
-The first is measured in terms of the average quadratic loss in the probability space, $\sum_{k=1}^K \sum_{t=1}^{T_k} (\hat{p}_{t,k} - \logit^{-1}(X_{t,k}))^2 / \sum_{k=1}^K T_k$, while the latter is measured in terms of the average quadratic loss, $|| \boldsymbol{b} - \hat{\boldsymbol{b}}||^2/5$.
+where $\beta$ generates the bias vector  $\boldsymbol{b} = [0.50, 0.75, \dots, 1.50]^T \beta$.
+These values were considered to be realistic based on our experience with the real-world probability forecasting data.
+Each triple of parameter values,  i.e.
+a grid point, is used  40 times to generate a synthetic dataset.
+This dataset is then used to assess how accurately the approaches are able to estimate the hidden process and bias vector.
+The accuracy to estimate the hidden process is measured with the average quadratic loss in the probability space, $\sum_{k=1}^K \sum_{t=1}^{T_k} ( \logit^{-1}(\hat{X}_{t,k}) - \logit^{-1}(X_{t,k}))^2 / \sum_{k=1}^K T_k$, and the accuracy to capture the bias vector is quantified by the average quadratic loss, $|| \boldsymbol{b} - \hat{\boldsymbol{b}}||^2/5$.
 
@@
 This causes the estimation accuracy to degrade because  a high level of truncation results to a drastic deviation from the model assumptions.
+Increasing the number of questions improves the estimation accuracy but only very little.
 
 Based on the second row of plots, STC-Log and STC-Bri  capture the bias vector rather precisely as long as the true values of $\beta$ and $\sigma^2$ are not very large.
-This behavior can be again considered as being caused  to a large extent by the truncation of the extreme expert logit-probabilities.
+This behavior can be again considered as being caused by the truncation of the extreme expert logit-probabilities.
 Surprisingly, even though STC-Log is more accurate in estimating the bias vector, it is unable to outperform STC-Bri in estimating the hidden process.
@@
 Firstly, the model was shown to outperform other probability aggregators in terms of forecasting ability.
-In particular, the model was deemed a well-balanced compromise that avoids over-fitting without begin overly conservative.
+In particular, the model was deemed a well-balanced compromise that avoids over-fitting without being overly conservative.
 Secondly, the crowd belief was used as the no-bias reference point to study the bias among groups of experts with different levels of self-reported expertise.

### 2013-08-31 paper.tex
Edits

@@
 The updates are assumed to occur very infrequently, resulting into a highly sparse dataset that cannot be modeled by standard time-series procedures.
-In response to the lack of appropriate methodology, this paper presents a dynamic model that takes into account the forecaster's level of self-reported expertise and produces aggregate probabilities that are sharp and well-calibrated both in- and out-of-sample.
-The model is demonstrated on a real-world dataset that includes over 2,300 experts making multiple probability forecasts on different subsets of 166 different international political events.
+In response to the lack of appropriate methodology, this paper presents a hierarchical model that takes into account the expert's level of self-reported expertise and produces aggregate probabilities that are sharp and well-calibrated both in- and out-of-sample.
+The model is demonstrated on a real-world dataset that includes over 2,300 experts making multiple probability forecasts on different subsets of 166 international political events.
 \end{abstract}
@@
 Individual experts can differ radically from one another in their abilities to assess probabilities of future events.
-The quality of a probability assessment is typically evaluated in regards to \textit{calibration}, which is an attribute measuring the degree to which the frequency of occurrence of the events agree with the given probabilities.
-This means that, for instance, the proportion of occurred events is 60\% for all those events for which a well-calibrated expert assessed a probability of 0.60.
-Even though several experiments have been conducted to show that experts are generally poorly calibrated (see, e.g., \citet{cooke1991experts, shlyakhter1994quantifying}), relative differences can occur between different types of experts.
-In particular, \citet{wright1994coherence} argue that a higher level of self-reported expertise on the subject-matter can be associated with better calibration.
+Their probability assessments are often evaluated and compared in regards to \textit{calibration} that measures how closely the frequency of occurrence of the events agree with the given probabilities.
+For instance, the proportion of occurred events is 60\% for all those events for which a well-calibrated expert assessed a probability of 0.60.
+Even though several experiments have been conducted to show that experts are generally poorly calibrated (see, e.g., \citet{cooke1991experts, shlyakhter1994quantifying}), relative differences can occur among different types of experts.
+In particular, \citet{wright1994coherence} argue that a higher level of self-reported expertise can be associated with better calibration.
 
 Calibration by itself, however, is not sufficient for a useful probability assessment.
-To see this, consider a relatively stationary process, such rain on different days in some geographic region, whose empirical frequency of occurrence in the last 10 years is 45\%.
+To see this, consider a relatively stationary process, such as rain on different days in some geographic region, whose empirical frequency of occurrence in the last 10 years is 45\%.
 In this setting an expert can assess a constant probability of 0.45 and be well-calibrated.
@@
 For this reason the long-term frequency is often considered as the baseline probability -- a naive assessment that provides the decision-maker very little extra information.
-A consulting expert should therefore aim to make probability assessments that are as far from the baseline as possible.
+Therefore a consulting expert should aim to make probability assessments that are as far from the baseline as possible.
 The extent to which his probabilities differ from the baseline is measured with an important attribute called \textit{sharpness} (\citet{gneiting2008rejoinder, winkler2008comments}).
 If the expert is both sharp and well-calibrated, he is able to forecast the behavior of the process with high certainty and accuracy.
-From the decision-maker's perspective, his advice is highly useful as it involves much information and very little risk.
-Therefore useful probability estimation should aim to maximize sharpness subject to calibration (see, e.g., \citet{raftery2005using, murphy1987general}).
+From the decision-maker's perspective his advice is highly useful as it involves much information and very little risk.
+Therefore, to summarize, useful probability estimation should aim to maximize sharpness subject to calibration (see, e.g., \citet{raftery2005using, murphy1987general}).
 This is a well-defined goal that has led to a wide range of novel and insightful observations in probability forecasting.
@@
 After summarizing the dataset, the paper introduces a dynamic hierarchical model for capturing the crowd belief.
-The model is estimated in a two-step procedure.
-First, a sampling step produces constrained parameter estimates via Gibbs sampling (see \citet{geman1984stochastic} for the original introduction of Gibbs sampling).
-Second, a calibration step transforms these estimates to their unconstrained equivalents via a one-dimension optimization procedure.
+The model is estimated in a two-step procedure: first, a sampling step produces constrained parameter estimates via Gibbs sampling (see \citet{geman1984stochastic} for the original introduction of Gibbs sampling); second, a calibration step transforms these estimates to their unconstrained equivalents via a one-dimension optimization procedure.
 An extension of this model to polychotomous outcomes is briefly discussed before model evaluation.
 The first evaluation section uses synthetic data to study the accuracy to which the two-step procedure is able to estimate parameter values.
-The second evaluation section discusses different aspects of the model by applying it to our real-world probability forecasting data.
+The second evaluation section illustrates different aspects of the model by applying it to our real-world probability forecasting data.
 The paper concludes with a discussion on future research directions and model limitations.
@@
 
-The data collection began with a recruitment of 2,365 forecasters ranging from graduate students to political science faculty and practitioners.
+The data collection began with a recruitment of 2,365 experts ranging from graduate students to political science faculty and practitioners.
 The recruiting was made from professional societies, research centers, alumni associations, science bloggers, and word of mouth.
@@
 Each question was active for a period of time during which the participating experts were allowed to update their forecasts as frequently as they liked without being penalized.
-The forecasters knew that their probability estimates would be assessed for accuracy using Brier scores\footnote{The Brier score is the squared distance between the probability forecast and the event indicator that equals 1.0 or 0.0 depending on whether the event happened or not, respectively.
+The experts knew that their probability estimates would be assessed for accuracy using Brier scores\footnote{The Brier score is the squared distance between the probability forecast and the event indicator that equals 1.0 or 0.0 depending on whether the event happened or not, respectively.
 See \citet{Brier} for the original introduction.}.
 This incentivized them to report their true beliefs instead of attempting to game the system (\citet{winkler1968good}).
-In addition to receiving \$150 for meeting minimum participation requirements that did not depend on prediction accuracy, the forecasters also received status rewards for their performance via leader-boards displaying Brier scores for the top 20 forecasters.
-Since a typical expert participated only in a small subset of the 166 questions, the forecasters are considered indistinguishable conditional on the level of self-reported expertise.
+In addition to receiving \$150 for meeting minimum participation requirements that did not depend on prediction accuracy, the experts received status rewards for their performance via leader-boards displaying Brier scores for the top 20 experts.
+Since a typical expert participated only in a small subset of the 166 questions, the experts are considered indistinguishable conditional on the level of self-reported expertise.
+
+Overall, the data is quite sparse.
+As mentioned before, the experts were allowed to update their forecasts as long as the question remained open.
+Updating, however, was done on a very infrequent basis.
+An expert gave, on average, only 0.016 forecasts per day.
+Therefore a typical expert made a forecast about every 62 days, resulting to an average response rate of around 13.5 forecasts per day from a very large and diverse group of experts.
 Tables \ref{DataStats} and \ref{ExpertiseTable} provide more relevant summary statistics on the data.
@@
 \caption[Optional caption for list of figures]{Scatterplots of the probability forecasts given for two questions in our dataset.
-The shadings represents the self-reported expertise of the forecaster who provided the probability forecast.}
+The shadings represents the self-reported expertise of the expert who provided the probability forecast.}
 \label{ExamplePlots}
@@
 The darkness of the points is positively associated with the self-reported expertise.
-Given that the European bailout fund was ratified before 1 November 2011 and that the Nikkei 225 index finished trading at around 8,700 on 30 September 2011, the general trends of the probability forecasts converge to the correct answers.
-There is, however, much disagreement among the experts.
-This disagreement persists even near the closing dates of the questions.
-Unfortunately, the data is not only noisy but also quite sparse.
-As mentioned before, the forecasters were allowed to update their forecast as long as the questions remained open.
-Updating, however, was done on a very infrequent basis.
-An expert gave, on average, only 0.016 forecasts per day, and so a typical expert made a forecast about every 62 days.
-As a result the average response rate was around 13.5 forecasts per day from a large group of experts.
+Given that the European bailout fund was ratified before November 1, 2011 and that the Nikkei 225 index finished trading at around 8,700 on September 30, 2011, the general trend of the probability forecasts tends to converge towards the correct answers.
+The individual experts, however, disagree rather strongly, with the disagreement persisting even near the closing dates of the questions.
 
@@
 Partition the experts into $J$ groups based on some individual feature, such as self-reported expertise, with each group sharing a common multiplicative bias term $b_{j} \in \R$ for $j = 1, \dots, J$.
-Collect these bias terms into a bias vector $\boldsymbol{b} = [b_{1}\ b_{2}\ \dots\ b_{J}]^T$ and let $\boldsymbol{M}_k$ be a $I_k \times J$ matrix denoting  the group-memberships of the forecasters in question $k$.
-That is, if the $i$th expert participating in the $k$th question belongs to the $j$th group, then the $i$th row of $\boldsymbol{M}_k$ is the $j$th standard basis vector, $\boldsymbol{e}_j$.
+Collect these bias terms into a bias vector $\boldsymbol{b} = [b_{1}\ b_{2}\ \dots\ b_{J}]^T$.
+Let $\boldsymbol{M}_k$ be a $I_k \times J$ matrix denoting  the group-memberships of the experts in question $k$; that is, if the $i$th expert participating in the $k$th question belongs to the $j$th group, then the $i$th row of $\boldsymbol{M}_k$ is the $j$th standard basis vector $\boldsymbol{e}_j$.
 The bias vector $\boldsymbol{b}$ does not include a subindex because it is considered shared among all the $K$ questions.
@@
 \end{eqnarray}
-where Equation (\ref{observedpr}) denotes the observed process, Equation (\ref{hiddenpr}) shows the hidden process that is driven by the constant $\gamma_k \in \R$.
+where Equation (\ref{observedpr}) denotes the observed process and Equation (\ref{hiddenpr}) shows the hidden process that is driven by the constant $\gamma_k \in \R$.
 The error terms are independent and identically distributed normal random variables with mean zero
@@
 To give intuitive interpretations of the other model parameters, notice that the error term in the observed process has mean zero.
-Therefore the forecasters are assumed to be, on average, a multiplicative constant $\boldsymbol{b}$ away from the calibrated crowd belief.
+Therefore the experts are assumed to be, on average, a multiplicative constant $\boldsymbol{b}$ away from the calibrated crowd belief.
 An individual element of $\boldsymbol{b}$ can be interpreted as a group-specific \textit{systematic bias} that labels the group either as over-confident $\left( b_j \in (1, \infty)\right)$ or as under-confident $\left(b_j \in (0,1)\right)$.
@@
 
-The most challenging part of our model estimation is to incorporate the event indicators $\{Z_k\}_{k=1}^K$ such that the hidden process is well-calibrated and the interpretability of our model is not compromised.
-This section introduces a two-step procedure, called \textit{Sample-Then-Calibrate} or simply \textit{STC}, that achieves this goal in a flexible and efficient manner by first estimating  the model parameters under a constraint (\textit{Sampling Step}) and then performing a one-dimension optimization procedure to transform the constrained estimates into their unconstrained counterparts (\textit{Calibration Step}).
+The main challenge is to capture a well-calibrated estimate of the hidden process without sacrificing the interpretability of our model.
+This section introduces a two-step procedure, called \textit{Sample-Then-Calibrate} (STC), that achieves this goal in a flexible and efficient manner: the first step estimates the model parameters under a constraint (\textit{Sampling Step}), and the second step performs a one-dimension optimization procedure to transform the constrained estimates into their unconstrained counterparts (\textit{Calibration Step}).
 
@@
 \label{sampling_step}
-Since $\left(a \boldsymbol{b}, X_{t,k}/a, a^2 \tau_k^2\right) \neq \left(\boldsymbol{b}, X_{t,k}, \tau_k^2\right)$ for any $a > 0$ yield the same  likelihood for $\boldsymbol{Y}_{t,k}$, the model as described by Equations (\ref{observedpr}) and (\ref{hiddenpr}) is not identifiable and the parameter estimates tend to drift during the sampling process.
+Since $\left(a \boldsymbol{b}, X_{t,k}/a, a^2 \tau_k^2\right) \neq \left(\boldsymbol{b}, X_{t,k}, \tau_k^2\right)$ for any $a > 0$ yield the same  likelihood for $\boldsymbol{Y}_{t,k}$, the model as described by Equations (\ref{observedpr}) and (\ref{hiddenpr}) is not identifiable.
+As a result, the parameter estimates tend to drift during the sampling process.
 A well-known solution is to choose one of the elements of $\boldsymbol{b}$, say $b_3$, as the reference point and fix $b_3 = 1$.
@@
 These estimates are found via Gibbs sampling that only makes use of standard distributions.
-See Appendix \ref{appendix} for the technical details of our sampler, and, e.g., \citet{gelman2003bayesian} for a discussion on the general principles of Gibbs sampling.
+See Appendix \ref{appendix} for the technical details of the sampling step, and, e.g., \citet{gelman2003bayesian} for a discussion on the general principles of Gibbs sampling.
 
@@
 \label{calibration_step}
-Since the model parameters can be estimated under a constrained version of $b_3$, the next step is to determine how $b_3$ should be constrained such that both the sharpness and calibration of the estimated hidden process are maximized.
+Since the model parameters can be estimated by fixing $b_3$ to any constant, the next step is to search for the constant that gives an optimally sharp and calibrated estimate of the hidden process.
 This section introduces an efficient procedure that finds the optimal constraint without requiring any additional runs of the sampling step.
@@
 This means that $X_{t,k} = X_{t,k}(1)/\beta$ when $\beta$ is equal to the true value of $b_3$.
-Since the hidden process $X_{t, k}$ is assumed to be sharp and well-calibrated, $b_3$ can be estimated by the value of $\beta$ that simultaneously maximizes the sharpness and calibration of $\hat{X}_{t, k}(1) / \beta$.
-A natural criterion for this maximization is provided by the class of \textit{proper scoring rules} that combine sharpness and calibration (\citet{gneiting2008rejoinder, buja2005loss}).
+Since the hidden process $X_{t, k}$ is assumed to be sharp and well-calibrated, $b_3$ can be estimated with the value of $\beta$ that simultaneously maximizes the sharpness and calibration of $\hat{X}_{t, k}(1) / \beta$.
+A natural criterion for this maximization is given by the class of \textit{proper scoring rules} that combine sharpness and calibration (\citet{gneiting2008rejoinder, buja2005loss}).
 Due to the possibility of \textit{complete separation} in any one question (see, e.g., \citet{gelman2008weakly}), the maximization must be performed over multiple questions.
-Therefore we estimate
+Therefore,
 \begin{eqnarray}
@@
 \end{eqnarray*}
-Since it is not clear which rule should be used for predicting geopolitical events, such as the ones in our dataset, the \textit{Sample-Then-Calibrate} procedure is evaluated  separately under both rules in Sections \ref{syntheticData} and \ref{realData}.
+Since it is not clear which rule should be used for predicting geopolitical events, the \textit{Sample-Then-Calibrate} procedure is evaluated  separately under both rules in Sections \ref{syntheticData} and \ref{realData}.
 Once $\hat{\beta}$ has been computed, estimates of the unconstrained model parameters are given by
@@
 \end{eqnarray}
-Notice that the parameters $\sigma^2_k$ and $\gamma_k$ are not affected by the constraint. Therefore their constrained and unconstrained estimates are the same.
+Notice that estimates of $\sigma^2_k$ and $\gamma_k$ are not affected by the constraint. Therefore their constrained and unconstrained versions are the same.
 
 \subsection{Discussion}
-If the class labels in the data are balanced with respect to the time points, the calibration step with the logarithmic scoring rule is approximately equivalent to \textit{Platt calibration}, which has been shown to yield good calibration under various modeling scenarios (see, e.g., \citet{platt1999probabilistic, niculescu2005obtaining}).
+If the class labels in the data are balanced with respect to the time points, the calibration step under the logarithmic scoring rule is approximately equivalent to \textit{Platt calibration}, which has been shown to yield good calibration under various modeling scenarios (see, e.g., \citet{platt1999probabilistic, niculescu2005obtaining}).
 To see this, recall that the Platt calibrated logit-probabilities are given by $\hat{A} + \hat{B}\hat{X}_{t,k}(1)$, where
@@
 %model with $Z_k$ as the response and $\hat{X}_{t,k}(1)$ as the explanatory variable. 
-To understand the behavior of the coefficients $A$ and $B$, express the logistic regression as linear regression
+To understand the behavior of the coefficients $A$ and $B$, express the logistic regression as a linear regression model
 \begin{eqnarray*}
@@
 with $e_{t,k} \stackrel{i.i.d.}{\sim} \mathcal{N}(0,\sigma^2)$.
-If the data are balanced with respect to the time points, then exactly half of the summands in Equation (\ref{platt}) have $Z_k = 1$ and the average response logit-probability $\logit(\P(Z_k = 1 | \hat{X}_{t,k})) $ is close to zero.
-Since the values of $\hat{X}_{t,k}$ are estimated logit-probabilities of the same $K$ events across different time points, their overall average is also around zero.
+If the data are balanced with respect to the time points, then exactly half of the summands in Equation (\ref{platt}) have $Z_k = 1$ and the average response logit-probability $\logit(\P(Z_k = 1 | \hat{X}_{t,k})) $ is expected to be close to zero.
+Since the values of $\hat{X}_{t,k}$ are estimated logit-probabilities of the same $K$ events across different time points, their overall average is also expected to be around zero.
 Therefore both the response and explanatory variables are approximately centered.
 This means that the intercept term $A$ is near zero reducing Platt calibration to Equation (\ref{OSE}) under the logarithmic scoring rule.
-If the data is not balanced, Platt calibration can be easily incorporated into our model via an additional intercept parameter.
+If the data are not balanced, Platt calibration can be easily incorporated into our model via an additional intercept parameter.
 This, however, reduces the interpretability of our model.
@@
 This is akin to the real-world data setting, where extreme probabilities must be truncated (see Section \ref{realData}).
-Therefore, to understand the effect that truncation has on the real-world data results, it should be first analyzed and understood under the controlled environment of synthetic data.
-
-Throughout the evaluation data is generated under different choices of parameter values. More specifically, the evaluation procedure iterates over a three-dimensional grid of parameter values
+Therefore, to understand the effect that truncation has on the real-world data results, truncation should be first analyzed and understood under the controlled environment of synthetic data.
+
+Throughout the evaluation, data is generated under different choices of parameter values. More specifically, the evaluation procedure iterates over a three-dimensional grid of parameter values
 \begin{eqnarray*}
@@
 These values were considered to be realistic based on our experience with the real-world probability forecasting data.
-Each triple of parameter values,  i.e.
+Each triple of the parameter values,  i.e.
 a grid point, is used  40 times to generate a synthetic dataset.
-This dataset is then used to assess how accurately the approaches are able to estimate the hidden process and bias vector.
-The accuracy to estimate the hidden process is measured with the average quadratic loss in the probability space, $\sum_{k=1}^K \sum_{t=1}^{T_k} ( \logit^{-1}(\hat{X}_{t,k}) - \logit^{-1}(X_{t,k}))^2 / \sum_{k=1}^K T_k$, and the accuracy to capture the bias vector is quantified by the average quadratic loss, $|| \boldsymbol{b} - \hat{\boldsymbol{b}}||^2/5$.
+Each dataset is then used to assess how accurately \textit{Sample-Then-Calibrate} is able to estimate the hidden process and bias vector.
+The accuracy of the estimated hidden process is measured with the average quadratic loss in the probability space, $\sum_{k=1}^K \sum_{t=1}^{T_k} ( \logit^{-1}(\hat{X}_{t,k}) - \logit^{-1}(X_{t,k}))^2 / \sum_{k=1}^K T_k$, and the accuracy of the estimtated bias vector is quantified by the average quadratic loss, $|| \boldsymbol{b} - \hat{\boldsymbol{b}}||^2/5$.
 
@@
 Based on the first row of plots, STC-Log and STC-Bri estimate the hidden process to a similar accuracy.
-Even though some relative differences are visible, the loss values are very small in absolute terms; hence suggesting good overall accuracy to capture the hidden process.
+Even though some relative differences are visible, the loss values are very small in absolute terms, hence suggesting good overall accuracy to capture the hidden process.
 Estimating the hidden process, however, becomes more difficult as either $\sigma^2$ or $\beta$ gets larger.
@@
 Transforming such forecasts into the logit-space yields infinities that are highly influential and cause problems in model estimation.
-To avoid this, \citet{Ariely00theeffects} suggest changing $p = 0.0$ and $1.0$ to $p = 0.02$ and $0.98$, respectively.
+To avoid this, \citet{Ariely00theeffects} suggest changing $p = 0.00$ and $1.00$ to $p = 0.02$ and $0.98$, respectively.
 This is similar to \textit{winsorising} that sets the extreme probabilities to a specified percentile of the data (see, e.g., \citet{hastings1947low} for more details on winsorising).
 \citet{Geo}, on the other hand, consider only probabilities that fall within a constrained interval, say $[0.001, 0.999]$, and discard the rest.
-Since this implies ignoring a portion of the data, we have decided to adopt the first approach by changing $p = 0.0$ and $1.0$ to $p = 0.01$ and $0.99$, respectively.
+Since this implies ignoring a portion of the data, we have decided to adopt the first approach by changing $p = 0.00$ and $1.00$ to $p = 0.01$ and $0.99$, respectively.
 
@@
 Fortunately, it is often possible to use a well-balanced version of the data.
-The first step is to find a partition, $S_0$ and $S_1$, of the question indices, $\{1, 2, \dots, K\}$, such that the equality $\sum_{k \in S_0} T_k  = \sum_{k \in S_1} T_k$ is as closely approximated as possible.
+The first step is to find a partition $S_0$ and $S_1$ of the question indices $\{1, 2, \dots, K\}$ such that the equality $\sum_{k \in S_0} T_k  = \sum_{k \in S_1} T_k$ is as closely approximated as possible.
 This is equivalent to an NP-hard problem known in computer science as the \textit{Partition Problem}:  determine whether a given set of positive integers can be partitioned into two sets such that the sums of the two sets equal to each other (see, e.g., \citet{karmarkar1982differencing, hayes2002easiest}).
-A simple solution is to use a greedy algorithm, which iterates through the values of $T_k$ in descending order, assigning each $T_k$ to the subset that currently has the smaller sum (see, e.g.
+A simple solution is to use a greedy algorithm that iterates through the values of $T_k$ in descending order, assigning each $T_k$ to the subset that currently has the smaller sum (see, e.g.
 \citet{kellerer2004knapsack, gent1996phase} for more details on the Partition Problem).
-After finding a well-balanced partition, the next step is to assign the class labels such that the labels for the questions in $S_x$ are equal to $x$ with $x = 0$ or $1$.
-Recall from section \ref{calibration_step} that $Z_k$ represents the event indicator for the event associated with the $k$th question.
-To define a balanced set of indicators, $ \tilde{Z}_k$ for all $k \in S_x$, let
+After finding a well-balanced partition, the next step is to assign the class labels such that the labels for the questions in $S_x$ are equal to $x$ for $x = 0$ or $1$.
+Recall from section \ref{calibration_step} that $Z_k$ represents the event indicator associated with the $k$th question.
+To define a balanced set of indicators $ \tilde{Z}_k$ for all $k \in S_x$, let
 \begin{align*}
@@
 1-p_{i,t,k} & \text{if } Z_k = 1-x\\
-p_{i,t,k}& \text{if } Z_k = x 
+p_{i,t,k}& \text{if } Z_k = x,
 \end{cases}
 \end{align*}
-with $i = 1, \dots, I_k$, and $t = 1, \dots, T_k$.
+where $i = 1, \dots, I_k$, and $t = 1, \dots, T_k$.
 The resulting set $$\left\{\left(\tilde{Z}_k, \{\tilde{p}_{i,t,k} | i = 1, \dots, I_k, t = 1, \dots, T_k \}\right) \right\}_{k=1}^K$$ is a balanced version of the data.
@@
 In order to make predictions, the sampler is run for 500 iterations of which the first 200 are used for burn-in.
-The remaining 300 iterations is thinned by discarding every other observation, leaving a final predictive sample of 150 observations.
+The remaining 300 iterations are thinned by discarding every other observation, leaving a final predictive sample of 150 observations.
 
@@
 After thinning by only keeping every fifth observation, the calibration step is performed for the remaining 500 observations.
-The out-of-sample prediction is done by running the sampling step for 500 iterations with each consecutive iteration reading in and conditioning on the next value of $\beta$ and $\boldsymbol{b}$ that were found during the training period.
+The out-of-sample prediction is done by running the sampling step for 500 iterations with each consecutive iteration reading in and conditioning on the next value of $\beta$ and $\boldsymbol{b}$ found during the training period.
 The first 200 iterations are used for burn-in.
@@
 \item \textit{A fully Bayesian version of STC-Log (BSTC-Log)}.
-Denote the (constrained) calibrated logit-probabilities and  the event indicators across all $K$ questions with $\boldsymbol{X}(1)$ and $\boldsymbol{Z}$, respectively.
-The posterior distribution of $\beta$ conditional on $\boldsymbol{X}(1)$ is then given by $p(\beta | \boldsymbol{X}(1), \boldsymbol{Z}) \propto p( \boldsymbol{Z} | \beta, \boldsymbol{X}(1)) p(\beta | \boldsymbol{X}(1))$.
+Denote the calibrated logit-probabilities and  the event indicators across all $K$ questions with $\boldsymbol{X}(1)$ and $\boldsymbol{Z}$, respectively.
+The posterior distribution of $\beta$ conditional on $\boldsymbol{X}(1)$ is given by $p(\beta | \boldsymbol{X}(1), \boldsymbol{Z}) \propto p( \boldsymbol{Z} | \beta, \boldsymbol{X}(1)) p(\beta | \boldsymbol{X}(1))$.
 Recall that the calibration step under $S_{LOG}$ is equivalent to fitting a logistic regression model with $Z_k$ as the response and $\hat{X}_{k,t}(1)$ as the explanatory variable.
@@
 Therefore it is equivalent to the bias vector found under \textit{Sample-Then-Calibrate}.
-The term $\boldsymbol{e}_{i,k}$ is a vector of length 5 indicating which level of self-reported expertise the $i$th forecaster in the $k$th question belongs to.
+The term $\boldsymbol{e}_{i,k}$ is a vector of length 5 indicating which level of self-reported expertise the $i$th expert in the $k$th question belongs to.
 For instance, if $\boldsymbol{e}_{i,k} = [0, 1, 0, 0, 0]$, then the expert identifies himself with the expertise level two.
@@
 The evaluation then iterates 10 times, each time using one of the 10 sets as the testing set and the remaining 9 sets as the training set.
-Each question is therefore used nine times for training and exactly once for testing.
+Therefore each question is used nine times for training and exactly once for testing.
 The testing proceeds sequentially one testing question at a time as follows: First, for a question with a time horizon of $T_k$, make a prediction based on the first two days.
 Compute the Brier score for the aggregate forecast of the second day.
-Next, make a prediction based on the first three days and  compute the Brier score for the most recent day, namely, the third day.
+Next make a prediction based on the first three days and  compute the Brier score for the most recent day, namely, the third day.
 Repeat this process until the prediction is made on all of the $T_k-1$ days.
@@
 The second option, denoted by \textit{Scores by Problem}, gives each question an equal weight regardless how long the question remained open.
-This is done by first averaging the scores within a question, and then averaging the average scores across all the questions.
+This is done by first averaging the scores within a question and then averaging the average scores across all the questions.
 Both scores can be further broken down into subcategories by considering the length of the questions.
@@
 The bolded scores indicate the lowest score in each column.
-The values in the parenthesis quantify the variability in the scores.
-Under \textit{Scores by Day} the values give the standard errors of all the scores.
+The values in the parenthesis quantify the variability in the scores: Under \textit{Scores by Day} the values give the standard errors of all the scores.
 Under \textit{Scores by Problem}, on other hand, the values represent the standard errors of the average scores of the different questions.
@@
 
-Overall, STC-Bri and STC-Log achieve the lowest average scores across all columns except \textit{Short} where they are slightly outperformed by BSTC-Log.
+Overall, STC-Bri and STC-Log achieve the lowest average scores across all columns except \textit{Short}, where they are slightly outperformed by BSTC-Log.
 BSTC-Log, however, turns out to be overconfident (see Section \ref{calibration}).
@@
 Instead, it merely smooths the forecasts given by the experts.
-The resulting aggregate forecasts are therefore necessarily conservative and result into high average scores with low variability.
+The resulting aggregate forecasts are therefore necessarily conservative, resulting into high average scores with low variability.
 
@@
 Overall, these two aggregators perform very similarly.
-They both suffer from over-confidence and are hence unable to outperform the \textit{Sample-Then-Calibrate}-procedure.
+Unfortunately, they both suffer from over-confidence and are hence unable to outperform the \textit{Sample-Then-Calibrate}-procedure.
 
@@
 The idea is to plot the probability forecasts against the observed empirical frequencies.
-Any deviation from the diagonal line therefore suggests poor calibration.
+Therefore any deviation from the diagonal line suggests poor calibration.
 A model is considered under-confident (or over-confidence) if the points follow an S-shaped (or \reflectbox{S}-shaped) trend.
 In order to assess sharpness of the model, it is common practice to place a histogram of the given forecasts in the corner of the plot.
-Since the data were balanced, any deviation from the the baseline probability at 0.5 suggests improved sharpness.
+Since the data were balanced, any deviation from the the baseline probability of 0.5 suggests improved sharpness.
 
@@
 \label{ExpertBias}
-Recall from Section \ref{data} that the experts were asked to self-assess their level of expertise (on a 1-to-5 scale with 1 = Not At All Expert to 5 = Extremely Expert) on any questions that they participated in.
-The self-reported expertise then divides the experts into 5  groups, with each group assigned a separate multiplicative bias term.
-This section uses the \textit{Sample-Then-Calibrate}-procedure to explore the posterior distribution of these multiplicative bias terms.
-Figure \ref{Biases} presents the posterior distributions of the bias terms with side-by-side box plots.
-Since the distributions fall completely below the \textit{no-bias} reference-line at 1.0, all the groups are deemed under-confident.
-
-The under-confidence, however, decreases as the level of  expertise increases.
-For instance, the posterior probability that the most expert group is the least under-confident is approximately equal to $1.0$, and the posterior probability of a strictly decreasing level of under-confidence is approximately 0.87.
-The latter probability is driven down by the inseparability of the two groups associated with the lowest levels of self-reported expertise.
-The fact that these groups are very similar suggests that the forecasters are poor at assessing how little they know about a subject that is strange to them.
-If these groups are combined into a single group, the posterior probability of a strictly decreasing level of under-confidence is approximately 1.0.
-
-The decreasing trend in under-confidence can be reasoned by viewing the process of making a subjective probability  as Bayesian updating: At first a completely ignorant expert aiming to minimize a reasonable loss function, such as the Brier score, has no reason to give anything but 0.5 as his probability forecast.
-However, as soon as the expert gains some knowledge about the event, he produces an updated forecast that is a compromise between his initial forecast and the new information acquired.
-The updated forecast is therefore conservative and too close to 0.5 as long as the expert remains only partially informed about the event.
-If most experts fall somewhere on this spectrum between ignorance and full information, their average forecast tends to fall strictly between 0.5 and the most-informed probability forecast (see \citet{Baron} for more details).
-Since expertise is to a large extent determined by subject-matter knowledge, the level of under-confidence can be expected to decrease as a function of the group's level of self-reported expertise.
-
-Finding under-confidence in all the groups is a rather surprising result given that many previous studies have been conducted to show that experts are likely to be over-confident.
-For instance,  \citet{lichtenstein1977calibration, morgan1992uncertainty, bier2004implications} summarize the results from numerous calibration studies, and conclude that experts are systematically over-conﬁdent about their probability assessments.
-Our result, however, is a statement about groups of experts and hence does not invalidate the possibility of the individual experts being overconfident.
-To make conclusions at the individual-level based on the group-level bias terms would be considered an \textit{ecological inference fallacy} (see, e.g., \citet{lubinski1996seeing}).
 
@@
 
+Recall from Section \ref{data} that the experts were asked to self-assess their level of expertise (on a 1-to-5 scale with 1 = Not At All Expert to 5 = Extremely Expert) on any questions in which they participated.
+The self-reported expertise then divides the experts into 5  groups, with each group assigned a separate multiplicative bias term.
+This section uses the \textit{Sample-Then-Calibrate}-procedure to explore the posterior distributions of these multiplicative bias terms.
+Figure \ref{Biases} presents the posterior distributions of the bias terms with side-by-side box plots.
+Since the distributions fall completely below the \textit{no-bias} reference-line at 1.0, all the groups are deemed under-confident.
+
+The under-confidence, however, decreases as the level of  expertise increases.
+For instance, the posterior probability that the most expert group is the least under-confident is approximately equal to $1.0$, and the posterior probability of a strictly decreasing level of under-confidence is approximately 0.87.
+The latter probability is driven down by the inseparability of the two groups with the lowest levels of self-reported expertise.
+The fact that these groups are very similar suggests that the experts are poor at assessing how little they know about a subject that is strange to them.
+If these groups are combined into a single group, the posterior probability of a strictly decreasing level of under-confidence is approximately 1.0.
+
+The decreasing trend in under-confidence can be reasoned by viewing the process of making a subjective probability  as Bayesian updating: A completely ignorant expert aiming to minimize a reasonable loss function, such as the Brier score, has no reason to give anything but 0.5 as his probability forecast.
+However, as soon as the expert gains some knowledge about the event, he produces an updated forecast that is a compromise between his initial forecast and the new information acquired.
+The updated forecast is therefore conservative and too close to 0.5 as long as the expert remains only partially informed about the event.
+If most experts fall somewhere on this spectrum between ignorance and full information, their average forecast tends to fall strictly between 0.5 and the most-informed probability forecast (see \citet{Baron} for more details).
+Since expertise is to a large extent determined by subject-matter knowledge, the level of under-confidence can be expected to decrease as a function of the group's level of self-reported expertise.
+
+Finding under-confidence in all the groups is a rather surprising result given that many previous studies have been conducted to show that experts are likely to be over-confident.
+For instance,  \citet{lichtenstein1977calibration, morgan1992uncertainty, bier2004implications} summarize the results from numerous calibration studies, and conclude that experts are systematically over-conﬁdent about their probability assessments.
+Our result, however, is a statement about groups of experts and hence does not invalidate the possibility of the individual experts being overconfident.
+To make conclusions at the individual-level based on the group-level bias terms would be considered an \textit{ecological inference fallacy} (see, e.g., \citet{lubinski1996seeing}).
+
 %More specific group-level inference can be performed by computing probabilities based the estimated posterior distributions.
@@
 The latter probability is driven down by the inseparability of the two groups associated with the lowest levels of self-reported expertise.
-The fact that these groups are very similar suggests that the forecasters are poor at assessing how little they know about a subject that is strange to them.
+The fact that these groups are very similar suggests that the experts are poor at assessing how little they know about a subject that is strange to them.
 If these groups are combined into a single group, the posterior probability of a strictly decreasing level of under-confidence is approximately 1.0.
@@
 
-As a final illustration of our model, consider the two questions used for illustrative purposes in Section \ref{data}.
-Figure \ref{ExamplePlotsFinal} is a copy of Figure \ref{ExamplePlots} with the addition of a solid line surrounded by a dashed band.
-The solid line represents the posterior mean of the calibrated crowd belief as estimated by STC-Log.
-The dashed lines connect the point-wise 95\% posterior intervals across different time points.
-Since $\hat{\sigma}_k^2 = 2.43$  and $\hat{\sigma}_k^2 = 1.77$ for the questions depicted in Figures \ref{Example1b} and \ref{Example2b}, respectively, the first question provokes more disagreement among the experts than the second one.
-Intuitively this makes sense because the target event in Figure \ref{Example1b}  is determined by several conditions that may change radically from one day to the next while the target event in Figure \ref{Example2b} is determined by a relatively steady stock market index.
-Therefore it is not surprising to find that the first question has $\hat{\tau}_k^2 = 1.38$ while the second one has  $\hat{\tau}_k^2 = 0.198$.
-We conclude that the first question can be considered inherently more difficult than the second one.
-
 \begin{figure}[h!]
@@
 \caption[Optional caption for list of figures]{Scatterplots of the probability forecasts given for two questions in our dataset.
-The shadings represents the self-reported expertise of the forecaster who provided the probability forecast.
+The shadings represents the self-reported expertise of the expert who provided the probability forecast.
 The solid line gives the posterior mean of the calibrated crowd belief as estimated by STC-Log.
@@
 
+As a final illustration of our model, consider the two questions used for illustrative purposes in Section \ref{data}.
+Figure \ref{ExamplePlotsFinal} is a copy of Figure \ref{ExamplePlots} with the addition of a solid line surrounded by a dashed band.
+The solid line represents the posterior mean of the calibrated crowd belief as estimated by STC-Log.
+The dashed lines connect the point-wise 95\% posterior intervals across different time points.
+Since $\hat{\sigma}_k^2 = 2.43$  and $\hat{\sigma}_k^2 = 1.77$ for the questions depicted in Figures \ref{Example1b} and \ref{Example2b}, respectively, the first question provokes more disagreement among the experts than the second one.
+Intuitively this makes sense because the target event in Figure \ref{Example1b}  is determined by several conditions that may change radically from one day to the next while the target event in Figure \ref{Example2b} is determined by a relatively steady stock market index.
+Therefore it is not surprising to find that the first question has $\hat{\tau}_k^2 = 1.38$ and the second one has  $\hat{\tau}_k^2 = 0.198$.
+We conclude that the first question can be considered inherently more difficult than the second one.
+
 \section{Discussion}
@@
 In response to this lack of appropriate modeling procedures, our work introduces an interpretable time-series model that incorporates self-reported expertise to capture a sharp and well-calibrated estimate of the crowd belief.
-The model estimation is performed in two steps.
-The first step, known as the \textit{sampling step}, samples constrained versions of the model parameters via a Gibbs sampler.
+The model estimation is performed in two steps: The first step, known as the \textit{sampling step}, samples constrained versions of the model parameters via Gibbs sampling.
 The sampling is done from standard distributions with fast convergence to the target distribution (see Appendix \ref{appendix} for technical details).
 The second step, known as the \textit{calibration step}, uses a one-dimensional optimization procedure to transforms the constrained parameter values to their unconstrained counterparts.
-With minimal sacrifices in model interpretability and efficiency, this procedure is able to extend forecasting literature towards rather unexplored areas of probability aggregation.
+To the best of our knowledge, this procedure extends forecasting literature towards rather unexplored areas of probability aggregation.
 
@@
 Therefore it can be viewed as a basis for many future extensions.
-To illustrate some potential directions, recall that most of the model parameters are assumed to remain constant over time.
-It is intuitively reasonable, however, that these model parameters behave differently during different time intervals of the question.
-For instance, the level of disagreement (represented by $\sigma^2_k$ in our model) among the experts can be expected to decrease towards the final time point when the the question finally resolves.
+To give some ideas, recall that most of the model parameters were assumed constant over time.
+It is intuitively reasonable, however, that these parameters behave differently during different time intervals of the question.
+For instance, the level of disagreement (represented by $\sigma^2_k$ in our model) among the experts can be expected to decrease towards the final time point when the question  resolves.
 This hypothesis could be explored by letting $\sigma^2_{t,k}$ evolve dynamically as a function of the previous term $\sigma^2_{t-1,k}$ and random noise.
 
-If personal information on the experts were inquired during the data collection process, it may be of interest to study the forecasting behavior of different kinds of experts.
+If the experts were asked to provide personal information during the data collection process, it may be of interest to study the forecasting behavior of different kinds of experts.
 For instance, this paper modeled the bias separately within each expertise group.

### 2013-09-01 paper.tex
Edits (some AOAS guidelines)

@@
 In response to the lack of appropriate modeling procedures, this paper introduces a time-series model that incorporates self-reported expertise and captures a sharp and well-calibrated estimate of the crowd belief.
-The model is highly interpretable and can be used for
+The model is highly interpretable and can be used for:
 \begin{itemize}
@@
 \tilde{p}_{i,t,k} &=  \begin{cases} 
-1-p_{i,t,k} & \text{if } Z_k = 1-x\\
-p_{i,t,k}& \text{if } Z_k = x,
+1-p_{i,t,k}, & \text{if } Z_k = 1-x,\\
+p_{i,t,k}, & \text{if } Z_k = x,
 \end{cases}
@@
 This means that the $k$th question is in the training set if and only if $k \in S_{train}$.
-The competing models are as follows.
+The competing models are as follows:
 
@@
 \begin{cases}
- \bar{p}_{1,k} & \text{ for } t  = 1 \\
-\alpha \bar{p}_{t,k}  + (1-\alpha) \hat{p}_{t-1,k}(\alpha) & \text{ for } t > 1
+ \bar{p}_{1,k}, & \text{ for } t  = 1, \\
+\alpha \bar{p}_{t,k}  + (1-\alpha) \hat{p}_{t-1,k}(\alpha), & \text{ for } t > 1,
 \end{cases}
@@
 \begin{cases}
-G_{t,k}(\boldsymbol{b} ) & \text{ for } t  = 1 \\
-\alpha G_{t,k}(\boldsymbol{b} )  + (1-\alpha) \hat{p}_{t-1,k}(\alpha, \boldsymbol{b})  & \text{ for } t > 1
+G_{t,k}(\boldsymbol{b} ), & \text{ for } t  = 1, \\
+\alpha G_{t,k}(\boldsymbol{b} )  + (1-\alpha) \hat{p}_{t-1,k}(\alpha, \boldsymbol{b}),  & \text{ for } t > 1,
 \end{cases}
@@
 This has direct implications  on predictive power and model interpretability.
-Firstly, the model was shown to outperform other probability aggregators in terms of forecasting ability.
+First, the model was shown to outperform other probability aggregators in terms of forecasting ability.
 In particular, the model was deemed a well-balanced compromise that avoids over-fitting without being overly conservative.
-Secondly, the crowd belief was used as the no-bias reference point to study the bias among groups of experts with different levels of self-reported expertise.
+Second, the crowd belief was used as the no-bias reference point to study the bias among groups of experts with different levels of self-reported expertise.
 All the groups were found to be under-confident.

### 2013-09-02 paper.tex
Edits

@@
  \hat{\boldsymbol{b}}_{t,k}&=& \hat{\boldsymbol{b}}(1) \hat{\beta} \nonumber\\
-   \hat{\tau}_{k}^2&=& \hat{\tau}_{k}^2(1) \hat{\beta}^2\nonumber\\
+   \hat{\tau}_{k}^2&=& \hat{\tau}_{k}^2(1)/ \hat{\beta}^2\nonumber\\
   \hat{\sigma}_{k}^2&=& \hat{\sigma}_{k}^2(1)\nonumber\\
@@
 
-\section{Extension: Polychotomous Outcomes}
-If the future event can take upon $M > 2$ possible outcomes, the hidden state $X_{t,k}$ must be extended to a vector of size $M-1$.
+If the future event can take upon $M > 2$ possible outcomes, the hidden state $X_{t,k}$ must extended to a vector of size $M-1$ and one of the outcomes, e.g., the $M$th one, must be chosen as the base-case to ensure that the probabilities will sum to one at any given time point.
+Each of the remaining  $M-1$ possible outcomes is represented by an observed process similar to Equation (\ref{observedpr}).
+Since this multinomial extension is equivalent to having $M-1$ independent binary-outcome models, the estimation and properties of the model are easily extended to the multi-outcome case.
+This paper focuses on binary-outcomes because it is the simplest and most commonly encountered setting in practice.
+
+%
+%\section{Extension: Polychotomous Outcomes}
+%If the future event can take upon $M > 2$ possible outcomes, the hidden state $X_{t,k}$ must be extended to a vector of size $M-1$.
 One of the outcomes, e.g., the $M$th one, is chosen as the base-case.
 This gives us a total of $M-1$ observed processes and one hidden process
-\begin{eqnarray*}
-\boldsymbol{Y}_{1,t,k} &=&  \boldsymbol{M}_k\boldsymbol{b}_{1}X_{1,t,k} + \boldsymbol{v}_{1,t,k}\\
-\boldsymbol{Y}_{2,t,k} &=&  \boldsymbol{M}_k\boldsymbol{b}_{2}X_{2,t,k} + \boldsymbol{v}_{2,t,k}\\
-&\vdots&\\
-\boldsymbol{Y}_{M-1,t,k} &=&  \boldsymbol{M}_k\boldsymbol{b}_{M-1}X_{M-1,t,k} + \boldsymbol{v}_{M-1,t, k}\\
-\boldsymbol{X}_{t,k} &=& \boldsymbol{\gamma}^T_k \boldsymbol{I}_{M-1} \boldsymbol{X}_{t-1,k} + \boldsymbol{w}_{t,k}
-\end{eqnarray*}
-where $\boldsymbol{X}_{t,k}$ is a $(M-1) \times 1$ matrix of calibrated logit-probabilities.
+%\begin{eqnarray*}
+%\boldsymbol{Y}_{1,t,k} &=&  \boldsymbol{M}_k\boldsymbol{b}_{1}X_{1,t,k} + \boldsymbol{v}_{1,t,k}\\
+%\boldsymbol{Y}_{2,t,k} &=&  \boldsymbol{M}_k\boldsymbol{b}_{2}X_{2,t,k} + \boldsymbol{v}_{2,t,k}\\
+%&\vdots&\\
+%\boldsymbol{Y}_{M-1,t,k} &=&  \boldsymbol{M}_k\boldsymbol{b}_{M-1}X_{M-1,t,k} + \boldsymbol{v}_{M-1,t, k}\\
+%\boldsymbol{X}_{t,k} &=& \boldsymbol{\gamma}^T_k \boldsymbol{I}_{M-1} \boldsymbol{X}_{t-1,k} + \boldsymbol{w}_{t,k}
+%\end{eqnarray*}
+%where $\boldsymbol{X}_{t,k}$ is a $(M-1) \times 1$ matrix of calibrated logit-probabilities.
 Notice that each process has a separate bias vector.
 The error terms are independent and identically distributed multivariate normal random variables
-\begin{eqnarray*}
-\boldsymbol{v}_{m, t, k} | \sigma^2_{j,k} &\stackrel{i.i.d.}{\sim}& \mathcal{N}_{I_k} \left(\boldsymbol{0}, \sigma^2_{j,k} \boldsymbol{I}_{I_k} \right)\\
-\boldsymbol{w}_{t,k} | \tau_k^2 &\stackrel{i.i.d.}{\sim}& \mathcal{N}_{M-1} \left(\boldsymbol{0}, \tau_k^2 \boldsymbol{I}_{M-1} \right)
-\end{eqnarray*}
-On the left-hand side of observed process
-\begin{eqnarray*}
-\boldsymbol{Y}_{m,t,k} &=&  \left[ \begin{matrix} Y_{m,1,t,k} & Y_{m,2,t,k} & \dots & Y_{m,I_k,t,k} \end{matrix} \right]^T\\
-&=&  \left[ \begin{matrix} \log \left( \frac{p_{m,1,t,k}}{p_{M,1,t,k}} \right) &  \log \left( \frac{p_{m,2,t,k}}{p_{M,2,t,k}} \right) & \dots &  \log \left( \frac{p_{m,I_k,t}}{p_{M,I_k,t,k}} \right) \end{matrix} \right]^T
-\end{eqnarray*}
-is a $I_k \times 1$ matrix of the expert logit-probabilities for the $m$th outcome of the $k$th question at time $t$, and on the left-hand side of the hidden process
-\begin{eqnarray*}
-\boldsymbol{X}_{t,k} &=&  \left[ \begin{matrix}X_{1,t,k} &X_{2,t,k} & \dots &X_{M-1,t,k} \end{matrix} \right]^T
-\end{eqnarray*}
-is the matrix of sharp and calibrated logit-probabilities at time $t$.
+%\begin{eqnarray*}
+%\boldsymbol{v}_{m, t, k} | \sigma^2_{j,k} &\stackrel{i.i.d.}{\sim}& \mathcal{N}_{I_k} \left(\boldsymbol{0}, \sigma^2_{j,k} \boldsymbol{I}_{I_k} \right)\\
+%\boldsymbol{w}_{t,k} | \tau_k^2 &\stackrel{i.i.d.}{\sim}& \mathcal{N}_{M-1} \left(\boldsymbol{0}, \tau_k^2 \boldsymbol{I}_{M-1} \right)
+%\end{eqnarray*}
+%On the left-hand side of observed process
+%\begin{eqnarray*}
+%\boldsymbol{Y}_{m,t,k} &=&  \left[ \begin{matrix} Y_{m,1,t,k} & Y_{m,2,t,k} & \dots & Y_{m,I_k,t,k} \end{matrix} \right]^T\\
+%&=&  \left[ \begin{matrix} \log \left( \frac{p_{m,1,t,k}}{p_{M,1,t,k}} \right) &  \log \left( \frac{p_{m,2,t,k}}{p_{M,2,t,k}} \right) & \dots &  \log \left( \frac{p_{m,I_k,t}}{p_{M,I_k,t,k}} \right) \end{matrix} \right]^T
+%\end{eqnarray*}
+%is a $I_k \times 1$ matrix of the expert logit-probabilities for the $m$th outcome of the $k$th question at time $t$, and on the left-hand side of the hidden process
+%\begin{eqnarray*}
+%\boldsymbol{X}_{t,k} &=&  \left[ \begin{matrix}X_{1,t,k} &X_{2,t,k} & \dots &X_{M-1,t,k} \end{matrix} \right]^T
+%\end{eqnarray*}
+%is the matrix of sharp and calibrated logit-probabilities at time $t$.
 Choosing one of the outcomes as the base-case, ensures that the probabilities will sum to one at any given time point.
@@
 \label{calibration}
-A calibration plot is a simple tool for assessing the sharpness and calibration of a model.
+ \ville{Add equations of sharpness and calibration}
+A calibration plot is a simple tool for visually assessing the sharpness and calibration of a model.
 The idea is to plot the probability forecasts against the observed empirical frequencies.
@@
 
-As a final illustration of our model, consider the two questions used for illustrative purposes in Section \ref{data}.
+As a final illustration of our model, consider the two example questions introduced in Section \ref{data}.
 Figure \ref{ExamplePlotsFinal} is a copy of Figure \ref{ExamplePlots} with the addition of a solid line surrounded by a dashed band.
@@
 Therefore it is not surprising to find that the first question has $\hat{\tau}_k^2 = 1.38$ and the second one has  $\hat{\tau}_k^2 = 0.198$.
-We conclude that the first question can be considered inherently more difficult than the second one.
+\ville{Elaborate more}We conclude that the first question can be considered inherently more difficult than the second one.
 

### 2013-09-04 paper.tex
Changed then name from STC --> SAC

@@
 \usepackage{rotating}
+%\usepackage[margin=1.2in]{geometry}
 \usepackage{geometry}
@@
 %
+\newtheorem{mydef}{Definition}
 \newcommand{\myN}{\hbox{N\hspace*{-.9em}I\hspace*{.4em}}}
@@
 
-%\begin{keyword}[class=MSC]
-%\kwd[Primary ]{}
-%\kwd{}
-%\kwd[; secondary ]{}
-%\end{keyword}
-
 \begin{keyword}
@@
 \end{frontmatter}
-
-%
-%\title{Dynamic Hierarchical Model for Belief Streams}
-%\author{by\\
-%\\
-%Ville A. Satop\"a\"a\\
-%Shane T. Jensen and Lyle H. Ungar, Faculty Advisors\\
-%\\
-% \small Department of Statistics\\
-% \small The Wharton School of the University of Pennsylvania\\
-% \small Philadelphia, PA 19104- 6340, USA\\ [-0.25in]} \date{}
-%\maketitle
-%
-%\pagestyle{myheadings}
-%\markboth{Ville A. Satop\"a\"a}{Calibrated Crowd Belief}
-%\thispagestyle{empty}
 
@@
 and $(\mu_0, \sigma_0^2) \in (\R, \R^+)$ are hyper-parameters chosen \textit{a priori}.
-The hidden state $X_{t,k}$ represents the sharp and well-calibrated logit-probability for the $k$th event given the information available up to and including time $t$.
+
+The hidden state $X_{t,k}$ represents the logit-probability for the $k$th event given all the information available up to and including time $t$.
 To make this more specific, let $Z_k \in \{0, 1\}$ indicate whether the event associated with the $k$th question happened $\left(Z_k = 1\right)$ or did not happen $\left(Z_k = 0\right)$.
 If $\{\mathcal{F}_{t,k}\}_{t=1}^{T_k}$ is a filtration representing the information available up to and including a given time point, then $\E[Z_k | \mathcal{F}_{t,k}] = \P(Z_k = 1| \mathcal{F}_{t,k}) = \logit^{-1}(X_{t,k})$.
-These probabilities are made estimable by assuming that the observed process is associated with the filtration $\{\mathcal{F}_{t,k}\}_{t=1}^{T_k}$ and hence can be used as a proxy for the amount of information available at any given time point.
-%Therefore it is not unreasonable to interpreted the hidden process as the crowd belief across time.
-
-To give intuitive interpretations of the other model parameters, notice that the error term in the observed process has mean zero.
-Therefore the experts are assumed to be, on average, a multiplicative constant $\boldsymbol{b}$ away from the calibrated crowd belief.
-An individual element of $\boldsymbol{b}$ can be interpreted as a group-specific \textit{systematic bias} that labels the group either as over-confident $\left( b_j \in (1, \infty)\right)$ or as under-confident $\left(b_j \in (0,1)\right)$.
-Unfortunately, due to the high sparsity of our data, estimating a bias term and analyzing the bias separately for each expert is not possible.
+It is reasonable to assume that an expert, who has access to all available information and reports $\logit^{-1}(X_{t,k})$ as the probability forecast, maximizes sharpness subject to calibration, where calibration and sharpness are understood as follows:
+\begin{mydef}
+A probability forecast $p$ for the $k$th question is calibrated if $\P(Z_k = 1 | p) = \E[Z_k | p] = p$ almost surely (see, e.g., \citet{murphy1987general}).
+\end{mydef}
+\begin{mydef}
+A probability forecast $p$ is sharper than a probability forecast $q$ if $\E[(p-p_0)^2] > \E[(q-p_0)^2]$, where $p_0$ is the baseline probability for the $k$th question (see, e.g., \citet{ranjan2009combining}).
+\end{mydef}
+\noindent
+Even though it is unlikely that any single expert has all the available information, a large and diverse group of experts may share a considerable portion of this information.
+The so called \textit{wisdom of the crowd} is therefore an attractive proxy for the amount of information available at any given time point.
+
+If a model is to efficiently use this proxy for estimating the hidden process $X_{t,k}$, it must be able to discriminate between noise and signal as experts may believe in false information, hide their true beliefs on an ethical basis, or be biased for many other reasons.
+Since in our model the experts are assumed to be, on average, a multiplicative constant $\boldsymbol{b}$ away from the calibrated crowd belief, an individual element of $\boldsymbol{b}$ can be interpreted as a group-specific \textit{systematic bias} that labels the group either as over-confident $\left( b_j \in (1, \infty)\right)$ or as under-confident $\left(b_j \in (0,1)\right)$.
+Due to the high sparsity of our data, estimating a bias term and analyzing the bias separately for each expert is not possible.
 See Section \ref{ExpertBias} for an analysis and discussion on the bias terms.
@@
 The main challenge is to capture a well-calibrated estimate of the hidden process without sacrificing the interpretability of our model.
-This section introduces a two-step procedure, called \textit{Sample-Then-Calibrate} (STC), that achieves this goal in a flexible and efficient manner: the first step estimates the model parameters under a constraint (\textit{Sampling Step}), and the second step performs a one-dimension optimization procedure to transform the constrained estimates into their unconstrained counterparts (\textit{Calibration Step}).
+This section introduces a two-step procedure, called \textit{Sample-And-Calibrate} (SAC), that achieves this goal in a flexible and efficient manner: the first step estimates the model parameters under a constraint (\textit{Sampling Step}), and the second step performs a one-dimension optimization procedure to transform the constrained estimates into their unconstrained counterparts (\textit{Calibration Step}).
 
@@
 \end{eqnarray*}
-Since it is not clear which rule should be used for predicting geopolitical events, the \textit{Sample-Then-Calibrate} procedure is evaluated  separately under both rules in Sections \ref{syntheticData} and \ref{realData}.
+Since it is not clear which rule should be used for predicting geopolitical events, the \textit{Sample-And-Calibrate} procedure is evaluated  separately under both rules in Sections \ref{syntheticData} and \ref{realData}.
 Once $\hat{\beta}$ has been computed, estimates of the unconstrained model parameters are given by
@@
 \label{syntheticData}
-
 The synthetic data is not generated directly from the model for two reasons: (i) showing good performance on a dataset directly generated from the model assumptions is hardly any news, and (ii) generating data from the dynamic model description does not produce well-calibrated hidden states.
@@
 gives a sequence of $T_k$ calibrated logit-probabilities for the event $Z_k = 1$.
-
-We use this procedure to generate the hidden processes for $K$ questions with $T_k = 100$ for $k = 1, \dots, K$.
-Each question involves 50 experts allocated evenly among the five expertise groups.
-The expert logit-probabilities $Y_{i,t,k}$ are generated from Equation (\ref{observedpr}) by applying bias and noise to the hidden process $X_{t,k}$.
-Any extreme logit-probabilities are truncated to avoid overly influential observations; that is, any $p_{i,t,k} = \logit^{-1}(Y_{i,t,k})$ less than 0.01 (or greater than 0.99) is truncated to 0.01 (or to 0.99).
-This is akin to the real-world data setting, where extreme probabilities must be truncated (see Section \ref{realData}).
-Therefore, to understand the effect that truncation has on the real-world data results, truncation should be first analyzed and understood under the controlled environment of synthetic data.
-
-Throughout the evaluation, data is generated under different choices of parameter values. More specifically, the evaluation procedure iterates over a three-dimensional grid of parameter values
-\begin{eqnarray*}
+Repeating this procedure for $K$ times, we generate the hidden process for $K$ questions, each with a time horizon of $T_k = 101$.
+The questions involve 25 experts allocated evenly among five expertise groups.
+Since the event resolves at time $t = 101$, the expert logit-probabilities $Y_{i,t,k}$ are only given for times $t = 1$ to $100$.
+These are generated from Equation (\ref{observedpr}) by applying bias and noise to the hidden process $X_{t,k}$.
+The simulation then iterates over a three-dimensional grid of parameter values
+\begin{eqnarray*}
+\sigma^2 &\in& \{1/2, 1, 3/2, 2, 5/2\}\\
 \beta &\in& \{1/2, 3/4, 1, 4/3, 2/1\}\\
-\sigma^2 &\in& \{0.5, 1.0, 1.5, 2.0, 2.5\}\\
-K &\in& \{20, 35, 50, 65, 80\},
-\end{eqnarray*}
-where $\beta$ generates the bias vector  $\boldsymbol{b} = [0.50, 0.75, \dots, 1.50]^T \beta$.
-These values were considered to be realistic based on our experience with the real-world probability forecasting data.
-Each triple of the parameter values,  i.e.
-a grid point, is used  40 times to generate a synthetic dataset.
-Each dataset is then used to assess how accurately \textit{Sample-Then-Calibrate} is able to estimate the hidden process and bias vector.
-The accuracy of the estimated hidden process is measured with the average quadratic loss in the probability space, $\sum_{k=1}^K \sum_{t=1}^{T_k} ( \logit^{-1}(\hat{X}_{t,k}) - \logit^{-1}(X_{t,k}))^2 / \sum_{k=1}^K T_k$, and the accuracy of the estimtated bias vector is quantified by the average quadratic loss, $|| \boldsymbol{b} - \hat{\boldsymbol{b}}||^2/5$.
+K &\in& \{20, 40, 60, 80, 100\},
+\end{eqnarray*}
+where $\beta$ gives the value of the bias vector by letting $\boldsymbol{b} = [1/2, 3/4, 1, 4/3, 2/1]^T \beta$.
+Each grid point is used  40 times to generate a synthetic dataset.
+This dataset is then used to assess how accurately \textit{Sample-And-Calibrate} estimates the hidden process and bias vector.
+The accuracy of the estimated hidden process is measured with the average quadratic loss in the probability space, $\sum_{k=1}^K \sum_{t=1}^{T_k} ( \logit^{-1}(\hat{X}_{t,k}) - \logit^{-1}(X_{t,k}))^2 / \sum_{k=1}^K T_k$, and the accuracy of the estimated bias vector is measured with the average quadratic loss, $|| \boldsymbol{b} - \hat{\boldsymbol{b}}||^2/5$.
 
@@
 
-\caption[Optional caption for list of figures]{Comparing \textit{Sample-Then-Calibrate} that optimizes over the logarithmic score (STC-Log) and \textit{Sample-Then-Calibrate} that optimizes over the Brier score (STC-Bri) under synthetic data.
+\caption[Optional caption for list of figures]{Comparing \textit{Sample-And-Calibrate} that optimizes over the logarithmic score ($\text{SAC}_{\text{LOG}}$) and \textit{Sample-And-Calibrate} that optimizes over the Brier score (SAC-Bri) under synthetic data.
 The top row presents the accuracy to capture the hidden process.
@@
 Figure \ref{Synthetic} summarizes the results in a series of plots.
-The first row shows the marginal effects of the three grid variables on the accuracy to which the \textit{Sample-Then-Calibrate} under the logarithmic loss (STC-Log) and the Brier score (STC-Bri) are able to estimate the hidden process.
+The first row shows the marginal effects of the three grid variables on the accuracy to which the \textit{Sample-And-Calibrate} under the logarithmic loss ($\text{SAC}_{\text{LOG}}$) and the Brier score (SAC-Bri) are able to estimate the hidden process.
 The second row is similar in nature but measures the accuracy of estimating the bias vector instead.
-The plots were computed by choosing one grid variable (e.g.
+Since the plots were computed by choosing one grid variable (e.g.
 $\beta$) at a time and averaging over the remaining two variables  (e.g.
-$K$, and $\sigma^2$).
-Therefore each value in the plots represents an average of a $40 \times 5^2 = 1,000$ values.
-
-Based on the first row of plots, STC-Log and STC-Bri estimate the hidden process to a similar accuracy.
+$K$, and $\sigma^2$), each value in the plots represents an average of a $40 \times 5^2 = 1,000$ values.
+
+Based on the first row of plots, $\text{SAC}_{\text{LOG}}$ and SAC-Bri estimate the hidden process to a similar accuracy.
 Even though some relative differences are visible, the loss values are very small in absolute terms, hence suggesting good overall accuracy to capture the hidden process.
@@
 
-Based on the second row of plots, STC-Log and STC-Bri  capture the bias vector rather precisely as long as the true values of $\beta$ and $\sigma^2$ are not very large.
+Based on the second row of plots, $\text{SAC}_{\text{LOG}}$ and SAC-Bri  capture the bias vector rather precisely as long as the true values of $\beta$ and $\sigma^2$ are not very large.
 This behavior can be again considered as being caused by the truncation of the extreme expert logit-probabilities.
-Surprisingly, even though STC-Log is more accurate in estimating the bias vector, it is unable to outperform STC-Bri in estimating the hidden process.
+Surprisingly, even though $\text{SAC}_{\text{LOG}}$ is more accurate in estimating the bias vector, it is unable to outperform SAC-Bri in estimating the hidden process.
 
@@
 \noindent
-This section discusses results related to the real-world data described in Section \ref{data}.
-The goal is to provide application specific insight by discussing the specific research objectives itemized in Section \ref{intro} and also to evaluate the \textit{Sample-Then-Calibrate}-procedure in terms of predictive power and calibration.
+This section presents results related to the real-world data described in Section \ref{data}.
+The goal is to provide application specific insight by discussing the specific research objectives itemized in Section \ref{intro} and also to evaluate the \textit{Sample-And-Calibrate}-procedure in terms of predictive power and calibration.
 Before presenting the results, however, we discuss two practical matters that must be taken into account when aggregating real-world  probability forecasts.
@@
 \subsection{Incoherent and Imbalanced Data}
+\label{practicalMatters}
 The first matter regards human experts making probability forecasts of 0.0 or 1.0 even if they are not completely sure of the outcome of the event.
@@
 \citet{Geo}, on the other hand, consider only probabilities that fall within a constrained interval, say $[0.001, 0.999]$, and discard the rest.
-Since this implies ignoring a portion of the data, we have decided to adopt the first approach by changing $p = 0.00$ and $1.00$ to $p = 0.01$ and $0.99$, respectively.
+Since this implies ignoring a portion of the data, we adopt the first approach by truncating $p = 0.00$ and $1.00$ to $p = 0.01$ and $0.99$, respectively.
+Our results remain insensitive to the exact choice of truncation as long as this is done in a reasonable manner to keep the extreme probabilities from becoming highly influential in the logit-space.
 
@@
 \label{forecasting}
-This section evaluates the out-of-sample predictive performance of \textit{Sample-Then-Calibrate} against several other probability aggregation procedures.
+This section is motivated by decision-making. 
+%For instance, recall our example with an investor who consults a group of experts on a daily basis in order to assess whether a stock index will end up trading above some threshold on a given date. 
+The goal is to evaluate the out-of-sample predictive performance of \textit{Sample-And-Calibrate} against several other probability aggregation procedures.
 The models are allowed to utilize a training set before making predictions on an independent testing set.
@@
 
-\item \textit{The Sample-Then-Calibrate procedure both under the Brier (STC-Bri) and the logarithmic score (STC-Log)}.
+\item \textit{The Sample-And-Calibrate procedure both under the Brier (SAC-Bri) and the logarithmic score ($\text{SAC}_{\text{LOG}}$)}.
 The model is first fit on the training set by running the sampling step for 3,000 iterations of which the first 500 iterations are used for burn-in.
@@
 
-\item \textit{A fully Bayesian version of STC-Log (BSTC-Log)}.
+\item \textit{A fully Bayesian version of $\text{SAC}_{\text{LOG}}$ ($\text{BSAC}_{\text{LOG}}$)}.
 Denote the calibrated logit-probabilities and  the event indicators across all $K$ questions with $\boldsymbol{X}(1)$ and $\boldsymbol{Z}$, respectively.
@@
 Therefore the Metropolis algorithm is also used for sampling the hidden states.
-Predictions are made with the same choices of thinning and burn-in as described under \textit{Sample-Then-Calibrate}.
+Predictions are made with the same choices of thinning and burn-in as described under \textit{Sample-And-Calibrate}.
 
@@
 The vector $\boldsymbol{b}$ collects the bias terms of the different expertise groups.
-Therefore it is equivalent to the bias vector found under \textit{Sample-Then-Calibrate}.
+Therefore it is equivalent to the bias vector found under \textit{Sample-And-Calibrate}.
 The term $\boldsymbol{e}_{i,k}$ is a vector of length 5 indicating which level of self-reported expertise the $i$th expert in the $k$th question belongs to.
@@
 SDLM & 0.100 (0.156) & 0.066 (0.116) & 0.098 (0.154) & 0.102 (0.157)\\ 
-BSTC-Log & 0.097 (0.213) & \textbf{0.053} (0.147) & 0.100 (0.215) & 0.098 (0.215)\\ 
-STC-Bri & 0.096 (0.190) & 0.056 (0.134) & 0.097 (0.190) & 0.098 (0.192)\\ 
-STC-Log & \textbf{0.096} (0.191) & 0.056 (0.134) & \textbf{0.096} (0.189) & \textbf{0.098} (0.193)\\ 
+$\text{BSAC}_{\text{LOG}}$ & 0.097 (0.213) & \textbf{0.053} (0.147) & 0.100 (0.215) & 0.098 (0.215)\\ 
+$\text{SAC}_{\text{BRI}}$ & 0.096 (0.190) & 0.056 (0.134) & 0.097 (0.190) & 0.098 (0.192)\\ 
+$\text{SAC}_{\text{LOG}}$ & \textbf{0.096} (0.191) & 0.056 (0.134) & \textbf{0.096} (0.189) & \textbf{0.098} (0.193)\\ 
 EWMBA & 0.102 (0.203) & 0.060 (0.124) & 0.110 (0.201) & 0.103 (0.206)\\ 
@@
 SDLM & 0.089 (0.116) & 0.064 (0.085) & 0.106 (0.141) & 0.092 (0.117)\\ 
-BSTC-Log & 0.083 (0.160) & \textbf{0.052} (0.103) & 0.110 (0.198) & 0.085 (0.162)\\ 
-STC-Bri & 0.083 (0.142) & 0.055 (0.096) & 0.106 (0.174) & 0.085 (0.144)\\ 
-STC-Log & \textbf{0.082} (0.142) & 0.055 (0.096) & \textbf{0.105} (0.174) & \textbf{0.085} (0.144)\\ 
+$\text{BSAC}_{\text{LOG}}$ & 0.083 (0.160) & \textbf{0.052} (0.103) & 0.110 (0.198) & 0.085 (0.162)\\ 
+$\text{SAC}_{\text{BRI}}$& 0.083 (0.142) & 0.055 (0.096) & 0.106 (0.174) & 0.085 (0.144)\\ 
+$\text{SAC}_{\text{LOG}}$ & \textbf{0.082} (0.142) & 0.055 (0.096) & \textbf{0.105} (0.174) & \textbf{0.085} (0.144)\\ 
 EWMBA & 0.090 (0.156) & 0.063 (0.101) & 0.118 (0.186) & 0.091 (0.161)\\ 
@@
 
-Overall, STC-Bri and STC-Log achieve the lowest average scores across all columns except \textit{Short}, where they are slightly outperformed by BSTC-Log.
-BSTC-Log, however, turns out to be overconfident (see Section \ref{calibration}).
-This means that BSTC-Log underestimates the uncertainty in the events and outputs probability forecasts that are typically too near 0.0 or 1.0.
+Overall, SAC-Bri and $\text{SAC}_{\text{LOG}}$ achieve the lowest average scores across all columns except \textit{Short}, where they are slightly outperformed by $\text{BSAC}_{\text{LOG}}$.
+$\text{BSAC}_{\text{LOG}}$, however, turns out to be overconfident (see Section \ref{calibration}).
+This means that $\text{BSAC}_{\text{LOG}}$ underestimates the uncertainty in the events and outputs probability forecasts that are typically too near 0.0 or 1.0.
 As a result, the aggregate forecasts are either very close to the correct answer or very far from it.
@@
 
-In comparison to BSTC-Log, the baseline SDLM-model lacks sharpness and is highly under-confident (see Section \ref{calibration}).
+In comparison to $\text{BSAC}_{\text{LOG}}$, the baseline SDLM-model lacks sharpness and is highly under-confident (see Section \ref{calibration}).
 This behavior is expected as the experts are under-confident at the group-level (see Section \ref{ExpertBias}) and the SDLM-procedure does not use the training set to explicitly calibrate its forecasts.
@@
 Overall, these two aggregators perform very similarly.
-Unfortunately, they both suffer from over-confidence and are hence unable to outperform the \textit{Sample-Then-Calibrate}-procedure.
+Unfortunately, they both suffer from over-confidence and are hence unable to outperform the \textit{Sample-And-Calibrate}-procedure.
 
@@
 \label{calibration}
- \ville{Add equations of sharpness and calibration}
 A calibration plot is a simple tool for visually assessing the sharpness and calibration of a model.
@@
 }
-\subfigure[\scriptsize STC-Log In-Sample]{
+\subfigure[\scriptsize $\text{SAC}_{\text{LOG}}$ In-Sample]{
 \hspace*{-1.15em} \includegraphics{Figures/CalibrationLOG10}
@@
 }
-\subfigure[\scriptsize STC-Bri In-Sample]{
+\subfigure[\scriptsize $\text{SAC}_{\text{BRI}}$ In-Sample]{
 \hspace*{-1.15em} \includegraphics{Figures/CalibrationBRI10}
@@
 }
-\subfigure[\scriptsize BSTC-Log In-Sample]{
+\subfigure[\scriptsize $\text{BSAC}_{\text{LOG}}$ In-Sample]{
 \hspace*{-1.15em} \includegraphics{Figures/CalibrationFULL10}
@@
 }
-\subfigure[ \scriptsize STC-Log Out-of-Sample]{
+\subfigure[ \scriptsize $\text{SAC}_{\text{LOG}}$ Out-of-Sample]{
 \hspace*{-1.15em} \includegraphics{Figures/CalibrationSTO-Log}
@@
 }
-\subfigure[ \scriptsize STC-Bri Out-of-Sample]{
+\subfigure[ \scriptsize $\text{SAC}_{\text{BRI}}$ Out-of-Sample]{
 \hspace*{-1.15em} \includegraphics{Figures/CalibrationSTO-Bri}
@@
 }
-\subfigure[\scriptsize  BSTC-Log Out-of-Sample]{
+\subfigure[\scriptsize  $\text{BSAC}_{\text{LOG}}$ Out-of-Sample]{
 \hspace*{-1.15em} \includegraphics{Figures/CalibrationSO}
@@
 \caption[Optional caption for list of figures]{The top and bottom rows show in- and out-of-sample calibration and sharpness, respectively.
-The models are Simple Dynamic Linear Model (SDLM), the \textit{Sample-Then-Calibrate} approach that optimizes over the logarithmic score (STC-Log), the \textit{Sample-Then-Calibrate} approach that optimizes over the Brier score (STC-Bri), and the fully-Bayesian version of the \textit{Sample-Then-Calibrate} approach that optimizes over the logarithmic score (BSTC-Log).
+The models are Simple Dynamic Linear Model (SDLM), the \textit{Sample-And-Calibrate} approach that optimizes over the logarithmic score ($\text{SAC}_{\text{LOG}}$), the \textit{Sample-And-Calibrate} approach that optimizes over the Brier score (SAC-Bri), and the fully-Bayesian version of the \textit{Sample-And-Calibrate} approach that optimizes over the logarithmic score ($\text{BSAC}_{\text{LOG}}$).
 }
@@
 
-The top and bottom rows of Figure \ref{Calibration-Out} present calibration plots for SDLM, STC-Log, STC-Bri, and BSTC-Log under in- and out-of-sample probability estimation, respectively.
+The top and bottom rows of Figure \ref{Calibration-Out} present calibration plots for SDLM, $\text{SAC}_{\text{LOG}}$, SAC-Bri, and $\text{BSAC}_{\text{LOG}}$ under in- and out-of-sample probability estimation, respectively.
 Each setting is of interest in its own right:  Good in-sample calibration is crucial for model interpretability.
@@
 
-Overall, the \textit{Sample-Then-Calibrate}-procedure is sharp and well-calibrated both in- and out-of-sample with only a few points barely falling outside the \textit{point-wise} critical values.
-Since the calibration does not change drastically from the top to the bottom row, the \textit{Sample-Then-Calibrate}-procedure can be considered to present robustness against over-fitting.
-This is, however, not the case with BSTC-Log that is well-calibrated in-sample but presents over-confidence out-of-sample.
+Overall, the \textit{Sample-And-Calibrate}-procedure is sharp and well-calibrated both in- and out-of-sample with only a few points barely falling outside the \textit{point-wise} critical values.
+Since the calibration does not change drastically from the top to the bottom row, the \textit{Sample-And-Calibrate}-procedure can be considered to present robustness against over-fitting.
+This is, however, not the case with $\text{BSAC}_{\text{LOG}}$ that is well-calibrated in-sample but presents over-confidence out-of-sample.
 Figures \ref{CalibrationNONEa} and \ref{CalibrationNONEb} serve as baselines by showing the reliability plots for the SDLM-model.
@@
 Furthermore, the inset histogram reveals drastic lack of sharpness.
-Therefore the \textit{Sample-Then-Calibrate}-model can be viewed as a well-performing compromise between SDLM and BSTC-Log that avoids over-confidence without being too conservative.
+Therefore the \textit{Sample-And-Calibrate}-model can be viewed as a well-performing compromise between SDLM and $\text{BSAC}_{\text{LOG}}$ that avoids over-confidence without being too conservative.
 
@@
 \caption[Optional caption for list of figures]{Comparing the bias-levels across self-reported expertise under different approaches. 
-Posterior distributions of $b_j$ for $j = 1, \dots, 5$ under the \textit{Sample-Then-Calibrate} approach that optimizes over the logarithmic score and the \textit{Sample-Then-Calibrate} approach that optimizes over the Brier score.
+Posterior distributions of $b_j$ for $j = 1, \dots, 5$ under the \textit{Sample-And-Calibrate} approach that optimizes over the logarithmic score and the \textit{Sample-And-Calibrate} approach that optimizes over the Brier score.
 }
@@
 The self-reported expertise then divides the experts into 5  groups, with each group assigned a separate multiplicative bias term.
-This section uses the \textit{Sample-Then-Calibrate}-procedure to explore the posterior distributions of these multiplicative bias terms.
+This section uses the \textit{Sample-And-Calibrate}-procedure to explore the posterior distributions of these multiplicative bias terms.
 Figure \ref{Biases} presents the posterior distributions of the bias terms with side-by-side box plots.
 Since the distributions fall completely below the \textit{no-bias} reference-line at 1.0, all the groups are deemed under-confident.
-
-The under-confidence, however, decreases as the level of  expertise increases.
+Even though the exact level of under-confidence is affected slightly by the extent to which the extreme probabilities are truncated (see Section \ref{practicalMatters}), the qualitative results in this section remain insensitive to different levels of truncation.
+
+The under-confidence decreases as the level of  expertise increases.
 For instance, the posterior probability that the most expert group is the least under-confident is approximately equal to $1.0$, and the posterior probability of a strictly decreasing level of under-confidence is approximately 0.87.
@@
 The shadings represents the self-reported expertise of the expert who provided the probability forecast.
-The solid line gives the posterior mean of the calibrated crowd belief as estimated by STC-Log.
+The solid line gives the posterior mean of the calibrated crowd belief as estimated by $\text{SAC}_{\text{LOG}}$.
 The surrounding dashed lines connect the point-wise 95\% posterior intervals.}
@@
 
-As a final illustration of our model, consider the two example questions introduced in Section \ref{data}.
+As a final illustration of our model, we return to the two example questions introduced in Section \ref{data}.
 Figure \ref{ExamplePlotsFinal} is a copy of Figure \ref{ExamplePlots} with the addition of a solid line surrounded by a dashed band.
-The solid line represents the posterior mean of the calibrated crowd belief as estimated by STC-Log.
+The solid line represents the posterior mean of the calibrated crowd belief as estimated by $\text{SAC}_{\text{LOG}}$.
 The dashed lines connect the point-wise 95\% posterior intervals across different time points.
@@
 Intuitively this makes sense because the target event in Figure \ref{Example1b}  is determined by several conditions that may change radically from one day to the next while the target event in Figure \ref{Example2b} is determined by a relatively steady stock market index.
-Therefore it is not surprising to find that the first question has $\hat{\tau}_k^2 = 1.38$ and the second one has  $\hat{\tau}_k^2 = 0.198$.
-\ville{Elaborate more}We conclude that the first question can be considered inherently more difficult than the second one.
+Therefore it is not surprising to find that in Figure \ref{Example1b} $\hat{\tau}_k^2 = 0.269$, which is much higher than $\hat{\tau}_k^2 = 0.039$ in Figure \ref{Example2b}.
+We may conclude that the first question is inherently more difficult than the second one.
 

### 2013-09-04 paper.tex
Edits

@@
 
-Overall, the data is quite sparse.
-As mentioned before, the experts were allowed to update their forecasts as long as the question remained open.
-Updating, however, was done on a very infrequent basis.
-An expert gave, on average, only 0.016 forecasts per day.
-Therefore a typical expert made a forecast about every 62 days, resulting to an average response rate of around 13.5 forecasts per day from a very large and diverse group of experts.
+The experts made updates on a very infrequent basis.
+An expert gave, on average, only 0.016 forecasts per day; that is, a forecast in about every 63 days.
+The average group-level response rate was around 13.5 forecasts per day.
+Given that the group of experts is large and diverse, the data is very sparse.
 Tables \ref{DataStats} and \ref{ExpertiseTable} provide more relevant summary statistics on the data.
@@
 If $\{\mathcal{F}_{t,k}\}_{t=1}^{T_k}$ is a filtration representing the information available up to and including a given time point, then $\E[Z_k | \mathcal{F}_{t,k}] = \P(Z_k = 1| \mathcal{F}_{t,k}) = \logit^{-1}(X_{t,k})$.
-It is reasonable to assume that an expert, who has access to all available information and reports $\logit^{-1}(X_{t,k})$ as the probability forecast, maximizes sharpness subject to calibration, where calibration and sharpness are understood as follows:
+It is reasonable to assume that this probability forecast
+%that makes uses all available information and hence is equivalent to $\logit^{-1}(X_{t,k})$ 
+%an expert, who has access to all available information and reports $\logit^{-1}(X_{t,k})$ as the probability forecast, 
+maximizes sharpness subject to calibration, where calibration and sharpness are understood as follows:
 \begin{mydef}
@@
 \noindent
-Even though it is unlikely that any single expert has all the available information, a large and diverse group of experts may share a considerable portion of this information.
-The so called \textit{wisdom of the crowd} is therefore an attractive proxy for the amount of information available at any given time point.
-
-If a model is to efficiently use this proxy for estimating the hidden process $X_{t,k}$, it must be able to discriminate between noise and signal as experts may believe in false information, hide their true beliefs on an ethical basis, or be biased for many other reasons.
-Since in our model the experts are assumed to be, on average, a multiplicative constant $\boldsymbol{b}$ away from the calibrated crowd belief, an individual element of $\boldsymbol{b}$ can be interpreted as a group-specific \textit{systematic bias} that labels the group either as over-confident $\left( b_j \in (1, \infty)\right)$ or as under-confident $\left(b_j \in (0,1)\right)$.
+Any single expert is unlikely to have enough information to consistently report $\logit^{-1}(X_{t,k})$.
+He may also believe in false information, hide his true belief for ethical bases, or be biased for many other reasons.
+A large and diverse group of experts, on the other hand, may share a considerable portion of the available information.
+But if they are not collaborating and sharing information efficiently, their beliefs should be aggregated via a model that can detect potential bias, separate signal from noise, and use the collective opinion to estimate $X_{t,k}$.
+
+%A model should therefore aim to take into account the collective opinion of a group of experts. 
+%
+%By taking into account the collective opinion the group of experts may be able to 
+%
+%
+%. The collective wisdom of this group is therefore an attractive proxy for the amount of information available at any given time point. 
+%
+%The experts, however, may not be collaborating and sharing information efficiently with each other.
+
+%Therefore, if a model is to efficiently use the collective wisdom for estimating $X_{t,k}$, it must be able to discriminate between noise and signal.
+
+ In our model the experts are assumed to be, on average, a multiplicative constant $\boldsymbol{b}$ away from the calibrated crowd belief.
+An individual element of $\boldsymbol{b}$ can be interpreted as a group-specific \textit{systematic bias} that labels the group either as over-confident $\left( b_j \in (1, \infty)\right)$ or as under-confident $\left(b_j \in (0,1)\right)$.
 Due to the high sparsity of our data, estimating a bias term and analyzing the bias separately for each expert is not possible.
@@
 
- Similarly to the observed process, the hidden process  has a systematic and a random component.
-The \textit{random fluctuations} are measured in terms of $\tau^2_k$ and can be assumed to represent changes or shocks to the underlying circumstances that ultimately decide the outcome of the event.
+ The \textit{random fluctuations} in the hidden process are measured in terms of $\tau^2_k$ and can be assumed to represent changes or shocks to the underlying circumstances that ultimately decide the outcome of the event.
 The \textit{systematic component} $\gamma_k$, on the other hand, allows the model to incorporate a constant signal stream that drifts the hidden process to infinity (when $\gamma_k \in (1, \infty)$) or zero (when $\gamma_k \in (0, 1)$).

### 2013-09-05 paper.tex
Getting close...

@@
 Statistic & Min. & $Q_1$ & Median & Mean & $Q_3$ & Max.\\ \hline
-\# of Days a Problem is Active  & 4 &  35.6 &  72.0 & 106.3 & 145.20 & 418 \\
-\# of Experts per Problem & 212 &   543.2 &  693.5 &  783.7 &  983.2 & 1690\\ 
-\# Forecasts given by each Expert on a Problem & 1 &  1.0 &   1.0  &  1.8 &  2.0 & 131 \\
-\# Problems participated by an Expert & 1 &  14.0 &   36.0  &  55.0 &  90.0 & 166 \\ \hline
+\# of Days a Question is Active  & 4 &  35.6 &  72.0 & 106.3 & 145.20 & 418 \\
+\# of Experts per Question & 212 &   543.2 &  693.5 &  783.7 &  983.2 & 1690\\ 
+\# Forecasts given by each Expert on a Question & 1 &  1.0 &   1.0  &  1.8 &  2.0 & 131 \\
+\# Questions participated by an Expert & 1 &  14.0 &   36.0  &  55.0 &  90.0 & 166 \\ \hline
 \end{tabular}
@@
 \noindent
-Any single expert is unlikely to have enough information to consistently report $\logit^{-1}(X_{t,k})$.
-He may also believe in false information, hide his true belief for ethical bases, or be biased for many other reasons.
-A large and diverse group of experts, on the other hand, may share a considerable portion of the available information.
-But if they are not collaborating and sharing information efficiently, their beliefs should be aggregated via a model that can detect potential bias, separate signal from noise, and use the collective opinion to estimate $X_{t,k}$.
-
-%A model should therefore aim to take into account the collective opinion of a group of experts. 
-%
-%By taking into account the collective opinion the group of experts may be able to 
-%
-%
-%. The collective wisdom of this group is therefore an attractive proxy for the amount of information available at any given time point. 
-%
-%The experts, however, may not be collaborating and sharing information efficiently with each other.
-
-%Therefore, if a model is to efficiently use the collective wisdom for estimating $X_{t,k}$, it must be able to discriminate between noise and signal.
-
- In our model the experts are assumed to be, on average, a multiplicative constant $\boldsymbol{b}$ away from the calibrated crowd belief.
-An individual element of $\boldsymbol{b}$ can be interpreted as a group-specific \textit{systematic bias} that labels the group either as over-confident $\left( b_j \in (1, \infty)\right)$ or as under-confident $\left(b_j \in (0,1)\right)$.
-Due to the high sparsity of our data, estimating a bias term and analyzing the bias separately for each expert is not possible.
+Even though it is unlikely that any single expert has access to all the available information, a large and diverse group of experts may share a considerable portion of the available information.
+The collective wisdom of the group therefore provides an attractive proxy for $\mathcal{F}_{t,k}$.
+
+Since the experts may believe in false information, hide their true beliefs for ethical bases, or be biased for many other reasons, their probability assessments should be aggregated via a model that can detect potential bias, separate signal from noise, and use the collective opinion to estimate $X_{t,k}$.
+In our model the experts are assumed to be, on average, a multiplicative constant $\boldsymbol{b}$ away from $X_{t,k}$.
+Therefore an individual element of $\boldsymbol{b}$ can be interpreted as a group-specific \textit{systematic bias} that labels the group either as over-confident $\left( b_j \in (1, \infty)\right)$ or as under-confident $\left(b_j \in (0,1)\right)$.
+Due to the high sparsity of our data, estimating a bias term separately for each expert is not possible.
 See Section \ref{ExpertBias} for an analysis and discussion on the bias terms.
-Any other deviation from the calibrated crowd belief is considered \textit{random noise}.
+Any other deviation from $X_{t,k}$ is considered \textit{random noise}.
 This noise is measured in terms of $\sigma^2_k$ and can be assumed to be caused by momentary over-optimism (or pessimism), false beliefs, or other misconceptions.
 
- The \textit{random fluctuations} in the hidden process are measured in terms of $\tau^2_k$ and can be assumed to represent changes or shocks to the underlying circumstances that ultimately decide the outcome of the event.
-The \textit{systematic component} $\gamma_k$, on the other hand, allows the model to incorporate a constant signal stream that drifts the hidden process to infinity (when $\gamma_k \in (1, \infty)$) or zero (when $\gamma_k \in (0, 1)$).
+ The \textit{random fluctuations} in the hidden process are measured in terms of $\tau^2_k$ and are assumed to represent changes or shocks to the underlying circumstances that ultimately decide the outcome of the event.
+The \textit{systematic component} $\gamma_k$ allows the model to incorporate a constant signal stream that drifts the hidden process to infinity (when $\gamma_k \in (1, \infty)$) or zero (when $\gamma_k \in (0, 1)$).
 If the uncertainty in the question diminishes as the current time point $t$ approaches $T_k$, the hidden process drifts to infinity.
@@
  \hat{X}_{t,k}&=& \hat{X}_{k,t}(1) / \hat{\beta} \nonumber\\
- \hat{\boldsymbol{b}}_{t,k}&=& \hat{\boldsymbol{b}}(1) \hat{\beta} \nonumber\\
+ \hat{\boldsymbol{b}}&=& \hat{\boldsymbol{b}}(1) \hat{\beta} \nonumber\\
    \hat{\tau}_{k}^2&=& \hat{\tau}_{k}^2(1)/ \hat{\beta}^2\nonumber\\
@@
 \label{syntheticData}
+
+The goal in this section is to evaluate the ability of the SAC-procedure to capture true parameter values.
 The synthetic data is not generated directly from the model for two reasons: (i) showing good performance on a dataset directly generated from the model assumptions is hardly any news, and (ii) generating data from the dynamic model description does not produce well-calibrated hidden states.
 The latter is important for our interpretation of the hidden process as a sharp and well-calibrated version of the crowd belief.
-To ensure that the hidden process for question $k$ is well-calibrated, generate a path of the standard Brownian motion until time $T_k$.
-If $Z_{t,k}$ denotes the value of the path at time $t$, then
+
+To ensure that the hidden process for question $k$ is well-calibrated, generate a path of the standard Brownian motion until time $T_k$. If $Z_{t,k}$ denotes the value of the path at time $t$, then
 \begin{eqnarray*}
@@
 gives a sequence of $T_k$ calibrated logit-probabilities for the event $Z_k = 1$.
-Repeating this procedure for $K$ times, we generate the hidden process for $K$ questions, each with a time horizon of $T_k = 101$.
+Using this procedure we generate the hidden process for $K$ questions, each with a time horizon of $T_k = 101$.
 The questions involve 25 experts allocated evenly among five expertise groups.
-Since the event resolves at time $t = 101$, the expert logit-probabilities $Y_{i,t,k}$ are only given for times $t = 1$ to $100$.
-These are generated from Equation (\ref{observedpr}) by applying bias and noise to the hidden process $X_{t,k}$.
-The simulation then iterates over a three-dimensional grid of parameter values
+Each expert gives one probability forecast per day with the exception of time $t = 101$, when the event resolves.
+These are generated by applying bias and noise to the hidden process $X_{t,k}$ for $t = 1, \dots, 100$ (see Equation (\ref{observedpr})).
+In this fashion, the simulation iterates over a three-dimensional grid of parameter values:
 \begin{eqnarray*}
@@
 \end{eqnarray*}
-where $\beta$ gives the value of the bias vector by letting $\boldsymbol{b} = [1/2, 3/4, 1, 4/3, 2/1]^T \beta$.
-Each grid point is used  40 times to generate a synthetic dataset.
-This dataset is then used to assess how accurately \textit{Sample-And-Calibrate} estimates the hidden process and bias vector.
-The accuracy of the estimated hidden process is measured with the average quadratic loss in the probability space, $\sum_{k=1}^K \sum_{t=1}^{T_k} ( \logit^{-1}(\hat{X}_{t,k}) - \logit^{-1}(X_{t,k}))^2 / \sum_{k=1}^K T_k$, and the accuracy of the estimated bias vector is measured with the average quadratic loss, $|| \boldsymbol{b} - \hat{\boldsymbol{b}}||^2/5$.
+where $\beta$ gives the bias vector by letting $\boldsymbol{b} = [1/2, 3/4, 1, 4/3, 2/1]^T \beta$.
+Every grid point is used  40 times to generate a synthetic dataset.
+The SAC-procedure is run on each dataset for 200 iterations of which the first 100 are used for burn-in.
+The accuracy of the estimated hidden process and bias vector are measured in  with the average quadratic loss in the probability space, $\sum_{k=1}^K \sum_{t=1}^{100} ( \logit^{-1}(\hat{X}_{t,k}) - \logit^{-1}(X_{t,k}))^2 / (100K)$, and the average quadratic loss, $|| \log(\boldsymbol{b}) - \log(\hat{\boldsymbol{b}})||^2/5$, respectively.
+The latter is measured in the log-space to allow a fair comparison of any discrepancies between under- and over-confident bias-vectors.
 
 \begin{figure}[h!]
-\vspace{-5em}
+\vspace{-4.5em}
 \centering
- \hspace{-28em}	\includegraphics{Figures/LegendSynthetic} % requires the graphicx package
-\vspace{-0.5em}
+ \hspace{-31em}	\includegraphics{Figures/LegendSynthetic} % requires the graphicx package
+\vspace{-1.5em}
 
@@
 
-\caption[Optional caption for list of figures]{Comparing \textit{Sample-And-Calibrate} that optimizes over the logarithmic score ($\text{SAC}_{\text{LOG}}$) and \textit{Sample-And-Calibrate} that optimizes over the Brier score (SAC-Bri) under synthetic data.
+\caption[Optional caption for list of figures]{Comparing SAC that optimizes over the logarithmic score ($\text{SAC}_{\text{LOG}}$) and SAC that optimizes over the Brier score ($\text{SAC}_{\text{BRI}}$) under synthetic data.
 The top row presents the accuracy to capture the hidden process.
@@
 Figure \ref{Synthetic} summarizes the results in a series of plots.
-The first row shows the marginal effects of the three grid variables on the accuracy to which the \textit{Sample-And-Calibrate} under the logarithmic loss ($\text{SAC}_{\text{LOG}}$) and the Brier score (SAC-Bri) are able to estimate the hidden process.
-The second row is similar in nature but measures the accuracy of estimating the bias vector instead.
-Since the plots were computed by choosing one grid variable (e.g.
-$\beta$) at a time and averaging over the remaining two variables  (e.g.
-$K$, and $\sigma^2$), each value in the plots represents an average of a $40 \times 5^2 = 1,000$ values.
-
-Based on the first row of plots, $\text{SAC}_{\text{LOG}}$ and SAC-Bri estimate the hidden process to a similar accuracy.
+The first row presents the marginal effects of the three grid variables on the accuracy to which the SAC-procedure estimates the hidden process.
+The second row is similar in nature but measures the accuracy in estimating the bias vector instead.
+Since the plots were computed by choosing one grid variable (e.g., $\beta$) at a time and averaging over the remaining two variables  (e.g., $K$ and $\sigma^2$), each value in the plots represents an average of a $40 \times 5^2 = 1,000$ values.
+Based on these results, both $\text{SAC}_{\text{LOG}}$ and  $\text{SAC}_{\text{BRI}}$ estimate the hidden process and the bias vector very accurately.
+Apart from $\text{SAC}_{\text{LOG}}$ outperforming $\text{SAC}_{\text{BRI}}$ uniformly across all parameter values, the two approaches behave very similarly.
+They make great use of the increasing number of questions, suffer slightly from the increasing level of noise in the expert logit-probabilities, and perform better when $\beta$ is large.
+
+%The latter can be seen as a side-effect of measuring the accuracy an estimated multiplicative factor in the absolute scale. The effect is actually
+
+%
+%Since the hidden state is not explicitly engineered to maximize sharpness subject to calibration, the calibration step in our procedure tends to overestimate $\beta$; that is, the estimated hidden process tends to be a bit sharper (more extreme) than the actual hidden process.
+%
+%When $\beta$ is large most of the experts are over-confident.
+As a result, the sampling step produces an estimated hidden process that is too extreme.
+This estimate therefore must be pulled back towards zero by the calibration step.
+If the
+%
+%
+%The story for the estimating the bias vector is a bit different: the discrepancy between the estimate and the true bias vector may be the same in the log-scale when the $\beta$ is small versus when $\beta$ is large.
+This, however, translates into much smaller differences in the absolute scale that the differences are ultimately measured in our case.
+For instance, estimating 1/2 with 3/4 is the same as estimating 2 with 4/3.
+The difference in the first one is 0.25 while in the second 0.6666667.
+To make this even worse, the distance is squared which emphasizes large differences.
+
+%
+%Based on the first row of plots, $\text{SAC}_{\text{LOG}}$ and $\text{SAC}_{\text{BRI}}$ estimate the hidden process to a similar accuracy.
 Even though some relative differences are visible, the loss values are very small in absolute terms, hence suggesting good overall accuracy to capture the hidden process.
@@
 Increasing the number of questions improves the estimation accuracy but only very little.
-
-Based on the second row of plots, $\text{SAC}_{\text{LOG}}$ and SAC-Bri  capture the bias vector rather precisely as long as the true values of $\beta$ and $\sigma^2$ are not very large.
+%
+%
+%Based on the second row of plots, $\text{SAC}_{\text{LOG}}$ and $\text{SAC}_{\text{BRI}}$  capture the bias vector rather precisely as long as the true values of $\beta$ and $\sigma^2$ are not very large.
 This behavior can be again considered as being caused by the truncation of the extreme expert logit-probabilities.
-Surprisingly, even though $\text{SAC}_{\text{LOG}}$ is more accurate in estimating the bias vector, it is unable to outperform SAC-Bri in estimating the hidden process.
+Surprisingly, even though $\text{SAC}_{\text{LOG}}$ is more accurate in estimating the bias vector, it is unable to outperform $\text{SAC}_{\text{BRI}}$ in estimating the hidden process.
+%
+%
+%
 
@@
 
-\item \textit{The Sample-And-Calibrate procedure both under the Brier (SAC-Bri) and the logarithmic score ($\text{SAC}_{\text{LOG}}$)}.
+\item \textit{The Sample-And-Calibrate procedure both under the Brier ($\text{SAC}_{\text{BRI}}$) and the logarithmic score ($\text{SAC}_{\text{LOG}}$)}.
 The model is first fit on the training set by running the sampling step for 3,000 iterations of which the first 500 iterations are used for burn-in.
@@
 
-Overall, SAC-Bri and $\text{SAC}_{\text{LOG}}$ achieve the lowest average scores across all columns except \textit{Short}, where they are slightly outperformed by $\text{BSAC}_{\text{LOG}}$.
+Overall, $\text{SAC}_{\text{BRI}}$ and $\text{SAC}_{\text{LOG}}$ achieve the lowest average scores across all columns except \textit{Short}, where they are slightly outperformed by $\text{BSAC}_{\text{LOG}}$.
 $\text{BSAC}_{\text{LOG}}$, however, turns out to be overconfident (see Section \ref{calibration}).
@@
 \caption[Optional caption for list of figures]{The top and bottom rows show in- and out-of-sample calibration and sharpness, respectively.
-The models are Simple Dynamic Linear Model (SDLM), the \textit{Sample-And-Calibrate} approach that optimizes over the logarithmic score ($\text{SAC}_{\text{LOG}}$), the \textit{Sample-And-Calibrate} approach that optimizes over the Brier score (SAC-Bri), and the fully-Bayesian version of the \textit{Sample-And-Calibrate} approach that optimizes over the logarithmic score ($\text{BSAC}_{\text{LOG}}$).
+The models are Simple Dynamic Linear Model (SDLM), the \textit{Sample-And-Calibrate} approach that optimizes over the logarithmic score ($\text{SAC}_{\text{LOG}}$), the \textit{Sample-And-Calibrate} approach that optimizes over the Brier score ($\text{SAC}_{\text{BRI}}$), and the fully-Bayesian version of the \textit{Sample-And-Calibrate} approach that optimizes over the logarithmic score ($\text{BSAC}_{\text{LOG}}$).
 }
@@
 
-The top and bottom rows of Figure \ref{Calibration-Out} present calibration plots for SDLM, $\text{SAC}_{\text{LOG}}$, SAC-Bri, and $\text{BSAC}_{\text{LOG}}$ under in- and out-of-sample probability estimation, respectively.
+The top and bottom rows of Figure \ref{Calibration-Out} present calibration plots for SDLM, $\text{SAC}_{\text{LOG}}$, $\text{SAC}_{\text{BRI}}$, and $\text{BSAC}_{\text{LOG}}$ under in- and out-of-sample probability estimation, respectively.
 Each setting is of interest in its own right:  Good in-sample calibration is crucial for model interpretability.

### 2013-09-06 paper.tex
Edits: plots, annals of applied statistics format, tables, etc.

@@
 \documentclass[aoas, preprint]{imsart}
-\RequirePackage{natbib}
 
@@
 %\usepackage[section]{placeins}
-\usepackage{amsmath} 
+%\usepackage{amsmath} 
 \usepackage{times}
@@
 \usepackage{amssymb}
+\usepackage{array}
 \usepackage{url}
@@
 %\usepackage{algorithm}
-\usepackage{algorithmic}
+%\usepackage{algorithmic}
 %\usepackage{cite}
@@
 %\usepackage[margin=1.2in]{geometry}
-\usepackage{geometry}
-\usepackage{fix-cm}
+%\usepackage{geometry}
+%\usepackage{fix-cm}
 \usepackage{subfigure}
@@
 \usepackage{color} 
-\usepackage{natbib}
+%\usepackage{natbib}
 \usepackage{xcolor,xparse}
@@
 %\usepackage{setspace}
-\usepackage{booktabs}
+%\usepackage{booktabs}
 %\usepackage{subfig}
 %\usepackage{caption, subcaption}
+\usepackage[OT1]{fontenc}
+\usepackage[colorlinks,citecolor=blue,urlcolor=blue]{hyperref}
+\usepackage{natbib}
+
+\startlocaldefs
+\numberwithin{equation}{section}
+\theoremstyle{plain}
+\newtheorem{thm}{Theorem}[section]
+\endlocaldefs
 
@@
 
-\begin{table}
-\begin{center}
-\begin{tabular}{c c c c c c c c }
+\begin{table*}
+\caption{Five-number summaries of our real-world data.}
+\label{DataStats}
+\begin{tabular}{m{2in}	lllllll }
 \hline
@@
 \end{tabular}
-\caption{Five-number summaries of our real-world data.
-%: \textit{\# of Days Active} represents the number of days that the experts were given to update their forecast before the correct answer was revealed; \textit{\# of Experts per Problem } denotes the number of experts participating in a question; \textit{\# Forecasts given by each Expert on a Problem} gives the number of forecasts for a question made by an active expert; \textit{\# Problems participated by an Expert} represents the number of different questions a given expert participated in.
-}
-\label{DataStats}
-\end{center}
-\end{table}
-
-\begin{table}
+\end{table*}
+
+\begin{table*}
+\caption{Frequencies of the self-reported expertise (1 = Not At All Expert and 5 = Extremely Expert) levels across all the 166 questions in our real-world data.}
+\label{ExpertiseTable}
 \begin{center}
-\begin{tabular}{c c c c c c c c }
+\begin{tabular}{llllllll }
 \hline
@@
 \end{tabular}
-\caption{Frequencies of the self-reported expertise (1 = Not At All Expert and 5 = Extremely Expert) levels across all the 166 questions in our real-world data.}
-\label{ExpertiseTable}
 \end{center}
-\end{table}
+\end{table*}
 
@@
 \centering
-\hspace*{0em} 	\includegraphics{Figures/LegendExamplePlot} % requires the graphicx package
+\hspace*{-3.5em} 	\includegraphics{Figures/LegendExamplePlot} % requires the graphicx package
 \vspace{-3.5em}
@@
 \subfigure{
-\hspace{-1em} \includegraphics{Figures/SyntheticSigma2}
+\hspace{-1em} \includegraphics[width = 0.32\textwidth]{Figures/SyntheticSigma2}
 \label{SyntheticSigma2}
@@
 \subfigure{
-\hspace{-1.5em} \includegraphics{Figures/SyntheticBeta}
+\hspace{-1.5em} \includegraphics[width = 0.32\textwidth]{Figures/SyntheticBeta}
 \label{SyntheticBeta}
@@
 \subfigure{
-\hspace{-1.5em} \includegraphics{Figures/SyntheticK}
+\hspace{-1.5em} \includegraphics[width = 0.32\textwidth]{Figures/SyntheticK}
 \label{SyntheticK}
@@
 \subfigure{
-\hspace{-1em}  \includegraphics{Figures/SyntheticSigma2bias}
+\hspace{-1em}  \includegraphics[width = 0.32\textwidth]{Figures/SyntheticSigma2bias}
 \label{SyntheticSigma2b}
@@
 \subfigure{
-\hspace{-1.5em}  \includegraphics{Figures/SyntheticBetabias}
+\hspace{-1.5em}  \includegraphics[width = 0.32\textwidth]{Figures/SyntheticBetabias}
 \label{SyntheticBetab}
@@
 \subfigure{
-\hspace{-1.5em}   \includegraphics{Figures/SyntheticKbias}
+\hspace{-1.5em}   \includegraphics[width = 0.32\textwidth]{Figures/SyntheticKbias}
 \label{SyntheticKb}
@@
 
-\begin{table}[htbp]
+\begin{table*}
+   \caption{Brier Scores based on 10-fold cross-validation.
+\textit{Scores by Day} weighs a question by the number of days the question remained open.
+\textit{Scores by Problem} gives each question an equal weight regardless how long the question remained open.
+The bolded values indicate the lowest scores in each column.
+The values in the parenthesis represent standard errors in the scores.
+}
+   \label{prediction}
    \centering
       \begin{tabular}{l llll} % Column formatting, @{} suppresses leading/trailing space
+      \hline
  \multicolumn{5}{ c }{Scores by Day}\\
@@
 EWMA & 0.111 (0.142) & 0.089 (0.100) & 0.111 (0.136) & 0.112 (0.144)\\ 
-&&&&\\
+\hline
 \multicolumn{5}{ c }{Scores by Problem}\\
@@
 EWMA & 0.104 (0.105) & 0.092 (0.081) & 0.119 (0.125) & 0.103 (0.107)\\ 
+\hline
    \end{tabular}
-   \caption{Brier Scores based on 10-fold cross-validation.
-\textit{Scores by Day} weighs a question by the number of days the question remained open.
-\textit{Scores by Problem} gives each question an equal weight regardless how long the question remained open.
-The bolded values indicate the lowest scores in each column.
-The values in the parenthesis represent standard errors in the scores.
-}
-   \label{prediction}
-\end{table}
+\end{table*}
 
@@
 \begin{figure}[h!]
-\newgeometry{left=5cm, right = 1cm, bottom=0.1cm}
+%\newgeometry{left=5cm, right = 1cm, bottom=0.1cm}
 
 \subfigure[\scriptsize SDLM In-Sample]{
-\hspace*{-3em}  \includegraphics{Figures/CalibrationNONE10}
+\hspace*{-2em}  \includegraphics[width = 0.286\textwidth]{Figures/CalibrationNONE10}
 \label{CalibrationNONEa}
@@
 \subfigure[\scriptsize $\text{SAC}_{\text{LOG}}$ In-Sample]{
-\hspace*{-1.15em} \includegraphics{Figures/CalibrationLOG10}
+\hspace*{-1.15em} \includegraphics[width = 0.235\textwidth]{Figures/CalibrationLOG10}
 \label{CalibrationPLATTLOG10}
@@
 \subfigure[\scriptsize $\text{SAC}_{\text{BRI}}$ In-Sample]{
-\hspace*{-1.15em} \includegraphics{Figures/CalibrationBRI10}
+\hspace*{-1.15em} \includegraphics[width = 0.235\textwidth]{Figures/CalibrationBRI10}
 \label{CalibrationPLATTBRI10}
@@
 \subfigure[\scriptsize $\text{BSAC}_{\text{LOG}}$ In-Sample]{
-\hspace*{-1.15em} \includegraphics{Figures/CalibrationFULL10}
+\hspace*{-1.15em} \includegraphics[width = 0.286\textwidth]{Figures/CalibrationFULL10}
 \hspace*{-2em} \label{CalibrationPLATTFULL10}
@@
 \subfigure[ \scriptsize SDLM Out-of-Sample]{
-\hspace*{-3em} \includegraphics{Figures/CalibrationNONE}
+ \hspace*{-2em} \includegraphics[width = 0.286\textwidth]{Figures/CalibrationNONE}
 \label{CalibrationNONEb}
 }
-\subfigure[ \scriptsize $\text{SAC}_{\text{LOG}}$ Out-of-Sample]{
-\hspace*{-1.15em} \includegraphics{Figures/CalibrationSTO-Log}
+\subfigure[\scriptsize $\text{SAC}_{\text{LOG}}$ Out-of-Sample]{
+\hspace*{-1.15em} \includegraphics[width = 0.235\textwidth]{Figures/CalibrationSTO-Log}
 \label{CalibrationSTC-Log}
@@
 \subfigure[ \scriptsize $\text{SAC}_{\text{BRI}}$ Out-of-Sample]{
-\hspace*{-1.15em} \includegraphics{Figures/CalibrationSTO-Bri}
+\hspace*{-1.15em} \includegraphics[width = 0.235\textwidth]{Figures/CalibrationSTO-Bri}
 \label{CalibrationSTC-Bri}
@@
 \subfigure[\scriptsize  $\text{BSAC}_{\text{LOG}}$ Out-of-Sample]{
-\hspace*{-1.15em} \includegraphics{Figures/CalibrationSO}
+\hspace*{-1.15em} \includegraphics[width = 0.286\textwidth]{Figures/CalibrationSO}
 \hspace*{-2em}\label{CalibrationBSTC-Log}
 }
-\restoregeometry
+%\restoregeometry
 \caption[Optional caption for list of figures]{The top and bottom rows show in- and out-of-sample calibration and sharpness, respectively.
@@
 \vspace*{-1em} 
-\includegraphics{Figures/BiasesBoxplots}
+%\includegraphics[width = \textwidth]{Figures/BiasesBoxplots}
+\includegraphics[width = \textwidth]{Figures/BiasesBoxplots}
 \label{BiasesLOGBoxplots}
@@
 \begin{figure}[h!]
-\vspace{-1em}
+\vspace{-2em}
 \centering
-\hspace*{0em} 	\includegraphics{Figures/LegendExamplePlot} % requires the graphicx package
+\hspace*{-3.5em} 	\includegraphics{Figures/LegendExamplePlot} % requires the graphicx package
 \vspace{-3.5em}
@@
 \subfigure[\textit{Will the expansion of the European bailout fund be ratified by all 17 Eurozone nations before 1 November 2011?}]{
-\hspace*{-1em}  \includegraphics{Figures/ExamplePlot1b}
+\hspace*{-1em}  \includegraphics[width= 0.49\textwidth]{Figures/ExamplePlot1b}
 \label{Example1b}
@@
 \subfigure[\textit{Will the Nikkei 225 index finish trading at or above 9,500 on 30 September 2011?}]{
-\hspace*{-1em}  \includegraphics{Figures/ExamplePlot2b}
+\hspace*{-1em}  \includegraphics[width= 0.49\textwidth]{Figures/ExamplePlot2b}
 \label{Example2b}
@@
 For instance, this paper modeled the bias separately within each expertise group.
-This, however, is by no means restricted to the study of bias or its relation to self-reported expertise.
+This is by no means restricted to the study of bias or its relation to self-reported expertise.
 Different parameter dependencies could be constructed based on many other expert characteristic, such as gender, education, or specialty, to produce a range of novel insight on the forecasting behavior across different groups of experts.
@@
 This would be of interest to the decision-maker who could use the information as a basis for consulting only a high-performing subset of the available experts.
+
+%It may also be of interest to study hidden processes with other goals besides maximizing sharpness subject to calibration.
+This can be easily achieved by changing the optimization criterion in Equation (\ref{OSE}) to an appropriate function that meets the new goal.
 

### 2013-09-12 paper.tex
Edits Whatever...

@@
 % "Title of the paper"
-\title{Probability Aggregation: Dynamic Hierarchical Modeling of Sparse Expert Beliefs}
+\title{Probability Aggregation in Time-Series: Dynamic Hierarchical Modeling of Sparse Expert Beliefs}
 \runtitle{Dynamic Hierarchical Modeling of Sparse Expert Beliefs}
@@
 \author{\fnms{Shane T.} \snm{Jensen}\ead[label=e3]{stjensen@wharton.upenn.edu}},
-\author{\fnms{Lyle H.} \snm{Ungar}\ead[label=e2]{ungar@cis.upenn.edu}},
 \author{\fnms{Barbara A.} \snm{Mellers}\ead[label=e4]{mellers@wharton.upenn.edu}},
+\author{\fnms{Phil E.} \snm{Tetlock}\ead[label=e5]{tetlock@wharton.upenn.edu}},
 \and
-\author{\fnms{Phil E.} \snm{Tetlock}\ead[label=e5]{tetlock@wharton.upenn.edu}}
+\author{\fnms{Lyle H.} \snm{Ungar}\ead[label=e2]{ungar@cis.upenn.edu}}
 
@@
  \affiliation{Department of Statistics, The Wharton School of the University of Pennsylvania\\   \printead{e1,e3}}
+ \affiliation{Department of Psychology, University of Pennsylvania\\ \printead{e4,e5}}
  \affiliation{Department of Computer and Information Science, University of Pennsylvania\\ \printead{e2}}
- \affiliation{Department of Psychology, University of Pennsylvania\\ \printead{e4,e5}}
  \address{Philadelphia, PA 19104- 6340, USA\\  \printead{e1}}
@@
 Most subjective probability aggregation procedures use a single probability judgement from each expert, even though it is common for experts studying real problems to update their probability estimates over time.
-This paper makes an advance towards unexplored areas of probability aggregation by considering a dynamic context in which experts are allowed to update their beliefs at random intervals.
-The updates are assumed to occur very infrequently, resulting into a highly sparse dataset that cannot be modeled by standard time-series procedures.
+This paper advances into unexplored areas of probability aggregation by considering a dynamic context in which experts can update their beliefs at random intervals.
+The updates occur very infrequently, resulting in a highly sparse dataset that cannot be modeled by standard time-series procedures.
 In response to the lack of appropriate methodology, this paper presents a hierarchical model that takes into account the expert's level of self-reported expertise and produces aggregate probabilities that are sharp and well-calibrated both in- and out-of-sample.
-The model is demonstrated on a real-world dataset that includes over 2,300 experts making multiple probability forecasts on different subsets of 166 international political events.
+The model is demonstrated on a real-world dataset that includes over 2,300 experts making multiple probability forecasts over a period of two years on different subsets of 166 international political events.
 \end{abstract}
@@
 Individual experts can differ radically from one another in their abilities to assess probabilities of future events.
-Their probability assessments are often evaluated and compared in regards to \textit{calibration} that measures how closely the frequency of occurrence of the events agree with the given probabilities.
-For instance, the proportion of occurred events is 60\% for all those events for which a well-calibrated expert assessed a probability of 0.60.
+Their probability assessments are often evaluated and compared on \textit{calibration}, which measures how closely the frequency of event occurrence agrees with the assigned probabilities.
+For instance, the proportion of occurrences is 60\% for all those events to which a well-calibrated expert assigned a probability of 0.60.
 Even though several experiments have been conducted to show that experts are generally poorly calibrated (see, e.g., \citet{cooke1991experts, shlyakhter1994quantifying}), relative differences can occur among different types of experts.
@@
 
-Calibration by itself, however, is not sufficient for a useful probability assessment.
-To see this, consider a relatively stationary process, such as rain on different days in some geographic region, whose empirical frequency of occurrence in the last 10 years is 45\%.
-In this setting an expert can assess a constant probability of 0.45 and be well-calibrated.
+Calibration by itself, however, is not sufficient for useful probability estimation.
+Consider a relatively stationary process, such as rain on different days in a given geographic region, where the observed frequency of occurrence in the last 10 years is 45\%.
+In this setting an expert could always  assign a constant probability of 0.45 and be well-calibrated.
 This assessment, however, can be made without any subject-matter expertise or special training.
 For this reason the long-term frequency is often considered as the baseline probability -- a naive assessment that provides the decision-maker very little extra information.
-Therefore a consulting expert should aim to make probability assessments that are as far from the baseline as possible.
-The extent to which his probabilities differ from the baseline is measured with an important attribute called \textit{sharpness} (\citet{gneiting2008rejoinder, winkler2008comments}).
-If the expert is both sharp and well-calibrated, he is able to forecast the behavior of the process with high certainty and accuracy.
-From the decision-maker's perspective his advice is highly useful as it involves much information and very little risk.
+Therefore experts should aim to make probability assessments that are as far from the baseline as possible.
+The extent to which their probabilities differ from the baseline is measured by \textit{sharpness} (\citet{gneiting2008rejoinder, winkler2008comments}).
+If the experts are both sharp and well-calibrated, they can forecast the behavior of the process with high certainty and accuracy.
+From the decision-maker's perspective their advice is highly useful as it involves much information and very little risk.
 Therefore, to summarize, useful probability estimation should aim to maximize sharpness subject to calibration (see, e.g., \citet{raftery2005using, murphy1987general}).
@@
 
-One such observation is related to the aggregation of multiple probabilities.
-There is strong empirical evidence that bringing together the strengths of different experts by combining their probability forecasts into a single consensus, known as the \textit{crowd belief},  results in better predictive performance.
-Being motivated by the long list of applications of probability forecasts, including medical diagnosis (\citet{wilson1998prediction, pepe2003statistical}, political and socio-economic foresight (\citet{tetlock2005expert}), and meteorology (\citet{sanders1963subjective, vislocky1995improved, baars2005performance}), researchers have proposed many different approaches to combining probability forecasts (see, e.g., \citet{Ranjan08, satopaa} for some recent studies, and \citet{Genest, Wallsten97evaluatingand, clemen2007aggregating, primo2009calibration} for a comprehensive overview).
+%One such observation is related to the aggregation of multiple probabilities. 
+There is strong empirical evidence that bringing together the strengths of different experts by combining their probability forecasts into a single consensus, known as the \textit{crowd belief},  improves predictive performance.
+Prompted by the many applications of probability forecasts, including medical diagnosis (\citet{wilson1998prediction, pepe2003statistical}, political and socio-economic foresight (\citet{tetlock2005expert}), and meteorology (\citet{sanders1963subjective, vislocky1995improved, baars2005performance}), researchers have proposed many approaches to combining probability forecasts (see, e.g., \citet{Ranjan08, satopaa} for some recent studies, and \citet{Genest, Wallsten97evaluatingand, clemen2007aggregating, primo2009calibration} for a comprehensive overview).
 The general focus, however, has been on developing one-time aggregation procedures that consult the expert's advice only once before the event resolves.
@@
 Consequently, many areas of probability aggregation still remain rather unexplored.
-For instance, consider an investor aiming to assess whether a stock index will finish trading above some threshold on a given date.
-In order to maximize his overall predictive accuracy, he may consult a group of experts repeatedly over a period of time and adjust his estimate of the aggregate probability accordingly.
+For instance, consider investors aiming to assess whether a stock index will finish trading above a threshold on a given date.
+To maximize their overall predictive accuracy, they may consult a group of experts repeatedly over a period of time and adjust their estimate of the aggregate probability accordingly.
 Since the experts are allowed to update their probability assessments, the aggregation should be performed by taking into account the temporal correlation in their advice.
@@
 The problem at hand is therefore strikingly different from many time-series estimation problems, where one has an observation at every time point -- or almost every time point.
-As a result, standard time-series procedures like ARIMA (see, e.g., \citet{mills1991time}) are not directly applicable to our setting.
-In response to the lack of appropriate modeling procedures, this paper introduces a time-series model that incorporates self-reported expertise and captures a sharp and well-calibrated estimate of the crowd belief.
+As a result, standard time-series procedures like ARIMA (see, e.g., \citet{mills1991time}) are not directly applicable.
+This paper introduces a time-series model that incorporates self-reported expertise and captures a sharp and well-calibrated estimate of the crowd belief.
 The model is highly interpretable and can be used for:
@@
 \item accurate probability forecasts, and
-\item many question-specific quantities that have easy interpretations and can be used to gain novel insight in the social sciences. 
+\item many question-specific quantities, such as expert disagreement and problem difficulty, that have easy interpretations and can be used to gain novel insight in the social sciences. 
 \end{itemize}
 
-The paper begins with a description of our probability forecasting data that were collected by asking over 2,300 experts to give probability forecasts and to self-assess their level of expertise on a subset of 166 geopolitical binary events.
-After summarizing the dataset, the paper introduces a dynamic hierarchical model for capturing the crowd belief.
+This paper begins by describing our geopolitical database.
+The paper then introduces a dynamic hierarchical model for capturing the crowd belief.
 The model is estimated in a two-step procedure: first, a sampling step produces constrained parameter estimates via Gibbs sampling (see \citet{geman1984stochastic} for the original introduction of Gibbs sampling); second, a calibration step transforms these estimates to their unconstrained equivalents via a one-dimension optimization procedure.
 An extension of this model to polychotomous outcomes is briefly discussed before model evaluation.
-The first evaluation section uses synthetic data to study the accuracy to which the two-step procedure is able to estimate parameter values.
-The second evaluation section illustrates different aspects of the model by applying it to our real-world probability forecasting data.
+The first evaluation section uses synthetic data to study how accurately the two-step procedure can estimate parameter values.
+The second evaluation section applies the model to real-world forecasting data.
 The paper concludes with a discussion on future research directions and model limitations.
@@
 The experts were asked to give probability forecasts (to the second decimal point) and to self-assess their level of expertise (on a 1-to-5 scale with 1 = Not At All Expert and 5 = Extremely Expert) on a number of 166 geopolitical binary events taking place between September 29, 2011 and May 8, 2013.
-Each question was active for a period of time during which the participating experts were allowed to update their forecasts as frequently as they liked without being penalized.
+Each question was active for a period during which the participating experts could update their forecasts as frequently as they liked without penalty.
 The experts knew that their probability estimates would be assessed for accuracy using Brier scores\footnote{The Brier score is the squared distance between the probability forecast and the event indicator that equals 1.0 or 0.0 depending on whether the event happened or not, respectively.
@@
 The average group-level response rate was around 13.5 forecasts per day.
-Given that the group of experts is large and diverse, the data is very sparse.
+Given that the group of experts is large and diverse, the data are very sparse.
 Tables \ref{DataStats} and \ref{ExpertiseTable} provide more relevant summary statistics on the data.
@@
 Given that the European bailout fund was ratified before November 1, 2011 and that the Nikkei 225 index finished trading at around 8,700 on September 30, 2011, the general trend of the probability forecasts tends to converge towards the correct answers.
-The individual experts, however, disagree rather strongly, with the disagreement persisting even near the closing dates of the questions.
+The individual experts, however, sometimes disagree strongly, with the disagreement persisting even near the closing dates of the questions.
 
@@
 
-The hidden state $X_{t,k}$ represents the logit-probability for the $k$th event given all the information available up to and including time $t$.
+The hidden state $X_{t,k}$ represents the aggregate logit-probability for the $k$th event given all the information available up to and including time $t$.
 To make this more specific, let $Z_k \in \{0, 1\}$ indicate whether the event associated with the $k$th question happened $\left(Z_k = 1\right)$ or did not happen $\left(Z_k = 0\right)$.
@@
 
-Since the experts may believe in false information, hide their true beliefs for ethical bases, or be biased for many other reasons, their probability assessments should be aggregated via a model that can detect potential bias, separate signal from noise, and use the collective opinion to estimate $X_{t,k}$.
+Given that the experts may believe in false information, hide their true beliefs, or be biased for many other reasons, their probability assessments should be aggregated via a model that can detect potential bias, separate signal from noise, and use the collective opinion to estimate $X_{t,k}$.
 In our model the experts are assumed to be, on average, a multiplicative constant $\boldsymbol{b}$ away from $X_{t,k}$.
@@
 
- The \textit{random fluctuations} in the hidden process are measured in terms of $\tau^2_k$ and are assumed to represent changes or shocks to the underlying circumstances that ultimately decide the outcome of the event.
+ The \textit{random fluctuations} in the hidden process are measured by $\tau^2_k$ and are assumed to represent changes or shocks to the underlying circumstances that ultimately decide the outcome of the event.
 The \textit{systematic component} $\gamma_k$ allows the model to incorporate a constant signal stream that drifts the hidden process to infinity (when $\gamma_k \in (1, \infty)$) or zero (when $\gamma_k \in (0, 1)$).
@@
 The main challenge is to capture a well-calibrated estimate of the hidden process without sacrificing the interpretability of our model.
-This section introduces a two-step procedure, called \textit{Sample-And-Calibrate} (SAC), that achieves this goal in a flexible and efficient manner: the first step estimates the model parameters under a constraint (\textit{Sampling Step}), and the second step performs a one-dimension optimization procedure to transform the constrained estimates into their unconstrained counterparts (\textit{Calibration Step}).
+This section introduces a two-step procedure, called \textit{Sample-And-Calibrate} (SAC), that achieves this goal in a flexible and efficient manner: the first step estimates the model parameters under a constraint (\textit{Sampling Step}), and the second step performs a one-dimension optimizational procedure to transform the constrained estimates into their unconstrained counterparts (\textit{Calibration Step}).
 
@@
 The goal in this section is to evaluate the ability of the SAC-procedure to capture true parameter values.
-The synthetic data is not generated directly from the model for two reasons: (i) showing good performance on a dataset directly generated from the model assumptions is hardly any news, and (ii) generating data from the dynamic model description does not produce well-calibrated hidden states.
+The synthetic data are not generated directly from the model for two reasons: (i) showing good performance on a dataset directly generated from the model assumptions is hardly any news, and (ii) generating data from the dynamic model description does not produce well-calibrated hidden states.
 The latter is important for our interpretation of the hidden process as a sharp and well-calibrated version of the crowd belief.
@@
 The SAC-procedure is run on each dataset for 200 iterations of which the first 100 are used for burn-in.
-The accuracy of the estimated hidden process and bias vector are measured in  with the average quadratic loss in the probability space, $\sum_{k=1}^K \sum_{t=1}^{100} ( \logit^{-1}(\hat{X}_{t,k}) - \logit^{-1}(X_{t,k}))^2 / (100K)$, and the average quadratic loss, $|| \log(\boldsymbol{b}) - \log(\hat{\boldsymbol{b}})||^2/5$, respectively.
+The accuracy of the estimated hidden process and bias vector are measured with the average quadratic loss in the probability space, $\sum_{k=1}^K \sum_{t=1}^{100} ( \logit^{-1}(\hat{X}_{t,k}) - \logit^{-1}(X_{t,k}))^2 / (100K)$, and the average quadratic loss, $|| \log(\boldsymbol{b}) - \log(\hat{\boldsymbol{b}})||^2/5$, respectively.
 The latter is measured in the log-space to allow a fair comparison of any discrepancies between under- and over-confident bias-vectors.
@@
 For instance, all the 166 questions in our dataset contained both a zero and a one.
-Transforming such forecasts into the logit-space yields infinities that are highly influential and cause problems in model estimation.
+Transforming such forecasts into the logit-space yields infinities that can cause problems in model estimation.
 To avoid this, \citet{Ariely00theeffects} suggest changing $p = 0.00$ and $1.00$ to $p = 0.02$ and $0.98$, respectively.
@@
 \citet{Geo}, on the other hand, consider only probabilities that fall within a constrained interval, say $[0.001, 0.999]$, and discard the rest.
-Since this implies ignoring a portion of the data, we adopt the first approach by truncating $p = 0.00$ and $1.00$ to $p = 0.01$ and $0.99$, respectively.
+Since this implies ignoring a portion of the data, we adopt an approach similar to \citet{Ariely00theeffects}  by truncating $p = 0.00$ and $1.00$ to $p = 0.01$ and $0.99$, respectively.
 Our results remain insensitive to the exact choice of truncation as long as this is done in a reasonable manner to keep the extreme probabilities from becoming highly influential in the logit-space.
@@
 The second matter is related to the distribution of the class labels in the data.
-If the set of occurred events is much larger than the set of events that did not occur (or \textit{vice versa}), the dataset is called \textit{imbalanced}.
+If the set of occurrences is much larger than the set of non-occurrences (or \textit{vice versa}), the dataset is called \textit{imbalanced}.
 On such data the modeling procedure can end up over-focusing on the larger class, and as a result, give very accurate forecast performance over the larger class at the cost of performing poorly over the smaller class (see, e.g., \citet{chen2009learning, wallace2012class}).
@@
 The models are allowed to utilize a training set before making predictions on an independent testing set.
-In order to clarify some of the upcoming notation, let $S_{train}$ and $S_{test}$ be index sets that partition the data into training and testing sets of sizes $|S_{train}| = N_{train}$ and $|S_{test}| = 166 - N_{test}$, respectively.
+To clarify some of the upcoming notation, let $S_{train}$ and $S_{test}$ be index sets that partition the data into training and testing sets of sizes $|S_{train}| = N_{train}$ and $|S_{test}| = 166 - N_{test}$, respectively.
 This means that the $k$th question is in the training set if and only if $k \in S_{train}$.
@@
 where $X_{t,k}$ is the logit-probability used for prediction.
-Since this model is not hierarchical, estimates of the hidden process can be obtained directly for the questions in the testing set without fitting the model first on the training set.
-In order to make predictions, the sampler is run for 500 iterations of which the first 200 are used for burn-in.
+Given that this model is not hierarchical, estimates of the hidden process can be obtained directly for the questions in the testing set without fitting the model first on the training set.
+To make predictions, the sampler is run for 500 iterations of which the first 200 are used for burn-in.
 The remaining 300 iterations are thinned by discarding every other observation, leaving a final predictive sample of 150 observations.
@@
 \begin{eqnarray}
-p( \boldsymbol{Z} | \beta, \boldsymbol{X}(1)) &\propto& \prod_{k=1}^K \prod_{t=1}^{T_k}  \logit^{-1} \left(X_{k,t}(1)/\beta  \right)^{Z_k} \left( 1-  \logit^{-1} \left( X_{k,t}(1)/\beta  \right) \right)^{1-Z_k} \label{OSE2}
+p( \boldsymbol{Z} | \beta, \boldsymbol{X}(1)) \propto \prod_{k=1}^K \prod_{t=1}^{T_k} && \logit^{-1} \left(X_{k,t}(1)/\beta  \right)^{Z_k}\times \label{OSE2}\\ 
+&& \left( 1-  \logit^{-1} \left( X_{k,t}(1)/\beta  \right) \right)^{1-Z_k} \nonumber
 \end{eqnarray}
-Similarly to  \citet{gelman2003bayesian}, the prior is chosen to be locally uniform, $p(1/\beta) \propto 1$.
+As in \citet{gelman2003bayesian}, the prior is chosen to be locally uniform, $p(1/\beta) \propto 1$.
 Posterior estimates of $\beta$ can be sampled from Equation (\ref{OSE2}) using generic sampling algorithms such as the Metropolis algorithm (\citet{metropolis1953equation}) or slice sampling (\citet{neal2003slice}).
@@
 where
+\small
 \begin{eqnarray*}
@@
 \end{eqnarray*}
+\normalsize
 The vector $\boldsymbol{b}$ collects the bias terms of the different expertise groups.
@@
 As a result, the aggregate forecasts are either very close to the correct answer or very far from it.
-As can be seen in Table \ref{prediction}, this results into highly variable forecasting performance.
+As Table \ref{prediction} shows, this results into highly variable forecasting performance.
 The short questions, however, involve very little uncertainty and were generally the easiest to forecast.
@@
  Similarly behavior is exhibited by EWMA that performs the worst of all the competing models.
-The other two exponentially weighted aggregators, namely EWMLA and EWMBA, on other hand, make efficient use of the training set and present good forecasting performance in most columns of Table \ref{prediction}.
+In the contrary, the other two exponentially weighted aggregators, EWMLA and EWMBA, make efficient use of the training set and present good forecasting performance in most columns of Table \ref{prediction}.
 Recall that EWMBA uses the cumulative distribution function of the Beta distribution to transform the average forecast probability.
 This transformation depends on two parameters and is more flexible than the transformation used by EWMLA.
-On other hand, only EWLMA is given access to the self-reported expertise information.
+Only EWLMA is given access to the self-reported expertise information.
 Based on the results, however, it is unable to use this information to outperform EWMBA.

### 2013-09-14 paper.tex
Almost final

@@
 \title{Probability Aggregation in Time-Series: Dynamic Hierarchical Modeling of Sparse Expert Beliefs}
-\runtitle{Dynamic Hierarchical Modeling of Sparse Expert Beliefs}
+\runtitle{Probability Aggregation in Time-Series}
 
@@
 To maximize their overall predictive accuracy, they may consult a group of experts repeatedly over a period of time and adjust their estimate of the aggregate probability accordingly.
-Since the experts are allowed to update their probability assessments, the aggregation should be performed by taking into account the temporal correlation in their advice.
+Given that the experts are allowed to update their probability assessments, the aggregation should be performed by taking into account the temporal correlation in their advice.
 Many standard time-series procedures could be used to perform this aggregation as long as the experts update their advice consistently on a daily basis.
@@
 \begin{itemize}
-\item the analysis of group-level under- and overconfidence across different levels of self-reported expertise,
+\item borrowing strength across the hierarchy to analyze the under- and overconfidence in different groups of experts,
 %\item confidence bands for the crowd belief,
 \item accurate probability forecasts, and
-\item many question-specific quantities, such as expert disagreement and problem difficulty, that have easy interpretations and can be used to gain novel insight in the social sciences. 
+\item many question-specific quantities that have easy interpretations, such as expert disagreement and problem difficulty, and can be used to gain novel insight in the social sciences. 
 \end{itemize}
@@
 In addition to receiving \$150 for meeting minimum participation requirements that did not depend on prediction accuracy, the experts received status rewards for their performance via leader-boards displaying Brier scores for the top 20 experts.
-Since a typical expert participated only in a small subset of the 166 questions, the experts are considered indistinguishable conditional on the level of self-reported expertise.
-
-The experts made updates on a very infrequent basis.
-An expert gave, on average, only 0.016 forecasts per day; that is, a forecast in about every 63 days.
-The average group-level response rate was around 13.5 forecasts per day.
-Given that the group of experts is large and diverse, the data are very sparse.
+Given that a typical expert participated only in a small subset of the 166 questions, the experts are considered indistinguishable conditional on the level of self-reported expertise.
+
+The experts made updates on a very infrequent basis: the average number of forecasts per expert was around 0.017 forecasts per day, and the average group-level response rate was around 13.5 forecasts per day.
+Given that the group of experts is large and diverse, the resulting dataset is very sparse.
 Tables \ref{DataStats} and \ref{ExpertiseTable} provide more relevant summary statistics on the data.
 Notice that the distribution of the self-reported expertise is skewed to the right and that some questions remained active longer than others.
-For more details on the data and its collection see \citet{ungar2012good}.
+For more details on the dataset and its collection see \citet{ungar2012good}.
 
@@
 Alternatively, the hidden process can drift to zero in which case any available information about the target event does not improve predictive accuracy.
-Since each of the $K$ questions in our dataset was resolved within a pre-specified timeframe, $\gamma_k$ is expected to fall within the interval $(1, \infty)$ for all $k = 1, \dots, K$.
+Given that each of the $K$ questions in our dataset was resolved within a pre-specified timeframe, $\gamma_k$ is expected to fall within the interval $(1, \infty)$ for all $k = 1, \dots, K$.
 
@@
 \label{sampling_step}
-Since $\left(a \boldsymbol{b}, X_{t,k}/a, a^2 \tau_k^2\right) \neq \left(\boldsymbol{b}, X_{t,k}, \tau_k^2\right)$ for any $a > 0$ yield the same  likelihood for $\boldsymbol{Y}_{t,k}$, the model as described by Equations (\ref{observedpr}) and (\ref{hiddenpr}) is not identifiable.
+Given that $\left(a \boldsymbol{b}, X_{t,k}/a, a^2 \tau_k^2\right) \neq \left(\boldsymbol{b}, X_{t,k}, \tau_k^2\right)$ for any $a > 0$ yield the same  likelihood for $\boldsymbol{Y}_{t,k}$, the model as described by Equations (\ref{observedpr}) and (\ref{hiddenpr}) is not identifiable.
 As a result, the parameter estimates tend to drift during the sampling process.
@@
 where the trailing input parameter emphasizes the constraint $b_{3} = 1$.
-Since this version is identifiable, estimates of the model parameters can be obtained.
+Given that this version is identifiable, estimates of the model parameters can be obtained.
 Denote the estimates by placing a hat on the parameter symbol.
@@
 \label{calibration_step}
-Since the model parameters can be estimated by fixing $b_3$ to any constant, the next step is to search for the constant that gives an optimally sharp and calibrated estimate of the hidden process.
+Given that the model parameters can be estimated by fixing $b_3$ to any constant, the next step is to search for the constant that gives an optimally sharp and calibrated estimate of the hidden process.
 This section introduces an efficient procedure that finds the optimal constraint without requiring any additional runs of the sampling step.
 First, assume that parameter estimates  $\hat{\boldsymbol{b}}(1)$ and $\hat{X}_{t, k}(1)$ have already been obtained via the constrained sampling step described in Section \ref{sampling_step}.
-Since for any $\beta \in \R / \{0\}$,
+Given that for any $\beta \in \R / \{0\}$,
 \begin{eqnarray*}
@@
 This means that $X_{t,k} = X_{t,k}(1)/\beta$ when $\beta$ is equal to the true value of $b_3$.
-Since the hidden process $X_{t, k}$ is assumed to be sharp and well-calibrated, $b_3$ can be estimated with the value of $\beta$ that simultaneously maximizes the sharpness and calibration of $\hat{X}_{t, k}(1) / \beta$.
+Given that the hidden process $X_{t, k}$ is assumed to be sharp and well-calibrated, $b_3$ can be estimated with the value of $\beta$ that simultaneously maximizes the sharpness and calibration of $\hat{X}_{t, k}(1) / \beta$.
 A natural criterion for this maximization is given by the class of \textit{proper scoring rules} that combine sharpness and calibration (\citet{gneiting2008rejoinder, buja2005loss}).
@@
 \end{eqnarray*}
-Since it is not clear which rule should be used for predicting geopolitical events, the \textit{Sample-And-Calibrate} procedure is evaluated  separately under both rules in Sections \ref{syntheticData} and \ref{realData}.
+Given that it is not clear which rule should be used for predicting geopolitical events, the \textit{Sample-And-Calibrate} procedure is evaluated  separately under both rules in Sections \ref{syntheticData} and \ref{realData}.
 Once $\hat{\beta}$ has been computed, estimates of the unconstrained model parameters are given by
@@
 If the data are balanced with respect to the time points, then exactly half of the summands in Equation (\ref{platt}) have $Z_k = 1$ and the average response logit-probability $\logit(\P(Z_k = 1 | \hat{X}_{t,k})) $ is expected to be close to zero.
-Since the values of $\hat{X}_{t,k}$ are estimated logit-probabilities of the same $K$ events across different time points, their overall average is also expected to be around zero.
+Given that the values of $\hat{X}_{t,k}$ are estimated logit-probabilities of the same $K$ events across different time points, their overall average is also expected to be around zero.
 Therefore both the response and explanatory variables are approximately centered.
@@
 Each of the remaining  $M-1$ possible outcomes is represented by an observed process similar to Equation (\ref{observedpr}).
-Since this multinomial extension is equivalent to having $M-1$ independent binary-outcome models, the estimation and properties of the model are easily extended to the multi-outcome case.
+Given that this multinomial extension is equivalent to having $M-1$ independent binary-outcome models, the estimation and properties of the model are easily extended to the multi-outcome case.
 This paper focuses on binary-outcomes because it is the simplest and most commonly encountered setting in practice.
@@
 The SAC-procedure is run on each dataset for 200 iterations of which the first 100 are used for burn-in.
-The accuracy of the estimated hidden process and bias vector are measured with the average quadratic loss in the probability space, $\sum_{k=1}^K \sum_{t=1}^{100} ( \logit^{-1}(\hat{X}_{t,k}) - \logit^{-1}(X_{t,k}))^2 / (100K)$, and the average quadratic loss, $|| \log(\boldsymbol{b}) - \log(\hat{\boldsymbol{b}})||^2/5$, respectively.
+The accuracy of the estimated hidden process and bias vector are measured with the average quadratic loss in the probability space, $\sum_{k=1}^K \sum_{t=1}^{100} ( \logit^{-1}(\hat{X}_{t,k}) - \logit^{-1}(X_{t,k}))^2 / (100K)$, and the average quadratic loss, $|| \boldsymbol{b} - \hat{\boldsymbol{b}}||^2/5$, respectively.
+
+%$|| \log(\boldsymbol{b}) - \log(\hat{\boldsymbol{b}})||^2/5$, respectively.
 The latter is measured in the log-space to allow a fair comparison of any discrepancies between under- and over-confident bias-vectors.
@@
 The second row is similar in nature but measures the accuracy in estimating the bias vector instead.
-Since the plots were computed by choosing one grid variable (e.g., $\beta$) at a time and averaging over the remaining two variables  (e.g., $K$ and $\sigma^2$), each value in the plots represents an average of a $40 \times 5^2 = 1,000$ values.
+Given that the plots were computed by choosing one grid variable (e.g., $\beta$) at a time and averaging over the remaining two variables  (e.g., $K$ and $\sigma^2$), each value in the plots represents an average of a $40 \times 5^2 = 1,000$ values.
 Based on these results, both $\text{SAC}_{\text{LOG}}$ and  $\text{SAC}_{\text{BRI}}$ estimate the hidden process and the bias vector very accurately.
 Apart from $\text{SAC}_{\text{LOG}}$ outperforming $\text{SAC}_{\text{BRI}}$ uniformly across all parameter values, the two approaches behave very similarly.
-They make great use of the increasing number of questions, suffer slightly from the increasing level of noise in the expert logit-probabilities, and perform better when $\beta$ is large.
+They make great use of the increasing number of questions, suffer slightly from the increasing level of noise in the expert logit-probabilities, and present a trade-off between estimating $X_{t,k}$ and $\boldsymbol{b}$ under different values of $\beta$.
+This trade-off arises from \ville{WHY DOES THIS HAPPEN?}
 
@@
 \citet{Geo}, on the other hand, consider only probabilities that fall within a constrained interval, say $[0.001, 0.999]$, and discard the rest.
-Since this implies ignoring a portion of the data, we adopt an approach similar to \citet{Ariely00theeffects}  by truncating $p = 0.00$ and $1.00$ to $p = 0.01$ and $0.99$, respectively.
+Given that this implies ignoring a portion of the data, we adopt an approach similar to \citet{Ariely00theeffects}  by truncating $p = 0.00$ and $1.00$ to $p = 0.01$ and $0.99$, respectively.
 Our results remain insensitive to the exact choice of truncation as long as this is done in a reasonable manner to keep the extreme probabilities from becoming highly influential in the logit-space.
@@
 Posterior estimates of $\beta$ can be sampled from Equation (\ref{OSE2}) using generic sampling algorithms such as the Metropolis algorithm (\citet{metropolis1953equation}) or slice sampling (\citet{neal2003slice}).
-Since the sampling procedure conditions on the event indicators, the full conditional distribution of the hidden states is not in a standard form.
+Given that the sampling procedure conditions on the event indicators, the full conditional distribution of the hidden states is not in a standard form.
 Therefore the Metropolis algorithm is also used for sampling the hidden states.
@@
  Similarly behavior is exhibited by EWMA that performs the worst of all the competing models.
-In the contrary, the other two exponentially weighted aggregators, EWMLA and EWMBA, make efficient use of the training set and present good forecasting performance in most columns of Table \ref{prediction}.
-Recall that EWMBA uses the cumulative distribution function of the Beta distribution to transform the average forecast probability.
-This transformation depends on two parameters and is more flexible than the transformation used by EWMLA.
-Only EWLMA is given access to the self-reported expertise information.
-Based on the results, however, it is unable to use this information to outperform EWMBA.
-Overall, these two aggregators perform very similarly.
-Unfortunately, they both suffer from over-confidence and are hence unable to outperform the \textit{Sample-And-Calibrate}-procedure.
+In the contrary, the other two exponentially weighted aggregators, EWMLA and EWMBA, make efficient use of the training set and present moderate forecasting performance in most columns of Table \ref{prediction}.
+Recall that EWMBA uses the cumulative distribution function of the Beta distribution that depends on two parameters and is more flexible than the transformation used by EWMLA.
+On other hand, only EWLMA is given access to the self-reported expertise information.
+Neither approach, however, appears to dominate.
+The aggregators perform very similarly.
+The high variability and average of their performance scores indicates that their performance suffers from over-confidence.
 
@@
 A model is considered under-confident (or over-confidence) if the points follow an S-shaped (or \reflectbox{S}-shaped) trend.
-In order to assess sharpness of the model, it is common practice to place a histogram of the given forecasts in the corner of the plot.
-Since the data were balanced, any deviation from the the baseline probability of 0.5 suggests improved sharpness.
+To assess sharpness of the model, it is common practice to place a histogram of the given forecasts in the corner of the plot.
+Given that the data were balanced, any deviation from the the baseline probability of 0.5 suggests improved sharpness.
 
@@
 
-\subfigure[\scriptsize SDLM In-Sample]{
+\subfigure[\scriptsize SDLM ]{
 \hspace*{-2em}  \includegraphics[width = 0.286\textwidth]{Figures/CalibrationNONE10}
@@
 }
-\subfigure[\scriptsize $\text{SAC}_{\text{LOG}}$ In-Sample]{
+\subfigure[\scriptsize $\text{SAC}_{\text{LOG}}$ ]{
 \hspace*{-1.15em} \includegraphics[width = 0.235\textwidth]{Figures/CalibrationLOG10}
@@
 }
-\subfigure[\scriptsize $\text{SAC}_{\text{BRI}}$ In-Sample]{
+\subfigure[\scriptsize $\text{SAC}_{\text{BRI}}$ ]{
 \hspace*{-1.15em} \includegraphics[width = 0.235\textwidth]{Figures/CalibrationBRI10}
@@
 }
-\subfigure[\scriptsize $\text{BSAC}_{\text{LOG}}$ In-Sample]{
+\subfigure[\scriptsize $\text{BSAC}_{\text{LOG}}$ ]{
 \hspace*{-1.15em} \includegraphics[width = 0.286\textwidth]{Figures/CalibrationFULL10}
@@
 
-\subfigure[ \scriptsize SDLM Out-of-Sample]{
+\subfigure[ \scriptsize SDLM]{
  \hspace*{-2em} \includegraphics[width = 0.286\textwidth]{Figures/CalibrationNONE}
@@
 }
-\subfigure[\scriptsize $\text{SAC}_{\text{LOG}}$ Out-of-Sample]{
+\subfigure[\scriptsize $\text{SAC}_{\text{LOG}}$]{
 \hspace*{-1.15em} \includegraphics[width = 0.235\textwidth]{Figures/CalibrationSTO-Log}
@@
 }
-\subfigure[ \scriptsize $\text{SAC}_{\text{BRI}}$ Out-of-Sample]{
+\subfigure[ \scriptsize $\text{SAC}_{\text{BRI}}$]{
 \hspace*{-1.15em} \includegraphics[width = 0.235\textwidth]{Figures/CalibrationSTO-Bri}
@@
 }
-\subfigure[\scriptsize  $\text{BSAC}_{\text{LOG}}$ Out-of-Sample]{
+\subfigure[\scriptsize  $\text{BSAC}_{\text{LOG}}$]{
 \hspace*{-1.15em} \includegraphics[width = 0.286\textwidth]{Figures/CalibrationSO}
@@
 Overall, the \textit{Sample-And-Calibrate}-procedure is sharp and well-calibrated both in- and out-of-sample with only a few points barely falling outside the \textit{point-wise} critical values.
-Since the calibration does not change drastically from the top to the bottom row, the \textit{Sample-And-Calibrate}-procedure can be considered to present robustness against over-fitting.
+Given that the calibration does not change drastically from the top to the bottom row, the \textit{Sample-And-Calibrate}-procedure can be considered to present robustness against over-fitting.
 This is, however, not the case with $\text{BSAC}_{\text{LOG}}$ that is well-calibrated in-sample but presents over-confidence out-of-sample.
 Figures \ref{CalibrationNONEa} and \ref{CalibrationNONEb} serve as baselines by showing the reliability plots for the SDLM-model.
-Since this model does not perform any explicit calibration, it is not surprising to see most points outside the critical values.
+Given that this model does not perform any explicit calibration, it is not surprising to see most points outside the critical values.
 The pattern in the deviations suggests drastic under-confidence.
@@
 Figure \ref{Biases} presents the posterior distributions of the bias terms with side-by-side box plots.
-Since the distributions fall completely below the \textit{no-bias} reference-line at 1.0, all the groups are deemed under-confident.
+Given that the distributions fall completely below the \textit{no-bias} reference-line at 1.0, all the groups are deemed under-confident.
 Even though the exact level of under-confidence is affected slightly by the extent to which the extreme probabilities are truncated (see Section \ref{practicalMatters}), the qualitative results in this section remain insensitive to different levels of truncation.
@@
 If most experts fall somewhere on this spectrum between ignorance and full information, their average forecast tends to fall strictly between 0.5 and the most-informed probability forecast (see \citet{Baron} for more details).
-Since expertise is to a large extent determined by subject-matter knowledge, the level of under-confidence can be expected to decrease as a function of the group's level of self-reported expertise.
-
-Finding under-confidence in all the groups is a rather surprising result given that many previous studies have been conducted to show that experts are likely to be over-confident.
-For instance,  \citet{lichtenstein1977calibration, morgan1992uncertainty, bier2004implications} summarize the results from numerous calibration studies, and conclude that experts are systematically over-conﬁdent about their probability assessments.
-Our result, however, is a statement about groups of experts and hence does not invalidate the possibility of the individual experts being overconfident.
+Given that expertise is to a large extent determined by subject-matter knowledge, the level of under-confidence can be expected to decrease as a function of the group's level of self-reported expertise.
+
+Finding under-confidence in all the groups is a rather surprising result given that many previous studies have shown that experts are often over-confident (see, e.g., \citet{lichtenstein1977calibration, morgan1992uncertainty, bier2004implications} for a summary of numerous calibration studies).
+It is worth emphasizing two points: First, our result is a statement about groups of experts and hence does not invalidate the possibility of the individual experts being overconfident.
 To make conclusions at the individual-level based on the group-level bias terms would be considered an \textit{ecological inference fallacy} (see, e.g., \citet{lubinski1996seeing}).
+Second, the experts involved in our dataset are overall very well calibrated (\cite{mellers}).
+A group of well-calibrated experts, however, can produce an aggregate forecast that is under-confident.
 
@@
 To illustrate, recall that higher values of $\sigma^2_k$ suggest greater disagreement among the participating experts.
-Since experts are more likely to disagree over a difficult question than an easy one, it is reasonable to assume that $\sigma^2_k$ has a positive relationship with question difficulty.
+Given that experts are more likely to disagree over a difficult question than an easy one, it is reasonable to assume that $\sigma^2_k$ has a positive relationship with question difficulty.
 An alternative measure is given by $\tau_k$ that quantifies the volatility of the underlying circumstances that ultimately decide the outcome of the event.
-Therefore a high value of $\tau_k$ can cause the outcome of the event appear unstable and difficult to predict.
+Therefore a high value of $\tau_k$ can cause the outcome of the event to appear unstable and difficult to predict.
 
@@
 The dashed lines connect the point-wise 95\% posterior intervals across different time points.
-Since $\hat{\sigma}_k^2 = 2.43$  and $\hat{\sigma}_k^2 = 1.77$ for the questions depicted in Figures \ref{Example1b} and \ref{Example2b}, respectively, the first question provokes more disagreement among the experts than the second one.
+Given that $\hat{\sigma}_k^2 = 2.43$  and $\hat{\sigma}_k^2 = 1.77$ for the questions depicted in Figures \ref{Example1b} and \ref{Example2b}, respectively, the first question provokes more disagreement among the experts than the second one.
 Intuitively this makes sense because the target event in Figure \ref{Example1b}  is determined by several conditions that may change radically from one day to the next while the target event in Figure \ref{Example2b} is determined by a relatively steady stock market index.
@@
 The sampling is done from standard distributions with fast convergence to the target distribution (see Appendix \ref{appendix} for technical details).
-The second step, known as the \textit{calibration step}, uses a one-dimensional optimization procedure to transforms the constrained parameter values to their unconstrained counterparts.
-To the best of our knowledge, this procedure extends forecasting literature towards rather unexplored areas of probability aggregation.
+The second step, known as the \textit{calibration step}, uses a one-dimensional optimization procedure to transform the constrained parameter values to their unconstrained counterparts.
+To the best of our knowledge, this procedure extends the forecasting literature into rather unexplored areas of probability aggregation.
 
@@
 \subsection{Limitations and Directions for Future Research}
-Our model pertains model parsimony while addressing the main challenges that arise from modeling sparse probability forecasting data.
+Our model preserves parsimony while addressing the main challenges in modeling sparse probability forecasting data.
 Therefore it can be viewed as a basis for many future extensions.
@@
 This hypothesis could be explored by letting $\sigma^2_{t,k}$ evolve dynamically as a function of the previous term $\sigma^2_{t-1,k}$ and random noise.
+Furthermore, this parameter along with the bias term can be explored at an individual level if the experts updated their forecasts very frequently.
+Estimates of $\sigma_{i,k}^2$ and $b_{i,k}$ could then be used to separate the accurate from the inaccurate forecasters.
 
@@
 Different parameter dependencies could be constructed based on many other expert characteristic, such as gender, education, or specialty, to produce a range of novel insight on the forecasting behavior across different groups of experts.
-Furthermore, expert characteristics could be interacted with question types, such as economic, domestic, or international, to learn about the kind of experts who generally perform well on certain types of questions.
-This would be of interest to the decision-maker who could use the information as a basis for consulting only a high-performing subset of the available experts.
-
-%It may also be of interest to study hidden processes with other goals besides maximizing sharpness subject to calibration.
+It would also be useful to know how expert characteristics interact with question types, such as economic, domestic, or international.
+This way the researcher can gain insight about the kind of experts who generally perform well on certain types of questions.
+The results would be of interest to the decision-maker who could use the information as a basis for consulting only a high-performing subset of the available experts.
+
+%If the experts updated their probability forecasts very frequently, it is possible to use the model to estimate expert-specific parameters such as $\sigma_{k,i}^2$ or $b_{i,k}$.
+Together these estimates can be used to separate the accurate from the inaccurate forecasters.
+The estimates can also provide insight on the responsiveness and
+
+Given that decision-makers can have different preferences on calibration and sharpness, it may also be of interest to study hidden processes with other goals besides maximizing sharpness subject to calibration.
+For instance, a government official may want to sacrifice some calibration for extra sharpness.
 This can be easily achieved by changing the optimization criterion in Equation (\ref{OSE}) to an appropriate function that meets the new goal.
@@
 A more complex model, however, is likely to sacrifice interpretability.
-Given that our model is able to detect very intricate patterns in the crowd belief  (see Figure \ref{ExamplePlotsFinal}), compromising interpretability for the sake of facilitating non-linear growth is hardly necessary.
+Given that our model can detect very intricate patterns in the crowd belief  (see Figure \ref{ExamplePlotsFinal}), compromising interpretability for the sake of facilitating non-linear growth is hardly necessary.
+
+\section{Acknowledgements}
+\noindent
+This research was supported by a research contract to the University
+of Pennsylvania and the University of California from the Intelligence Advanced Research Projects Activity
+(IARPA) via the Department of Interior National Business Center
+contract number D11PC20061. The U.S. Government is authorized to reproduce and
+distribute reprints for Government purposes notwithstanding any
+copyright annotation thereon. Disclaimer: The views and conclusions
+expressed herein are those of the authors and should not be
+interpreted as necessarily representing the official policies or
+endorsements, either expressed or implied, of IARPA, DoI/NBC, or the
+U.S. Government.
+
+We deeply appreciate the project management
+skills and work of Terry Murray and David Wayrynen, which went far
+beyond the call-of-duty on this project.
 
@@
 The Gibbs sampler (\citet{geman1984stochastic}) iteratively samples all the unknown parameters from their full-conditional posterior distributions one block of parameters at a time.
-Since this is performed under the constraint $b_3 = 1$ to ensure model identifiability, the constrained parameter estimates should be denoted with a trailing $(1)$ to maintain consistency with earlier notation.
+Given that this is performed under the constraint $b_3 = 1$ to ensure model identifiability, the constrained parameter estimates should be denoted with a trailing $(1)$ to maintain consistency with earlier notation.
 For instance, the constrained estimate of $\gamma_k$ should be denoted by $\hat{\gamma}_k(1)$ while the unconstrained estimate is denoted by $\hat{\gamma}_k$.
@@
 First, vectorize all the response vectors $\boldsymbol{Y}_{t,k}$ into a single vector denoted $\boldsymbol{Y}_k = \left[\boldsymbol{Y}_{1,k}^T, \dots, \boldsymbol{Y}_{T_k,k}^T\right]^T$.
-Since each $\boldsymbol{Y}_{t,k}$ is matched with $X_{t,k}$ via the time index $t$, we can form a $|\boldsymbol{Y}_k| \times J$ design-matrix by letting $\boldsymbol{X}_k = \left[ (\boldsymbol{M}_kX_{1,k})^T, \dots, (\boldsymbol{M}_kX_{T_k,k})^T \right]^T$.
+Given that each $\boldsymbol{Y}_{t,k}$ is matched with $X_{t,k}$ via the time index $t$, we can form a $|\boldsymbol{Y}_k| \times J$ design-matrix by letting $\boldsymbol{X}_k = \left[ (\boldsymbol{M}_kX_{1,k})^T, \dots, (\boldsymbol{M}_kX_{T_k,k})^T \right]^T$.
 Given that the goal is to borrow strength across questions by assuming a common bias vector $\boldsymbol{b}$, the parameter values must be estimated in parallel for each question such that the matrices $\boldsymbol{X}_k$ can be further concatenated into $\boldsymbol{X} = [\boldsymbol{X}_1^T, \dots, \boldsymbol{X}_K^T]^T$ during every iteration.
@@
 \end{eqnarray}
-Since the covariance matrix in Equation (\ref{BiasVector}) is diagonal, the constraint is enforced at this point by letting $b_3 = 1$. The variance parameter is then sampled from
+Given that the covariance matrix in Equation (\ref{BiasVector}) is diagonal, the constraint is enforced at this point by letting $b_3 = 1$. The variance parameter is then sampled from
 \begin{eqnarray*}
@@
 where the distribution is a scaled inverse-$\chi^2$ (see, e.g., \citet{gelman2003bayesian}).
-Since the experts are not required to give a new prediction at every time unit, the design matrices must be trimmed accordingly such that their dimensions match up with the dimensions of the observed matrices.
+Given that the experts are not required to give a new prediction at every time unit, the design matrices must be trimmed accordingly such that their dimensions match up with the dimensions of the observed matrices.
 

### 2013-09-24 paper.tex
Final edits

@@
 \begin{itemize}
-\item borrowing strength across the hierarchy to analyze the under- and overconfidence in different groups of experts,
+\item borrowing strength across the hierarchy to analyze under- and overconfidence in different groups of experts,
 %\item confidence bands for the crowd belief,
@@
 The model is estimated in a two-step procedure: first, a sampling step produces constrained parameter estimates via Gibbs sampling (see \citet{geman1984stochastic} for the original introduction of Gibbs sampling); second, a calibration step transforms these estimates to their unconstrained equivalents via a one-dimension optimization procedure.
-An extension of this model to polychotomous outcomes is briefly discussed before model evaluation.
-The first evaluation section uses synthetic data to study how accurately the two-step procedure can estimate parameter values.
+The model introduction is followed by the first evaluation section that uses synthetic data to study how accurately the two-step procedure can estimate parameter values.
 The second evaluation section applies the model to real-world forecasting data.
@@
 
-The experts made updates on a very infrequent basis: the average number of forecasts per expert was around 0.017 forecasts per day, and the average group-level response rate was around 13.5 forecasts per day.
+The experts made updates on a very infrequent basis: the average number of forecasts made by a single expert in one day was around 0.017, and the average group-level response rate was around 13.5 forecasts per day.
 Given that the group of experts is large and diverse, the resulting dataset is very sparse.
-Tables \ref{DataStats} and \ref{ExpertiseTable} provide more relevant summary statistics on the data.
+Tables \ref{DataStats} and \ref{ExpertiseTable} provide relevant summary statistics on the data.
 Notice that the distribution of the self-reported expertise is skewed to the right and that some questions remained active longer than others.
@@
 A well-known solution is to choose one of the elements of $\boldsymbol{b}$, say $b_3$, as the reference point and fix $b_3 = 1$.
+In Section \ref{syntheticData} we provide a guideline for choosing the reference point.
 Denote the constrained version of the model by
@@
 \end{eqnarray*}
-Given that it is not clear which rule should be used for predicting geopolitical events, the \textit{Sample-And-Calibrate} procedure is evaluated  separately under both rules in Sections \ref{syntheticData} and \ref{realData}.
+Given that it is not clear which rule should be used for predicting geopolitical events, the SAC-procedure is evaluated  separately under both rules in Sections \ref{syntheticData} and \ref{realData}.
 Once $\hat{\beta}$ has been computed, estimates of the unconstrained model parameters are given by
@@
 
-If the future event can take upon $M > 2$ possible outcomes, the hidden state $X_{t,k}$ must extended to a vector of size $M-1$ and one of the outcomes, e.g., the $M$th one, must be chosen as the base-case to ensure that the probabilities will sum to one at any given time point.
+If the future event can take upon $M > 2$ possible outcomes, the hidden state $X_{t,k}$ is extended to a vector of size $M-1$ and one of the outcomes, e.g., the $M$th one, is chosen as the base-case to ensure that the probabilities will sum to one at any given time point.
 Each of the remaining  $M-1$ possible outcomes is represented by an observed process similar to Equation (\ref{observedpr}).
@@
 
-The goal in this section is to evaluate the ability of the SAC-procedure to capture true parameter values.
-The synthetic data are not generated directly from the model for two reasons: (i) showing good performance on a dataset directly generated from the model assumptions is hardly any news, and (ii) generating data from the dynamic model description does not produce well-calibrated hidden states.
+The goal in this section is to evaluate the ability of the \textit{Sample-And-Calibrate}-procedure to capture true parameter values.
+The synthetic data are not generated directly from the model for two reasons: (i) showing good performance on a dataset directly generated from the model assumptions is hardly news, and (ii) it is unclear how the dynamic part of the model can be used to produce well-calibrated hidden states.
 The latter is important for our interpretation of the hidden process as a sharp and well-calibrated version of the crowd belief.
 
-To ensure that the hidden process for question $k$ is well-calibrated, generate a path of the standard Brownian motion until time $T_k$. If $Z_{t,k}$ denotes the value of the path at time $t$, then
+To ensure that the hidden process for question $k$ is well-calibrated, we generate a path of standard Brownian motion until time $T_k$. If $Z_{t,k}$ denotes the value of the path at time $t$, then
 \begin{eqnarray*}
@@
 Using this procedure we generate the hidden process for $K$ questions, each with a time horizon of $T_k = 101$.
-The questions involve 25 experts allocated evenly among five expertise groups.
-Each expert gives one probability forecast per day with the exception of time $t = 101$, when the event resolves.
+The questions involve 50 experts allocated evenly among five expertise groups.
+Each expert gives one probability forecast per day with the exception of time $t = 101$ when the event resolves.
 These are generated by applying bias and noise to the hidden process $X_{t,k}$ for $t = 1, \dots, 100$ (see Equation (\ref{observedpr})).
-In this fashion, the simulation iterates over a three-dimensional grid of parameter values:
+Our simulation study considers a three-dimensional grid of parameter values:
 \begin{eqnarray*}
@@
 where $\beta$ gives the bias vector by letting $\boldsymbol{b} = [1/2, 3/4, 1, 4/3, 2/1]^T \beta$.
-Every grid point is used  40 times to generate a synthetic dataset.
-The SAC-procedure is run on each dataset for 200 iterations of which the first 100 are used for burn-in.
+Forty synthetic datasets are generated for each combination of $\sigma^2$, $\beta$, and $K$ values.
+The SAC-procedure runs for 200 iterations of which the first 100 are used for burn-in.
 The accuracy of the estimated hidden process and bias vector are measured with the average quadratic loss in the probability space, $\sum_{k=1}^K \sum_{t=1}^{100} ( \logit^{-1}(\hat{X}_{t,k}) - \logit^{-1}(X_{t,k}))^2 / (100K)$, and the average quadratic loss, $|| \boldsymbol{b} - \hat{\boldsymbol{b}}||^2/5$, respectively.
@@
 \begin{figure}[h!]
-\vspace{-4.5em}
+\vspace{-5.5em}
 \centering
- \hspace{-31em}	\includegraphics{Figures/LegendSynthetic} % requires the graphicx package
-\vspace{-1.5em}
+ \hspace{-25em}	\includegraphics{Figures/LegendSynthetic} % requires the graphicx package
+
+\vspace{0em}
+\hspace{2.5em}Hidden Process
+\vspace{-0.5em}
 
@@
 
-\vspace{-1em}
+\vspace{0.5em}
+\hspace{2.5em}Bias Vector
+\vspace{-0.5em}
 
@@
 
-\caption[Optional caption for list of figures]{Comparing SAC that optimizes over the logarithmic score ($\text{SAC}_{\text{LOG}}$) and SAC that optimizes over the Brier score ($\text{SAC}_{\text{BRI}}$) under synthetic data.
+\caption[Optional caption for list of figures]{Comparing SAC that optimizes over the logarithmic score ($\text{SAC}_{\text{LOG}}$), SAC that optimizes over the Brier score ($\text{SAC}_{\text{BRI}}$), and the Exponentially Weighted Moving Average (EWMA) under synthetic data.
 The top row presents the accuracy to capture the hidden process.
@@
 Figure \ref{Synthetic} summarizes the results in a series of plots.
-The first row presents the marginal effects of the three grid variables on the accuracy to which the SAC-procedure estimates the hidden process.
-The second row is similar in nature but measures the accuracy in estimating the bias vector instead.
+The first row presents the marginal effects of the three grid variables on the accuracy to which SAC and the Exponentially Weighted Moving Average (EWMA) (see Section \ref{forecasting} for more details on EWMA) estimate the hidden process.
+EWMA serves as a baseline for comparisons.
+Unfortunately, EWMA does not produce an estimate of the bias vector.
+Therefore it cannot be used as a baseline in the second row that measures the accuracy in estimating the bias vector.
 Given that the plots were computed by choosing one grid variable (e.g., $\beta$) at a time and averaging over the remaining two variables  (e.g., $K$ and $\sigma^2$), each value in the plots represents an average of a $40 \times 5^2 = 1,000$ values.
-Based on these results, both $\text{SAC}_{\text{LOG}}$ and  $\text{SAC}_{\text{BRI}}$ estimate the hidden process and the bias vector very accurately.
+
+Both $\text{SAC}_{\text{LOG}}$ and  $\text{SAC}_{\text{BRI}}$ estimate the hidden process and the bias vector very accurately.
 Apart from $\text{SAC}_{\text{LOG}}$ outperforming $\text{SAC}_{\text{BRI}}$ uniformly across all parameter values, the two approaches behave very similarly.
-They make great use of the increasing number of questions, suffer slightly from the increasing level of noise in the expert logit-probabilities, and present a trade-off between estimating $X_{t,k}$ and $\boldsymbol{b}$ under different values of $\beta$.
-This trade-off arises from \ville{WHY DOES THIS HAPPEN?}
+They suffer  from the increasing level of noise in the expert logit-probabilities, estimate the bias vector less accurately when $\beta$ is large, and, unlike EWMA, make greater use of the increasing number of questions.
+EWMA is able to outperform SAC only when the bias among the experts or the number of problems is very small.
+
+Interestingly, SAC estimates the hidden process more accurately when the experts are over-confident (large $\beta$) rather under-confident (small $\beta$).
+To understand this result, assume that the experts in the third group are highly under-confident.
+Therefore their logit-probabilities are expected to be closer to zero than the corresponding hidden states.
+After adding white noise to these expected logit-probabilities, they are likely to cross to the other side of zero.
+If the sampling step fixes $b_3 = 1$, as it does in our case, the third group is treated unbiased and some of the constrained estimates of the hidden states are likely to be on the other side of zero as well.
+Unfortunately, this discrepancy cannot be corrected by the calibration step that is restricted to shifting the constrained estimates either closer or further away from zero but not across it.
+To maximize the likelihood of having all the constrained estimates on the right side of zero and hence avoiding the discrepancy, the reference point in the sampling step should be chosen with care.
+A helpful guideline is to fix the element of $\boldsymbol{b}$ that is apriori believed  to be the largest.
 
@@
 \noindent
-This section presents results related to the real-world data described in Section \ref{data}.
-The goal is to provide application specific insight by discussing the specific research objectives itemized in Section \ref{intro} and also to evaluate the \textit{Sample-And-Calibrate}-procedure in terms of predictive power and calibration.
+This section presents results for the real-world data described in Section \ref{data}.
+The goal is to provide application specific insight by discussing the specific research objectives itemized in Section \ref{intro} and also to evaluate SAC in terms of predictive power and calibration.
 Before presenting the results, however, we discuss two practical matters that must be taken into account when aggregating real-world  probability forecasts.
@@
 A simple solution is to use a greedy algorithm that iterates through the values of $T_k$ in descending order, assigning each $T_k$ to the subset that currently has the smaller sum (see, e.g.
-\citet{kellerer2004knapsack, gent1996phase} for more details on the Partition Problem).
+\citet{kellerer2004knapsack, gent1996phase} for more details on the \textit{Partition Problem}).
 After finding a well-balanced partition, the next step is to assign the class labels such that the labels for the questions in $S_x$ are equal to $x$ for $x = 0$ or $1$.
@@
 %For instance, recall our example with an investor who consults a group of experts on a daily basis in order to assess whether a stock index will end up trading above some threshold on a given date. 
-The goal is to evaluate the out-of-sample predictive performance of \textit{Sample-And-Calibrate} against several other probability aggregation procedures.
+The goal is to evaluate the out-of-sample predictive performance of SAC against several other probability aggregation procedures.
 The models are allowed to utilize a training set before making predictions on an independent testing set.
@@
 The model is first fit on the training set by running the sampling step for 3,000 iterations of which the first 500 iterations are used for burn-in.
-After thinning by only keeping every fifth observation, the calibration step is performed for the remaining 500 observations.
+The remaining 2,500 observations are thinned by keeping every fifth observation.
+The calibration step is performed for the final 500 observations.
 The out-of-sample prediction is done by running the sampling step for 500 iterations with each consecutive iteration reading in and conditioning on the next value of $\beta$ and $\boldsymbol{b}$ found during the training period.
@@
 The posterior distribution of $\beta$ conditional on $\boldsymbol{X}(1)$ is given by $p(\beta | \boldsymbol{X}(1), \boldsymbol{Z}) \propto p( \boldsymbol{Z} | \beta, \boldsymbol{X}(1)) p(\beta | \boldsymbol{X}(1))$.
-Recall that the calibration step under $S_{LOG}$ is equivalent to fitting a logistic regression model with $Z_k$ as the response and $\hat{X}_{k,t}(1)$ as the explanatory variable.
+Recall that the calibration step under $S_{LOG}$ is equivalent to fitting a logistic regression model with $Z_k$ as the response and $\hat{X}_{t,k}(1)$ as the explanatory variable.
 Therefore the likelihood for the Bayesian version is
 \begin{eqnarray}
-p( \boldsymbol{Z} | \beta, \boldsymbol{X}(1)) \propto \prod_{k=1}^K \prod_{t=1}^{T_k} && \logit^{-1} \left(X_{k,t}(1)/\beta  \right)^{Z_k}\times \label{OSE2}\\ 
-&& \left( 1-  \logit^{-1} \left( X_{k,t}(1)/\beta  \right) \right)^{1-Z_k} \nonumber
+p( \boldsymbol{Z} | \beta, \boldsymbol{X}(1)) \propto \prod_{k=1}^K \prod_{t=1}^{T_k} && \logit^{-1} \left(X_{t,k}(1)/\beta  \right)^{Z_k}\times \label{OSE2}\\ 
+&& \left( 1-  \logit^{-1} \left( X_{t,k}(1)/\beta  \right) \right)^{1-Z_k} \nonumber
 \end{eqnarray}
@@
 \begin{cases}
-G_{t,k}(\boldsymbol{b} ), & \text{ for } t  = 1, \\
+G_{1,k}(\boldsymbol{b} ), & \text{ for } t  = 1, \\
 \alpha G_{t,k}(\boldsymbol{b} )  + (1-\alpha) \hat{p}_{t-1,k}(\alpha, \boldsymbol{b}),  & \text{ for } t > 1,
@@
 The short questions, however, involve very little uncertainty and were generally the easiest to forecast.
-On such easy questions, overconfidence can pay off frequently enough to compensate for a few large scores arising from the overconfident and incorrect forecasts.
-
-In comparison to $\text{BSAC}_{\text{LOG}}$, the baseline SDLM-model lacks sharpness and is highly under-confident (see Section \ref{calibration}).
+On such easy questions, overconfidence can pay off frequently enough to compensate for a few large losses arising from the overconfident and incorrect forecasts.
+
+In contrast to $\text{BSAC}_{\text{LOG}}$, the baseline SDLM-model lacks sharpness and is highly under-confident (see Section \ref{calibration}).
 This behavior is expected as the experts are under-confident at the group-level (see Section \ref{ExpertBias}) and the SDLM-procedure does not use the training set to explicitly calibrate its forecasts.
@@
 On other hand, only EWLMA is given access to the self-reported expertise information.
-Neither approach, however, appears to dominate.
+Neither approach, however, appears to dominate the other.
 The aggregators perform very similarly.
@@
 
-Overall, the \textit{Sample-And-Calibrate}-procedure is sharp and well-calibrated both in- and out-of-sample with only a few points barely falling outside the \textit{point-wise} critical values.
-Given that the calibration does not change drastically from the top to the bottom row, the \textit{Sample-And-Calibrate}-procedure can be considered to present robustness against over-fitting.
+Overall, SAC is sharp and well-calibrated both in- and out-of-sample with only a few points barely falling outside the \textit{point-wise} critical values.
+Given that the calibration does not change drastically from the top to the bottom row, SAC can be considered to present robustness against over-fitting.
 This is, however, not the case with $\text{BSAC}_{\text{LOG}}$ that is well-calibrated in-sample but presents over-confidence out-of-sample.
-Figures \ref{CalibrationNONEa} and \ref{CalibrationNONEb} serve as baselines by showing the reliability plots for the SDLM-model.
+Figures \ref{CalibrationNONEa} and \ref{CalibrationNONEb} serve as baselines by showing the reliability plots for SDLM.
 Given that this model does not perform any explicit calibration, it is not surprising to see most points outside the critical values.
@@
 Furthermore, the inset histogram reveals drastic lack of sharpness.
-Therefore the \textit{Sample-And-Calibrate}-model can be viewed as a well-performing compromise between SDLM and $\text{BSAC}_{\text{LOG}}$ that avoids over-confidence without being too conservative.
+Therefore SAC can be viewed as a well-performing compromise between SDLM and $\text{BSAC}_{\text{LOG}}$ that avoids over-confidence without being too conservative.
 
@@
 \caption[Optional caption for list of figures]{Comparing the bias-levels across self-reported expertise under different approaches. 
-Posterior distributions of $b_j$ for $j = 1, \dots, 5$ under the \textit{Sample-And-Calibrate} approach that optimizes over the logarithmic score and the \textit{Sample-And-Calibrate} approach that optimizes over the Brier score.
+Posterior distributions of $b_j$ for $j = 1, \dots, 5$ under the \textit{Sample-And-Calibrate} approach that optimizes over the logarithmic score ($\text{SAC}_{\text{LOG}}$) and the \textit{Sample-And-Calibrate} approach that optimizes over the Brier score ($\text{SAC}_{\text{BRI}}$).
 }
@@
 The self-reported expertise then divides the experts into 5  groups, with each group assigned a separate multiplicative bias term.
-This section uses the \textit{Sample-And-Calibrate}-procedure to explore the posterior distributions of these multiplicative bias terms.
+This section uses SAC to explore the posterior distributions of these multiplicative bias terms.
 Figure \ref{Biases} presents the posterior distributions of the bias terms with side-by-side box plots.
@@
 Furthermore, this parameter along with the bias term can be explored at an individual level if the experts updated their forecasts very frequently.
-Estimates of $\sigma_{i,k}^2$ and $b_{i,k}$ could then be used to separate the accurate from the inaccurate forecasters.
+Estimates of $\sigma_{i,k}^2$ and $b_{i,k}$ could then be used to separate the accurate forecasters from the inaccurate ones.
 
@@
 For each such repetition of the update step, the previous posterior values, $X_{t|t, k}$ and $P_{t|t, k}$, should be considered as the new prior values, $X_{t|t-1, k}$ and $P_{t|t-1, k}$.
+If the observation $\boldsymbol{Y}_{t,k}$ is completely missing at time $t$, the update step is skipped and
+\begin{align*}
+X_{t|t, k} &= X_{t|t-1, k}\\
+P_{t|t,k} &= P_{t|t-1,k}
+\end{align*}
 After running the Kalman Filter up to the final time point at $t = T_k$, the final hidden state is sampled from $X_{T_k,k} \sim \mathcal{N}(X_{T_k|T_k, k}, P_{T_k|T_k, k})$.
@@
 This can be viewed as backward updating that considers the Kalman Filter estimates as additional observations at each given time point.
-If the observation $\boldsymbol{Y}_{t,k}$ is completely missing at time $t$, the update step is skipped and the state estimates are sampled from
-\begin{align*}
-\mathcal{N}\left(\gamma_k X_{t-1|t-1,k}, \gamma_k^2 P_{t-1|t-1,k} + \tau_k^2\right)
-\end{align*}
+
+%
+%If the observation $\boldsymbol{Y}_{t,k}$ is completely missing at time $t$, the update step is skipped and the state estimates are sampled from
+%\begin{align*}
+%\mathcal{N}\left(\gamma_k X_{t-1|t-1,k}, \gamma_k^2 P_{t-1|t-1,k} + \tau_k^2\right)
+%\end{align*}
 
@@
 First, vectorize all the response vectors $\boldsymbol{Y}_{t,k}$ into a single vector denoted $\boldsymbol{Y}_k = \left[\boldsymbol{Y}_{1,k}^T, \dots, \boldsymbol{Y}_{T_k,k}^T\right]^T$.
-Given that each $\boldsymbol{Y}_{t,k}$ is matched with $X_{t,k}$ via the time index $t$, we can form a $|\boldsymbol{Y}_k| \times J$ design-matrix by letting $\boldsymbol{X}_k = \left[ (\boldsymbol{M}_kX_{1,k})^T, \dots, (\boldsymbol{M}_kX_{T_k,k})^T \right]^T$.
-Given that the goal is to borrow strength across questions by assuming a common bias vector $\boldsymbol{b}$, the parameter values must be estimated in parallel for each question such that the matrices $\boldsymbol{X}_k$ can be further concatenated into $\boldsymbol{X} = [\boldsymbol{X}_1^T, \dots, \boldsymbol{X}_K^T]^T$ during every iteration.
+Given that each $\boldsymbol{Y}_{t,k}$ is matched with $X_{t,k}$ via the time index $t$, we can form a $|\boldsymbol{Y}_k| \times J$ design-matrix by letting $$\boldsymbol{X}_k = \left[ (\boldsymbol{M}_kX_{1,k})^T, \dots, (\boldsymbol{M}_kX_{T_k,k})^T \right]^T$$ Given that the goal is to borrow strength across questions by assuming a common bias vector $\boldsymbol{b}$, the parameter values must be estimated in parallel for each question such that the matrices $\boldsymbol{X}_k$ can be further concatenated into $\boldsymbol{X} = [\boldsymbol{X}_1^T, \dots, \boldsymbol{X}_K^T]^T$ during every iteration.
 Similarly, $\boldsymbol{Y}_k$ must be further vectorized into a vector $\boldsymbol{Y} = [\boldsymbol{Y}_1^T, \dots, \boldsymbol{Y}_K^T]^T$.
 The question-specific variance terms are taken into account by letting $\boldsymbol{\Sigma} = \text{diag}(\sigma^2_1 \boldsymbol{1}_{1 \times T_1}, \dots, \sigma^2_K \boldsymbol{1}_{1 \times T_K})$.
-After adopting the non-informative prior $p(\boldsymbol{b}, \sigma_k^2 | \boldsymbol{X}_k) \propto \sigma_k^{-2}$ for each $k = 1, \dots, K$, the bias vector are sampled from
+After adopting the non-informative prior $p(\boldsymbol{b}, \sigma_k^2 | \boldsymbol{X}_k) \propto \sigma_k^{-2}$ for each $k = 1, \dots, K$, the bias vector is sampled from
 \begin{eqnarray}
@@
 \end{eqnarray}
-Given that the covariance matrix in Equation (\ref{BiasVector}) is diagonal, the constraint is enforced at this point by letting $b_3 = 1$. The variance parameter is then sampled from
+Given that the covariance matrix in Equation (\ref{BiasVector}) is diagonal, the identifiability constraint can be enforced after sampling a new value of $\boldsymbol{b}$ by letting $b_3 = 1$.
+The variance parameters are then sampled from
 \begin{eqnarray*}
@@
 where the distribution is a scaled inverse-$\chi^2$ (see, e.g., \citet{gelman2003bayesian}).
-Given that the experts are not required to give a new prediction at every time unit, the design matrices must be trimmed accordingly such that their dimensions match up with the dimensions of the observed matrices.
+Given that the experts are not required to give a new forecast at every time unit, the design matrices must be trimmed accordingly such that their dimensions match up with the dimensions of the observed matrices.
 
@@
 \end{center}
-Estimating the parameters related to the hidden process are estimated via a regression setup.
+The parameters of the hidden process are estimated via a regression setup.
 More specifically, after adopting the non-informative prior $p(\gamma_k, \tau_k^2 | \boldsymbol{X}_k) \propto \tau_k^{-2}$, the parameter values are sampled from

### 2013-09-29 paper.tex
Final commit before submit!

@@
 \author{\fnms{Barbara A.} \snm{Mellers}\ead[label=e4]{mellers@wharton.upenn.edu}},
-\author{\fnms{Phil E.} \snm{Tetlock}\ead[label=e5]{tetlock@wharton.upenn.edu}},
+\author{\fnms{Philip E.} \snm{Tetlock}\ead[label=e5]{tetlock@wharton.upenn.edu}},
 \and
@@
 The paper then introduces a dynamic hierarchical model for capturing the crowd belief.
-The model is estimated in a two-step procedure: first, a sampling step produces constrained parameter estimates via Gibbs sampling (see \citet{geman1984stochastic} for the original introduction of Gibbs sampling); second, a calibration step transforms these estimates to their unconstrained equivalents via a one-dimension optimization procedure.
+The model is estimated in a two-step procedure: first, a sampling step produces constrained parameter estimates via Gibbs sampling (see, e.g., \citet{geman1984stochastic}); second, a calibration step transforms these estimates to their unconstrained equivalents via a one-dimension optimization procedure.
 The model introduction is followed by the first evaluation section that uses synthetic data to study how accurately the two-step procedure can estimate parameter values.
-The second evaluation section applies the model to real-world forecasting data.
+The second evaluation section applies the model to our real-world geopolitical forecasting database.
 The paper concludes with a discussion on future research directions and model limitations.
@@
 The points have been jittered slightly to make overlaps visible.
-The darkness of the points is positively associated with the self-reported expertise.
+The darkness of the points is positively correlated with the self-reported expertise.
 Given that the European bailout fund was ratified before November 1, 2011 and that the Nikkei 225 index finished trading at around 8,700 on September 30, 2011, the general trend of the probability forecasts tends to converge towards the correct answers.
@@
 These estimates are found via Gibbs sampling that only makes use of standard distributions.
-See Appendix \ref{appendix} for the technical details of the sampling step, and, e.g., \citet{gelman2003bayesian} for a discussion on the general principles of Gibbs sampling.
+See Appendix \ref{appendix} for the technical details of the sampling steps, and, e.g., \citet{gelman2003bayesian} for a discussion on the general principles of Gibbs sampling.
 
@@
 \end{eqnarray}
-Notice that estimates of $\sigma^2_k$ and $\gamma_k$ are not affected by the constraint. Therefore their constrained and unconstrained versions are the same.
+Notice that estimates of $\sigma^2_k$ and $\gamma_k$ are not affected by the constraint.
 
@@
 They suffer  from the increasing level of noise in the expert logit-probabilities, estimate the bias vector less accurately when $\beta$ is large, and, unlike EWMA, make greater use of the increasing number of questions.
-EWMA is able to outperform SAC only when the bias among the experts or the number of problems is very small.
+EWMA is able to outperform SAC only when the bias among the experts is very small.
 
@@
 The vector $\boldsymbol{b}$ collects the bias terms of the different expertise groups.
-Therefore it is equivalent to the bias vector found under \textit{Sample-And-Calibrate}.
+Therefore it serves a similar role to the bias vector found under \textit{Sample-And-Calibrate}.
 The term $\boldsymbol{e}_{i,k}$ is a vector of length 5 indicating which level of self-reported expertise the $i$th expert in the $k$th question belongs to.
@@
 The number of questions in these subcategories were 36, 32 and 98, respectively.
-The bolded scores indicate the lowest score in each column.
+The bolded scores indicate the best score in each column.
 The values in the parenthesis quantify the variability in the scores: Under \textit{Scores by Day} the values give the standard errors of all the scores.
@@
 \textit{Scores by Problem} gives each question an equal weight regardless how long the question remained open.
-The bolded values indicate the lowest scores in each column.
+The bolded values indicate the best scores in each column.
 The values in the parenthesis represent standard errors in the scores.
@@
 Overall, $\text{SAC}_{\text{BRI}}$ and $\text{SAC}_{\text{LOG}}$ achieve the lowest average scores across all columns except \textit{Short}, where they are slightly outperformed by $\text{BSAC}_{\text{LOG}}$.
-$\text{BSAC}_{\text{LOG}}$, however, turns out to be overconfident (see Section \ref{calibration}).
+However, $\text{BSAC}_{\text{LOG}}$ turns out to be overconfident (see Section \ref{calibration}).
 This means that $\text{BSAC}_{\text{LOG}}$ underestimates the uncertainty in the events and outputs probability forecasts that are typically too near 0.0 or 1.0.
@@
 
- Similarly behavior is exhibited by EWMA that performs the worst of all the competing models.
-In the contrary, the other two exponentially weighted aggregators, EWMLA and EWMBA, make efficient use of the training set and present moderate forecasting performance in most columns of Table \ref{prediction}.
+ Similar behavior is exhibited by EWMA that performs the worst of all the competing models.
+In contrast, the other two exponentially weighted aggregators, EWMLA and EWMBA, make efficient use of the training set and present moderate forecasting performance in most columns of Table \ref{prediction}.
 Recall that EWMBA uses the cumulative distribution function of the Beta distribution that depends on two parameters and is more flexible than the transformation used by EWMLA.
@@
 The aggregators perform very similarly.
-The high variability and average of their performance scores indicates that their performance suffers from over-confidence.
+The high variability and average of their performance scores indicate that their performance suffers from over-confidence.
 
@@
 Therefore any deviation from the diagonal line suggests poor calibration.
-A model is considered under-confident (or over-confidence) if the points follow an S-shaped (or \reflectbox{S}-shaped) trend.
+A model is considered under-confident (or over-confident) if the points follow an S-shaped (or \reflectbox{S}-shaped) trend.
 To assess sharpness of the model, it is common practice to place a histogram of the given forecasts in the corner of the plot.
@@
 
-The decreasing trend in under-confidence can be reasoned by viewing the process of making a subjective probability  as Bayesian updating: A completely ignorant expert aiming to minimize a reasonable loss function, such as the Brier score, has no reason to give anything but 0.5 as his probability forecast.
+The decreasing trend in under-confidence can be reasoned by viewing the process of making a subjective probability  as Bayesian updating: a completely ignorant expert aiming to minimize a reasonable loss function, such as the Brier score, has no reason to give anything but 0.5 as his probability forecast.
 However, as soon as the expert gains some knowledge about the event, he produces an updated forecast that is a compromise between his initial forecast and the new information acquired.
@@
 The update step is repeated sequentially for each observation $Y_{i,t,k}$ given at time $t$.
-For each such repetition of the update step, the previous posterior values, $X_{t|t, k}$ and $P_{t|t, k}$, should be considered as the new prior values, $X_{t|t-1, k}$ and $P_{t|t-1, k}$.
+For each such repetition, the previous posterior values, $X_{t|t, k}$ and $P_{t|t, k}$, are considered as the new prior values, $X_{t|t-1, k}$ and $P_{t|t-1, k}$.
 If the observation $\boldsymbol{Y}_{t,k}$ is completely missing at time $t$, the update step is skipped and

### 2013-10-20 Presentation/Presentation.tex
Added Presentation

\documentclass{beamer}
%\usetheme{Madrid} % My favorite!
%\usetheme{Boadilla} % Pretty neat, soft color.
%\usetheme{default}
%\usetheme{Warsaw}
%\usetheme{Bergen} % This template has nagivation on the left
\usetheme{Frankfurt} % Similar to the default 
%with an extra region at the top.
%\usecolortheme{seahorse} % Simple and clean template
%\usetheme{Darmstadt} % not so good
% Uncomment the following line if you want %
% page numbers and using Warsaw theme%
% \setbeamertemplate{footline}[page number]
%\setbeamercovered{transparent}

\newcommand{\argmin}{\operatornamewithlimits{arg\ min}}
\newcommand{\argmax}{\operatornamewithlimits{arg\ max}}
\newcommand\ville[1]{{\color{red} #1 }}
%    \mymarginpar{\raggedright\hbadness=10000\tiny\it #1\par}}

\newcommand{\myN}{\hbox{N\hspace*{-.9em}I\hspace*{.4em}}}
\newcommand{\myZ}{\hbox{Z}^+}
\newcommand{\R}{\mathbb{R}}
\renewcommand{\P}{\mathbb{P}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\logit}{\text{logit}}

\setbeamercovered{invisible}
% To remove the navigation symbols from 
% the bottom of slides%
\setbeamertemplate{navigation symbols}{} 
%
\usepackage{graphicx}
\usepackage{amsmath,amsthm,amssymb}
\usepackage{subfigure}
%\usepackage{bm}         % For typesetting bold math (not \mathbold)
%\logo{\includegraphics[height=0.6cm]{yourlogo.eps}}
%
\title[Probability Aggregation]{Probability Aggregation: Dynamic Hierarchical Modeling of Sparse Expert Beliefs}
\author{Ville A. Satop\"a\"a,  Shane T. Jensen, Lyle H. Ungar, Barbara A. Mellers, Phil E. Tetlock}
\institute[University of  Pennsylvania]
{
University of Pennsylvania \\
\medskip
{\emph{satopaa@wharton.upenn.edu}}
}
\date{\today}
% \today will show current date. 
% Alternatively, you can specify a date.
%
\begin{document}
%
\begin{frame}
\titlepage
\end{frame}
%

\begin{frame}
\frametitle{Motivation}
\begin{enumerate}
\item Probability aggregation is important in decision-making (medical diagnosis, political and socio-economic foresight, meteorology, etc.)
\item Focus has been on developing one-time aggregation procedures that consult the expert's advice only once.
\item It is common for experts studying real problems to update their probability estimates over time.\\
\vspace{1em}
 $\Rightarrow$ Time-series probability aggregation
\end{enumerate}
\end{frame}

\begin{frame}
\frametitle{Data Example}

\begin{figure}[h!]
\centering
\vspace{-1.5em}
	\includegraphics[width=  \textwidth]{LegendExamplePlot} % requires the graphicx package
\vspace{-1.5em}

\subfigure{
\hspace*{-1em}  \includegraphics[width= 2.4in]{Figures/ExamplePlot2}
%\caption{\textit{Sample-Then-Optimize} with logarithmic score}
\label{Example2}
}

\caption{Will the Nikkei 225 index finish trading at or above 9,500 on 30 September 2011?}
\label{ExamplePlots}
\end{figure}
\end{frame}

\begin{frame}
\frametitle{Data Summary}
\begin{itemize}
\item $166$ questions about international political events.
\item Length of a question ranged from 4 to 418 days with an average of 73 days.
\item  $\sim 2,400$ experts participating.
\item On average, an expert participated in 55 questions.
\item On average, an expert gave 1 probability forecast per question.\\
\vspace{1em}
 $\Rightarrow$ The data is noisy and sparse.
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Data Example Returned}

\begin{figure}[h!]
\centering
\vspace{-1.5em}
	\includegraphics[width=  \textwidth]{LegendExamplePlot} % requires the graphicx package
\vspace{-1.5em}

\subfigure{
\hspace*{-1em}  \includegraphics[width= 2.4in]{Figures/ExamplePlot2b}
%\caption{\textit{Sample-Then-Optimize} with logarithmic score}
\label{Example2}
}

\caption{Will the expansion of the European bailout fund be ratified by all 17 Eurozone nations before 1 November 2011?}
\label{ExamplePlots}
\end{figure}
\end{frame}

\begin{frame}
\frametitle{Research Goals}
\textbf{Core Research Problem:} Construct an \textit{interpretable} model for capturing a sharp and well-calibrated \textit{crowd belief} across time.
\vspace{1em}

\begin{block}{Calibration}
A probability forecast $p$ for the $k$th question is calibrated if $\P(Z_k = 1 | p) = \E[Z_k | p] = p$ almost surely.
\end{block}
\begin{block}{Sharpness}
A probability forecast $p$ is sharper than a probability forecast $q$ if $\E[(p-p_0)^2] > \E[(q-p_0)^2]$, where $p_0$ is the baseline probability for the $k$th question.
\end{block}

\end{frame}

\begin{frame}
\frametitle{Research Goals}
\textbf{Specific Research Objectives:}
\begin{enumerate}
\item Study bias across different levels of self-reported expertise.
\item Make predictions.
\item Give many question-specific quantities that have easy interpretations. 
\end{enumerate}
\vspace{1em}
\end{frame}
%

\begin{frame}
\frametitle{Model}
Our model is
\begin{eqnarray*}
\boldsymbol{Y}_{t,k}  &=& \text{log-odds for the $k$th question at time }t\\
\boldsymbol{Y}_{t,k}  &\sim&  \mathcal{N}_{n_k}\left( \boldsymbol{M}_k \boldsymbol{b} X_{t,k}, \sigma^2_k\boldsymbol{I}_{n_k} \right)\\
X_{t,k} &\sim& \mathcal{N} \left( \gamma_k X_{t-1,k}, \tau^2_k\right)
\end{eqnarray*}
where
\begin{itemize}
\item Only $\boldsymbol{Y}_{t,k}$ is observed.
\item $X_{t,k}$ is a sharp and well-calibrated logit-probability for the $k$th question at time $t$.
\item $\boldsymbol{b}$ is a vector of bias terms for the expertise groups.
\item $\gamma_k$ drives the hidden process toward extremes as $t \to T$. 
\item $\sigma^2_k$ quantifies random noise around expert forecasts. 
\item $\tau^2_k$ quantifies random shocks to the hidden process.
\end{itemize}

\end{frame}
%
\begin{frame}
\frametitle{Identifiability}
\textbf{This model is not identifiable:} Let $a$ be some non-negative constant. Then  $(\boldsymbol{b}/a, aX_{t,k}) \neq (\boldsymbol{b}, X_{t,k})$ give the same likelihood for $\boldsymbol{Y}_{t,k}$.

\vspace{1em}
$\Rightarrow$ Parameter estimates drift!
\end{frame}

 \begin{frame}
\frametitle{Sample-And-Calibrate ($\text{SAC}_{LOG}$)}

\begin{block}{Sampling Step}
\begin{enumerate}
\item Fix $b_3 = 1$.
\item Sample constrained estimates: $\hat{X}_{t,k}(1)$, $\hat{\boldsymbol{b}}(1)$, etc.
\end{enumerate}
\end{block}

\begin{block}{Calibration Step}
\begin{enumerate}
\item Estimate $ \hat{b}_3 =  \argmax_{b \in \R / \{0\}} \sum_{k=1}^K \sum_{t=1}^{T_k}  S\left(Z_k, \hat{X}_{k,t}(1) / b \right)$
where $S$ is a strictly proper scoring rule such as the logarithmic score $Z \log\left(\logit^{-1}(X)\right) + (1-Z) \log\left(1-\logit^{-1}(X)\right)$

\item Adjust the constrained estimates:
\small
\vspace{-1em}
\begin{eqnarray}
 \hat{X}_{t,k}&=& \hat{X}_{k,t}(1) / \hat{b}_3 \nonumber\\
 \hat{\boldsymbol{b}}&=& \hat{\boldsymbol{b}}(1) \hat{b}_3 \nonumber\\
   \hat{\tau}_{k}^2&=& \hat{\tau}_{k}^2(1)/ \hat{b}_3^2\nonumber\\
  \hat{\sigma}_{k}^2&=& \hat{\sigma}_{k}^2(1)\nonumber\\
  \hat{\gamma}_{k}&=& \hat{\gamma}_{k}(1)\nonumber
\end{eqnarray}

\end{enumerate}
\end{block}
\end{frame}

\begin{frame}
\frametitle{Why Does This Work?}
\begin{itemize}
\item The maximization is really over different choices of $\hat{X}_{t,k}$. Since this process is assumed to be sharp and well-calibrated, it is natural to estimate it by maximizing a proper scoring rule.
\item If $S$ is the logarithmic scoring rule and the data is \textit{balanced}, then this process is approximately equivalent to Platt Calibration, which has been shown to yield good calibration under various modeling scenarios. \cite{platt1999probabilistic, niculescu2005obtaining}
\end{itemize}
\end{frame}

%  
%  \begin{frame}
%\frametitle{Balancing the Data}
%\begin{enumerate}
%\item Find a partition, $S_0$ and $S_1$, of the problem indices, $\{1, 2, \dots, 166\}$, such that the condition
%\begin{eqnarray*}
%\sum_{k \in S_0} T_k  = \sum_{k \in S_1} T_k  = \frac{1}{2}\sum_{k=1}^{166} T_k
%\end{eqnarray*}
%is as closely approximated as possible. 
%\item Let $\tilde{Z}_k = x$ with $x = 0,1$, and
%\begin{align*}
%\{\tilde{\boldsymbol{p}}_{t,k}\}_{t=1}^{T_k} &=  \begin{cases} 
%\{1-\boldsymbol{p}_{t,k}\}_{t=1}^{T_k} & \text{if } Z_k = 1-x\\
%\{\boldsymbol{p}_{t,k}\}_{t=1}^{T_k} & \text{if } Z_k = x 
%\end{cases}
%\end{align*}
%for $k \in S_x$. 
%\item $\left\{\left(\tilde{Z}_k, \{\tilde{\boldsymbol{p}}_{t,k} \}_{t=1}^{T_k}\right) \right\}_{k=1}^{166}$ is a balanced version of the data.
%\end{enumerate}
%\end{frame}

\begin{frame}
\frametitle{Calibration and Sharpness: in-sample}

\begin{figure}[h!]
\subfigure[\scriptsize SDLM In-Sample]{
\hspace*{-3em} \includegraphics{Figures/CalibrationNONE10}
\label{CalibrationNONEa}
}
\subfigure[\scriptsize $\text{SAC}_{\text{LOG}}$ In-Sample]{
\includegraphics{Figures/CalibrationLOG10}
\label{CalibrationPLATTLOG10}
}
\caption[Optional caption for list of figures]{Comparing the approaches in terms of sharpness and calibration.}
\label{Calibration-Out}
\end{figure}

\end{frame}

\begin{frame}
\frametitle{Calibration and Sharpness: out-of-sample}

\begin{figure}[h!]
\subfigure[ \scriptsize SDLM Out-of-Sample]{
\hspace*{-3em} \includegraphics{Figures/CalibrationNONE}
\label{CalibrationNONEb}
}
\subfigure[ \scriptsize $\text{SAC}_{\text{LOG}}$ Out-of-Sample]{
 \includegraphics{Figures/CalibrationSTO-Log}
\label{CalibrationSTC-Log}
}
\caption[Optional caption for list of figures]{Comparing the approaches in terms of sharpness and calibration.}
\label{Calibration-Out}
\end{figure}
\end{frame}

\begin{frame}
\frametitle{Out-of-sample Prediction}
\begin{table}[htbp]
   \centering
   \scriptsize
      \begin{tabular}{l llll} % Column formatting, @{} suppresses leading/trailing space
 \multicolumn{5}{ c }{Scores by Day}\\
Model & \multicolumn{1}{ c }{All} & \multicolumn{1}{ c }{Short} & \multicolumn{1}{ c }{Medium} & \multicolumn{1}{ c }{Long}\\ \hline
SDLM & 0.100 (0.156) & 0.066 (0.116) & 0.098 (0.154) & 0.102 (0.157)\\ 
$\text{BSAC}_{\text{LOG}}$ & 0.097 (0.213) & \textbf{0.053} (0.147) & 0.100 (0.215) & 0.098 (0.215)\\ 
$\text{SAC}_{\text{BRI}}$ & 0.096 (0.190) & 0.056 (0.134) & 0.097 (0.190) & 0.098 (0.192)\\ 
$\text{SAC}_{\text{LOG}}$ & \textbf{0.096} (0.191) & 0.056 (0.134) & \textbf{0.096} (0.189) & \textbf{0.098} (0.193)\\ 
EWMBA & 0.102 (0.203) & 0.060 (0.124) & 0.110 (0.201) & 0.103 (0.206)\\ 
EWMLA & 0.102 (0.199) & 0.061 (0.130) & 0.111 (0.214) & 0.103 (0.200)\\ 
EWMA & 0.111 (0.142) & 0.089 (0.100) & 0.111 (0.136) & 0.112 (0.144)\\ 
&&&&\\
\multicolumn{5}{ c }{Scores by Problem}\\
Model & \multicolumn{1}{ c }{All} & \multicolumn{1}{ c }{Short} & \multicolumn{1}{ c }{Medium} & \multicolumn{1}{ c }{Long}\\ \hline
SDLM & 0.089 (0.116) & 0.064 (0.085) & 0.106 (0.141) & 0.092 (0.117)\\ 
$\text{BSAC}_{\text{LOG}}$ & 0.083 (0.160) & \textbf{0.052} (0.103) & 0.110 (0.198) & 0.085 (0.162)\\ 
$\text{SAC}_{\text{BRI}}$& 0.083 (0.142) & 0.055 (0.096) & 0.106 (0.174) & 0.085 (0.144)\\ 
$\text{SAC}_{\text{LOG}}$ & \textbf{0.082} (0.142) & 0.055 (0.096) & \textbf{0.105} (0.174) & \textbf{0.085} (0.144)\\ 
EWMBA & 0.090 (0.156) & 0.063 (0.101) & 0.118 (0.186) & 0.091 (0.161)\\ 
EWMLA & 0.090 (0.159) & 0.064 (0.109) & 0.120 (0.200) & 0.090 (0.159)\\ 
EWMA & 0.104 (0.105) & 0.092 (0.081) & 0.119 (0.125) & 0.103 (0.107)\\ 
   \end{tabular}
   \caption{Brier Scores based on 10-fold cross-validation.}
   \label{prediction}
\end{table}

\end{frame}

\begin{frame}
\frametitle{Group-level bias}
\vspace{-0.9em}
\begin{figure}[!ht]
\includegraphics[width = 4in]{Figures/BiasesBoxplots}
\label{BiasesLOGBoxplots}
\vspace{-1em}
\caption[Optional caption for list of figures]{Comparing the bias-levels across self-reported expertise under different approaches.}
\label{Biases}
\end{figure}
\end{frame}

\begin{frame}
\frametitle{Conclusion}
\begin{itemize}
\item Extended probability aggregation towards unexplored areas.
\item Gave an interpretable model that offers a solution to the main problems in time-series probability aggregation.
\item Applied this to a large probability forecasting dataset.
\item Many future extensions.
\end{itemize}

\end{frame}

\begin{frame}
\centering 
\hspace{12em}Questions?
\end{frame}

       \begin{frame}
\frametitle{References}
\tiny
 \bibliographystyle{plainnat}
%\bibliographystyle{imsart-nameyear}
\bibliography{biblio}		% expects file "myrefs.bib"
\end{frame}

\end{document}

### 2014-01-31 Revision/Response/response.tex
added revision

\documentclass[aoas, preprint]{imsart}

%\long\def\authornote#1{%
%        \leavevmode\unskip\raisebox{-3.5pt}{\rlap{$\scriptstyle\diamond$}}%
%        \marginpar{\raggedright\hbadness=10000
%        \def\baselinestretch{0.8}\tiny
%        \it #1\par}}
%\newcommand{\ville}[1]{\authornote{NOTE TO SELF: #1}}

%\usepackage{amsthm,amsmath,natbib}
%\RequirePackage[colorlinks,citecolor=blue,urlcolor=blue]{hyperref}

% provide arXiv number if available:
%\arxiv{arXiv:0000.0000}

% put your definitions there:
\startlocaldefs
\endlocaldefs

\newcommand{\argmin}{\operatornamewithlimits{arg\ min}}
\newcommand{\argmax}{\operatornamewithlimits{arg\ max}}
\newcommand\ville[1]{{\color{red} #1 }}
%    \mymarginpar{\raggedright\hbadness=10000\tiny\it #1\par}}

%\usepackage[section]{placeins}
%\usepackage{amsmath} 
\usepackage{times}
\usepackage{amsmath,amsthm,amssymb}
\usepackage{fancyhdr}
\usepackage{moreverb}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{array}
\usepackage{url}
\usepackage{multirow} 
\usepackage[boxed, section]{algorithm}
%\usepackage{algorithm}
%\usepackage{algorithmic}
%\usepackage{cite}
\usepackage{multirow} 
\usepackage{rotating}
%\usepackage[margin=1.2in]{geometry}
%\usepackage{geometry}
%\usepackage{fix-cm}
\usepackage{subfigure}
\usepackage{bbm}
\usepackage{color} 
%\usepackage{natbib}
\usepackage{xcolor,xparse}
\usepackage{xparse}
%\usepackage{setspace}
%\usepackage{booktabs}
%\usepackage{subfig}
%\usepackage{caption, subcaption}
\usepackage[OT1]{fontenc}
\usepackage[colorlinks,citecolor=blue,urlcolor=blue]{hyperref}
\usepackage{natbib}

\startlocaldefs
\numberwithin{equation}{section}
\theoremstyle{plain}
\newtheorem{thm}{Theorem}[section]
\endlocaldefs

%\renewcommand{\baselinestretch}{1.2}
%\setlength{\topmargin}{-0.3in}
%\setlength{\textwidth}{6in}
%\setlength{\textheight}{8.5in}
%\setlength{\oddsidemargin}{0.25in}
%\setlength{\evensidemargin}{0.25in}
%\raggedbottom

%\doublespacing

\allowdisplaybreaks

% Math Macros.  It would be better to use the AMS LaTeX package,
% including the Bbb fonts, but I'm showing how to get by with the most
% primitive version of LaTeX.  I follow the naming convention to begin
% user-defined macro and variable names with the prefix "my" to make it
% easier to distiguish user-defined macros from LaTeX commands.
%
\newtheorem{mydef}{Definition}
\newcommand{\myN}{\hbox{N\hspace*{-.9em}I\hspace*{.4em}}}
\newcommand{\myZ}{\hbox{Z}^+}
\newcommand{\R}{\mathbb{R}}
\renewcommand{\P}{\mathbb{P}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\logit}{\text{logit}}
\NewDocumentCommand{\framecolorbox}{oommm}
 {% #1 = width (optional)
  % #2 = inner alignment (optional)
  % #3 = frame color
  % #4 = background color
  % #5 = text
  \IfValueTF{#1}
   {%
    \IfValueTF{#2}
     {\fcolorbox{#3}{#4}{\makebox[#1][#2]{#5}}}
     {\fcolorbox{#3}{#4}{\makebox[#1]{#5}}}%
   }
   {\fcolorbox{#3}{#4}{#5}}%
 }

\newcommand{\myfunction}[3]
{${#1} : {#2} \rightarrow {#3}$ }

\newcommand{\myzrfunction}[1]
{\myfunction{#1}{{\myZ}}{{\myR}}}

\newcommand{\mysection}[1]
{\noindent {\bf {#1}}}

%%%%%% Begin document with header and title %%%%%%%%%

\begin{document}
\begin{frontmatter}
% "Title of the paper"
\title{Probability Aggregation in Time-Series: Dynamic Hierarchical Modeling of Sparse Expert Beliefs\\
\\
(Revision of Manuscript AOAS1309-039)\\
\\
 POINT-BY-POINT RESPONSE}
\runtitle{Probability Aggregation in Time-Series}

% indicate corresponding author with \corref{}
% \author{\fnms{John} \snm{Smith}\corref{}\ead[label=e1]{smith@foo.com}\thanksref{t1}}
% \thankstext{t1}{Thanks to somebody} 
% \address{line 1\\ line 2\\ printead{e1}}
% \affiliation{Some University}

%\begin{aug}
%  \author{\fnms{First}  \snm{Author}\corref{}\thanksref{t2}\ead[label=e1]{first@somewhere.com}},
%  \author{\fnms{Second} \snm{Author}\ead[label=e2]{second@somewhere.com}}
%  \and
%  \author{\fnms{Third}  \snm{Author}%
%  \ead[label=e3]{third@somewhere.com}%
%  \ead[label=u1,url]{http://www.foo.com}}
%
%  \thankstext{t2}{Footnote to the first author with the `thankstext' command.}
%
%  \runauthor{F. Author et al.}
%
%  \affiliation{Some University and Another University}
%
%  \address{Address of the First and Second authors,\\ 
%          \printead{e1,e2}}
%
%  \address{Address of the Third author,\\
%          \printead{e3,u1}}
%
%\end{aug}

\begin{aug}
\author{\fnms{Ville A.} \snm{Satop\"a\"a}\corref{}\ead[label=e1]{satopaa@wharton.upenn.edu}},
\author{\fnms{Shane T.} \snm{Jensen}\ead[label=e3]{stjensen@wharton.upenn.edu}},
\author{\fnms{Barbara A.} \snm{Mellers}\ead[label=e4]{mellers@wharton.upenn.edu}},
\author{\fnms{Philip E.} \snm{Tetlock}\ead[label=e5]{tetlock@wharton.upenn.edu}},
\and
\author{\fnms{Lyle H.} \snm{Ungar}\ead[label=e2]{ungar@cis.upenn.edu}}

\runauthor{Satop\"a\"a et al.}

 \affiliation{Department of Statistics, The Wharton School of the University of Pennsylvania\\   \printead{e1,e3}}
 \affiliation{Department of Psychology, University of Pennsylvania\\ \printead{e4,e5}}
 \affiliation{Department of Computer and Information Science, University of Pennsylvania\\ \printead{e2}}
 \address{Philadelphia, PA 19104- 6340, USA\\  \printead{e1}}
\end{aug}
\end{frontmatter}

We thank both the editor and reviewers for their helpful comments on our manuscript.  We have addressed each of these comments in our revision, as well as providing a point-by-point response below. The comments are in boldface. Our comments are in normal typeface. 
\clearpage
\section*{\Large Editor}
\textbf{Concerning the out-of-sampling applicability of your methods, the cross-validation technique used in Section 6 indeed does not yield time-forward predictions, as would be required in practice.  If you could modify your technique so that it can be used for time-forward predictions, that would be most useful.  At a minimum, you should be very clear about the limitations of your approach, and you ought to discuss in detail how these issues could be remedied in future work.}\\
We have added a brief paragraph on time-forward propagation. In practice, however, social scientists tend to be more interested in the single ``best" estimate of the probability given all the information that is available at the moment. This is what our model addresses.
\\
\\
\textbf{In my own reading, I also felt that your paper could be usefully shortened.  For example, Figures 1 and 5 are redundant - so please remove Figure 1, and move Figure 5 forward to Section 1.  Also, please move the Appendix to an Online Supplement.  Throughout, please check for redundancies and remove them, unless the exposition requires otherwise.}
\\
The appendix has been moved to an Online Supplement.  Figure 1 has been replaced with Figure 5. Some sections, including the discussion on Platt calibration and summary of findings, have been removed. We also have removed many redundancies. The synthetic data section has also been shortened and now discusses only the relevant results therein.\\
\\
\textbf{As regards Referee 3's second specific comment, I agree that the definitions on page 6 are unlikely to be appreciated, unless you introduce the (rather technical) prediction space setting that is implicit in Ranjan and Gneiting (2010) and explicit in Gneiting and Ranjan ("Combining predictive distributions", Electronic Journal of Statistics 7, 1747-1782, 2013).  Therefore, let me suggest that you discuss these definitions informally (and within the text, rather than using displayed environments) and refer to the aforementioned references for technical details.}
\\
The definitions are now only discussed in the text with a reference to the aforementioned references for technical details.\\
\\
\textbf{In the references list, please unify the style used therein, capitalize all journal titles, expand acronyms such as AAAI, ECAI, and UAI, provide an update for Mellers et al. (2013), add the missing page ranges for Metropolis et al. (1953) and Niculescu-Mizil and Caruana (2005), the missing publisher for Pepe (2003), the missing authors, editors, and booktitle for Platt et al. (1999), and the multiply missing information for Wilson and Wilson (1994) and Wright et
al. (1994).}\\
- All acronyms have been expanded. We have aimed to make the style more unified.  All journal names have been capitalized.\\
- Mellers et al. (2013): This paper is currently under revision after receiving a friendly reject \& revise.\\
- Metropolis et al. (1953): The missing page ranges have been added.\\
- Niculescu-Mizil and Caruana (2005): No more referenced in the paper.\\
- Pepe (2003): Thank you for noticing this. It had been listed as an article in the .bibtex file and hence skipped the publisher info.\\
- Platt et al. (1999): No more referenced in our paper.\\
- Wilson and Wilson (1994): This has been corrected.  As this is a book, only the publisher was added.\\
- Wright et al. (1994): This reference seems to be complete.\\

\clearpage
\section*{\Large Referee 1}
\textbf{This is an interesting paper describing a model for probabilistic forecast aggregation over time.  The model is novel and has the ability to handle sparse, irregular data.  I have a few major comments and many smaller comments.\\
\\
Motivation: I think that more could be done to relate the proposed model to existing aggregation models.  As the authors state, the usual time series models do not apply to these data.  However, the proposed model can be viewed as an extension of a few psychometric models.  In particular, the model presented in Section 3 is a hierarchical factor analysis model with the traditional roles of subjects and items reversed.  Batchelder \& Romney (1988) developed a non-time-series version of this model to aggregate judgments on problems where no outcome Z exists (under the name "Cultural Consensus Theory").  Karabatsos \& Batchelder (2003) further developed Bayesian versions of this model.  Thus, if we stop after Section 3, the proposed model is an extension of cultural consensus models to time series.\\}
Thank you for pointing us to these papers. We were not aware of this work and it has definitely been an interesting read. There are so many static aggregation papers in the machine learning literature that missing some papers is unavoidable. To our understanding, the model described in Karabatsos \& Batchelder (2003) is applied to binary responses to estimate the solution key (which is also considered binary). Therefore the setup is slightly different from the probability aggregation paradigm considered in our paper. However, a more recent paper by Batchelder et al. called \textit{Cultural consensus theory: aggregating continuous responses in a finite interval} aims to use CCT to perform probability aggregation. We have cited this work in our updated introduction. This work, however, also considers static aggregation and is therefore very different in nature to our work. For this reason, and also for the sake of shortening the already too long paper, we have decided to not include an extensive literature review of the CCT. 
\\
\\
\textbf{Section 4 highlights the rotation issues inherent in factor analysis models.  The authors fit the model under one constraint, then rotate the parameters to an equivalent set of parameters that minimize a scoring rule.  So, following rotation, this model is no longer a strict consensus model because the outcomes Z impact the rotation.\\
\\
In general, I believe the relationship to factor analysis and cultural consensus provides a nice motivation for the model and its rotation issues (i.e., the calibration step).
\\
\\
Staying with the rotation issue, there are alternative constraints one may adopt instead of fixing one element of b to 1.  For example, I believe the model can be identified by restricting the sum of the elements of b to equal one (and possibly restricting the variance to 1).  This restriction may avoid the issue described on p 12, first full paragraph.}
\\
It is enough to constrain the sum to gain identifiability, i.e. no additional constraint on the variance is needed. Constraining the mean of the elements of $b$ to equal to one (or equivalently, their sum to equal to five) was considered at early stages of our implementation. 
%This implies that the groups are, on average, unbiased during the sampling step. Based on our experience this is an unreasonable assumption. 
%Unlike restricting, say, $b_3 = 1$, constraining the sum imposes unnecessary dependencies among the $b_i$'s. 
Unfortunately, the problem discussed in synthetic data section is more due to the calibration step than the choice of restriction. In essence the problem occurs because the calibration cannot move any prediction from one side of zero to the other. To see how this happens even under the sum restriction, notice that by enforcing the bias terms to sum to, say five, implies that the groups are, on average, unbiased during the sampling step. If, however, the experts are on average very under-confident, then it is likely that the resulting estimates of the hidden states are on the wrong side of zero. This is again a problem that our one-dimensional calibration cannot fix.
\\
\\
\textbf{This is a complex model, and there were a few issues that could be described in more detail.  For example, I did not see details on how the parameters $\mu_0$ and $\sigma^2_0$ were set.  This seems to be a place where one might artificially improve model performance.  }
\\
Thank you for pointing this out. In our studies we always set $\mu_0$ and $\sigma^2_0$ to 0 and 1, respectively. This is now mentioned in the text. Based on our experience, as long as these values are reasonable, they do not influence the results significantly. These values are mere baseline initializations that are rapidly washed out by the data. 
\\
\\
\textbf{Additionally, there could be further discussion about the model's handling of data sparsity.  Currently, the only details appear in the appendix (stating that an update is made for each observation at time $t$), even though data sparsity is offered as a motivation for this model.}
\\
We have added `` The first step of our Gibbs sampler is to sample the hidden states via the \textit{Forward-Filtering-Backward-Sampling} (FFBS) algorithm. FFBS first predicts the hidden states using a Kalman filter and then performs a backward sampling procedure that treats these predicted states as additional observations (see, e.g., \cite{carter1994gibbs, migon2005dynamic} for details on FFBS). Given that the Kalman filter can handle varying numbers or even no forecasts at different time points, it plays a very crucial role in our probability aggregation under sparse data."  Without describing the technical details of the Kalman Filter in the main text, it is difficult to explain how sparse data is handled. These details, however, are still kept in the Online Supplement. 
\\
\\
\textbf{Implementation: I wonder how the models were implemented, along with the speed of the model estimation.  I guess the models may take some time to run.  It would be nice if model estimation code were shared.}\\
We have added: "Our implementation of the sampling step is written in C++ and runs fast. For instance, to obtain a posterior sample of size 1,000 for 50 questions each with 100 time points and 50 experts takes about 215 seconds on the first author's computer (1.7 GHz Intel Core i5). "  The calibration step is currently run separately in R once the samples have been obtained from the sampling step. Given that the calibration requires a univariate optimization, it can be performed very fast. We plan to post our implementation upon publication.
\\
\\
\textbf{$BSAC_{log}$: I don't understand why we should expect this estimation method to be any different from the $SAC_{log}$ method, especially because flat priors are used.  In the Table 3 and Figure 3 results, it looks like there are some minor differences but not enough to be notable.  Hence, I am not convinced that the discussion of $BSAC_{log}$'s overconfidence is meaningful.}
\\
The main difference between $BSAC_{log}$ and $SAC_{log}$ is that $BSAC_{log}$ estimates the hidden states $\{X_{t,k}(1)\}$ and $\beta$ simultaneously (alternating between the two until convergence), whereas  $SAC_{log}$ first estimates $\{X_{t,k}(1)\}$ and then calibrates them by finding $\beta$. Therefore $BSAC_{log}$ is slightly more flexible than $SAC_{log}$. Unfortunately, it tends to overfit the data. This results into the overconfidence that it seen in our analyses. We have added ``Given that this model estimates $X_{t,k}(1)$ and $\beta$ simultaneously, it is a little more flexible than $\text{SAC}$."
\\
\\
\textbf{Finally, there were a variety of small comments and questions:\\
\\
Top p 6: I do not understand the phrase "the entire b vector is shared".  Does this mean that b has a single element?}
\\
This means that each of the elements of $b$ are considered the same across the 166 questions. We have modified the text to make this more clear. 
\\
\\
\textbf{Bottom p 9, expressing logistic regression as linear regression: I don't understand the normal error in this equation.  I would think that, if this has some equivalence to logistic regression, the error term should follow the logistic distribution.}
\\
 The error term was not needed. This entire paragraph has been removed for the sake of brevity. \\
\\
\textbf{p. 11, third line: I believe should read "$166 - N_{train}$".}
\\
This has been corrected. Thanks for catching it!
%Only a single dataset is generated under the synthetic data section. The algorithms are told the experts' forecasts and whether the events happened or not. After this we can estimate their accuracy to estimate the true hidden states, namely $X_{t,k}$. Therefore this evaluation is a little different from the real-world data section, where we do not have access to the true hidden states $X_{t,k}$ and hence must use the event indicators for performance evaluation.
\\
\\
\textbf{p. 11: Please define the EWMA method before describing results associated with it.}
\\
The definition of EWMA has been moved.\\
\\
\textbf{Figure 2: Would be nice to have common y-axis limits for each row of the graph.}
\\
This is a good idea. For the sake of brevity, however, we have ended up removing most of these plots. We only kept the one that tells something non-obvious about our model, and then briefly mention the other results within the text. After all this section is a mere sanity check.
\\
\\
\textbf{Appendix, p 26, top: I was initially confused because, e.g., e is missing an i subscript even though i appears in the equation.  I guess this is related to the text immediately below, stating that a new update is repeated for each observation.  But perhaps the notation could be clearer.}
\\
 The Appendix has been moved to Online Supplement, where the update step is not described in an algorithmic way and with improved notation. We believe that it is clearer now.

\clearpage
\section*{\Large Referee 2}

\textbf{This paper attempts a difficult problem of combining sparse expert probabilities which evolve over time. Three characteristics of the approach are significant }
\begin{itemize}
	\item \textbf{Experts are allowed to give multiple forecasts over time till the geo-political event materializes}
	\item \textbf{Experts are can update forecasts at irregular intervals of their liking during the same period}
	\item \textbf{Experts are grouped into multiple categories based on their level of self-acclaimed expertise}
\end{itemize}

\textbf{The extant literature doesn’t address the issue of dynamic aggregation over time.  The combination methods developed in the literature won’t be able to do justice to evolving expert opinions over time as they will treat them as one combination problem at a given point of time.  Also, the small no. of expert forecasts at any given time would make the combined forecast at a given time weak.  The alternative is to borrow strength from past forecasts in a sensible way – which is explored in the present paper. }

\textbf{The other contribution of the paper is study various groups of experts based on their self-assessed expertise. The parameters used in the framework directly lend themselves to the study of group specific over-confidence/under-confidence, disagreement between experts, question difficulty and evolving expert opinions.   The combined forecast is shown to be empirically calibrated and sharp.   The estimation involves two stages a sample step and a calibrate step. In the sample step the model parameters are sampled using a Gibbs sampling paradigm. In the calibrate step a one –parameter optimization is used to calibrate the experts forecasts with the truth using a proper scoring rule.}

\textbf{The methodology presented in the paper for combination has been compared against simple exponentially weighted generalizations of existing methods like equally weighted probability forecasts and the beta-transformed probability forecast method proposed by Ranjan and Gneiting (2010).  However, these generalizations have been restricted to single combination parameters only. One wonders why these methods were not generalized to include group specific combination parameters. This seems to me the natural thing to do otherwise, the comparison to other methods won’t be fair.}\\
Given that EWMLA already includes group specific combination parameters, only EWMA and EWMLA need to be generalized. In the revised manuscript EWMA and EWMLA make use the expertise information by incorporating a weighted mean (instead of the simple equally weighted mean used in the previous version) of the forecasts. The results, however, look very similar.\\
\\

\textbf{Another limitation of the present work is that it masks question specific and individual specific bias terms and only estimates group specific bias irrespective of question and individual.   This won’t allow one to understand any outcome specific bias that may exist in the groups or individuals.  For example certain geo-political outcomes may be more desirable for people in general and individuals might be biased for that outcome.  It would be interesting to see if authors can come up with an framework in which bias can be studied at multiple levels, be it question/ outcome specific, individual specific or group of experts specific or even time-specific.  The present work only looks at group specific bias.  Perhaps a hierarchical approach to bias decomposition may be more suitable.}\\
We have added a sub-section called \textit{Bias Structure} that discusses a hierarchical approach to bias decomposition. In this subsection we briefly explain how the bias can be estimated at  question-specific and individual-specific. An extension to a time-specific bias structure is briefly mentioned in the conclusion.
\\

\textbf{Overall I am happy with the work as it explores a difficult and unexplored problem and offers some directions.  The modeling of the problem exposes various parameters which have an intuitive interpretation for the researcher and helps understand the expert opinions better.  However I would like to see the authors update Table-3 with group specific parameters used for other methods as well (in particular, EWMA, EWMBA, EWMLA).  Also, I would like to see some discussion on how bias (under-forecasting and over-forecasting) can be estimated and studied more generally at multiple levels – individuals, questions, groups and even time specific.}
\\

\clearpage
\section*{\Large Referee 3}
\textbf{This manuscript describes a model for aggregating probabilities judgements made by experts. It allows for the experts to update their beliefs at different points in time and it allows for different biases between different groups of experts. The general idea is very interesting, but the implementation is lacking. The model makes strong, simplistic assumptions, and the applicability in practice is questionable.}
\\
\\
\textbf{General Comments}
\begin{enumerate}
\item \textbf{The manuscript is generally well-written, but it is way too long. The writing is too leisurely throughout, and many issues and concepts are unnecessarily discussed two or more times.}\\
The paper is now much shorter. We have moved material to the Online Supplement, removed some of the unnecessary sections and graphs, and cut many redundancies.\\

\item \textbf{In my opinion, the model is unnecessarily simplistic. The authors assume that experts are indistinguishable beyond (self-reported!) expertise. Bias and variance in their forecasts are assumed to be constant over experts, questions, and time. The authors fit a Bayesian hierarchical model that would allow for sophisticated shrinkage and borrowing of strength between experts and between questions, but this is basically not exploited in the model at all. In fact, the authors claim (incorrectly) that estimating a bias term separately for each expert is not possible. For example, why not specify a prior for the ith expert’s bias such as $\beta_i | b_{j(i)},\sigma_2 \sim N(b_{j(i)},\sigma^2)$, where $j(i)$ is the group membership of the $i$th expert?}\\
Thank you for pointing this out. We have added a brief subsection that discusses question-specific and individual-specific bias structures. The choice made in this paper was partly driven by our real-world dataset and the preference of the social scientists working with it. One goal of this paper is to describe a basic framework that can be extended rather easily for any such analysis. \\

\item \textbf{ I wonder how applicable this methodology actually is in practice. While I cannot comment on how interesting the fitted model parameters are to social scientists in terms interpreting things like question difficulty, the results to me offer little in terms of interpretation beyond what would have been clear from the outset or what can be learned from simple plots of the data.
\\
What I consider the most interesting is to make predictions for future events, which I am not sure is really possible in real-world situations using the model. Since the methodology is retrospective (i.e., you do not do on-line, filtering inference as I think would be more appropriate), it seems to be necessary to wait until a large study like the one described has been completed, with all outcomes having been observed. Then, you can fit the model. Only after this can the methodology be used to make forecasts for other, future events of interest, but it is not clear that a similar data situation could arise, and even if so, if the assumptions and parameters from your fitted model are still applicable to this new situation. For example, are the new experts comparable to the old ones? Since there seem to be large differences in question difficulties, what difficulty does a particular new question have?}\\
The model is interesting for three different reasons:
\begin{enumerate}
\item Model inference: This is inherently interesting to social scientist.
\item Time-Forward Forecasting: Our methodology does not take long to fit in a retrospect fashion. Once the estimates have been obtained, the estimate of the final hidden state can be propagated forward. We have included a paragraph about this in the revised version of the paper.
\item Aggregating at Current Time Point: The goal is to obtain the ``best" probability estimate given all the information available now. This is the most interesting to social sciences as it reflects the current consensus about the event and its associated uncertainty. This is what our model was built to address. We evaluate this ability in our out-of-sample aggregation section.
\end{enumerate}
The new experts tend to be very similar to new ones. They do not need to be the same individuals as long as they, as a group, look similar. In terms of determining the difficulty of a new problem, this is an active and very difficult research problem. One can use old questions to model the difficulty. With no information but the description of the problem, the best one can do is to assign the new problem an average difficulty level.

\end{enumerate}

\textbf{Specific Comments}
\begin{enumerate}
\item \textbf{p. 2: The punctuation seems to be off in the second part of the second paragraph.}\\
This has been corrected.\\

\item \textbf{p. 6: Can you explain why it is reasonable to assume that the forecast maximizes sharpness subject to calibration? Also, in Definition 2, it seems like p, q, and p0 are all fixed numbers on the unit interval, so I am not sure with respect to which random variables the expectations are meant to be taken.}\\
$logit^{-1}(X_{t,k})$ represents the ``best" forecast under all the relevant information available at time $t$. This is an idealized forecast that should be assumed to be well-calibrated. Furthermore, given that it is based on all the information available, it ought to be as close to 0 or 1 as possible; that is, as sharp as possible. Due to the editor's suggestions, the technical definitions have been removed. 
\\

\item \textbf{p. 7: Instead of writing “drifts to inifity,” I think things would be more easily understandable if you wrote “drifts to positive or negative infinity."}\\
 This has been corrected.\\

\item \textbf{p. 8: For completeness, you should briefly mention in the main text which parameters are assumed to be random and what their priors are. Also, even after reading the appendix, it is still not clear to me what you mean by “parameter estimates.” All you are obtaining with the Gibbs sampler are samples from the posterior distribution. A loss function would be needed to obtain (point) estimates of the parameters.}\\
 We added ``Therefore the parameters of the model are $\boldsymbol{b}$, $\sigma^2_k$, $\gamma_k$, and $\tau^2_k$ for $k = 1, \dots, K$. Their prior distributions are chosen to be non-informative,  $p(\boldsymbol{b}, \sigma^2_k| \boldsymbol{X}_k) \propto \sigma^2_k$ and $p(\gamma_k, \tau^2_k| \boldsymbol{X}_k) \propto \tau^2_k$." and "These estimates are obtained by first computing a posterior sample via Gibbs sampling and then taking the average of the posterior sample."
\\

\item \textbf{p. 10: While the simulated data are not generated exactly from your model, it does seem like the data-generating model is much closer to your model then to the model underlying EWMA.}\\
The underlying mechanism to generate the hidden states is very different from the one given in our dynamic linear model. The only similarity is in how the noise and bias are added to the hidden states. The noise in this case has mean zero and can be reduced by averaging. Therefore if the model can determine the bias terms accurately; that is, determine the extent to which each expertise group should be shifted for improved calibration, the model's performance ought to be very good.

Notice that this section has been updated to include an expertise-weighted version of EWMA. This procedure uses 6 parameters to calibrate the group-specific average probabilities. This is a much more flexible calibration function than the one in our SAC model. It is therefore expected that under an easy forecasting setup, such as the one in our paper, EWMA performs better than SAC. It is not entirely clear whether this comparison is completely fair. It serves, however, only a sanity check. The more important comparison is done under the real-world data. 
\\

\item  \textbf{p. 11: EWMA should be defined here, not much later in the manuscript.}\\
 The definition of EWMA has been moved.\\

\item \textbf{p. 12: EWMA outperforms SAC roughly when $\beta \in [.95, 1.5]$, which in my opinion does not correspond to a “very small” bias. If I understand correctly, $\beta = 1$, for example, actually means that the bias varies from 0.5 to 2.}\\
That is correct: $\beta$ is a multiplier for the vector $[1/2, 3/4, 1, 4/3, 2/1]$. When $\beta  = 1$, the average bias among the experts is around 1.12. This bias is not very large.
\\
\item \textbf{p. 14: Why is the SDLM model not hierarchical? The process model can clearly be written in a hierarchical way, and there are also parameters that (I believe) have prior distributions.}\\
We wanted to say that the model does not have any parameters that are shared across different questions. This has now been changed to "Given that this model does not share any parameters across different questions, estimates of the hidden process can be obtained directly for the questions in the testing set without fitting the model first on the training set."\\

\item \textbf{p. 14/15: For all models, how did you choose the number of iterations in the Gibbs sampler? Should the number of MCMC iterations be larger for the fully Bayesian model, since the hidden state is sampled using the Metropolis algorithm? By the way, if you assumed a probit-link instead of a logit-link, the full conditional distributions are available in closed form and no Metropolis steps are necessary.}\\
We have added: "Before introducing the competing models, note that all choices of thinning and burn-in made in this section are conservative and have been made based on pilot runs of the models.  This was done to ensure a posterior sample that has low autocorrelation and arises from a converged chain." The number of iterations for the fully Bayesian model should be a bit larger. We first made choices for this model and then used the same choice for SAC.
\end{enumerate}

\bibliographystyle{imsart-nameyear}
\bibliography{biblio}		% expects file "myrefs.bib"

\end{document}

### 2014-01-31 Revision/Paper/paper.tex
added revision

--- 2013-09-29 paper.tex
@@
 \usepackage{natbib}
+\usepackage{algorithmicx}
+\usepackage{algpseudocode}
 
@@
 Even though several experiments have been conducted to show that experts are generally poorly calibrated (see, e.g., \citet{cooke1991experts, shlyakhter1994quantifying}), relative differences can occur among different types of experts.
-In particular, \citet{wright1994coherence} argue that a higher level of self-reported expertise can be associated with better calibration.
+In particular, \citet{wright1994coherence} argue that higher self-reported expertise can be associated with better calibration.
 
@@
 In this setting an expert could always  assign a constant probability of 0.45 and be well-calibrated.
-This assessment, however, can be made without any subject-matter expertise or special training.
+His assessment, however, can be made without any subject-matter expertise.
 For this reason the long-term frequency is often considered as the baseline probability -- a naive assessment that provides the decision-maker very little extra information.
-Therefore experts should aim to make probability assessments that are as far from the baseline as possible.
+Experts should aim to make probability assessments that are as far from the baseline as possible.
 The extent to which their probabilities differ from the baseline is measured by \textit{sharpness} (\citet{gneiting2008rejoinder, winkler2008comments}).
 If the experts are both sharp and well-calibrated, they can forecast the behavior of the process with high certainty and accuracy.
-From the decision-maker's perspective their advice is highly useful as it involves much information and very little risk.
-Therefore, to summarize, useful probability estimation should aim to maximize sharpness subject to calibration (see, e.g., \citet{raftery2005using, murphy1987general}).
-This is a well-defined goal that has led to a wide range of novel and insightful observations in probability forecasting.
+Therefore useful probability estimation should aim to maximize sharpness subject to calibration (see, e.g., \citet{raftery2005using, murphy1987general}).
 
@@
 There is strong empirical evidence that bringing together the strengths of different experts by combining their probability forecasts into a single consensus, known as the \textit{crowd belief},  improves predictive performance.
-Prompted by the many applications of probability forecasts, including medical diagnosis (\citet{wilson1998prediction, pepe2003statistical}, political and socio-economic foresight (\citet{tetlock2005expert}), and meteorology (\citet{sanders1963subjective, vislocky1995improved, baars2005performance}), researchers have proposed many approaches to combining probability forecasts (see, e.g., \citet{Ranjan08, satopaa} for some recent studies, and \citet{Genest, Wallsten97evaluatingand, clemen2007aggregating, primo2009calibration} for a comprehensive overview).
-The general focus, however, has been on developing one-time aggregation procedures that consult the expert's advice only once before the event resolves.
+Prompted by the many applications of probability forecasts, including medical diagnosis (\citet{wilson1998prediction, pepe2003statistical}), political and socio-economic foresight (\citet{tetlock2005expert}), and meteorology (\citet{sanders1963subjective, vislocky1995improved, baars2005performance}), researchers have proposed many approaches to combining probability forecasts
+(see, e.g., \citet{Ranjan08, satopaa, batchelder2010cultural} for some recent studies, and \citet{Genest, Wallsten97evaluatingand, clemen2007aggregating, primo2009calibration} for a comprehensive overview).
+The general focus, however, has been on developing one-time aggregation procedures that consult the experts' advice only once before the event resolves.
 
@@
 Given that the experts are allowed to update their probability assessments, the aggregation should be performed by taking into account the temporal correlation in their advice.
-Many standard time-series procedures could be used to perform this aggregation as long as the experts update their advice consistently on a daily basis.
 
@@
 The model is estimated in a two-step procedure: first, a sampling step produces constrained parameter estimates via Gibbs sampling (see, e.g., \citet{geman1984stochastic}); second, a calibration step transforms these estimates to their unconstrained equivalents via a one-dimension optimization procedure.
-The model introduction is followed by the first evaluation section that uses synthetic data to study how accurately the two-step procedure can estimate parameter values.
+The model introduction is followed by the first evaluation section that uses synthetic data to study how accurately the two-step procedure can estimate the crowd belief.
 The second evaluation section applies the model to our real-world geopolitical forecasting database.
@@
 
-The experts made updates on a very infrequent basis: the average number of forecasts made by a single expert in one day was around 0.017, and the average group-level response rate was around 13.5 forecasts per day.
+The average number of forecasts made by a single expert in one day was around 0.017, and the average group-level response rate was around 13.5 forecasts per day.
 Given that the group of experts is large and diverse, the resulting dataset is very sparse.
@@
 For more details on the dataset and its collection see \citet{ungar2012good}.
+
+%\begin{figure}[h!]
+%\vspace{-2em}
+%\centering
+%\hspace*{-3.5em} 	\includegraphics{Figures/LegendExamplePlot} % requires the graphicx package
+%\vspace{-3.5em}
+%
+%\hspace{-0.5em}
+%\subfigure[\textit{Will the expansion of the European bailout fund be ratified by all 17 Eurozone nations before 1 November 2011?}]{
+%\hspace*{-1em}  \includegraphics[width= 0.49\textwidth]{Figures/ExamplePlot1}
+%\label{Example1}
+%} \hspace{0.5em}
+%\subfigure[\textit{Will the Nikkei 225 index finish trading at or above 9,500 on 30 September 2011?}]{
+%\hspace*{-1em}  \includegraphics[width= 0.49\textwidth]{Figures/ExamplePlot2}
+%\label{Example2}
+%}
+%
+%\caption[Optional caption for list of figures]{Scatterplots of the probability forecasts given for two questions in our dataset.
+The shadings represents the self-reported expertise of the expert who provided the probability forecast.}
+%\label{ExamplePlots}
+%\end{figure}
 
@@
 \subfigure[\textit{Will the expansion of the European bailout fund be ratified by all 17 Eurozone nations before 1 November 2011?}]{
-\hspace*{-1em}  \includegraphics[width= 0.49\textwidth]{Figures/ExamplePlot1}
-\label{Example1}
+\hspace*{-1em}  \includegraphics[width= 0.49\textwidth]{Figures/ExamplePlot1b}
+\label{Example1b}
 } \hspace{0.5em}
 \subfigure[\textit{Will the Nikkei 225 index finish trading at or above 9,500 on 30 September 2011?}]{
-\hspace*{-1em}  \includegraphics[width= 0.49\textwidth]{Figures/ExamplePlot2}
-\label{Example2}
-}
-
-\caption[Optional caption for list of figures]{Scatterplots of the probability forecasts given for two questions in our dataset.
-The shadings represents the self-reported expertise of the expert who provided the probability forecast.}
-\label{ExamplePlots}
+\hspace*{-1em}  \includegraphics[width= 0.49\textwidth]{Figures/ExamplePlot2b}
+\label{Example2b}
+}
+
+\caption[Optional caption for list of figures]{Scatterplots of the probability forecasts given for two questions in our dataset. }
+\label{ExamplePlotsFinal}
 \end{figure}
 
-To illustrate the nature of the data with some concrete examples, Figures \ref{Example1} and \ref{Example2} show scatterplots of the probability forecasts given for (a) \textit{Will the expansion of the European bailout fund be ratified by all 17 Eurozone nations before 1 November 2011?}, and (b) \textit{Will the Nikkei 225 index finish trading at or above 9,500 on 30 September 2011?}.
-The points have been jittered slightly to make overlaps visible.
-The darkness of the points is positively correlated with the self-reported expertise.
+To illustrate the nature of the data with some concrete examples, Figures \ref{Example1b} and \ref{Example2b} show scatterplots of the probability forecasts given for (a) \textit{Will the expansion of the European bailout fund be ratified by all 17 Eurozone nations before 1 November 2011?}, and (b) \textit{Will the Nikkei 225 index finish trading at or above 9,500 on 30 September 2011?}.
+The points have been shaded according to the level of self-reported expertise and jittered slightly to make overlaps visible.
+The solid line gives the posterior mean of the calibrated crowd belief as estimated by our model.
+The surrounding dashed lines connect the point-wise 95\% posterior intervals.
 Given that the European bailout fund was ratified before November 1, 2011 and that the Nikkei 225 index finished trading at around 8,700 on September 30, 2011, the general trend of the probability forecasts tends to converge towards the correct answers.
@@
 \end{eqnarray*}
-and collect the logit-probability forecasts given for question $k$ at time $t$ into a vector $\boldsymbol{Y}_{t,k} = [Y_{1,t,k}\ Y_{2,t,k}\ \dots\ Y_{I_{k},t,k}]^T$.
+and collect the logit-probabilities for question $k$ at time $t$ into a vector $\boldsymbol{Y}_{t,k} = [Y_{1,t,k}\ Y_{2,t,k}\ \dots\ Y_{I_{k},t,k}]^T$.
 Partition the experts into $J$ groups based on some individual feature, such as self-reported expertise, with each group sharing a common multiplicative bias term $b_{j} \in \R$ for $j = 1, \dots, J$.
@@
 Let $\boldsymbol{M}_k$ be a $I_k \times J$ matrix denoting  the group-memberships of the experts in question $k$; that is, if the $i$th expert participating in the $k$th question belongs to the $j$th group, then the $i$th row of $\boldsymbol{M}_k$ is the $j$th standard basis vector $\boldsymbol{e}_j$.
-The bias vector $\boldsymbol{b}$ does not include a subindex because it is considered shared among all the $K$ questions.
-To secure model identifiability, it is sufficient to share only one of the elements of $\boldsymbol{b}$ among the questions.
-This element defines a baseline under which it is possible to estimate the remaining $J-1$ bias terms separately within each of the questions.
-In this paper, however, the entire vector $\boldsymbol{b}$ is shared because some of the questions in our real-world data set involve very few experts with the highest level of self-reported expertise.
+The bias vector $\boldsymbol{b}$ is assumed to be identical across all $K$ questions.
 Under this notation, the model for the $k$th question can be expressed as
@@
 \end{eqnarray}
-where Equation (\ref{observedpr}) denotes the observed process and Equation (\ref{hiddenpr}) shows the hidden process that is driven by the constant $\gamma_k \in \R$.
-The error terms are independent and identically distributed normal random variables with mean zero
+where (\ref{observedpr}) denotes the observed process, (\ref{hiddenpr}) shows the hidden process that is driven by the constant $\gamma_k \in \R$, and $(\mu_0, \sigma_0^2) \in (\R, \R^+)$ are hyper-parameters fixed \textit{a priori} to 0 and 1, respectively.
+The error terms follow
 \begin{eqnarray*}
@@
 \end{eqnarray*}
-and $(\mu_0, \sigma_0^2) \in (\R, \R^+)$ are hyper-parameters chosen \textit{a priori}.
+Therefore the parameters of the model are $\boldsymbol{b}$, $\sigma^2_k$, $\gamma_k$, and $\tau^2_k$ for $k = 1, \dots, K$.
+Their prior distributions are chosen to be non-informative,  $p(\boldsymbol{b}, \sigma^2_k| \boldsymbol{X}_k) \propto \sigma^2_k$ and $p(\gamma_k, \tau^2_k| \boldsymbol{X}_k) \propto \tau^2_k$.
 
@@
 If $\{\mathcal{F}_{t,k}\}_{t=1}^{T_k}$ is a filtration representing the information available up to and including a given time point, then $\E[Z_k | \mathcal{F}_{t,k}] = \P(Z_k = 1| \mathcal{F}_{t,k}) = \logit^{-1}(X_{t,k})$.
-It is reasonable to assume that this probability forecast
+It is reasonable to assume that this probability
 %that makes uses all available information and hence is equivalent to $\logit^{-1}(X_{t,k})$ 
 %an expert, who has access to all available information and reports $\logit^{-1}(X_{t,k})$ as the probability forecast, 
-maximizes sharpness subject to calibration, where calibration and sharpness are understood as follows:
-\begin{mydef}
-A probability forecast $p$ for the $k$th question is calibrated if $\P(Z_k = 1 | p) = \E[Z_k | p] = p$ almost surely (see, e.g., \citet{murphy1987general}).
-\end{mydef}
-\begin{mydef}
-A probability forecast $p$ is sharper than a probability forecast $q$ if $\E[(p-p_0)^2] > \E[(q-p_0)^2]$, where $p_0$ is the baseline probability for the $k$th question (see, e.g., \citet{ranjan2009combining}).
-\end{mydef}
-\noindent
-Even though it is unlikely that any single expert has access to all the available information, a large and diverse group of experts may share a considerable portion of the available information.
+maximizes sharpness subject to calibration (for technical definitions of calibration and sharpness see \cite{Ranjan08, gneiting2011combining})
+%\begin{mydef}
+%A probability forecast $p$ for the $k$th question is calibrated if $\P(Z_k = 1 | p) = \E[Z_k | p] = p$ almost surely (see, e.g., \citet{murphy1987general}).
+%\end{mydef}
+%\begin{mydef}
+%A probability forecast $p$ is sharper than a probability forecast $q$ if $\E[(p-p_0)^2] > \E[(q-p_0)^2]$, where $p_0$ is the baseline probability for the $k$th question (see, e.g., \citet{ranjan2009combining}).
+%\end{mydef}
+%\noindent
+Even though a single expert is unlikely to have access to all the available information, a large and diverse group of experts may share a considerable portion of the available information.
 The collective wisdom of the group therefore provides an attractive proxy for $\mathcal{F}_{t,k}$.
@@
 Therefore an individual element of $\boldsymbol{b}$ can be interpreted as a group-specific \textit{systematic bias} that labels the group either as over-confident $\left( b_j \in (1, \infty)\right)$ or as under-confident $\left(b_j \in (0,1)\right)$.
-Due to the high sparsity of our data, estimating a bias term separately for each expert is not possible.
-See Section \ref{ExpertBias} for an analysis and discussion on the bias terms.
+See Section \ref{BiasStr} for a brief discussion on different bias structures.
 Any other deviation from $X_{t,k}$ is considered \textit{random noise}.
@@
  The \textit{random fluctuations} in the hidden process are measured by $\tau^2_k$ and are assumed to represent changes or shocks to the underlying circumstances that ultimately decide the outcome of the event.
-The \textit{systematic component} $\gamma_k$ allows the model to incorporate a constant signal stream that drifts the hidden process to infinity (when $\gamma_k \in (1, \infty)$) or zero (when $\gamma_k \in (0, 1)$).
-If the uncertainty in the question diminishes as the current time point $t$ approaches $T_k$, the hidden process drifts to infinity.
-Alternatively, the hidden process can drift to zero in which case any available information about the target event does not improve predictive accuracy.
-Given that each of the $K$ questions in our dataset was resolved within a pre-specified timeframe, $\gamma_k$ is expected to fall within the interval $(1, \infty)$ for all $k = 1, \dots, K$.
+The \textit{systematic component} $\gamma_k$ allows the model to incorporate a constant signal stream that drifts the hidden process.
+If the uncertainty in the question diminishes ($\gamma_k \in (1, \infty)$), the hidden process drifts to positive or negative infinity.
+Alternatively, the hidden process can drift to zero in which case any available information does not improve predictive accuracy ($\gamma_k \in (0, 1)$).
+Given that all the questions in our dataset were resolved within a pre-specified timeframe, we expect $\gamma_k \in (1, \infty)$ for all $k = 1, \dots, K$.
+
+As for any future time $T^{*}\geq t$
+\begin{eqnarray*}
+X_{T^*,k} &=&  \gamma_k^{T^*-t}X_t + \sum_{i=t+1}^{T^*} \gamma_k^{T^*-i} w_{i}\\
+  &\sim& \mathcal{N}\left( \gamma_k^{T^*-t}X_{t,k}, \tau_k^2 \sum_{i=t+1}^{T^*} \gamma_k^{T^*-i}  \right),
+\end{eqnarray*}
+the model can be used for time-forward prediction as well.
+The prediction for the aggregate logit-probability at time $T^*$ is given by an estimate of $\gamma^{T^*-t}X_{t,k}$.
+Naturally the uncertainty in this prediction grows in $T$.
+To make such time-forward predictions it is necessary to assume that the past population of experts is representative of the future population.
+This is a reasonable assumption because even though the future population may consist of entirely different individuals, on average the population is likely to look very similar to the past population.
+In practice, however, social scientists are generally more interested in an estimate of the current probability  than the probability under unknown conditions in the future.
+For this reason, our analysis focuses on probability aggregation only up to the current time $t$.
+
+\subsection{Bias Structure}
+\label{BiasStr}
+For the sake of model identifiability, it is sufficient to share only one of the elements of $\boldsymbol{b}$ among the $K$ questions.
+In this paper, however, all the elements of $\boldsymbol{b}$ are assumed to be identical across the questions because some of the questions in our real-world data set involve very few experts with the highest level of self-reported expertise.
+The model can be extended rather easily to estimate bias at a more general level.
+For instance, by assuming a hierarchical structure $b_{ik} \sim \mathcal{N}\left(b_{j(i,k)}, \sigma^2_{j(i,k)}\right)$, where $j(i,k)$ denotes the  self-reported expertise of the $i$th expert in question $k$, the bias can be estimated at an individual-level.
+These estimates can then be compared across questions.
+Individual-level analysis was not performed in our analysis for two reasons.
+First, most experts gave only a single prediction per problem, which makes accurate bias estimation at the individual-level very difficult.
+Second, it is unclear how the individually estimated bias terms can be validated.
+%If the analysis requires a time-dependent $\boldsymbol{b}$, a hidden process similar to (\ref{hiddenpr}) can be defined for $\boldsymbol{b}$.
+
+%As a final note, the bias term $\boldsymbol{b}$ can be made time-dependent by adding another regression equation $\boldsymbol{b}_t \sim \mathcal{N}(\delta \boldsymbol{b}_{t-1}, \sigma^2_b)$ to the model.
+
+\subsection{Multinomial Outcomes}
+%If the class labels in the data are balanced with respect to the time points, the calibration step under the logarithmic scoring rule is approximately equivalent to \textit{Platt calibration}, which has been shown to yield good calibration under various modeling scenarios (see, e.g., \citet{platt1999probabilistic, niculescu2005obtaining}).
+To see this, recall that the Platt calibrated logit-probabilities are given by $\hat{A} + \hat{B}\hat{X}_{t,k}(1)$, where
+%\begin{eqnarray}
+%\left(\hat{A}, \hat{B} \right) =  \argmax_{A, B \in \R} \sum_{k=1}^K \sum_{t=1}^{T_k}  S_{LOG}(Z_k, A + B\hat{X}_{t,k}(1) ) \label{platt}
+%\end{eqnarray}
+%This is equivalent to fitting a logistic regression model with $Z_k$ as the response and $\hat{X}_{t,k}(1)$ as the explanatory variable.
+%%model with $Z_k$ as the response and $\hat{X}_{t,k}(1)$ as the explanatory variable. 
+%To understand the behavior of the coefficients $A$ and $B$, express the logistic regression as a generalized linear model
+%\begin{eqnarray*}
+%\logit(\P(Z_k = 1 | \hat{X}_{t,k})) = A + B\hat{X}_{t,k}
+%\end{eqnarray*}
+%If the data are balanced with respect to the time points, then exactly half of the summands in  (\ref{platt}) have $Z_k = 1$ and the average response logit-probability $\logit(\P(Z_k = 1 | \hat{X}_{t,k})) $ is expected to be close to zero.
+Given that the values of $\hat{X}_{t,k}$ are estimated logit-probabilities of the same $K$ events across different time points, their overall average is also expected to be around zero.
+Therefore both the response and explanatory variables are approximately centered.
+This means that the intercept term $A$ is near zero reducing Platt calibration to  (\ref{OSE}) under the logarithmic scoring rule.
+If the data are not balanced, Platt calibration can be easily incorporated into our model via an additional intercept parameter.
+This, however, reduces the interpretability of our model.
+Fortunately, compromising interpretability is rarely necessary because it is often possible to use the data in a well-balanced form.
+One procedure to attain this is described in the beginning of Section \ref{realData}.
+
+If the future event can take upon $M > 2$ possible outcomes, the hidden state $X_{t,k}$ is extended to a vector of size $M-1$ and one of the outcomes, e.g., the $M$th one, is chosen as the base-case to ensure that the probabilities will sum to one at any given time point.
+Each of the remaining  $M-1$ possible outcomes is represented by an observed process similar to  (\ref{observedpr}).
+Given that this multinomial extension is equivalent to having $M-1$ independent binary-outcome models, the estimation and properties of the model are easily extended to the multi-outcome case.
+This paper focuses on binary-outcomes because it is the most commonly encountered setting in practice.
 
@@
 \label{identifiability}
-
-The main challenge is to capture a well-calibrated estimate of the hidden process without sacrificing the interpretability of our model.
-This section introduces a two-step procedure, called \textit{Sample-And-Calibrate} (SAC), that achieves this goal in a flexible and efficient manner: the first step estimates the model parameters under a constraint (\textit{Sampling Step}), and the second step performs a one-dimension optimizational procedure to transform the constrained estimates into their unconstrained counterparts (\textit{Calibration Step}).
+This section introduces a two-step procedure, called \textit{Sample-And-Calibrate} (SAC), that captures a well-calibrated estimate of the hidden process without sacrificing the interpretability of our model.
 
@@
 \label{sampling_step}
-Given that $\left(a \boldsymbol{b}, X_{t,k}/a, a^2 \tau_k^2\right) \neq \left(\boldsymbol{b}, X_{t,k}, \tau_k^2\right)$ for any $a > 0$ yield the same  likelihood for $\boldsymbol{Y}_{t,k}$, the model as described by Equations (\ref{observedpr}) and (\ref{hiddenpr}) is not identifiable.
-As a result, the parameter estimates tend to drift during the sampling process.
+Given that $\left(a \boldsymbol{b}, X_{t,k}/a, a^2 \tau_k^2\right) \neq \left(\boldsymbol{b}, X_{t,k}, \tau_k^2\right)$ for any $a > 0$ yield the same  likelihood for $\boldsymbol{Y}_{t,k}$, the model as described by (\ref{observedpr}) and (\ref{hiddenpr}) is not identifiable.
 A well-known solution is to choose one of the elements of $\boldsymbol{b}$, say $b_3$, as the reference point and fix $b_3 = 1$.
@@
 \end{eqnarray*}
-where the trailing input parameter emphasizes the constraint $b_{3} = 1$.
+where the trailing input notation, $(a)$, signifies the value under the constraint $b_{3} = a$.
 Given that this version is identifiable, estimates of the model parameters can be obtained.
@@
 For instance, $\hat{\boldsymbol{b}}(1)$ and $\hat{X}_{t, k}(1)$ represent the estimates of $\boldsymbol{b}(1)$ and $X_{t, k}(1)$, respectively.
-These estimates are found via Gibbs sampling that only makes use of standard distributions.
-See Appendix \ref{appendix} for the technical details of the sampling steps, and, e.g., \citet{gelman2003bayesian} for a discussion on the general principles of Gibbs sampling.
+
+These estimates are obtained by first computing a posterior sample via Gibbs sampling and then taking the average of the posterior sample.
+The first step of our Gibbs sampler is to sample the hidden states via the \textit{Forward-Filtering-Backward-Sampling} (FFBS) algorithm.
+FFBS first predicts the hidden states using a Kalman filter and then performs a backward sampling procedure that treats these predicted states as additional observations (see, e.g., \cite{carter1994gibbs, migon2005dynamic} for details on FFBS).
+Given that the Kalman filter can handle varying numbers or even no forecasts at different time points, it plays a very crucial role in our probability aggregation under sparse data.
+%
+%The update step is repeated sequentially for each observation $Y_{i,t,k}$ given at time $t$.
+For each such repetition, the previous posterior values, $X_{t|t, k}$ and $P_{t|t, k}$, are considered as the new prior values, $X_{t|t-1, k}$ and $P_{t|t-1, k}$.
+If the observation $\boldsymbol{Y}_{t,k}$ is completely missing at time $t$, the update step is skipped and
+%\begin{align*}
+%X_{t|t, k} &= X_{t|t-1, k}\\
+%P_{t|t,k} &= P_{t|t-1,k}
+%\end{align*}
+%
+%If the observation $\boldsymbol{Y}_{t,k}$ is completely missing at time $t$, the update step is skipped and the state estimates are sampled from
+%\begin{align*}
+%\mathcal{N}\left(\gamma_k X_{t-1|t-1,k}, \gamma_k^2 P_{t-1|t-1,k} + \tau_k^2\right)
+%\end{align*}
+
+Our implementation of the sampling step is written in C++ and runs fast.
+For instance, to obtain a posterior sample of size 1,000 for 50 questions each with 100 time points and 50 experts takes about 215 seconds on the first author's computer (1.7 GHz Intel Core i5).
+See the Online Supplement for the technical details of the sampling steps, and, e.g., \citet{gelman2003bayesian} for a discussion on the general principles of Gibbs sampling.
 
@@
 Given that the model parameters can be estimated by fixing $b_3$ to any constant, the next step is to search for the constant that gives an optimally sharp and calibrated estimate of the hidden process.
-This section introduces an efficient procedure that finds the optimal constraint without requiring any additional runs of the sampling step.
-First, assume that parameter estimates  $\hat{\boldsymbol{b}}(1)$ and $\hat{X}_{t, k}(1)$ have already been obtained via the constrained sampling step described in Section \ref{sampling_step}.
+This section introduces an efficient procedure that finds the optimal constant without requiring any additional runs of the sampling step.
+First, assume that parameter estimates  $\hat{\boldsymbol{b}}(1)$ and $\hat{X}_{t, k}(1)$ have already been obtained via the sampling step described in Section \ref{sampling_step}.
 Given that for any $\beta \in \R / \{0\}$,
@@
 \end{eqnarray*}
-the parameter values under $b_{3} = \beta$ can be obtained from $\boldsymbol{b} (\beta) = \boldsymbol{b} (1) \beta $ and $X_{t, k}(\beta) = X_{t, k}(1)/\beta$. 
+we have that  $\boldsymbol{b} (\beta) = \boldsymbol{b} (1) \beta $ and $X_{t, k}(\beta) = X_{t, k}(1)/\beta$. 
 %Furthermore, an estimate of $X_{t, k}(\beta)$ is given by $\hat{X}_{t, k}(1)/\beta$. 
-This means that $X_{t,k} = X_{t,k}(1)/\beta$ when $\beta$ is equal to the true value of $b_3$.
-Given that the hidden process $X_{t, k}$ is assumed to be sharp and well-calibrated, $b_3$ can be estimated with the value of $\beta$ that simultaneously maximizes the sharpness and calibration of $\hat{X}_{t, k}(1) / \beta$.
+%This means that $X_{t,k} = X_{t,k}(1)/\beta$ when $\beta$ is equal to the true value of $b_3$. 
+Recall that the hidden process $X_{t, k}$ is assumed to be sharp and well-calibrated.
+Therefore $b_3$ can be estimated with the value of $\beta$ that simultaneously maximizes the sharpness and calibration of $\hat{X}_{t, k}(1) / \beta$.
 A natural criterion for this maximization is given by the class of \textit{proper scoring rules} that combine sharpness and calibration (\citet{gneiting2008rejoinder, buja2005loss}).
@@
 \end{eqnarray}
-where $Z_k \in \{0, 1\}$ indicate whether the event associated with the $k$th question happened $\left(Z_k = 1\right)$ or did not happen $\left(Z_k = 0\right)$.
-The function $S$ is a strictly proper scoring rule such as the negative Brier score (\citet{Brier})
+where $Z_k \in \{0, 1\}$ is the event indicator for question $k$. The function $S$ is a strictly proper scoring rule such as the negative Brier score (\citet{Brier})
 \begin{eqnarray*}
@@
 \end{eqnarray*}
-Given that it is not clear which rule should be used for predicting geopolitical events, the SAC-procedure is evaluated  separately under both rules in Sections \ref{syntheticData} and \ref{realData}.
-Once $\hat{\beta}$ has been computed, estimates of the unconstrained model parameters are given by
+%Given that it is not clear which rule should be used for predicting geopolitical events, the SAC-procedure is evaluated  separately under both rules in Sections \ref{syntheticData} and \ref{realData}.
+Once $\hat{\beta}$ has been computed,
+The estimates of the unconstrained model parameters are then given by
 \begin{eqnarray}
@@
 Notice that estimates of $\sigma^2_k$ and $\gamma_k$ are not affected by the constraint.
-
-\subsection{Discussion}
-If the class labels in the data are balanced with respect to the time points, the calibration step under the logarithmic scoring rule is approximately equivalent to \textit{Platt calibration}, which has been shown to yield good calibration under various modeling scenarios (see, e.g., \citet{platt1999probabilistic, niculescu2005obtaining}).
-To see this, recall that the Platt calibrated logit-probabilities are given by $\hat{A} + \hat{B}\hat{X}_{t,k}(1)$, where
-\begin{eqnarray}
-\left(\hat{A}, \hat{B} \right) =  \argmax_{A, B \in \R} \sum_{k=1}^K \sum_{t=1}^{T_k}  S_{LOG}(Z_k, A + B\hat{X}_{t,k}(1) ) \label{platt}
-\end{eqnarray}
-This is equivalent to fitting a logistic regression model with $Z_k$ as the response and $\hat{X}_{t,k}(1)$ as the explanatory variable.
-%model with $Z_k$ as the response and $\hat{X}_{t,k}(1)$ as the explanatory variable. 
-To understand the behavior of the coefficients $A$ and $B$, express the logistic regression as a linear regression model
-\begin{eqnarray*}
-\logit(\P(Z_k = 1 | \hat{X}_{t,k})) = A + B\hat{X}_{t,k} + e_{t,k}
-\end{eqnarray*}
-with $e_{t,k} \stackrel{i.i.d.}{\sim} \mathcal{N}(0,\sigma^2)$.
-If the data are balanced with respect to the time points, then exactly half of the summands in Equation (\ref{platt}) have $Z_k = 1$ and the average response logit-probability $\logit(\P(Z_k = 1 | \hat{X}_{t,k})) $ is expected to be close to zero.
-Given that the values of $\hat{X}_{t,k}$ are estimated logit-probabilities of the same $K$ events across different time points, their overall average is also expected to be around zero.
-Therefore both the response and explanatory variables are approximately centered.
-This means that the intercept term $A$ is near zero reducing Platt calibration to Equation (\ref{OSE}) under the logarithmic scoring rule.
-If the data are not balanced, Platt calibration can be easily incorporated into our model via an additional intercept parameter.
-This, however, reduces the interpretability of our model.
-Fortunately, compromising interpretability is rarely necessary because it is often possible to use the data in a well-balanced form.
-One procedure to attain this is described in the beginning of Section \ref{realData}.
-
-If the future event can take upon $M > 2$ possible outcomes, the hidden state $X_{t,k}$ is extended to a vector of size $M-1$ and one of the outcomes, e.g., the $M$th one, is chosen as the base-case to ensure that the probabilities will sum to one at any given time point.
-Each of the remaining  $M-1$ possible outcomes is represented by an observed process similar to Equation (\ref{observedpr}).
-Given that this multinomial extension is equivalent to having $M-1$ independent binary-outcome models, the estimation and properties of the model are easily extended to the multi-outcome case.
-This paper focuses on binary-outcomes because it is the simplest and most commonly encountered setting in practice.
 
@@
 
-The goal in this section is to evaluate the ability of the \textit{Sample-And-Calibrate}-procedure to capture true parameter values.
-The synthetic data are not generated directly from the model for two reasons: (i) showing good performance on a dataset directly generated from the model assumptions is hardly news, and (ii) it is unclear how the dynamic part of the model can be used to produce well-calibrated hidden states.
+This section uses synthetic data to evaluate how accurately the SAC-procedure captures the hidden states and bias vector.
+The analysis is therefore not reflective of real-world performance but serves rather as a sanity check.
+%The synthetic data are not generated directly from the model for two reasons: (i) showing good performance on a dataset directly generated from the model assumptions is hardly news, and (ii) it is unclear how the dynamic part of the model can be used to produce well-calibrated hidden states.
 The latter is important for our interpretation of the hidden process as a sharp and well-calibrated version of the crowd belief.
-
-To ensure that the hidden process for question $k$ is well-calibrated, we generate a path of standard Brownian motion until time $T_k$. If $Z_{t,k}$ denotes the value of the path at time $t$, then
+%To ensure that the hidden process for question $k$ is well-calibrated, 
+The hidden process is generated from standard Brownian motion.  More specifically, if $Z_{t,k}$ denotes the value of a path at time $t$, then
 \begin{eqnarray*}
@@
 gives a sequence of $T_k$ calibrated logit-probabilities for the event $Z_k = 1$.
-Using this procedure we generate the hidden process for $K$ questions, each with a time horizon of $T_k = 101$.
+A hidden process is generated for $K$ questions with a time horizon of $T_k = 101$.
 The questions involve 50 experts allocated evenly among five expertise groups.
 Each expert gives one probability forecast per day with the exception of time $t = 101$ when the event resolves.
-These are generated by applying bias and noise to the hidden process $X_{t,k}$ for $t = 1, \dots, 100$ (see Equation (\ref{observedpr})).
+The forecasts are generated by applying bias and noise to the hidden process as described by  (\ref{observedpr}).
 Our simulation study considers a three-dimensional grid of parameter values:
@@
 \end{eqnarray*}
-where $\beta$ gives the bias vector by letting $\boldsymbol{b} = [1/2, 3/4, 1, 4/3, 2/1]^T \beta$.
+where $\beta$ varies the bias vector by $\boldsymbol{b} = [1/2, 3/4, 1, 4/3, 2/1]^T \beta$.
 Forty synthetic datasets are generated for each combination of $\sigma^2$, $\beta$, and $K$ values.
 The SAC-procedure runs for 200 iterations of which the first 100 are used for burn-in.
-The accuracy of the estimated hidden process and bias vector are measured with the average quadratic loss in the probability space, $\sum_{k=1}^K \sum_{t=1}^{100} ( \logit^{-1}(\hat{X}_{t,k}) - \logit^{-1}(X_{t,k}))^2 / (100K)$, and the average quadratic loss, $|| \boldsymbol{b} - \hat{\boldsymbol{b}}||^2/5$, respectively.
+
+\begin{figure}[h!]
+\vspace{-5em}
+\hspace{-25em}	\includegraphics{Figures/LegendSynthetic} % requires the graphicx package
+ \includegraphics[width= 0.49\textwidth]{Figures/SyntheticBeta}
+\caption[Optional caption for list of figures]{
+%Comparing SAC that optimizes over the logarithmic score ($\text{SAC}_{\text{LOG}}$), SAC that optimizes over the Brier score ($\text{SAC}_{\text{BRI}}$), and the Exponentially Weighted Moving Average (EWMA) under synthetic data.
+The marginal effect of $\beta$ on the average quadratic loss.}
+\label{Synthetic}
+\end{figure}
 
@@
 
-\begin{figure}[h!]
-\vspace{-5.5em}
-\centering
- \hspace{-25em}	\includegraphics{Figures/LegendSynthetic} % requires the graphicx package
-
-\vspace{0em}
-\hspace{2.5em}Hidden Process
-\vspace{-0.5em}
-
-\subfigure{
-\hspace{-1em} \includegraphics[width = 0.32\textwidth]{Figures/SyntheticSigma2}
-\label{SyntheticSigma2}
-}
-\subfigure{
-\hspace{-1.5em} \includegraphics[width = 0.32\textwidth]{Figures/SyntheticBeta}
-\label{SyntheticBeta}
-}
-\subfigure{
-\hspace{-1.5em} \includegraphics[width = 0.32\textwidth]{Figures/SyntheticK}
-\label{SyntheticK}
-}
-
-\vspace{0.5em}
-\hspace{2.5em}Bias Vector
-\vspace{-0.5em}
-
-\subfigure{
-\hspace{-1em}  \includegraphics[width = 0.32\textwidth]{Figures/SyntheticSigma2bias}
-\label{SyntheticSigma2b}
-}
-\subfigure{
-\hspace{-1.5em}  \includegraphics[width = 0.32\textwidth]{Figures/SyntheticBetabias}
-\label{SyntheticBetab}
-}
-\subfigure{
-\hspace{-1.5em}   \includegraphics[width = 0.32\textwidth]{Figures/SyntheticKbias}
-\label{SyntheticKb}
-}
-
-\caption[Optional caption for list of figures]{Comparing SAC that optimizes over the logarithmic score ($\text{SAC}_{\text{LOG}}$), SAC that optimizes over the Brier score ($\text{SAC}_{\text{BRI}}$), and the Exponentially Weighted Moving Average (EWMA) under synthetic data.
-The top row presents the accuracy to capture the hidden process.
-The bottom row presents the accuracy to capture the bias vector.
-}
-\label{Synthetic}
-\end{figure}
-
-Figure \ref{Synthetic} summarizes the results in a series of plots.
-The first row presents the marginal effects of the three grid variables on the accuracy to which SAC and the Exponentially Weighted Moving Average (EWMA) (see Section \ref{forecasting} for more details on EWMA) estimate the hidden process.
-EWMA serves as a baseline for comparisons.
-Unfortunately, EWMA does not produce an estimate of the bias vector.
-Therefore it cannot be used as a baseline in the second row that measures the accuracy in estimating the bias vector.
-Given that the plots were computed by choosing one grid variable (e.g., $\beta$) at a time and averaging over the remaining two variables  (e.g., $K$ and $\sigma^2$), each value in the plots represents an average of a $40 \times 5^2 = 1,000$ values.
-
-Both $\text{SAC}_{\text{LOG}}$ and  $\text{SAC}_{\text{BRI}}$ estimate the hidden process and the bias vector very accurately.
-Apart from $\text{SAC}_{\text{LOG}}$ outperforming $\text{SAC}_{\text{BRI}}$ uniformly across all parameter values, the two approaches behave very similarly.
+%\begin{figure}[h!]
+%\vspace{-5.5em}
+%\centering
+% \hspace{-25em}	\includegraphics{Figures/LegendSynthetic} % requires the graphicx package
+%
+%\vspace{0em}
+%\hspace{2.5em}Hidden Process
+%\vspace{-0.5em}
+%
+%\subfigure{
+%\hspace{-1em} \includegraphics[width = 0.32\textwidth]{Figures/SyntheticSigma2}
+%\label{SyntheticSigma2}
+%}
+%\subfigure{
+%\hspace{-1.5em} \includegraphics[width = 0.32\textwidth]{Figures/SyntheticBeta}
+%\label{SyntheticBeta}
+%}
+%\subfigure{
+%\hspace{-1.5em} \includegraphics[width = 0.32\textwidth]{Figures/SyntheticK}
+%\label{SyntheticK}
+%}
+%
+%\vspace{0.5em}
+%\hspace{2.5em}Bias Vector
+%\vspace{-0.5em}
+%
+%\subfigure{
+%\hspace{-1em}  \includegraphics[width = 0.32\textwidth]{Figures/SyntheticSigma2bias}
+%\label{SyntheticSigma2b}
+%}
+%\subfigure{
+%\hspace{-1.5em}  \includegraphics[width = 0.32\textwidth]{Figures/SyntheticBetabias}
+%\label{SyntheticBetab}
+%}
+%\subfigure{
+%\hspace{-1.5em}   \includegraphics[width = 0.32\textwidth]{Figures/SyntheticKbias}
+%\label{SyntheticKb}
+%}
+%
+%
+%\caption[Optional caption for list of figures]{
+%%Comparing SAC that optimizes over the logarithmic score ($\text{SAC}_{\text{LOG}}$), SAC that optimizes over the Brier score ($\text{SAC}_{\text{BRI}}$), and the Exponentially Weighted Moving Average (EWMA) under synthetic data.
+%The top and bottom rows present the accuracies to capture the hidden process and the bias vector, respectively. }
+%\label{Synthetic}
+%\end{figure}
+
+%Figure \ref{Synthetic} summarizes the results in a series of plots. The first row presents the marginal effects of the three grid variables on
+
+SAC under the Brier ($\text{SAC}_{\text{BRI}}$) and logarithm score ($\text{SAC}_{\text{LOG}}$) are compared with \textit{Exponentially Weighted Moving Average} (EWMA).
+EWMA, which serves as a baseline, can be understood by first denoting the (expertise-weighted) average forecast at time $t$ for the $k$th question by
+\begin{eqnarray}
+\bar{p}_{t,k}  = \sum_{j=1}^J \omega_j \left( \frac{1}{|E_j|}\sum_{i \in E_j} p_{i,t,k} \right) \label{weighted mean}
+\end{eqnarray}
+where $E_j$ refers to an index set of all experts in the $j$th expertise group, and $\omega_j$ denotes the weight associated with the $j$th expertise group.
+The EWMA forecasts for the $k$th problem are then constructed recursively from
+\begin{eqnarray*}
+\hat{p}_{t,k}(\alpha) =
+\begin{cases}
+ \bar{p}_{1,k}, & \text{ for } t  = 1, \\
+\alpha \bar{p}_{t,k}  + (1-\alpha) \hat{p}_{t-1,k}(\alpha), & \text{ for } t > 1,
+\end{cases}
+\end{eqnarray*}
+where $\alpha$ and $\boldsymbol{\omega}$ are learned from the training set by 
+\begin{eqnarray}
+(\hat{\alpha}, \hat{\boldsymbol{\omega}}) &=& \argmin_{\alpha, \omega_j \in [0,1]} \sum_{k =1}^K \sum_{t=1}^{T_k} \left(Z_k - \hat{p}_{t,k}(\alpha, \boldsymbol{\omega}) \right)^2 \label{EWMA}  \hspace{1em} \text{ s.t.} \hspace{0.5em} \sum_{j=1}^J \omega_j = 1 \nonumber
+\end{eqnarray}
+
+The accuracy of the estimated hidden process is measured with the quadratic loss in the probability space
+\begin{eqnarray}
+\sum_{k=1}^K \sum_{t=1}^{100} ( \logit^{-1}(\hat{X}_{t,k}) - \logit^{-1}(X_{t,k}))^2 / (100K) \label{loss1}
+\end{eqnarray}
+The three competing methods, $\text{SAC}_{\text{BRI}}$, $\text{SAC}_{\text{LOG}}$, and EWMA, estimate the hidden process with great accuracy.
+Their quadratic losses (\ref{loss1}) averaged over all simulations are 0.00245, 0.00257, and 0.00211, respectively.
+Even though EWMA outperforms SAC marginally in this synthetic and hence rather easy setup, it performs much worse under real-world data (see Section \ref{realData}).
+Based on other performance measures that are not shown  for the sake of brevity, all three methods suffer from an increasing level of noise in the expert logit-probabilities but can make efficient use of extra data.
+
+%The accuracy of the estimated hidden process and bias vector are measured with the average quadratic loss in the probability space
+%\begin{eqnarray}
+%\sum_{k=1}^K \sum_{t=1}^{100} ( \logit^{-1}(\hat{X}_{t,k}) - \logit^{-1}(X_{t,k}))^2 / (100K) \label{loss1}
+%\end{eqnarray}
+% and the average quadratic loss, $|| \boldsymbol{b} - \hat{\boldsymbol{b}}||^2/5$, respectively.
+
+%Given that the plots were computed by choosing one grid variable (e.g., $\beta$) at a time and averaging over the remaining two variables  (e.g., $K$ and $\sigma^2$), each value in the plots represents an average of a $40 \times 5^2 = 1,000$ values.
+%
+%As the Y-scale in Figure \ref{Synthetic} is very small, all three methods estimate the hidden process very accurately.
+Generally EWMA performs slightly better than $\text{SAC}$.
+This result is due to its 6-parameter calibration function (\ref{EWMA}) that is more flexible than the 1-parameter calibration made by $\text{SAC}$.
+%
+%Apart from $\text{SAC}_{\text{LOG}}$ outperforming $\text{SAC}_{\text{BRI}}$ uniformly across all parameter values, the two approaches behave very similarly.
 They suffer  from the increasing level of noise in the expert logit-probabilities, estimate the bias vector less accurately when $\beta$ is large, and, unlike EWMA, make greater use of the increasing number of questions.
 EWMA is able to outperform SAC only when the bias among the experts is very small.
-
+Some interesting differences are found by inspecting Figure \ref{Synthetic} that shows the marginal effect of $\beta$ on the average quadratic loss.
+As can be expected, EWMA performs well when the experts are, on average, close to unbiased.
 Interestingly, SAC estimates the hidden process more accurately when the experts are over-confident (large $\beta$) rather under-confident (small $\beta$).
 To understand this result, assume that the experts in the third group are highly under-confident.
-Therefore their logit-probabilities are expected to be closer to zero than the corresponding hidden states.
+Their logit-probabilities are then expected to be closer to zero than the corresponding hidden states.
 After adding white noise to these expected logit-probabilities, they are likely to cross to the other side of zero.
@@
 %
+The accuracy of the estimated bias vector is measured with the quadratic loss, $|| \boldsymbol{b} - \hat{\boldsymbol{b}}||^2/5$. 
+Unfortunately, EWMA does not produce an estimate of the bias vector.
+Therefore it cannot be used as a baseline for the accuracy in estimating the bias vector.
+Given that the quadratic losses averaged over all simulations for $\text{SAC}_{\text{BRI}}$ and $\text{SAC}_{\text{LOG}}$ are 0.104 and 0.100, respectively, both methods estimate the bias vector accurately.
 
@@
 This section presents results for the real-world data described in Section \ref{data}.
-The goal is to provide application specific insight by discussing the specific research objectives itemized in Section \ref{intro} and also to evaluate SAC in terms of predictive power and calibration.
-Before presenting the results, however, we discuss two practical matters that must be taken into account when aggregating real-world  probability forecasts.
+The goal is to provide application specific insight by discussing the specific research objectives itemized in Section \ref{intro}.
+First, however, we discuss two practical matters that must be taken into account when aggregating real-world  probability forecasts.
 
@@
 The first matter regards human experts making probability forecasts of 0.0 or 1.0 even if they are not completely sure of the outcome of the event.
-For instance, all the 166 questions in our dataset contained both a zero and a one.
+For instance, all the 166 questions in our dataset contain both a zero and a one.
 Transforming such forecasts into the logit-space yields infinities that can cause problems in model estimation.
@@
 This is equivalent to an NP-hard problem known in computer science as the \textit{Partition Problem}:  determine whether a given set of positive integers can be partitioned into two sets such that the sums of the two sets equal to each other (see, e.g., \citet{karmarkar1982differencing, hayes2002easiest}).
-A simple solution is to use a greedy algorithm that iterates through the values of $T_k$ in descending order, assigning each $T_k$ to the subset that currently has the smaller sum (see, e.g.
-\citet{kellerer2004knapsack, gent1996phase} for more details on the \textit{Partition Problem}).
+A simple solution is to use a greedy algorithm that iterates through the values of $T_k$ in descending order, assigning each $T_k$ to the subset that currently has the smaller sum (see, e.g., \citet{kellerer2004knapsack, gent1996phase} for more details on the \textit{Partition Problem}).
 After finding a well-balanced partition, the next step is to assign the class labels such that the labels for the questions in $S_x$ are equal to $x$ for $x = 0$ or $1$.
-Recall from section \ref{calibration_step} that $Z_k$ represents the event indicator associated with the $k$th question.
+Recall from Section \ref{calibration_step} that $Z_k$ represents the event indicator for the $k$th question.
 To define a balanced set of indicators $ \tilde{Z}_k$ for all $k \in S_x$, let
@@
 
-\subsection{Out-of-Sample Forecasting}
+\subsection{Out-of-Sample Aggregation}
 \label{forecasting}
@@
 %For instance, recall our example with an investor who consults a group of experts on a daily basis in order to assess whether a stock index will end up trading above some threshold on a given date. 
-The goal is to evaluate the out-of-sample predictive performance of SAC against several other probability aggregation procedures.
-The models are allowed to utilize a training set before making predictions on an independent testing set.
-To clarify some of the upcoming notation, let $S_{train}$ and $S_{test}$ be index sets that partition the data into training and testing sets of sizes $|S_{train}| = N_{train}$ and $|S_{test}| = 166 - N_{test}$, respectively.
+The goal is to evaluate the accuracy of the aggregate probabilities made by SAC and several other procedures.
+The models are allowed to utilize a training set before making aggregations on an independent testing set.
+To clarify some of the upcoming notation, let $S_{train}$ and $S_{test}$ be index sets that partition the data into training and testing sets of sizes $|S_{train}| = N_{train}$ and $|S_{test}| = 166 - N_{train}$, respectively.
 This means that the $k$th question is in the training set if and only if $k \in S_{train}$.
-The competing models are as follows:
+Before introducing the competing models, note that all choices of thinning and burn-in made in this section are conservative and have been made based on pilot runs of the models.
+This was done to ensure a posterior sample that has low autocorrelation and arises from a converged chain.
+The competing models are as follows.
 
@@
 \end{eqnarray*}
-where $X_{t,k}$ is the logit-probability used for prediction.
-Given that this model is not hierarchical, estimates of the hidden process can be obtained directly for the questions in the testing set without fitting the model first on the training set.
-To make predictions, the sampler is run for 500 iterations of which the first 200 are used for burn-in.
-The remaining 300 iterations are thinned by discarding every other observation, leaving a final predictive sample of 150 observations.
+where $X_{t,k}$ is the aggregate logit-probability.
+Given that this model does not share any parameters across different questions, estimates of the hidden process can be obtained directly for the questions in the testing set without fitting the model first on the training set.
+The Gibbs sampler is run for 500 iterations of which the first 200 are used for burn-in.
+The remaining 300 iterations are thinned by discarding every other observation, leaving a final posterior sample of 150 observations.
+The mean of this sample gives the final estimates.
 
@@
 The calibration step is performed for the final 500 observations.
-The out-of-sample prediction is done by running the sampling step for 500 iterations with each consecutive iteration reading in and conditioning on the next value of $\beta$ and $\boldsymbol{b}$ found during the training period.
+The out-of-sample aggregation is done by running the sampling step for 500 iterations with each consecutive iteration reading in and conditioning on the next value of $\beta$ and $\boldsymbol{b}$ found during the training period.
 The first 200 iterations are used for burn-in.
-The remaining 300 iterations are thinned by discarding every other observation, leaving a final predictive sample of 150 observations.
+The remaining 300 iterations are thinned by discarding every other observation, leaving a final posterior sample of 150 observations.
+The mean of this sample gives the final estimates.
 
 \item \textit{A fully Bayesian version of $\text{SAC}_{\text{LOG}}$ ($\text{BSAC}_{\text{LOG}}$)}.
-Denote the calibrated logit-probabilities and  the event indicators across all $K$ questions with $\boldsymbol{X}(1)$ and $\boldsymbol{Z}$, respectively.
+Denote the calibrated logit-probabilities and event indicators across all $K$ questions with $\boldsymbol{X}(1)$ and $\boldsymbol{Z}$, respectively.
 The posterior distribution of $\beta$ conditional on $\boldsymbol{X}(1)$ is given by $p(\beta | \boldsymbol{X}(1), \boldsymbol{Z}) \propto p( \boldsymbol{Z} | \beta, \boldsymbol{X}(1)) p(\beta | \boldsymbol{X}(1))$.
-Recall that the calibration step under $S_{LOG}$ is equivalent to fitting a logistic regression model with $Z_k$ as the response and $\hat{X}_{t,k}(1)$ as the explanatory variable.
-Therefore the likelihood for the Bayesian version is
+The likelihood is
 \begin{eqnarray}
@@
 \end{eqnarray}
-As in \citet{gelman2003bayesian}, the prior is chosen to be locally uniform, $p(1/\beta) \propto 1$.
-Posterior estimates of $\beta$ can be sampled from Equation (\ref{OSE2}) using generic sampling algorithms such as the Metropolis algorithm (\citet{metropolis1953equation}) or slice sampling (\citet{neal2003slice}).
+As in \citet{gelman2003bayesian}, the prior for $\beta$ is chosen to be locally uniform, $p(1/\beta) \propto 1$.
+Given that this model estimates $X_{t,k}(1)$ and $\beta$ simultaneously, it is a little more flexible than $\text{SAC}$.
+Posterior estimates of $\beta$ can be sampled from (\ref{OSE2}) using generic sampling algorithms such as the Metropolis algorithm (\citet{metropolis1953equation}) or slice sampling (\citet{neal2003slice}).
 Given that the sampling procedure conditions on the event indicators, the full conditional distribution of the hidden states is not in a standard form.
 Therefore the Metropolis algorithm is also used for sampling the hidden states.
-Predictions are made with the same choices of thinning and burn-in as described under \textit{Sample-And-Calibrate}.
+Estimation is made with the same choices of thinning and burn-in as described under \textit{Sample-And-Calibrate}.
 
@@
 
-\item \textit{Exponentially Weighted Moving Average (EWMA)}. If 
-\begin{eqnarray*}
-\bar{p}_{t,k}  = \frac{1}{N_{t,k}} \sum_{i=1}^{N_{t,k}} p_{i,t,k},
-\end{eqnarray*}
-is the average probability forecast given at time $t$ for the $k$th question, then the EWMA forecasts for the $k$th problem are obtained recursively from
-\begin{eqnarray*}
-\hat{p}_{t,k}(\alpha) =
-\begin{cases}
- \bar{p}_{1,k}, & \text{ for } t  = 1, \\
-\alpha \bar{p}_{t,k}  + (1-\alpha) \hat{p}_{t-1,k}(\alpha), & \text{ for } t > 1,
-\end{cases}
-\end{eqnarray*}
-where the input parameter $\alpha$ is learned from the training set by 
-\begin{eqnarray*}
-\hat{\alpha} &=& \argmin_{\alpha \in [0,1]} \sum_{k \in S_{train}} \sum_{t=1}^{T_k} \left(Z_k - \hat{p}_{t,k}(\alpha) \right)^2
-\end{eqnarray*}
-
-\item  \textit{Exponentially Weighted Moving Logit Aggregator (EWMLA)}.
-This is a moving version of the aggregator $\hat{p}_G(\boldsymbol{b})$ that was introduced in \cite{satopaa}.
-If $\boldsymbol{p}_{t,k}$ is a vector collecting all the probability forecasts made for the $k$th question at time $t$, then the EWMLA forecasts are found recursively from
+\item \textit{Exponentially Weighted Moving Average (EWMA)} as described in Section \ref{syntheticData}.
+
+\item  \textit{Exponentially Weighted Moving Logit Aggregator (EWMLA)}. This is a moving version of the aggregator $\hat{p}_G(\boldsymbol{b})$ that was introduced in \cite{satopaa}. 
+%If $\boldsymbol{p}_{t,k}$ is a vector collecting all the probability forecasts made for the $k$th question at time $t$, then 
+%
+The EWMLA aggregate probabilities are found recursively from
 \begin{eqnarray*}
@@
 \end{eqnarray*}
-where
-\small
-\begin{eqnarray*}
-G_{t,k}(\nu )  &=& \left( \prod\limits_{i=1}^{N_{t,k}} \left( \frac{p_{i, t, k}}{1-p_{i, t, k}} \right)^{ \frac{\boldsymbol{e}_{i,k}'\boldsymbol{b}}{N_{t,k}}} \right) \Bigg/ \left(1+ \prod\limits_{i=1}^{N_{t,k}} \left( \frac{p_{i, t, k}}{1-p_{i, t, k}} \right)^{\frac{\boldsymbol{e}_{i,k}' \boldsymbol{b}}{N_{t,k}}} \right)
+where the vector $\boldsymbol{b} \in \mathbb{R}^J$ collects the bias terms of the expertise groups, and
+\begin{eqnarray*}
+G_{t,k}(\nu )  = \left( \prod\limits_{i=1}^{N_{t,k}} \left( \frac{p_{i, t, k}}{1-p_{i, t, k}} \right)^{ \frac{b_{j(i,k)}}{N_{t,k}}} \right) \Bigg/ \left(1+ \prod\limits_{i=1}^{N_{t,k}} \left( \frac{p_{i, t, k}}{1-p_{i, t, k}} \right)^{\frac{ b_{j(i,k)}}{N_{t,k}}} \right)
 \end{eqnarray*}
 \normalsize
-The vector $\boldsymbol{b}$ collects the bias terms of the different expertise groups.
-Therefore it serves a similar role to the bias vector found under \textit{Sample-And-Calibrate}.
-The term $\boldsymbol{e}_{i,k}$ is a vector of length 5 indicating which level of self-reported expertise the $i$th expert in the $k$th question belongs to.
+%Therefore it serves a similar role to the bias vector found under \textit{Sample-And-Calibrate}. 
+%The term $\boldsymbol{e}_{i,k}$ is a vector of length 5 indicating which level of self-reported expertise the $i$th expert in the $k$th question belongs to.
 For instance, if $\boldsymbol{e}_{i,k} = [0, 1, 0, 0, 0]$, then the expert identifies himself with the expertise level two.
-The tuning parameters $(\alpha, \boldsymbol{b})$ are learned from the training set by
+The parameters $\alpha$ and $\boldsymbol{b}$ are learned from the training set by
 \begin{eqnarray*}
@@
 The static version of the Beta-transformed  aggregator was introduced in \cite{Ranjan08}.
-A dynamic version can be obtained by replacing $G_{t,k}(\nu )$ in the EWMLA description with
-\begin{eqnarray*}
-H_{\nu, \tau} \left( \bar{p}_{t,k}\right),
-\end{eqnarray*}
-where $H_{\nu, \tau}$ is the cumulative distribution function of the Beta distribution and $\bar{p}_{t,k}$ is the average probability forecast defined under EWMA.
-The tuning parameters $(\alpha, \nu, \tau)$ are learned from the training set by
-\begin{eqnarray*}
-(\hat{\alpha}, \hat{\nu}, \hat{\tau}) &=& \argmin_{\nu, \tau > 0\ \alpha \in [0,1]} \sum_{k \in S_{train}} \sum_{t=1}^{T_k} \left(Z_k - \hat{p}_{t,k}(\alpha, \nu, \tau)\right)^2
-\end{eqnarray*}
-
+A dynamic version can be obtained by replacing $G_{t,k}(\nu )$ in the EWMLA description with $H_{\nu, \tau} \left( \bar{p}_{t,k}\right)$, where $H_{\nu, \tau}$ is the cumulative distribution function of the Beta distribution and $\bar{p}_{t,k}$ is given by \ref{weighted mean}.
+The parameters $\alpha, \nu, \tau$, and  $\boldsymbol{\omega}$ are learned from the training set by
+\begin{eqnarray*}
+(\hat{\alpha}, \hat{\nu}, \hat{\tau}, \hat{\boldsymbol{\omega}}) &=& \argmin_{\nu, \tau > 0\ \alpha, \omega_j \in [0,1]} \sum_{k \in S_{train}} \sum_{t=1}^{T_k} \left(Z_k - \hat{p}_{t,k}(\alpha, \nu, \tau, \boldsymbol{\omega})\right)^2\\
+&& \text{s.t.} \sum_{j=1}^J \omega_j = 1
+\end{eqnarray*} 
 \end{enumerate}
@@
 The competing models are evaluated via a 10-fold cross-validation\footnote{A 5-fold cross-validation was also performed.
-The results were, however, very similar to the 10-fold cross-validation and hence not presented in the paper.} that first partitions the 166 questions into 10 sets.
-The partition is chosen such that each of the 10 sets has approximately the same number of questions (16 or 17 questions per set in our case) and the same number of time points (between 1760 and 1764 time points per set in our case).
+The results were, however, very similar to the 10-fold cross-validation and hence not presented in the paper.} that first partitions the 166 questions into 10 sets such that each set has approximately the same number of questions (16 or 17 questions in our case) and the same number of time points (between 1,760 and 1,764 time points our case).
 The evaluation then iterates 10 times, each time using one of the 10 sets as the testing set and the remaining 9 sets as the training set.
 Therefore each question is used nine times for training and exactly once for testing.
-The testing proceeds sequentially one testing question at a time as follows: First, for a question with a time horizon of $T_k$, make a prediction based on the first two days.
-Compute the Brier score for the aggregate forecast of the second day.
-Next make a prediction based on the first three days and  compute the Brier score for the most recent day, namely, the third day.
-Repeat this process until the prediction is made on all of the $T_k-1$ days.
+The testing proceeds sequentially one testing question at a time as follows: First, for a question with a time horizon of $T_k$, give an aggregate probability at time $t=2$ based on the first two days.
+Compute the Brier score for this probability.
+Next give an aggregate probability at time $t=3$ based on the first three days and  compute the Brier score for this probability.
+Repeat this process for all of the $T_k-1$ days.
 This leads to $T_k-1$ Brier scores per testing question and a total of  17,475 Brier scores across the entire dataset.
 
-Table \ref{prediction} summarizes different ways to aggregate these scores.
+Table \ref{prediction} summarizes these scores in different ways.
 The first option, denoted by \textit{Scores by Day}, weighs each question by the number of days the question remained open.
@@
 $\text{SAC}_{\text{LOG}}$ & \textbf{0.096} (0.191) & 0.056 (0.134) & \textbf{0.096} (0.189) & \textbf{0.098} (0.193)\\ 
-EWMBA & 0.102 (0.203) & 0.060 (0.124) & 0.110 (0.201) & 0.103 (0.206)\\ 
+%EWMBA & 0.102 (0.203) & 0.060 (0.124) & 0.110 (0.201) & 0.103 (0.206)\\ 
+EWMBA & 0.104 (0.204) & 0.057 (0.120) & 0.113 (0.205) & 0.105 (0.206)\\ 
 EWMLA & 0.102 (0.199) & 0.061 (0.130) & 0.111 (0.214) & 0.103 (0.200)\\ 
-EWMA & 0.111 (0.142) & 0.089 (0.100) & 0.111 (0.136) & 0.112 (0.144)\\ 
+%EWMA & 0.111 (0.142) & 0.089 (0.100) & 0.111 (0.136) & 0.112 (0.144)\\ 
+EWMA & 0.111 (0.146) & 0.080 (0.101) & 0.116 (0.152) & 0.112 (0.146)\\ 
 \hline
@@
 $\text{SAC}_{\text{LOG}}$ & \textbf{0.082} (0.142) & 0.055 (0.096) & \textbf{0.105} (0.174) & \textbf{0.085} (0.144)\\ 
-EWMBA & 0.090 (0.156) & 0.063 (0.101) & 0.118 (0.186) & 0.091 (0.161)\\ 
+EWMBA & 0.091 (0.157) & 0.057 (0.095) & 0.121 (0.187) & 0.093 (0.164)\\ 
+%EWMBA & 0.090 (0.156) & 0.063 (0.101) & 0.118 (0.186) & 0.091 (0.161)\\ 
 EWMLA & 0.090 (0.159) & 0.064 (0.109) & 0.120 (0.200) & 0.090 (0.159)\\ 
-EWMA & 0.104 (0.105) & 0.092 (0.081) & 0.119 (0.125) & 0.103 (0.107)\\ 
+EWMA & 0.102 (0.108) & 0.080 (0.075) & 0.123 (0.130) & 0.103 (0.110)\\ 
+%EWMA & 0.104 (0.105) & 0.092 (0.081) & 0.119 (0.125) & 0.103 (0.107)\\ 
 \hline
@@
 
-Overall, $\text{SAC}_{\text{BRI}}$ and $\text{SAC}_{\text{LOG}}$ achieve the lowest average scores across all columns except \textit{Short}, where they are slightly outperformed by $\text{BSAC}_{\text{LOG}}$.
-However, $\text{BSAC}_{\text{LOG}}$ turns out to be overconfident (see Section \ref{calibration}).
-This means that $\text{BSAC}_{\text{LOG}}$ underestimates the uncertainty in the events and outputs probability forecasts that are typically too near 0.0 or 1.0.
-As a result, the aggregate forecasts are either very close to the correct answer or very far from it.
-As Table \ref{prediction} shows, this results into highly variable forecasting performance.
-The short questions, however, involve very little uncertainty and were generally the easiest to forecast.
-On such easy questions, overconfidence can pay off frequently enough to compensate for a few large losses arising from the overconfident and incorrect forecasts.
-
-In contrast to $\text{BSAC}_{\text{LOG}}$, the baseline SDLM-model lacks sharpness and is highly under-confident (see Section \ref{calibration}).
-This behavior is expected as the experts are under-confident at the group-level (see Section \ref{ExpertBias}) and the SDLM-procedure does not use the training set to explicitly calibrate its forecasts.
+As can be seen in Table \ref{prediction}, $\text{SAC}_{\text{LOG}}$ achieves the lowest score across all columns except \textit{Short} where it is outperformed by $\text{BSAC}_{\text{LOG}}$.
+It turns out that $\text{BSAC}_{\text{LOG}}$ is overconfident (see Section \ref{calibration}).
+This means that $\text{BSAC}_{\text{LOG}}$ underestimates the uncertainty in the events and outputs aggregate probabilities that are typically too near 0.0 or 1.0.
+%As a result, the aggregate forecasts are either very close to the correct answer or very far from it.  
+This results into highly variable performance.
+The short questions generally involved very little uncertainty.
+On such easy questions, overconfidence can pay off frequently enough to compensate for a few large losses arising from the overconfident and drastically incorrect forecasts.
+
+SDLM, on other hand, lacks sharpness and is highly under-confident (see Section \ref{calibration}).
+This behavior is expected as the experts are under-confident at the group-level (see Section \ref{ExpertBias}) and SDLM does not use the training set to explicitly calibrate its aggregate probabilities.
 Instead, it merely smooths the forecasts given by the experts.
-The resulting aggregate forecasts are therefore necessarily conservative, resulting into high average scores with low variability.
+The resulting aggregate probabilities are therefore necessarily conservative, resulting into high average scores with low variability.
 
  Similar behavior is exhibited by EWMA that performs the worst of all the competing models.
-In contrast, the other two exponentially weighted aggregators, EWMLA and EWMBA, make efficient use of the training set and present moderate forecasting performance in most columns of Table \ref{prediction}.
-Recall that EWMBA uses the cumulative distribution function of the Beta distribution that depends on two parameters and is more flexible than the transformation used by EWMLA.
-On other hand, only EWLMA is given access to the self-reported expertise information.
+The other two exponentially weighted aggregators, EWMLA and EWMBA, make efficient use of the training set and present moderate forecasting performance in most columns of Table \ref{prediction}.
 Neither approach, however, appears to dominate the other.
-The aggregators perform very similarly.
 The high variability and average of their performance scores indicate that their performance suffers from over-confidence.
@@
 A calibration plot is a simple tool for visually assessing the sharpness and calibration of a model.
-The idea is to plot the probability forecasts against the observed empirical frequencies.
+The idea is to plot the aggregate probabilities against the observed empirical frequencies.
 Therefore any deviation from the diagonal line suggests poor calibration.
@@
 \caption[Optional caption for list of figures]{The top and bottom rows show in- and out-of-sample calibration and sharpness, respectively. 
-The models are Simple Dynamic Linear Model (SDLM), the \textit{Sample-And-Calibrate} approach that optimizes over the logarithmic score ($\text{SAC}_{\text{LOG}}$), the \textit{Sample-And-Calibrate} approach that optimizes over the Brier score ($\text{SAC}_{\text{BRI}}$), and the fully-Bayesian version of the \textit{Sample-And-Calibrate} approach that optimizes over the logarithmic score ($\text{BSAC}_{\text{LOG}}$).
+%The models are Simple Dynamic Linear Model (SDLM), the \textit{Sample-And-Calibrate} approach that optimizes over the logarithmic score ($\text{SAC}_{\text{LOG}}$), the \textit{Sample-And-Calibrate} approach that optimizes over the Brier score ($\text{SAC}_{\text{BRI}}$), and the fully-Bayesian version of the \textit{Sample-And-Calibrate} approach that optimizes over the logarithmic score ($\text{BSAC}_{\text{LOG}}$).
 }
@@
 
-The top and bottom rows of Figure \ref{Calibration-Out} present calibration plots for SDLM, $\text{SAC}_{\text{LOG}}$, $\text{SAC}_{\text{BRI}}$, and $\text{BSAC}_{\text{LOG}}$ under in- and out-of-sample probability estimation, respectively.
+The top and bottom rows of Figure \ref{Calibration-Out} present calibration plots for SDLM, $\text{SAC}_{\text{LOG}}$, $\text{SAC}_{\text{BRI}}$, and $\text{BSAC}_{\text{LOG}}$ under in- and out-of-sample probability aggregation, respectively.
 Each setting is of interest in its own right:  Good in-sample calibration is crucial for model interpretability.
 In particular, if the estimated crowd belief is well-calibrated, then the elements of the bias vector $\boldsymbol{b}$ can be used to study the amount of under- or over-confidence in the different expertise groups.
-Good out-of-sample calibration and sharpness, on other hand, are necessary properties in predicting future events with high accuracy.
+Good out-of-sample calibration and sharpness, on other hand, are necessary properties in decision making.
 To guide our assessment, the dashed bands around the diagonal connect the point-wise, Bonferroni-corrected (\citet{bonferroni}) 95\% lower and upper critical values under the null hypothesis of calibration.
@@
 This is, however, not the case with $\text{BSAC}_{\text{LOG}}$ that is well-calibrated in-sample but presents over-confidence out-of-sample.
-Figures \ref{CalibrationNONEa} and \ref{CalibrationNONEb} serve as baselines by showing the reliability plots for SDLM.
+Figures \ref{CalibrationNONEa} and \ref{CalibrationNONEb} serve as baselines by showing the calibration plots for SDLM.
 Given that this model does not perform any explicit calibration, it is not surprising to see most points outside the critical values.
-The pattern in the deviations suggests drastic under-confidence.
+The pattern in the deviations suggests strong under-confidence.
 Furthermore, the inset histogram reveals drastic lack of sharpness.
@@
 
-\caption[Optional caption for list of figures]{Comparing the bias-levels across self-reported expertise under different approaches. 
-Posterior distributions of $b_j$ for $j = 1, \dots, 5$ under the \textit{Sample-And-Calibrate} approach that optimizes over the logarithmic score ($\text{SAC}_{\text{LOG}}$) and the \textit{Sample-And-Calibrate} approach that optimizes over the Brier score ($\text{SAC}_{\text{BRI}}$).
-}
+\caption[Optional caption for list of figures]{
+%Comparing the bias-levels across self-reported expertise under different approaches. 
+Posterior distributions of $b_j$ for $j = 1, \dots, 5$.}
+% under $\text{SAC}_{\text{LOG}}$ and $\text{SAC}_{\text{BRI}}$.}
+%
+%the \textit{Sample-And-Calibrate} approach that optimizes over the logarithmic score ($\text{SAC}_{\text{LOG}}$) and the \textit{Sample-And-Calibrate} approach that optimizes over the Brier score ($\text{SAC}_{\text{BRI}}$).}
 \label{Biases}
@@
 
-Recall from Section \ref{data} that the experts were asked to self-assess their level of expertise (on a 1-to-5 scale with 1 = Not At All Expert to 5 = Extremely Expert) on any questions in which they participated.
+%Recall from Section \ref{data} that the experts were asked to self-assess their level of expertise (on a 1-to-5 scale with 1 = Not At All Expert to 5 = Extremely Expert) on any questions in which they participated.
 The self-reported expertise then divides the experts into 5  groups, with each group assigned a separate multiplicative bias term.
-This section uses SAC to explore the posterior distributions of these multiplicative bias terms.
-Figure \ref{Biases} presents the posterior distributions of the bias terms with side-by-side box plots.
-Given that the distributions fall completely below the \textit{no-bias} reference-line at 1.0, all the groups are deemed under-confident.
+This section explores the bias among the five expertise groups in our dataset.
+Figure \ref{Biases} compares the posterior distributions of the individual elements of $\boldsymbol{b}$ with side-by-side boxplots.
+Given that the distributions fall completely below the \textit{no-bias} reference-line at 1.0, all the expertise groups are deemed under-confident.
 Even though the exact level of under-confidence is affected slightly by the extent to which the extreme probabilities are truncated (see Section \ref{practicalMatters}), the qualitative results in this section remain insensitive to different levels of truncation.
 
-The under-confidence decreases as the level of  expertise increases.
-For instance, the posterior probability that the most expert group is the least under-confident is approximately equal to $1.0$, and the posterior probability of a strictly decreasing level of under-confidence is approximately 0.87.
+From Figure \ref{Biases} it is clear that the under-confidence decreases as expertise increases.
+The posterior probability that the most expert group is the least under-confident is approximately equal to $1.0$, and the posterior probability of a strictly decreasing level of under-confidence is approximately 0.87.
 The latter probability is driven down by the inseparability of the two groups with the lowest levels of self-reported expertise.
-The fact that these groups are very similar suggests that the experts are poor at assessing how little they know about a subject that is strange to them.
+This inseparability suggests that the experts are poor at assessing how little they know about a topic that is strange to them.
 If these groups are combined into a single group, the posterior probability of a strictly decreasing level of under-confidence is approximately 1.0.
@@
 Given that experts are more likely to disagree over a difficult question than an easy one, it is reasonable to assume that $\sigma^2_k$ has a positive relationship with question difficulty.
-An alternative measure is given by $\tau_k$ that quantifies the volatility of the underlying circumstances that ultimately decide the outcome of the event.
-Therefore a high value of $\tau_k$ can cause the outcome of the event to appear unstable and difficult to predict.
-
-\begin{figure}[h!]
-\vspace{-2em}
-\centering
-\hspace*{-3.5em} 	\includegraphics{Figures/LegendExamplePlot} % requires the graphicx package
-\vspace{-3.5em}
-
-\hspace{-0.5em}
-\subfigure[\textit{Will the expansion of the European bailout fund be ratified by all 17 Eurozone nations before 1 November 2011?}]{
-\hspace*{-1em}  \includegraphics[width= 0.49\textwidth]{Figures/ExamplePlot1b}
-\label{Example1b}
-} \hspace{0.5em}
-\subfigure[\textit{Will the Nikkei 225 index finish trading at or above 9,500 on 30 September 2011?}]{
-\hspace*{-1em}  \includegraphics[width= 0.49\textwidth]{Figures/ExamplePlot2b}
-\label{Example2b}
-}
-
-\caption[Optional caption for list of figures]{Scatterplots of the probability forecasts given for two questions in our dataset.
-The shadings represents the self-reported expertise of the expert who provided the probability forecast.
-The solid line gives the posterior mean of the calibrated crowd belief as estimated by $\text{SAC}_{\text{LOG}}$.
-The surrounding dashed lines connect the point-wise 95\% posterior intervals.}
-\label{ExamplePlotsFinal}
-\end{figure}
-
-As a final illustration of our model, we return to the two example questions introduced in Section \ref{data}.
-Figure \ref{ExamplePlotsFinal} is a copy of Figure \ref{ExamplePlots} with the addition of a solid line surrounded by a dashed band.
-The solid line represents the posterior mean of the calibrated crowd belief as estimated by $\text{SAC}_{\text{LOG}}$.
+An alternative measure is given by $\tau_k^2$ that quantifies the volatility of the underlying circumstances that ultimately decide the outcome of the event.
+Therefore a high value of $\tau_k^2$ can cause the outcome of the event to appear unstable and difficult to predict.
+
+As a final illustration of our model, we return to the two example questions introduced Figure \ref{ExamplePlotsFinal}. 
+%The solid line represents the posterior mean of the calibrated crowd belief as estimated by $\text{SAC}_{\text{LOG}}$.
 The dashed lines connect the point-wise 95\% posterior intervals across different time points.
@@
 In response to this lack of appropriate modeling procedures, our work introduces an interpretable time-series model that incorporates self-reported expertise to capture a sharp and well-calibrated estimate of the crowd belief.
-The model estimation is performed in two steps: The first step, known as the \textit{sampling step}, samples constrained versions of the model parameters via Gibbs sampling.
-The sampling is done from standard distributions with fast convergence to the target distribution (see Appendix \ref{appendix} for technical details).
+%The model estimation is performed in two steps: The first step, known as the \textit{sampling step}, samples constrained versions of the model parameters via Gibbs sampling.
+The sampling is done from standard distributions with fast convergence to the target distribution (see Online Supplement for technical details).
 The second step, known as the \textit{calibration step}, uses a one-dimensional optimization procedure to transform the constrained parameter values to their unconstrained counterparts.
 To the best of our knowledge, this procedure extends the forecasting literature into rather unexplored areas of probability aggregation.
-
-\subsection{Summary of Findings}
-The model was applied to an unusually large dataset on expert probability forecasts.
+%
+%\subsection{Summary of Findings}
+%The model was applied to an unusually large dataset on expert probability forecasts.
 The estimated crowd belief was found to be sharp and well-calibrated under both in- and out-of-sample settings.
@@
 
-\subsection{Limitations and Directions for Future Research}
+%\subsection{Limitations and Directions for Future Research}
 Our model preserves parsimony while addressing the main challenges in modeling sparse probability forecasting data.
@@
 This hypothesis could be explored by letting $\sigma^2_{t,k}$ evolve dynamically as a function of the previous term $\sigma^2_{t-1,k}$ and random noise.
-Furthermore, this parameter along with the bias term can be explored at an individual level if the experts updated their forecasts very frequently.
-Estimates of $\sigma_{i,k}^2$ and $b_{i,k}$ could then be used to separate the accurate forecasters from the inaccurate ones.
-
-If the experts were asked to provide personal information during the data collection process, it may be of interest to study the forecasting behavior of different kinds of experts.
-For instance, this paper modeled the bias separately within each expertise group.
+
+%If the experts were asked to provide personal information, it may be of interest to study the forecasting behavior of different kinds of experts. For instance, 
+%
+This paper modeled the bias separately within each expertise group.
 This is by no means restricted to the study of bias or its relation to self-reported expertise.
@@
 This way the researcher can gain insight about the kind of experts who generally perform well on certain types of questions.
-The results would be of interest to the decision-maker who could use the information as a basis for consulting only a high-performing subset of the available experts.
+The results would be of interest to the decision-maker who could use the information as a basis for hiring only a high-performing subset of the available experts.
 
@@
 For instance, a government official may want to sacrifice some calibration for extra sharpness.
-This can be easily achieved by changing the optimization criterion in Equation (\ref{OSE}) to an appropriate function that meets the new goal.
+This can be easily achieved by changing the optimization criterion in (\ref{OSE}) to an appropriate function that meets the new goal.
 
@@
 
-\appendix
-\section{Technical Details of the Sampling Step}
-\label{appendix}
-
-The Gibbs sampler (\citet{geman1984stochastic}) iteratively samples all the unknown parameters from their full-conditional posterior distributions one block of parameters at a time.
-Given that this is performed under the constraint $b_3 = 1$ to ensure model identifiability, the constrained parameter estimates should be denoted with a trailing $(1)$ to maintain consistency with earlier notation.
-For instance, the constrained estimate of $\gamma_k$ should be denoted by $\hat{\gamma}_k(1)$ while the unconstrained estimate is denoted by $\hat{\gamma}_k$.
-For the sake of clarity, however, the constraint suffix is omitted in this section.
-Nonetheless, it is important to keep in mind that all the estimates in this section are constrained.
-
-\begin{center}
-\framecolorbox[\textwidth]{gray}{gray!15}{Sample $X_{t,k}$}
-\end{center}
-The hidden states are sampled via the \textit{Forward-Filtering-Backward-Sampling} (FFBS) algorithm that first predicts the hidden states using a Kalman Filter and then performs a backward sampling procedure that treats these predicted states as additional observations (see, e.g., \cite{carter1994gibbs, migon2005dynamic} for details on FFBS).
-More specifically, the first part, namely the Kalman Filter, is deterministic and consists of a predict and an update step.
-Given all the other parameters except the hidden states, the predict step for the $k$th question is
-\begin{align*}
-X_{t|t-1,k} &= \gamma_k X_{t-1|t-1,k} \\
-P_{t|t-1, k} &= \gamma_k^2 P_{t-1|t-1, k} + \tau_k^2,
-\end{align*}
-where the initial values, $X_{0|0,k}$ and $P_{0|0, k}$, are equal to $0$ and $1$, respectively. The update step is 
-\begin{align*}
-e_{t,k} &= Y_{i,t,k} - b_{i,k} X_{t | t-1, k} \\
-S_{t,k} &=  \sigma_k^2 + b_{i,k}^2 P_{t|t-1, k}\\
-K_{t,k} &=  P_{t|t-1, k} b_{i,k} S_{t,k}^{-1} \\
-X_{t|t, k} &= X_{t|t-1, k} + K_{t,k} e_{t,k} \\
-P_{t|t,k} &= (1 - K_{t,k} b_{i,k}) P_{t|t-1,k},
-\end{align*}
-where $b_{i,k}$ is the corresponding bias term for the $i$th expert in the $k$th question.
-The update step is repeated sequentially for each observation $Y_{i,t,k}$ given at time $t$.
-For each such repetition, the previous posterior values, $X_{t|t, k}$ and $P_{t|t, k}$, are considered as the new prior values, $X_{t|t-1, k}$ and $P_{t|t-1, k}$.
-If the observation $\boldsymbol{Y}_{t,k}$ is completely missing at time $t$, the update step is skipped and
-\begin{align*}
-X_{t|t, k} &= X_{t|t-1, k}\\
-P_{t|t,k} &= P_{t|t-1,k}
-\end{align*}
-After running the Kalman Filter up to the final time point at $t = T_k$, the final hidden state is sampled from $X_{T_k,k} \sim \mathcal{N}(X_{T_k|T_k, k}, P_{T_k|T_k, k})$.
-The remaining states are obtained via the backward sampling that is performed in reverse from
-\begin{align*}
-X_{t-1, k} &\sim  \mathcal{N} \left(V\left( \frac{\gamma_kX_{t,k}}{\tau_k^2}  + \frac{X_{t|t,k}}{P_{t|t,k} } \right),  V\right),
-\end{align*}
-where
-\begin{align*}
-V &= \left( \frac{\gamma_k^2}{\tau_k^2} + \frac{1}{P_{t|t,k}}\right)^{-1}
-\end{align*}
-This can be viewed as backward updating that considers the Kalman Filter estimates as additional observations at each given time point.
-
-%
-%If the observation $\boldsymbol{Y}_{t,k}$ is completely missing at time $t$, the update step is skipped and the state estimates are sampled from
-%\begin{align*}
-%\mathcal{N}\left(\gamma_k X_{t-1|t-1,k}, \gamma_k^2 P_{t-1|t-1,k} + \tau_k^2\right)
-%\end{align*}
-
-\begin{center}
-\framecolorbox[\textwidth]{gray}{gray!15}{Sample $\boldsymbol{b}$ and $\sigma^2_k$}
-\end{center}
-First, vectorize all the response vectors $\boldsymbol{Y}_{t,k}$ into a single vector denoted $\boldsymbol{Y}_k = \left[\boldsymbol{Y}_{1,k}^T, \dots, \boldsymbol{Y}_{T_k,k}^T\right]^T$.
-Given that each $\boldsymbol{Y}_{t,k}$ is matched with $X_{t,k}$ via the time index $t$, we can form a $|\boldsymbol{Y}_k| \times J$ design-matrix by letting $$\boldsymbol{X}_k = \left[ (\boldsymbol{M}_kX_{1,k})^T, \dots, (\boldsymbol{M}_kX_{T_k,k})^T \right]^T$$ Given that the goal is to borrow strength across questions by assuming a common bias vector $\boldsymbol{b}$, the parameter values must be estimated in parallel for each question such that the matrices $\boldsymbol{X}_k$ can be further concatenated into $\boldsymbol{X} = [\boldsymbol{X}_1^T, \dots, \boldsymbol{X}_K^T]^T$ during every iteration.
-Similarly, $\boldsymbol{Y}_k$ must be further vectorized into a vector $\boldsymbol{Y} = [\boldsymbol{Y}_1^T, \dots, \boldsymbol{Y}_K^T]^T$.
-The question-specific variance terms are taken into account by letting $\boldsymbol{\Sigma} = \text{diag}(\sigma^2_1 \boldsymbol{1}_{1 \times T_1}, \dots, \sigma^2_K \boldsymbol{1}_{1 \times T_K})$.
-After adopting the non-informative prior $p(\boldsymbol{b}, \sigma_k^2 | \boldsymbol{X}_k) \propto \sigma_k^{-2}$ for each $k = 1, \dots, K$, the bias vector is sampled from
-\begin{eqnarray}
-\boldsymbol{b} | \dots &\sim& \mathcal{N}_J \left( (\boldsymbol{X}^T \boldsymbol{\Sigma}^{-1} \boldsymbol{X})^{-1} \boldsymbol{X}^T \boldsymbol{\Sigma}^{-1} \boldsymbol{Y}, (\boldsymbol{X}^T \boldsymbol{\Sigma}^{-1} \boldsymbol{X})^{-1} \right) \label{BiasVector}
-\end{eqnarray}
-Given that the covariance matrix in Equation (\ref{BiasVector}) is diagonal, the identifiability constraint can be enforced after sampling a new value of $\boldsymbol{b}$ by letting $b_3 = 1$.
-The variance parameters are then sampled from
-\begin{eqnarray*}
-\sigma^2_k | \dots &\sim& \text{Inv-}\chi^2\left(|\boldsymbol{Y}_k| - J, \frac{1}{|\boldsymbol{Y}_k| - J} (\boldsymbol{Y}_k - \boldsymbol{X}_k \boldsymbol{b})^T (\boldsymbol{Y}_k - \boldsymbol{X}_k \boldsymbol{b}) \right),
-\end{eqnarray*}
-where the distribution is a scaled inverse-$\chi^2$ (see, e.g., \citet{gelman2003bayesian}).
-Given that the experts are not required to give a new forecast at every time unit, the design matrices must be trimmed accordingly such that their dimensions match up with the dimensions of the observed matrices.
-
-\begin{center}
-\framecolorbox[\textwidth]{gray}{gray!15}{Sample $\gamma_k$ and $\tau^2_k$}
-\end{center}
-The parameters of the hidden process are estimated via a regression setup.
-More specifically, after adopting the non-informative prior $p(\gamma_k, \tau_k^2 | \boldsymbol{X}_k) \propto \tau_k^{-2}$, the parameter values are sampled from
-\begin{eqnarray*}
-\gamma_k | \dots  &\sim&  \mathcal{N} \left( \frac{\sum_{t=2}^{T_k} X_{t,k}X_{t-1,k}}{\sum_{t=1}^{T_k-1} X_{t,k}^2} , \frac{\tau_k^2}{\sum_{t=1}^{T_k-1} X_{t,k}^2} \right)\\
-\tau_k^2 | \dots &\sim& \text{Inv-}\chi^2 \left(T_k-1, \frac{1}{T_k-1} \sum_{t=2}^{T_k} \left(X_{t,k} - \gamma_k X_{t-1,k} \right)^2 \right),
-\end{eqnarray*}
-where the final distribution is a scaled inverse-$\chi^2$ (see, e.g., \citet{gelman2003bayesian}).
-
 \bibliographystyle{imsart-nameyear}

### 2014-01-31 Revision/Supplementary/supplementary.tex
added revision

--- 2013-09-29 paper.tex
@@
 % "Title of the paper"
-\title{Probability Aggregation in Time-Series: Dynamic Hierarchical Modeling of Sparse Expert Beliefs}
-\runtitle{Probability Aggregation in Time-Series}
+\title{Probability Aggregation in Time-Series: Dynamic Hierarchical Modeling of Sparse Expert Beliefs: Online Supplement}
+\runtitle{Probability Aggregation in Time-Series: Online Supplement}
 
@@
 
- \affiliation{Department of Statistics, The Wharton School of the University of Pennsylvania\\   \printead{e1,e3}}
- \affiliation{Department of Psychology, University of Pennsylvania\\ \printead{e4,e5}}
- \affiliation{Department of Computer and Information Science, University of Pennsylvania\\ \printead{e2}}
- \address{Philadelphia, PA 19104- 6340, USA\\  \printead{e1}}
-
+% \affiliation{Department of Statistics, The Wharton School of the University of Pennsylvania\\   \printead{e1,e3}}
+% \affiliation{Department of Psychology, University of Pennsylvania\\ \printead{e4,e5}}
+% \affiliation{Department of Computer and Information Science, University of Pennsylvania\\ \printead{e2}}
+% \address{Philadelphia, PA 19104- 6340, USA\\  \printead{e1}}
+%  
 \end{aug}
 
-\begin{abstract}
-Most subjective probability aggregation procedures use a single probability judgement from each expert, even though it is common for experts studying real problems to update their probability estimates over time.
-This paper advances into unexplored areas of probability aggregation by considering a dynamic context in which experts can update their beliefs at random intervals.
-The updates occur very infrequently, resulting in a highly sparse dataset that cannot be modeled by standard time-series procedures.
-In response to the lack of appropriate methodology, this paper presents a hierarchical model that takes into account the expert's level of self-reported expertise and produces aggregate probabilities that are sharp and well-calibrated both in- and out-of-sample.
-The model is demonstrated on a real-world dataset that includes over 2,300 experts making multiple probability forecasts over a period of two years on different subsets of 166 international political events.
-\end{abstract}
-
-\begin{keyword}
-\kwd{Probability Aggregation}
-\kwd{Dynamic Linear Model}
-\kwd{Hierarchical Modeling}
-\kwd{Expert Forecast}
-\kwd{Subjective Probability}
-\kwd{Bias Estimation}
-\kwd{Calibration}
-\kwd{Time Series}
-\end{keyword}
 \end{frontmatter}
 
-\section{Introduction}
-\label{intro}
-Individual experts can differ radically from one another in their abilities to assess probabilities of future events.
-Their probability assessments are often evaluated and compared on \textit{calibration}, which measures how closely the frequency of event occurrence agrees with the assigned probabilities.
-For instance, the proportion of occurrences is 60\% for all those events to which a well-calibrated expert assigned a probability of 0.60.
-Even though several experiments have been conducted to show that experts are generally poorly calibrated (see, e.g., \citet{cooke1991experts, shlyakhter1994quantifying}), relative differences can occur among different types of experts.
-In particular, \citet{wright1994coherence} argue that a higher level of self-reported expertise can be associated with better calibration.
+This supplementary material accompanies the paper ``Probability Aggregation in Time-Series: Dynamic Hierarchical Modeling of Sparse Expert Beliefs''.
+It provides a technical description of the sampling step of the   SAC-algorithm.
 
-Calibration by itself, however, is not sufficient for useful probability estimation.
-Consider a relatively stationary process, such as rain on different days in a given geographic region, where the observed frequency of occurrence in the last 10 years is 45\%.
-In this setting an expert could always  assign a constant probability of 0.45 and be well-calibrated.
-This assessment, however, can be made without any subject-matter expertise or special training.
-For this reason the long-term frequency is often considered as the baseline probability -- a naive assessment that provides the decision-maker very little extra information.
-Therefore experts should aim to make probability assessments that are as far from the baseline as possible.
-The extent to which their probabilities differ from the baseline is measured by \textit{sharpness} (\citet{gneiting2008rejoinder, winkler2008comments}).
-If the experts are both sharp and well-calibrated, they can forecast the behavior of the process with high certainty and accuracy.
-From the decision-maker's perspective their advice is highly useful as it involves much information and very little risk.
-Therefore, to summarize, useful probability estimation should aim to maximize sharpness subject to calibration (see, e.g., \citet{raftery2005using, murphy1987general}).
-This is a well-defined goal that has led to a wide range of novel and insightful observations in probability forecasting.
-
-%One such observation is related to the aggregation of multiple probabilities. 
-There is strong empirical evidence that bringing together the strengths of different experts by combining their probability forecasts into a single consensus, known as the \textit{crowd belief},  improves predictive performance.
-Prompted by the many applications of probability forecasts, including medical diagnosis (\citet{wilson1998prediction, pepe2003statistical}, political and socio-economic foresight (\citet{tetlock2005expert}), and meteorology (\citet{sanders1963subjective, vislocky1995improved, baars2005performance}), researchers have proposed many approaches to combining probability forecasts (see, e.g., \citet{Ranjan08, satopaa} for some recent studies, and \citet{Genest, Wallsten97evaluatingand, clemen2007aggregating, primo2009calibration} for a comprehensive overview).
-The general focus, however, has been on developing one-time aggregation procedures that consult the expert's advice only once before the event resolves.
-
-Consequently, many areas of probability aggregation still remain rather unexplored.
-For instance, consider investors aiming to assess whether a stock index will finish trading above a threshold on a given date.
-To maximize their overall predictive accuracy, they may consult a group of experts repeatedly over a period of time and adjust their estimate of the aggregate probability accordingly.
-Given that the experts are allowed to update their probability assessments, the aggregation should be performed by taking into account the temporal correlation in their advice.
-Many standard time-series procedures could be used to perform this aggregation as long as the experts update their advice consistently on a daily basis.
-
-This paper adds another layer of complexity by assuming a heterogeneous set of experts, most of whom only make one or two probability assessments over the hundred or so days before the event resolves.
-This means that the decision-maker faces a different group of experts  every day, with only a few experts returning later on for a second round of advice.
-The problem at hand is therefore strikingly different from many time-series estimation problems, where one has an observation at every time point -- or almost every time point.
-As a result, standard time-series procedures like ARIMA (see, e.g., \citet{mills1991time}) are not directly applicable.
-This paper introduces a time-series model that incorporates self-reported expertise and captures a sharp and well-calibrated estimate of the crowd belief.
-The model is highly interpretable and can be used for:
-\begin{itemize}
-\item borrowing strength across the hierarchy to analyze under- and overconfidence in different groups of experts,
-%\item confidence bands for the crowd belief,
-\item accurate probability forecasts, and
-\item many question-specific quantities that have easy interpretations, such as expert disagreement and problem difficulty, and can be used to gain novel insight in the social sciences. 
-\end{itemize}
-
-This paper begins by describing our geopolitical database.
-The paper then introduces a dynamic hierarchical model for capturing the crowd belief.
-The model is estimated in a two-step procedure: first, a sampling step produces constrained parameter estimates via Gibbs sampling (see, e.g., \citet{geman1984stochastic}); second, a calibration step transforms these estimates to their unconstrained equivalents via a one-dimension optimization procedure.
-The model introduction is followed by the first evaluation section that uses synthetic data to study how accurately the two-step procedure can estimate parameter values.
-The second evaluation section applies the model to our real-world geopolitical forecasting database.
-The paper concludes with a discussion on future research directions and model limitations.
-
-\section{Geopolitical Forecasting Data}
-\label{data}
-
-\begin{table*}
-\caption{Five-number summaries of our real-world data.}
-\label{DataStats}
-\begin{tabular}{m{2in}	lllllll }
-\hline
-Statistic & Min. & $Q_1$ & Median & Mean & $Q_3$ & Max.\\ \hline
-\# of Days a Question is Active  & 4 &  35.6 &  72.0 & 106.3 & 145.20 & 418 \\
-\# of Experts per Question & 212 &   543.2 &  693.5 &  783.7 &  983.2 & 1690\\ 
-\# Forecasts given by each Expert on a Question & 1 &  1.0 &   1.0  &  1.8 &  2.0 & 131 \\
-\# Questions participated by an Expert & 1 &  14.0 &   36.0  &  55.0 &  90.0 & 166 \\ \hline
-\end{tabular}
-\end{table*}
-
-\begin{table*}
-\caption{Frequencies of the self-reported expertise (1 = Not At All Expert and 5 = Extremely Expert) levels across all the 166 questions in our real-world data.}
-\label{ExpertiseTable}
-\begin{center}
-\begin{tabular}{llllllll }
-\hline
-Expertise Level & 1 & 2 & 3 & 4 & 5\\
-Frequency (\%) & 25.3 & 30.7 & 33.6 &  8.2 &  2.1 \\\hline
-\end{tabular}
-\end{center}
-\end{table*}
-
-The data collection began with a recruitment of 2,365 experts ranging from graduate students to political science faculty and practitioners.
-The recruiting was made from professional societies, research centers, alumni associations, science bloggers, and word of mouth.
-Requirements included at least a Bachelor's degree and completion of psychological and political tests that took roughly two hours.
-These measures assessed cognitive styles, cognitive abilities, personality traits, political attitudes, and real-world knowledge.
-The experts were asked to give probability forecasts (to the second decimal point) and to self-assess their level of expertise (on a 1-to-5 scale with 1 = Not At All Expert and 5 = Extremely Expert) on a number of 166 geopolitical binary events taking place between September 29, 2011 and May 8, 2013.
-Each question was active for a period during which the participating experts could update their forecasts as frequently as they liked without penalty.
-The experts knew that their probability estimates would be assessed for accuracy using Brier scores\footnote{The Brier score is the squared distance between the probability forecast and the event indicator that equals 1.0 or 0.0 depending on whether the event happened or not, respectively.
-See \citet{Brier} for the original introduction.}.
-This incentivized them to report their true beliefs instead of attempting to game the system (\citet{winkler1968good}).
-In addition to receiving \$150 for meeting minimum participation requirements that did not depend on prediction accuracy, the experts received status rewards for their performance via leader-boards displaying Brier scores for the top 20 experts.
-Given that a typical expert participated only in a small subset of the 166 questions, the experts are considered indistinguishable conditional on the level of self-reported expertise.
-
-The experts made updates on a very infrequent basis: the average number of forecasts made by a single expert in one day was around 0.017, and the average group-level response rate was around 13.5 forecasts per day.
-Given that the group of experts is large and diverse, the resulting dataset is very sparse.
-Tables \ref{DataStats} and \ref{ExpertiseTable} provide relevant summary statistics on the data.
-Notice that the distribution of the self-reported expertise is skewed to the right and that some questions remained active longer than others.
-For more details on the dataset and its collection see \citet{ungar2012good}.
-
-\begin{figure}[h!]
-\vspace{-2em}
-\centering
-\hspace*{-3.5em} 	\includegraphics{Figures/LegendExamplePlot} % requires the graphicx package
-\vspace{-3.5em}
-
-\hspace{-0.5em}
-\subfigure[\textit{Will the expansion of the European bailout fund be ratified by all 17 Eurozone nations before 1 November 2011?}]{
-\hspace*{-1em}  \includegraphics[width= 0.49\textwidth]{Figures/ExamplePlot1}
-\label{Example1}
-} \hspace{0.5em}
-\subfigure[\textit{Will the Nikkei 225 index finish trading at or above 9,500 on 30 September 2011?}]{
-\hspace*{-1em}  \includegraphics[width= 0.49\textwidth]{Figures/ExamplePlot2}
-\label{Example2}
-}
-
-\caption[Optional caption for list of figures]{Scatterplots of the probability forecasts given for two questions in our dataset.
-The shadings represents the self-reported expertise of the expert who provided the probability forecast.}
-\label{ExamplePlots}
-\end{figure}
-
-To illustrate the nature of the data with some concrete examples, Figures \ref{Example1} and \ref{Example2} show scatterplots of the probability forecasts given for (a) \textit{Will the expansion of the European bailout fund be ratified by all 17 Eurozone nations before 1 November 2011?}, and (b) \textit{Will the Nikkei 225 index finish trading at or above 9,500 on 30 September 2011?}.
-The points have been jittered slightly to make overlaps visible.
-The darkness of the points is positively correlated with the self-reported expertise.
-Given that the European bailout fund was ratified before November 1, 2011 and that the Nikkei 225 index finished trading at around 8,700 on September 30, 2011, the general trend of the probability forecasts tends to converge towards the correct answers.
-The individual experts, however, sometimes disagree strongly, with the disagreement persisting even near the closing dates of the questions.
-
-\section{Model}
-\label{model}
-Let $p_{i,t,k} \in (0,1)$ be the probability forecast given by the $i$th expert at time $t$ for the $k$th question, where $i = 1, \dots, I_k$, $t = 1, \dots, T_k$, and $k = 1, \dots, K$.
-Denote the logit-probabilities with
-\begin{eqnarray*}
-Y_{i,t,k} = \logit(p_{i,t,k}) = \log \left( \frac{p_{i,t,k}}{1-p_{i,t,k}}\right) \in \R
-\end{eqnarray*}
-and collect the logit-probability forecasts given for question $k$ at time $t$ into a vector $\boldsymbol{Y}_{t,k} = [Y_{1,t,k}\ Y_{2,t,k}\ \dots\ Y_{I_{k},t,k}]^T$.
-Partition the experts into $J$ groups based on some individual feature, such as self-reported expertise, with each group sharing a common multiplicative bias term $b_{j} \in \R$ for $j = 1, \dots, J$.
-Collect these bias terms into a bias vector $\boldsymbol{b} = [b_{1}\ b_{2}\ \dots\ b_{J}]^T$.
-Let $\boldsymbol{M}_k$ be a $I_k \times J$ matrix denoting  the group-memberships of the experts in question $k$; that is, if the $i$th expert participating in the $k$th question belongs to the $j$th group, then the $i$th row of $\boldsymbol{M}_k$ is the $j$th standard basis vector $\boldsymbol{e}_j$.
-The bias vector $\boldsymbol{b}$ does not include a subindex because it is considered shared among all the $K$ questions.
-To secure model identifiability, it is sufficient to share only one of the elements of $\boldsymbol{b}$ among the questions.
-This element defines a baseline under which it is possible to estimate the remaining $J-1$ bias terms separately within each of the questions.
-In this paper, however, the entire vector $\boldsymbol{b}$ is shared because some of the questions in our real-world data set involve very few experts with the highest level of self-reported expertise.
-Under this notation, the model for the $k$th question can be expressed as
-\begin{eqnarray}
-\boldsymbol{Y}_{t, k} &=&  \boldsymbol{M}_k \boldsymbol{b} X_{t, k} + \boldsymbol{v}_{t, k} \label{observedpr} \\
-X_{t, k} &=& \gamma_k X_{t-1, k} + w_{t, k} \label{hiddenpr}\\
-X_{0,k} &\sim& \mathcal{N}(\mu_0, \sigma^2_0) \nonumber
-\end{eqnarray}
-where Equation (\ref{observedpr}) denotes the observed process and Equation (\ref{hiddenpr}) shows the hidden process that is driven by the constant $\gamma_k \in \R$.
-The error terms are independent and identically distributed normal random variables with mean zero
-\begin{eqnarray*}
-\boldsymbol{v}_{t, k} | \sigma^2_k &\stackrel{i.i.d.}{\sim}& \mathcal{N}_{I_k}(\boldsymbol{0}, \sigma^2_k \boldsymbol{I}_{I_k})\\
-w_{t, k} | \tau^2_k &\stackrel{i.i.d.}{\sim}& \mathcal{N}(0, \tau^2_k),
-\end{eqnarray*}
-and $(\mu_0, \sigma_0^2) \in (\R, \R^+)$ are hyper-parameters chosen \textit{a priori}.
-
-The hidden state $X_{t,k}$ represents the aggregate logit-probability for the $k$th event given all the information available up to and including time $t$.
-To make this more specific, let $Z_k \in \{0, 1\}$ indicate whether the event associated with the $k$th question happened $\left(Z_k = 1\right)$ or did not happen $\left(Z_k = 0\right)$.
-If $\{\mathcal{F}_{t,k}\}_{t=1}^{T_k}$ is a filtration representing the information available up to and including a given time point, then $\E[Z_k | \mathcal{F}_{t,k}] = \P(Z_k = 1| \mathcal{F}_{t,k}) = \logit^{-1}(X_{t,k})$.
-It is reasonable to assume that this probability forecast
-%that makes uses all available information and hence is equivalent to $\logit^{-1}(X_{t,k})$ 
-%an expert, who has access to all available information and reports $\logit^{-1}(X_{t,k})$ as the probability forecast, 
-maximizes sharpness subject to calibration, where calibration and sharpness are understood as follows:
-\begin{mydef}
-A probability forecast $p$ for the $k$th question is calibrated if $\P(Z_k = 1 | p) = \E[Z_k | p] = p$ almost surely (see, e.g., \citet{murphy1987general}).
-\end{mydef}
-\begin{mydef}
-A probability forecast $p$ is sharper than a probability forecast $q$ if $\E[(p-p_0)^2] > \E[(q-p_0)^2]$, where $p_0$ is the baseline probability for the $k$th question (see, e.g., \citet{ranjan2009combining}).
-\end{mydef}
-\noindent
-Even though it is unlikely that any single expert has access to all the available information, a large and diverse group of experts may share a considerable portion of the available information.
-The collective wisdom of the group therefore provides an attractive proxy for $\mathcal{F}_{t,k}$.
-
-Given that the experts may believe in false information, hide their true beliefs, or be biased for many other reasons, their probability assessments should be aggregated via a model that can detect potential bias, separate signal from noise, and use the collective opinion to estimate $X_{t,k}$.
-In our model the experts are assumed to be, on average, a multiplicative constant $\boldsymbol{b}$ away from $X_{t,k}$.
-Therefore an individual element of $\boldsymbol{b}$ can be interpreted as a group-specific \textit{systematic bias} that labels the group either as over-confident $\left( b_j \in (1, \infty)\right)$ or as under-confident $\left(b_j \in (0,1)\right)$.
-Due to the high sparsity of our data, estimating a bias term separately for each expert is not possible.
-See Section \ref{ExpertBias} for an analysis and discussion on the bias terms.
-Any other deviation from $X_{t,k}$ is considered \textit{random noise}.
-This noise is measured in terms of $\sigma^2_k$ and can be assumed to be caused by momentary over-optimism (or pessimism), false beliefs, or other misconceptions.
-
- The \textit{random fluctuations} in the hidden process are measured by $\tau^2_k$ and are assumed to represent changes or shocks to the underlying circumstances that ultimately decide the outcome of the event.
-The \textit{systematic component} $\gamma_k$ allows the model to incorporate a constant signal stream that drifts the hidden process to infinity (when $\gamma_k \in (1, \infty)$) or zero (when $\gamma_k \in (0, 1)$).
-If the uncertainty in the question diminishes as the current time point $t$ approaches $T_k$, the hidden process drifts to infinity.
-Alternatively, the hidden process can drift to zero in which case any available information about the target event does not improve predictive accuracy.
-Given that each of the $K$ questions in our dataset was resolved within a pre-specified timeframe, $\gamma_k$ is expected to fall within the interval $(1, \infty)$ for all $k = 1, \dots, K$.
-
-\section{Model Estimation}
-\label{identifiability}
-
-The main challenge is to capture a well-calibrated estimate of the hidden process without sacrificing the interpretability of our model.
-This section introduces a two-step procedure, called \textit{Sample-And-Calibrate} (SAC), that achieves this goal in a flexible and efficient manner: the first step estimates the model parameters under a constraint (\textit{Sampling Step}), and the second step performs a one-dimension optimizational procedure to transform the constrained estimates into their unconstrained counterparts (\textit{Calibration Step}).
-
-\subsection{Sampling Step}
-\label{sampling_step}
-Given that $\left(a \boldsymbol{b}, X_{t,k}/a, a^2 \tau_k^2\right) \neq \left(\boldsymbol{b}, X_{t,k}, \tau_k^2\right)$ for any $a > 0$ yield the same  likelihood for $\boldsymbol{Y}_{t,k}$, the model as described by Equations (\ref{observedpr}) and (\ref{hiddenpr}) is not identifiable.
-As a result, the parameter estimates tend to drift during the sampling process.
-A well-known solution is to choose one of the elements of $\boldsymbol{b}$, say $b_3$, as the reference point and fix $b_3 = 1$.
-In Section \ref{syntheticData} we provide a guideline for choosing the reference point.
-Denote the constrained version of the model by
-\begin{eqnarray*}
-\boldsymbol{Y}_{t, k} &=&  \boldsymbol{M}_k \boldsymbol{b} (1) X_{t, k}(1)+ \boldsymbol{v}_{t, k}\\
-X_{t, k}(1) &=& \gamma_k(1) X_{t-1, k}(1) + w_{t, k}\\
-\boldsymbol{v}_{t, k} | \sigma^2_k(1) &\stackrel{i.i.d.}{\sim}& \mathcal{N}_{I_k}(\boldsymbol{0}, \sigma^2_k(1) \boldsymbol{I}_{I_k})\\
-w_{t, k} | \tau^2_k(1) &\stackrel{i.i.d.}{\sim}& \mathcal{N}\left(0, \tau^2_k(1)\right),
-\end{eqnarray*}
-where the trailing input parameter emphasizes the constraint $b_{3} = 1$.
-Given that this version is identifiable, estimates of the model parameters can be obtained.
-Denote the estimates by placing a hat on the parameter symbol.
-For instance, $\hat{\boldsymbol{b}}(1)$ and $\hat{X}_{t, k}(1)$ represent the estimates of $\boldsymbol{b}(1)$ and $X_{t, k}(1)$, respectively.
-These estimates are found via Gibbs sampling that only makes use of standard distributions.
-See Appendix \ref{appendix} for the technical details of the sampling steps, and, e.g., \citet{gelman2003bayesian} for a discussion on the general principles of Gibbs sampling.
-
-\subsection{Calibration Step}
-\label{calibration_step}
-Given that the model parameters can be estimated by fixing $b_3$ to any constant, the next step is to search for the constant that gives an optimally sharp and calibrated estimate of the hidden process.
-This section introduces an efficient procedure that finds the optimal constraint without requiring any additional runs of the sampling step.
-First, assume that parameter estimates  $\hat{\boldsymbol{b}}(1)$ and $\hat{X}_{t, k}(1)$ have already been obtained via the constrained sampling step described in Section \ref{sampling_step}.
-Given that for any $\beta \in \R / \{0\}$,
-\begin{eqnarray*}
-\boldsymbol{Y}_{t, k} &=&  \boldsymbol{M}_k \boldsymbol{b} (1) X_{t, k}(1)+ \boldsymbol{v}_{t, k}\\
- &=&   \boldsymbol{M}_k \left(\boldsymbol{b} (1) \beta \right) \left(X_{t, k}(1)/\beta\right)  + \boldsymbol{v}_{t, k}\\
-&=&   \boldsymbol{M}_k \boldsymbol{b} (\beta) X_{t, k}(\beta) + \boldsymbol{v}_{t, k},
-\end{eqnarray*}
-the parameter values under $b_{3} = \beta$ can be obtained from $\boldsymbol{b} (\beta) = \boldsymbol{b} (1) \beta $ and $X_{t, k}(\beta) = X_{t, k}(1)/\beta$. 
-%Furthermore, an estimate of $X_{t, k}(\beta)$ is given by $\hat{X}_{t, k}(1)/\beta$. 
-This means that $X_{t,k} = X_{t,k}(1)/\beta$ when $\beta$ is equal to the true value of $b_3$.
-Given that the hidden process $X_{t, k}$ is assumed to be sharp and well-calibrated, $b_3$ can be estimated with the value of $\beta$ that simultaneously maximizes the sharpness and calibration of $\hat{X}_{t, k}(1) / \beta$.
-A natural criterion for this maximization is given by the class of \textit{proper scoring rules} that combine sharpness and calibration (\citet{gneiting2008rejoinder, buja2005loss}).
-Due to the possibility of \textit{complete separation} in any one question (see, e.g., \citet{gelman2008weakly}), the maximization must be performed over multiple questions.
-Therefore,
-\begin{eqnarray}
- \hat{\beta}&=&  \argmax_{\beta \in \R / \{0\}} \sum_{k=1}^K \sum_{t=1}^{T_k}  S\left(Z_k, \hat{X}_{k,t}(1) / \beta \right) \label{OSE}
-\end{eqnarray}
-where $Z_k \in \{0, 1\}$ indicate whether the event associated with the $k$th question happened $\left(Z_k = 1\right)$ or did not happen $\left(Z_k = 0\right)$.
-The function $S$ is a strictly proper scoring rule such as the negative Brier score (\citet{Brier})
-\begin{eqnarray*}
-S_{BRI}(Z, X) &=& -(Z - \logit^{-1}(X))^2
-\end{eqnarray*}
-or the logarithmic score (\citet{good1952rational})
-\begin{eqnarray*}
-S_{LOG}(Z, X) &=& Z \log\left(\logit^{-1}(X)\right) + (1-Z) \log\left(1-\logit^{-1}(X)\right)
-\end{eqnarray*}
-Given that it is not clear which rule should be used for predicting geopolitical events, the SAC-procedure is evaluated  separately under both rules in Sections \ref{syntheticData} and \ref{realData}.
-Once $\hat{\beta}$ has been computed, estimates of the unconstrained model parameters are given by
-\begin{eqnarray}
- \hat{X}_{t,k}&=& \hat{X}_{k,t}(1) / \hat{\beta} \nonumber\\
- \hat{\boldsymbol{b}}&=& \hat{\boldsymbol{b}}(1) \hat{\beta} \nonumber\\
-   \hat{\tau}_{k}^2&=& \hat{\tau}_{k}^2(1)/ \hat{\beta}^2\nonumber\\
-  \hat{\sigma}_{k}^2&=& \hat{\sigma}_{k}^2(1)\nonumber\\
-  \hat{\gamma}_{k}&=& \hat{\gamma}_{k}(1)\nonumber
-\end{eqnarray}
-Notice that estimates of $\sigma^2_k$ and $\gamma_k$ are not affected by the constraint.
-
-\subsection{Discussion}
-If the class labels in the data are balanced with respect to the time points, the calibration step under the logarithmic scoring rule is approximately equivalent to \textit{Platt calibration}, which has been shown to yield good calibration under various modeling scenarios (see, e.g., \citet{platt1999probabilistic, niculescu2005obtaining}).
-To see this, recall that the Platt calibrated logit-probabilities are given by $\hat{A} + \hat{B}\hat{X}_{t,k}(1)$, where
-\begin{eqnarray}
-\left(\hat{A}, \hat{B} \right) =  \argmax_{A, B \in \R} \sum_{k=1}^K \sum_{t=1}^{T_k}  S_{LOG}(Z_k, A + B\hat{X}_{t,k}(1) ) \label{platt}
-\end{eqnarray}
-This is equivalent to fitting a logistic regression model with $Z_k$ as the response and $\hat{X}_{t,k}(1)$ as the explanatory variable.
-%model with $Z_k$ as the response and $\hat{X}_{t,k}(1)$ as the explanatory variable. 
-To understand the behavior of the coefficients $A$ and $B$, express the logistic regression as a linear regression model
-\begin{eqnarray*}
-\logit(\P(Z_k = 1 | \hat{X}_{t,k})) = A + B\hat{X}_{t,k} + e_{t,k}
-\end{eqnarray*}
-with $e_{t,k} \stackrel{i.i.d.}{\sim} \mathcal{N}(0,\sigma^2)$.
-If the data are balanced with respect to the time points, then exactly half of the summands in Equation (\ref{platt}) have $Z_k = 1$ and the average response logit-probability $\logit(\P(Z_k = 1 | \hat{X}_{t,k})) $ is expected to be close to zero.
-Given that the values of $\hat{X}_{t,k}$ are estimated logit-probabilities of the same $K$ events across different time points, their overall average is also expected to be around zero.
-Therefore both the response and explanatory variables are approximately centered.
-This means that the intercept term $A$ is near zero reducing Platt calibration to Equation (\ref{OSE}) under the logarithmic scoring rule.
-If the data are not balanced, Platt calibration can be easily incorporated into our model via an additional intercept parameter.
-This, however, reduces the interpretability of our model.
-Fortunately, compromising interpretability is rarely necessary because it is often possible to use the data in a well-balanced form.
-One procedure to attain this is described in the beginning of Section \ref{realData}.
-
-If the future event can take upon $M > 2$ possible outcomes, the hidden state $X_{t,k}$ is extended to a vector of size $M-1$ and one of the outcomes, e.g., the $M$th one, is chosen as the base-case to ensure that the probabilities will sum to one at any given time point.
-Each of the remaining  $M-1$ possible outcomes is represented by an observed process similar to Equation (\ref{observedpr}).
-Given that this multinomial extension is equivalent to having $M-1$ independent binary-outcome models, the estimation and properties of the model are easily extended to the multi-outcome case.
-This paper focuses on binary-outcomes because it is the simplest and most commonly encountered setting in practice.
-
-%
-%\section{Extension: Polychotomous Outcomes}
-%If the future event can take upon $M > 2$ possible outcomes, the hidden state $X_{t,k}$ must be extended to a vector of size $M-1$.
-One of the outcomes, e.g., the $M$th one, is chosen as the base-case.
-This gives us a total of $M-1$ observed processes and one hidden process
-%\begin{eqnarray*}
-%\boldsymbol{Y}_{1,t,k} &=&  \boldsymbol{M}_k\boldsymbol{b}_{1}X_{1,t,k} + \boldsymbol{v}_{1,t,k}\\
-%\boldsymbol{Y}_{2,t,k} &=&  \boldsymbol{M}_k\boldsymbol{b}_{2}X_{2,t,k} + \boldsymbol{v}_{2,t,k}\\
-%&\vdots&\\
-%\boldsymbol{Y}_{M-1,t,k} &=&  \boldsymbol{M}_k\boldsymbol{b}_{M-1}X_{M-1,t,k} + \boldsymbol{v}_{M-1,t, k}\\
-%\boldsymbol{X}_{t,k} &=& \boldsymbol{\gamma}^T_k \boldsymbol{I}_{M-1} \boldsymbol{X}_{t-1,k} + \boldsymbol{w}_{t,k}
-%\end{eqnarray*}
-%where $\boldsymbol{X}_{t,k}$ is a $(M-1) \times 1$ matrix of calibrated logit-probabilities.
-Notice that each process has a separate bias vector.
-The error terms are independent and identically distributed multivariate normal random variables
-%\begin{eqnarray*}
-%\boldsymbol{v}_{m, t, k} | \sigma^2_{j,k} &\stackrel{i.i.d.}{\sim}& \mathcal{N}_{I_k} \left(\boldsymbol{0}, \sigma^2_{j,k} \boldsymbol{I}_{I_k} \right)\\
-%\boldsymbol{w}_{t,k} | \tau_k^2 &\stackrel{i.i.d.}{\sim}& \mathcal{N}_{M-1} \left(\boldsymbol{0}, \tau_k^2 \boldsymbol{I}_{M-1} \right)
-%\end{eqnarray*}
-%On the left-hand side of observed process
-%\begin{eqnarray*}
-%\boldsymbol{Y}_{m,t,k} &=&  \left[ \begin{matrix} Y_{m,1,t,k} & Y_{m,2,t,k} & \dots & Y_{m,I_k,t,k} \end{matrix} \right]^T\\
-%&=&  \left[ \begin{matrix} \log \left( \frac{p_{m,1,t,k}}{p_{M,1,t,k}} \right) &  \log \left( \frac{p_{m,2,t,k}}{p_{M,2,t,k}} \right) & \dots &  \log \left( \frac{p_{m,I_k,t}}{p_{M,I_k,t,k}} \right) \end{matrix} \right]^T
-%\end{eqnarray*}
-%is a $I_k \times 1$ matrix of the expert logit-probabilities for the $m$th outcome of the $k$th question at time $t$, and on the left-hand side of the hidden process
-%\begin{eqnarray*}
-%\boldsymbol{X}_{t,k} &=&  \left[ \begin{matrix}X_{1,t,k} &X_{2,t,k} & \dots &X_{M-1,t,k} \end{matrix} \right]^T
-%\end{eqnarray*}
-%is the matrix of sharp and calibrated logit-probabilities at time $t$.
-Choosing one of the outcomes as the base-case, ensures that the probabilities will sum to one at any given time point.
-Since this multinomial extension is equivalent to having $M-1$ independent binary-outcome models (see Section \ref{model}), the estimation can be done separately for each outcome.
-Therefore, even though this paper focuses on modeling binary events, all the properties and discussion generalize to the multi-outcome case.
-
-\section{Synthetic Data Results}
-\label{syntheticData}
-
-The goal in this section is to evaluate the ability of the \textit{Sample-And-Calibrate}-procedure to capture true parameter values.
-The synthetic data are not generated directly from the model for two reasons: (i) showing good performance on a dataset directly generated from the model assumptions is hardly news, and (ii) it is unclear how the dynamic part of the model can be used to produce well-calibrated hidden states.
-The latter is important for our interpretation of the hidden process as a sharp and well-calibrated version of the crowd belief.
-
-To ensure that the hidden process for question $k$ is well-calibrated, we generate a path of standard Brownian motion until time $T_k$. If $Z_{t,k}$ denotes the value of the path at time $t$, then
-\begin{eqnarray*}
-Z_k &=& \mathbbm{1}\left( Z_{T_k, k}  > 0\right)\\
- X_{t,k} &=& \logit \left[ \Phi \left(\frac{Z_{t, k}}{\sqrt{T_k- t}} \right) \right]
-\end{eqnarray*}
-gives a sequence of $T_k$ calibrated logit-probabilities for the event $Z_k = 1$.
-Using this procedure we generate the hidden process for $K$ questions, each with a time horizon of $T_k = 101$.
-The questions involve 50 experts allocated evenly among five expertise groups.
-Each expert gives one probability forecast per day with the exception of time $t = 101$ when the event resolves.
-These are generated by applying bias and noise to the hidden process $X_{t,k}$ for $t = 1, \dots, 100$ (see Equation (\ref{observedpr})).
-Our simulation study considers a three-dimensional grid of parameter values:
-\begin{eqnarray*}
-\sigma^2 &\in& \{1/2, 1, 3/2, 2, 5/2\}\\
-\beta &\in& \{1/2, 3/4, 1, 4/3, 2/1\}\\
-K &\in& \{20, 40, 60, 80, 100\},
-\end{eqnarray*}
-where $\beta$ gives the bias vector by letting $\boldsymbol{b} = [1/2, 3/4, 1, 4/3, 2/1]^T \beta$.
-Forty synthetic datasets are generated for each combination of $\sigma^2$, $\beta$, and $K$ values.
-The SAC-procedure runs for 200 iterations of which the first 100 are used for burn-in.
-The accuracy of the estimated hidden process and bias vector are measured with the average quadratic loss in the probability space, $\sum_{k=1}^K \sum_{t=1}^{100} ( \logit^{-1}(\hat{X}_{t,k}) - \logit^{-1}(X_{t,k}))^2 / (100K)$, and the average quadratic loss, $|| \boldsymbol{b} - \hat{\boldsymbol{b}}||^2/5$, respectively.
-
-%$|| \log(\boldsymbol{b}) - \log(\hat{\boldsymbol{b}})||^2/5$, respectively.
-The latter is measured in the log-space to allow a fair comparison of any discrepancies between under- and over-confident bias-vectors.
-
-\begin{figure}[h!]
-\vspace{-5.5em}
-\centering
- \hspace{-25em}	\includegraphics{Figures/LegendSynthetic} % requires the graphicx package
-
-\vspace{0em}
-\hspace{2.5em}Hidden Process
-\vspace{-0.5em}
-
-\subfigure{
-\hspace{-1em} \includegraphics[width = 0.32\textwidth]{Figures/SyntheticSigma2}
-\label{SyntheticSigma2}
-}
-\subfigure{
-\hspace{-1.5em} \includegraphics[width = 0.32\textwidth]{Figures/SyntheticBeta}
-\label{SyntheticBeta}
-}
-\subfigure{
-\hspace{-1.5em} \includegraphics[width = 0.32\textwidth]{Figures/SyntheticK}
-\label{SyntheticK}
-}
-
-\vspace{0.5em}
-\hspace{2.5em}Bias Vector
-\vspace{-0.5em}
-
-\subfigure{
-\hspace{-1em}  \includegraphics[width = 0.32\textwidth]{Figures/SyntheticSigma2bias}
-\label{SyntheticSigma2b}
-}
-\subfigure{
-\hspace{-1.5em}  \includegraphics[width = 0.32\textwidth]{Figures/SyntheticBetabias}
-\label{SyntheticBetab}
-}
-\subfigure{
-\hspace{-1.5em}   \includegraphics[width = 0.32\textwidth]{Figures/SyntheticKbias}
-\label{SyntheticKb}
-}
-
-\caption[Optional caption for list of figures]{Comparing SAC that optimizes over the logarithmic score ($\text{SAC}_{\text{LOG}}$), SAC that optimizes over the Brier score ($\text{SAC}_{\text{BRI}}$), and the Exponentially Weighted Moving Average (EWMA) under synthetic data.
-The top row presents the accuracy to capture the hidden process.
-The bottom row presents the accuracy to capture the bias vector.
-}
-\label{Synthetic}
-\end{figure}
-
-Figure \ref{Synthetic} summarizes the results in a series of plots.
-The first row presents the marginal effects of the three grid variables on the accuracy to which SAC and the Exponentially Weighted Moving Average (EWMA) (see Section \ref{forecasting} for more details on EWMA) estimate the hidden process.
-EWMA serves as a baseline for comparisons.
-Unfortunately, EWMA does not produce an estimate of the bias vector.
-Therefore it cannot be used as a baseline in the second row that measures the accuracy in estimating the bias vector.
-Given that the plots were computed by choosing one grid variable (e.g., $\beta$) at a time and averaging over the remaining two variables  (e.g., $K$ and $\sigma^2$), each value in the plots represents an average of a $40 \times 5^2 = 1,000$ values.
-
-Both $\text{SAC}_{\text{LOG}}$ and  $\text{SAC}_{\text{BRI}}$ estimate the hidden process and the bias vector very accurately.
-Apart from $\text{SAC}_{\text{LOG}}$ outperforming $\text{SAC}_{\text{BRI}}$ uniformly across all parameter values, the two approaches behave very similarly.
-They suffer  from the increasing level of noise in the expert logit-probabilities, estimate the bias vector less accurately when $\beta$ is large, and, unlike EWMA, make greater use of the increasing number of questions.
-EWMA is able to outperform SAC only when the bias among the experts is very small.
-
-Interestingly, SAC estimates the hidden process more accurately when the experts are over-confident (large $\beta$) rather under-confident (small $\beta$).
-To understand this result, assume that the experts in the third group are highly under-confident.
-Therefore their logit-probabilities are expected to be closer to zero than the corresponding hidden states.
-After adding white noise to these expected logit-probabilities, they are likely to cross to the other side of zero.
-If the sampling step fixes $b_3 = 1$, as it does in our case, the third group is treated unbiased and some of the constrained estimates of the hidden states are likely to be on the other side of zero as well.
-Unfortunately, this discrepancy cannot be corrected by the calibration step that is restricted to shifting the constrained estimates either closer or further away from zero but not across it.
-To maximize the likelihood of having all the constrained estimates on the right side of zero and hence avoiding the discrepancy, the reference point in the sampling step should be chosen with care.
-A helpful guideline is to fix the element of $\boldsymbol{b}$ that is apriori believed  to be the largest.
-
-%The latter can be seen as a side-effect of measuring the accuracy an estimated multiplicative factor in the absolute scale. The effect is actually
-
-%
-%Since the hidden state is not explicitly engineered to maximize sharpness subject to calibration, the calibration step in our procedure tends to overestimate $\beta$; that is, the estimated hidden process tends to be a bit sharper (more extreme) than the actual hidden process.
-%
-%When $\beta$ is large most of the experts are over-confident.
-As a result, the sampling step produces an estimated hidden process that is too extreme.
-This estimate therefore must be pulled back towards zero by the calibration step.
-If the
-%
-%
-%The story for the estimating the bias vector is a bit different: the discrepancy between the estimate and the true bias vector may be the same in the log-scale when the $\beta$ is small versus when $\beta$ is large.
-This, however, translates into much smaller differences in the absolute scale that the differences are ultimately measured in our case.
-For instance, estimating 1/2 with 3/4 is the same as estimating 2 with 4/3.
-The difference in the first one is 0.25 while in the second 0.6666667.
-To make this even worse, the distance is squared which emphasizes large differences.
-
-%
-%Based on the first row of plots, $\text{SAC}_{\text{LOG}}$ and $\text{SAC}_{\text{BRI}}$ estimate the hidden process to a similar accuracy.
-Even though some relative differences are visible, the loss values are very small in absolute terms, hence suggesting good overall accuracy to capture the hidden process.
-Estimating the hidden process, however, becomes more difficult as either $\sigma^2$ or $\beta$ gets larger.
-The latter seems more unintuitive and can be understood better by observing that larger values of $\beta$ lead to more extreme expert logit-probabilities and hence to more frequent truncation.
-This causes the estimation accuracy to degrade because  a high level of truncation results to a drastic deviation from the model assumptions.
-Increasing the number of questions improves the estimation accuracy but only very little.
-%
-%
-%Based on the second row of plots, $\text{SAC}_{\text{LOG}}$ and $\text{SAC}_{\text{BRI}}$  capture the bias vector rather precisely as long as the true values of $\beta$ and $\sigma^2$ are not very large.
-This behavior can be again considered as being caused by the truncation of the extreme expert logit-probabilities.
-Surprisingly, even though $\text{SAC}_{\text{LOG}}$ is more accurate in estimating the bias vector, it is unable to outperform $\text{SAC}_{\text{BRI}}$ in estimating the hidden process.
-%
-%
-%
-
-\section{Geopolitical Data Results}
-\label{realData}
-\noindent
-This section presents results for the real-world data described in Section \ref{data}.
-The goal is to provide application specific insight by discussing the specific research objectives itemized in Section \ref{intro} and also to evaluate SAC in terms of predictive power and calibration.
-Before presenting the results, however, we discuss two practical matters that must be taken into account when aggregating real-world  probability forecasts.
-
-\subsection{Incoherent and Imbalanced Data}
-\label{practicalMatters}
-The first matter regards human experts making probability forecasts of 0.0 or 1.0 even if they are not completely sure of the outcome of the event.
-For instance, all the 166 questions in our dataset contained both a zero and a one.
-Transforming such forecasts into the logit-space yields infinities that can cause problems in model estimation.
-To avoid this, \citet{Ariely00theeffects} suggest changing $p = 0.00$ and $1.00$ to $p = 0.02$ and $0.98$, respectively.
-This is similar to \textit{winsorising} that sets the extreme probabilities to a specified percentile of the data (see, e.g., \citet{hastings1947low} for more details on winsorising).
-\citet{Geo}, on the other hand, consider only probabilities that fall within a constrained interval, say $[0.001, 0.999]$, and discard the rest.
-Given that this implies ignoring a portion of the data, we adopt an approach similar to \citet{Ariely00theeffects}  by truncating $p = 0.00$ and $1.00$ to $p = 0.01$ and $0.99$, respectively.
-Our results remain insensitive to the exact choice of truncation as long as this is done in a reasonable manner to keep the extreme probabilities from becoming highly influential in the logit-space.
-
-The second matter is related to the distribution of the class labels in the data.
-If the set of occurrences is much larger than the set of non-occurrences (or \textit{vice versa}), the dataset is called \textit{imbalanced}.
-On such data the modeling procedure can end up over-focusing on the larger class, and as a result, give very accurate forecast performance over the larger class at the cost of performing poorly over the smaller class (see, e.g., \citet{chen2009learning, wallace2012class}).
-Fortunately, it is often possible to use a well-balanced version of the data.
-The first step is to find a partition $S_0$ and $S_1$ of the question indices $\{1, 2, \dots, K\}$ such that the equality $\sum_{k \in S_0} T_k  = \sum_{k \in S_1} T_k$ is as closely approximated as possible.
-This is equivalent to an NP-hard problem known in computer science as the \textit{Partition Problem}:  determine whether a given set of positive integers can be partitioned into two sets such that the sums of the two sets equal to each other (see, e.g., \citet{karmarkar1982differencing, hayes2002easiest}).
-A simple solution is to use a greedy algorithm that iterates through the values of $T_k$ in descending order, assigning each $T_k$ to the subset that currently has the smaller sum (see, e.g.
-\citet{kellerer2004knapsack, gent1996phase} for more details on the \textit{Partition Problem}).
-After finding a well-balanced partition, the next step is to assign the class labels such that the labels for the questions in $S_x$ are equal to $x$ for $x = 0$ or $1$.
-Recall from section \ref{calibration_step} that $Z_k$ represents the event indicator associated with the $k$th question.
-To define a balanced set of indicators $ \tilde{Z}_k$ for all $k \in S_x$, let
-\begin{align*}
- \tilde{Z}_k &= x\\
-\tilde{p}_{i,t,k} &=  \begin{cases} 
-1-p_{i,t,k}, & \text{if } Z_k = 1-x,\\
-p_{i,t,k}, & \text{if } Z_k = x,
-\end{cases}
-\end{align*}
-where $i = 1, \dots, I_k$, and $t = 1, \dots, T_k$.
-The resulting set $$\left\{\left(\tilde{Z}_k, \{\tilde{p}_{i,t,k} | i = 1, \dots, I_k, t = 1, \dots, T_k \}\right) \right\}_{k=1}^K$$ is a balanced version of the data.
-This procedure was used to balance our real-world dataset both in terms of events and time points.
-The final output splits the events exactly in half ($|S_0| = |S_1| = 83$) such that number of time points in the first and second halves are 8,737 and 8,738, respectively.
-
-\subsection{Out-of-Sample Forecasting}
-\label{forecasting}
-This section is motivated by decision-making. 
-%For instance, recall our example with an investor who consults a group of experts on a daily basis in order to assess whether a stock index will end up trading above some threshold on a given date. 
-The goal is to evaluate the out-of-sample predictive performance of SAC against several other probability aggregation procedures.
-The models are allowed to utilize a training set before making predictions on an independent testing set.
-To clarify some of the upcoming notation, let $S_{train}$ and $S_{test}$ be index sets that partition the data into training and testing sets of sizes $|S_{train}| = N_{train}$ and $|S_{test}| = 166 - N_{test}$, respectively.
-This means that the $k$th question is in the training set if and only if $k \in S_{train}$.
-The competing models are as follows:
-
-\begin{enumerate}
-\item \textit{Simple Dynamic Linear Model (SDLM)}. This is equivalent to the dynamic model from Section \ref{model} but with $\boldsymbol{b} = \boldsymbol{1}$ and $\beta = 1$. Thus,
-\begin{eqnarray*}
-\boldsymbol{Y}_{t, k} &=&  X_{t, k} + \boldsymbol{v}_{t, k}  \\
-X_{t, k} &=& \gamma_k X_{t-1, k} + w_{t, k},
-\end{eqnarray*}
-where $X_{t,k}$ is the logit-probability used for prediction.
-Given that this model is not hierarchical, estimates of the hidden process can be obtained directly for the questions in the testing set without fitting the model first on the training set.
-To make predictions, the sampler is run for 500 iterations of which the first 200 are used for burn-in.
-The remaining 300 iterations are thinned by discarding every other observation, leaving a final predictive sample of 150 observations.
-
-\item \textit{The Sample-And-Calibrate procedure both under the Brier ($\text{SAC}_{\text{BRI}}$) and the logarithmic score ($\text{SAC}_{\text{LOG}}$)}.
-The model is first fit on the training set by running the sampling step for 3,000 iterations of which the first 500 iterations are used for burn-in.
-The remaining 2,500 observations are thinned by keeping every fifth observation.
-The calibration step is performed for the final 500 observations.
-The out-of-sample prediction is done by running the sampling step for 500 iterations with each consecutive iteration reading in and conditioning on the next value of $\beta$ and $\boldsymbol{b}$ found during the training period.
-The first 200 iterations are used for burn-in.
-The remaining 300 iterations are thinned by discarding every other observation, leaving a final predictive sample of 150 observations.
-
-\item \textit{A fully Bayesian version of $\text{SAC}_{\text{LOG}}$ ($\text{BSAC}_{\text{LOG}}$)}.
-Denote the calibrated logit-probabilities and  the event indicators across all $K$ questions with $\boldsymbol{X}(1)$ and $\boldsymbol{Z}$, respectively.
-The posterior distribution of $\beta$ conditional on $\boldsymbol{X}(1)$ is given by $p(\beta | \boldsymbol{X}(1), \boldsymbol{Z}) \propto p( \boldsymbol{Z} | \beta, \boldsymbol{X}(1)) p(\beta | \boldsymbol{X}(1))$.
-Recall that the calibration step under $S_{LOG}$ is equivalent to fitting a logistic regression model with $Z_k$ as the response and $\hat{X}_{t,k}(1)$ as the explanatory variable.
-Therefore the likelihood for the Bayesian version is
-\begin{eqnarray}
-p( \boldsymbol{Z} | \beta, \boldsymbol{X}(1)) \propto \prod_{k=1}^K \prod_{t=1}^{T_k} && \logit^{-1} \left(X_{t,k}(1)/\beta  \right)^{Z_k}\times \label{OSE2}\\ 
-&& \left( 1-  \logit^{-1} \left( X_{t,k}(1)/\beta  \right) \right)^{1-Z_k} \nonumber
-\end{eqnarray}
-As in \citet{gelman2003bayesian}, the prior is chosen to be locally uniform, $p(1/\beta) \propto 1$.
-Posterior estimates of $\beta$ can be sampled from Equation (\ref{OSE2}) using generic sampling algorithms such as the Metropolis algorithm (\citet{metropolis1953equation}) or slice sampling (\citet{neal2003slice}).
-Given that the sampling procedure conditions on the event indicators, the full conditional distribution of the hidden states is not in a standard form.
-Therefore the Metropolis algorithm is also used for sampling the hidden states.
-Predictions are made with the same choices of thinning and burn-in as described under \textit{Sample-And-Calibrate}.
-
-\item Due to the lack of previous literature on dynamic aggregation of expert probability forecasts,  the main competitors are exponentially weighted versions of procedures that have been proposed for static probability aggregation:
-
-\begin{enumerate}
-
-\item \textit{Exponentially Weighted Moving Average (EWMA)}. If 
-\begin{eqnarray*}
-\bar{p}_{t,k}  = \frac{1}{N_{t,k}} \sum_{i=1}^{N_{t,k}} p_{i,t,k},
-\end{eqnarray*}
-is the average probability forecast given at time $t$ for the $k$th question, then the EWMA forecasts for the $k$th problem are obtained recursively from
-\begin{eqnarray*}
-\hat{p}_{t,k}(\alpha) =
-\begin{cases}
- \bar{p}_{1,k}, & \text{ for } t  = 1, \\
-\alpha \bar{p}_{t,k}  + (1-\alpha) \hat{p}_{t-1,k}(\alpha), & \text{ for } t > 1,
-\end{cases}
-\end{eqnarray*}
-where the input parameter $\alpha$ is learned from the training set by 
-\begin{eqnarray*}
-\hat{\alpha} &=& \argmin_{\alpha \in [0,1]} \sum_{k \in S_{train}} \sum_{t=1}^{T_k} \left(Z_k - \hat{p}_{t,k}(\alpha) \right)^2
-\end{eqnarray*}
-
-\item  \textit{Exponentially Weighted Moving Logit Aggregator (EWMLA)}.
-This is a moving version of the aggregator $\hat{p}_G(\boldsymbol{b})$ that was introduced in \cite{satopaa}.
-If $\boldsymbol{p}_{t,k}$ is a vector collecting all the probability forecasts made for the $k$th question at time $t$, then the EWMLA forecasts are found recursively from
-\begin{eqnarray*}
-\hat{p}_{t,k}(\alpha, \boldsymbol{b}) =
-\begin{cases}
-G_{1,k}(\boldsymbol{b} ), & \text{ for } t  = 1, \\
-\alpha G_{t,k}(\boldsymbol{b} )  + (1-\alpha) \hat{p}_{t-1,k}(\alpha, \boldsymbol{b}),  & \text{ for } t > 1,
-\end{cases}
-\end{eqnarray*}
-where
-\small
-\begin{eqnarray*}
-G_{t,k}(\nu )  &=& \left( \prod\limits_{i=1}^{N_{t,k}} \left( \frac{p_{i, t, k}}{1-p_{i, t, k}} \right)^{ \frac{\boldsymbol{e}_{i,k}'\boldsymbol{b}}{N_{t,k}}} \right) \Bigg/ \left(1+ \prod\limits_{i=1}^{N_{t,k}} \left( \frac{p_{i, t, k}}{1-p_{i, t, k}} \right)^{\frac{\boldsymbol{e}_{i,k}' \boldsymbol{b}}{N_{t,k}}} \right)
-\end{eqnarray*}
-\normalsize
-The vector $\boldsymbol{b}$ collects the bias terms of the different expertise groups.
-Therefore it serves a similar role to the bias vector found under \textit{Sample-And-Calibrate}.
-The term $\boldsymbol{e}_{i,k}$ is a vector of length 5 indicating which level of self-reported expertise the $i$th expert in the $k$th question belongs to.
-For instance, if $\boldsymbol{e}_{i,k} = [0, 1, 0, 0, 0]$, then the expert identifies himself with the expertise level two.
-The tuning parameters $(\alpha, \boldsymbol{b})$ are learned from the training set by
-\begin{eqnarray*}
-(\hat{\alpha}, \hat{ \boldsymbol{b}}) &=& \argmin_{ \boldsymbol{b} \in \R^5, \alpha \in [0,1]} \sum_{k \in S_{train}} \sum_{t=1}^{T_k} \left(Z_k - \hat{p}_{t,k}(\alpha,  \boldsymbol{b})\right)^2
-\end{eqnarray*}
-
-\item  \textit{Exponentially Weighted Moving Beta-transformed Aggregator (EWMBA)}.
-The static version of the Beta-transformed  aggregator was introduced in \cite{Ranjan08}.
-A dynamic version can be obtained by replacing $G_{t,k}(\nu )$ in the EWMLA description with
-\begin{eqnarray*}
-H_{\nu, \tau} \left( \bar{p}_{t,k}\right),
-\end{eqnarray*}
-where $H_{\nu, \tau}$ is the cumulative distribution function of the Beta distribution and $\bar{p}_{t,k}$ is the average probability forecast defined under EWMA.
-The tuning parameters $(\alpha, \nu, \tau)$ are learned from the training set by
-\begin{eqnarray*}
-(\hat{\alpha}, \hat{\nu}, \hat{\tau}) &=& \argmin_{\nu, \tau > 0\ \alpha \in [0,1]} \sum_{k \in S_{train}} \sum_{t=1}^{T_k} \left(Z_k - \hat{p}_{t,k}(\alpha, \nu, \tau)\right)^2
-\end{eqnarray*}
-
-\end{enumerate}
-\end{enumerate}
-
-The competing models are evaluated via a 10-fold cross-validation\footnote{A 5-fold cross-validation was also performed.
-The results were, however, very similar to the 10-fold cross-validation and hence not presented in the paper.} that first partitions the 166 questions into 10 sets.
-The partition is chosen such that each of the 10 sets has approximately the same number of questions (16 or 17 questions per set in our case) and the same number of time points (between 1760 and 1764 time points per set in our case).
-The evaluation then iterates 10 times, each time using one of the 10 sets as the testing set and the remaining 9 sets as the training set.
-Therefore each question is used nine times for training and exactly once for testing.
-The testing proceeds sequentially one testing question at a time as follows: First, for a question with a time horizon of $T_k$, make a prediction based on the first two days.
-Compute the Brier score for the aggregate forecast of the second day.
-Next make a prediction based on the first three days and  compute the Brier score for the most recent day, namely, the third day.
-Repeat this process until the prediction is made on all of the $T_k-1$ days.
-This leads to $T_k-1$ Brier scores per testing question and a total of  17,475 Brier scores across the entire dataset.
-
-Table \ref{prediction} summarizes different ways to aggregate these scores.
-The first option, denoted by \textit{Scores by Day}, weighs each question by the number of days the question remained open.
-This is performed by computing the average of the 17,475 scores.
-The second option, denoted by \textit{Scores by Problem}, gives each question an equal weight regardless how long the question remained open.
-This is done by first averaging the scores within a question and then averaging the average scores across all the questions.
-Both scores can be further broken down into subcategories by considering the length of the questions.
-The final three columns of Table \ref{prediction} divide the questions into \textit{Short} questions (30 days or fewer), \textit{Medium} questions (between 31 and 59 days), and \textit{Long} Problems (60 days or more).
-The number of questions in these subcategories were 36, 32 and 98, respectively.
-The bolded scores indicate the best score in each column.
-The values in the parenthesis quantify the variability in the scores: Under \textit{Scores by Day} the values give the standard errors of all the scores.
-Under \textit{Scores by Problem}, on other hand, the values represent the standard errors of the average scores of the different questions.
-
-\begin{table*}
-   \caption{Brier Scores based on 10-fold cross-validation.
-\textit{Scores by Day} weighs a question by the number of days the question remained open.
-\textit{Scores by Problem} gives each question an equal weight regardless how long the question remained open.
-The bolded values indicate the best scores in each column.
-The values in the parenthesis represent standard errors in the scores.
-}
-   \label{prediction}
-   \centering
-      \begin{tabular}{l llll} % Column formatting, @{} suppresses leading/trailing space
-      \hline
- \multicolumn{5}{ c }{Scores by Day}\\
-Model & \multicolumn{1}{ c }{All} & \multicolumn{1}{ c }{Short} & \multicolumn{1}{ c }{Medium} & \multicolumn{1}{ c }{Long}\\ \hline
-SDLM & 0.100 (0.156) & 0.066 (0.116) & 0.098 (0.154) & 0.102 (0.157)\\ 
-$\text{BSAC}_{\text{LOG}}$ & 0.097 (0.213) & \textbf{0.053} (0.147) & 0.100 (0.215) & 0.098 (0.215)\\ 
-$\text{SAC}_{\text{BRI}}$ & 0.096 (0.190) & 0.056 (0.134) & 0.097 (0.190) & 0.098 (0.192)\\ 
-$\text{SAC}_{\text{LOG}}$ & \textbf{0.096} (0.191) & 0.056 (0.134) & \textbf{0.096} (0.189) & \textbf{0.098} (0.193)\\ 
-EWMBA & 0.102 (0.203) & 0.060 (0.124) & 0.110 (0.201) & 0.103 (0.206)\\ 
-EWMLA & 0.102 (0.199) & 0.061 (0.130) & 0.111 (0.214) & 0.103 (0.200)\\ 
-EWMA & 0.111 (0.142) & 0.089 (0.100) & 0.111 (0.136) & 0.112 (0.144)\\ 
-\hline
-\multicolumn{5}{ c }{Scores by Problem}\\
-Model & \multicolumn{1}{ c }{All} & \multicolumn{1}{ c }{Short} & \multicolumn{1}{ c }{Medium} & \multicolumn{1}{ c }{Long}\\ \hline
-SDLM & 0.089 (0.116) & 0.064 (0.085) & 0.106 (0.141) & 0.092 (0.117)\\ 
-$\text{BSAC}_{\text{LOG}}$ & 0.083 (0.160) & \textbf{0.052} (0.103) & 0.110 (0.198) & 0.085 (0.162)\\ 
-$\text{SAC}_{\text{BRI}}$& 0.083 (0.142) & 0.055 (0.096) & 0.106 (0.174) & 0.085 (0.144)\\ 
-$\text{SAC}_{\text{LOG}}$ & \textbf{0.082} (0.142) & 0.055 (0.096) & \textbf{0.105} (0.174) & \textbf{0.085} (0.144)\\ 
-EWMBA & 0.090 (0.156) & 0.063 (0.101) & 0.118 (0.186) & 0.091 (0.161)\\ 
-EWMLA & 0.090 (0.159) & 0.064 (0.109) & 0.120 (0.200) & 0.090 (0.159)\\ 
-EWMA & 0.104 (0.105) & 0.092 (0.081) & 0.119 (0.125) & 0.103 (0.107)\\ 
-\hline
-   \end{tabular}
-\end{table*}
-
-Overall, $\text{SAC}_{\text{BRI}}$ and $\text{SAC}_{\text{LOG}}$ achieve the lowest average scores across all columns except \textit{Short}, where they are slightly outperformed by $\text{BSAC}_{\text{LOG}}$.
-However, $\text{BSAC}_{\text{LOG}}$ turns out to be overconfident (see Section \ref{calibration}).
-This means that $\text{BSAC}_{\text{LOG}}$ underestimates the uncertainty in the events and outputs probability forecasts that are typically too near 0.0 or 1.0.
-As a result, the aggregate forecasts are either very close to the correct answer or very far from it.
-As Table \ref{prediction} shows, this results into highly variable forecasting performance.
-The short questions, however, involve very little uncertainty and were generally the easiest to forecast.
-On such easy questions, overconfidence can pay off frequently enough to compensate for a few large losses arising from the overconfident and incorrect forecasts.
-
-In contrast to $\text{BSAC}_{\text{LOG}}$, the baseline SDLM-model lacks sharpness and is highly under-confident (see Section \ref{calibration}).
-This behavior is expected as the experts are under-confident at the group-level (see Section \ref{ExpertBias}) and the SDLM-procedure does not use the training set to explicitly calibrate its forecasts.
-Instead, it merely smooths the forecasts given by the experts.
-The resulting aggregate forecasts are therefore necessarily conservative, resulting into high average scores with low variability.
-
- Similar behavior is exhibited by EWMA that performs the worst of all the competing models.
-In contrast, the other two exponentially weighted aggregators, EWMLA and EWMBA, make efficient use of the training set and present moderate forecasting performance in most columns of Table \ref{prediction}.
-Recall that EWMBA uses the cumulative distribution function of the Beta distribution that depends on two parameters and is more flexible than the transformation used by EWMLA.
-On other hand, only EWLMA is given access to the self-reported expertise information.
-Neither approach, however, appears to dominate the other.
-The aggregators perform very similarly.
-The high variability and average of their performance scores indicate that their performance suffers from over-confidence.
-
-%The dataset contained a few long questions with an abrupt change of the crowd belief from one answer option to another.
-Since these abrupt changes happened generally towards the end of each question, during most days the Brier scores were really high.
-Given that these problems were typically very long, they gained a higher weight under \textit{Scores by Day} than under \textit{Scores by Problem}.
-This explains why the \textit{Scores by Day} are higher than the \textit{Scores by Problem} under the columns \textit{All Problems} and \textit{Long}.
-
-\subsection{In- and Out-of-Sample Sharpness and Calibration}
-\label{calibration}
-A calibration plot is a simple tool for visually assessing the sharpness and calibration of a model.
-The idea is to plot the probability forecasts against the observed empirical frequencies.
-Therefore any deviation from the diagonal line suggests poor calibration.
-A model is considered under-confident (or over-confident) if the points follow an S-shaped (or \reflectbox{S}-shaped) trend.
-To assess sharpness of the model, it is common practice to place a histogram of the given forecasts in the corner of the plot.
-Given that the data were balanced, any deviation from the the baseline probability of 0.5 suggests improved sharpness.
-
-\begin{figure}[h!]
-%\newgeometry{left=5cm, right = 1cm, bottom=0.1cm}
-
-\subfigure[\scriptsize SDLM ]{
-\hspace*{-2em}  \includegraphics[width = 0.286\textwidth]{Figures/CalibrationNONE10}
-\label{CalibrationNONEa}
-}
-\subfigure[\scriptsize $\text{SAC}_{\text{LOG}}$ ]{
-\hspace*{-1.15em} \includegraphics[width = 0.235\textwidth]{Figures/CalibrationLOG10}
-\label{CalibrationPLATTLOG10}
-}
-\subfigure[\scriptsize $\text{SAC}_{\text{BRI}}$ ]{
-\hspace*{-1.15em} \includegraphics[width = 0.235\textwidth]{Figures/CalibrationBRI10}
-\label{CalibrationPLATTBRI10}
-}
-\subfigure[\scriptsize $\text{BSAC}_{\text{LOG}}$ ]{
-\hspace*{-1.15em} \includegraphics[width = 0.286\textwidth]{Figures/CalibrationFULL10}
-\hspace*{-2em} \label{CalibrationPLATTFULL10}
-}
-
-\subfigure[ \scriptsize SDLM]{
- \hspace*{-2em} \includegraphics[width = 0.286\textwidth]{Figures/CalibrationNONE}
-\label{CalibrationNONEb}
-}
-\subfigure[\scriptsize $\text{SAC}_{\text{LOG}}$]{
-\hspace*{-1.15em} \includegraphics[width = 0.235\textwidth]{Figures/CalibrationSTO-Log}
-\label{CalibrationSTC-Log}
-}
-\subfigure[ \scriptsize $\text{SAC}_{\text{BRI}}$]{
-\hspace*{-1.15em} \includegraphics[width = 0.235\textwidth]{Figures/CalibrationSTO-Bri}
-\label{CalibrationSTC-Bri}
-}
-\subfigure[\scriptsize  $\text{BSAC}_{\text{LOG}}$]{
-\hspace*{-1.15em} \includegraphics[width = 0.286\textwidth]{Figures/CalibrationSO}
-\hspace*{-2em}\label{CalibrationBSTC-Log}
-} 
-%\restoregeometry
-\caption[Optional caption for list of figures]{The top and bottom rows show in- and out-of-sample calibration and sharpness, respectively.
-The models are Simple Dynamic Linear Model (SDLM), the \textit{Sample-And-Calibrate} approach that optimizes over the logarithmic score ($\text{SAC}_{\text{LOG}}$), the \textit{Sample-And-Calibrate} approach that optimizes over the Brier score ($\text{SAC}_{\text{BRI}}$), and the fully-Bayesian version of the \textit{Sample-And-Calibrate} approach that optimizes over the logarithmic score ($\text{BSAC}_{\text{LOG}}$).
-}
-\label{Calibration-Out}
-\end{figure}
-
-The top and bottom rows of Figure \ref{Calibration-Out} present calibration plots for SDLM, $\text{SAC}_{\text{LOG}}$, $\text{SAC}_{\text{BRI}}$, and $\text{BSAC}_{\text{LOG}}$ under in- and out-of-sample probability estimation, respectively.
-Each setting is of interest in its own right:  Good in-sample calibration is crucial for model interpretability.
-In particular, if the estimated crowd belief is well-calibrated, then the elements of the bias vector $\boldsymbol{b}$ can be used to study the amount of under- or over-confidence in the different expertise groups.
-Good out-of-sample calibration and sharpness, on other hand, are necessary properties in predicting future events with high accuracy.
-To guide our assessment, the dashed bands around the diagonal connect the point-wise, Bonferroni-corrected (\citet{bonferroni}) 95\% lower and upper critical values under the null hypothesis of calibration.
-These have been computed by running the bootstrap technique described in \citet{brocker2007increasing} for 10,000 iterations.
-The in-sample predictions were obtained by running the models for 10,200 iterations, leading to a final posterior sample of 1,000 observations after thinning and using the first 200 iterations for burn-in.
-The out-of-sample predictions were given by the 10-fold cross-validation discussed in Section \ref{forecasting}.
-
-Overall, SAC is sharp and well-calibrated both in- and out-of-sample with only a few points barely falling outside the \textit{point-wise} critical values.
-Given that the calibration does not change drastically from the top to the bottom row, SAC can be considered to present robustness against over-fitting.
-This is, however, not the case with $\text{BSAC}_{\text{LOG}}$ that is well-calibrated in-sample but presents over-confidence out-of-sample.
-Figures \ref{CalibrationNONEa} and \ref{CalibrationNONEb} serve as baselines by showing the reliability plots for SDLM.
-Given that this model does not perform any explicit calibration, it is not surprising to see most points outside the critical values.
-The pattern in the deviations suggests drastic under-confidence.
-Furthermore, the inset histogram reveals drastic lack of sharpness.
-Therefore SAC can be viewed as a well-performing compromise between SDLM and $\text{BSAC}_{\text{LOG}}$ that avoids over-confidence without being too conservative.
-
-\subsection{Group-Level Expertise Bias}
-\label{ExpertBias}
-
-\begin{figure}[!ht]
-\vspace*{-1em} 
-%\includegraphics[width = \textwidth]{Figures/BiasesBoxplots}
-\includegraphics[width = \textwidth]{Figures/BiasesBoxplots}
-\label{BiasesLOGBoxplots}
-\vspace*{-2em}
-
-\caption[Optional caption for list of figures]{Comparing the bias-levels across self-reported expertise under different approaches. 
-Posterior distributions of $b_j$ for $j = 1, \dots, 5$ under the \textit{Sample-And-Calibrate} approach that optimizes over the logarithmic score ($\text{SAC}_{\text{LOG}}$) and the \textit{Sample-And-Calibrate} approach that optimizes over the Brier score ($\text{SAC}_{\text{BRI}}$).
-}
-\label{Biases}
-\end{figure}
-
-Recall from Section \ref{data} that the experts were asked to self-assess their level of expertise (on a 1-to-5 scale with 1 = Not At All Expert to 5 = Extremely Expert) on any questions in which they participated.
-The self-reported expertise then divides the experts into 5  groups, with each group assigned a separate multiplicative bias term.
-This section uses SAC to explore the posterior distributions of these multiplicative bias terms.
-Figure \ref{Biases} presents the posterior distributions of the bias terms with side-by-side box plots.
-Given that the distributions fall completely below the \textit{no-bias} reference-line at 1.0, all the groups are deemed under-confident.
-Even though the exact level of under-confidence is affected slightly by the extent to which the extreme probabilities are truncated (see Section \ref{practicalMatters}), the qualitative results in this section remain insensitive to different levels of truncation.
-
-The under-confidence decreases as the level of  expertise increases.
-For instance, the posterior probability that the most expert group is the least under-confident is approximately equal to $1.0$, and the posterior probability of a strictly decreasing level of under-confidence is approximately 0.87.
-The latter probability is driven down by the inseparability of the two groups with the lowest levels of self-reported expertise.
-The fact that these groups are very similar suggests that the experts are poor at assessing how little they know about a subject that is strange to them.
-If these groups are combined into a single group, the posterior probability of a strictly decreasing level of under-confidence is approximately 1.0.
-
-The decreasing trend in under-confidence can be reasoned by viewing the process of making a subjective probability  as Bayesian updating: a completely ignorant expert aiming to minimize a reasonable loss function, such as the Brier score, has no reason to give anything but 0.5 as his probability forecast.
-However, as soon as the expert gains some knowledge about the event, he produces an updated forecast that is a compromise between his initial forecast and the new information acquired.
-The updated forecast is therefore conservative and too close to 0.5 as long as the expert remains only partially informed about the event.
-If most experts fall somewhere on this spectrum between ignorance and full information, their average forecast tends to fall strictly between 0.5 and the most-informed probability forecast (see \citet{Baron} for more details).
-Given that expertise is to a large extent determined by subject-matter knowledge, the level of under-confidence can be expected to decrease as a function of the group's level of self-reported expertise.
-
-Finding under-confidence in all the groups is a rather surprising result given that many previous studies have shown that experts are often over-confident (see, e.g., \citet{lichtenstein1977calibration, morgan1992uncertainty, bier2004implications} for a summary of numerous calibration studies).
-It is worth emphasizing two points: First, our result is a statement about groups of experts and hence does not invalidate the possibility of the individual experts being overconfident.
-To make conclusions at the individual-level based on the group-level bias terms would be considered an \textit{ecological inference fallacy} (see, e.g., \citet{lubinski1996seeing}).
-Second, the experts involved in our dataset are overall very well calibrated (\cite{mellers}).
-A group of well-calibrated experts, however, can produce an aggregate forecast that is under-confident.
-
-%More specific group-level inference can be performed by computing probabilities based the estimated posterior distributions.
-For instance, the posterior probability that the most expert group is the least under-confident is approximately equal to $1.0$, and
-%the posterior probability of a strictly decreasing level of under-confidence by expertise ranking is approximately 0.87.
-The latter probability is driven down by the inseparability of the two groups associated with the lowest levels of self-reported expertise.
-The fact that these groups are very similar suggests that the experts are poor at assessing how little they know about a subject that is strange to them.
-If these groups are combined into a single group, the posterior probability of a strictly decreasing level of under-confidence is approximately 1.0.
-
-\subsection{Question Difficulty and Other Measures}
-\label{QuestionDifficulty}
-One advantage of our model arises from its ability to produce estimates of interpretable question-specific parameters $\gamma_k$, $\sigma^2_k$, and $\tau^2_k$.
-These quantities can be combined in many interesting ways to answer questions about different groups of experts or the questions themselves.
-For instance, being able to assess the difficulty of a question could lead to more principled ways of aggregating performance measures across questions or to novel insight on the kind of questions that are found difficult by experts (see, e.g., a discussion on the \textit{Hard-Easy Effect} in \cite{Wilson94cognitivefactors}).
-To illustrate, recall that higher values of $\sigma^2_k$ suggest greater disagreement among the participating experts.
-Given that experts are more likely to disagree over a difficult question than an easy one, it is reasonable to assume that $\sigma^2_k$ has a positive relationship with question difficulty.
-An alternative measure is given by $\tau_k$ that quantifies the volatility of the underlying circumstances that ultimately decide the outcome of the event.
-Therefore a high value of $\tau_k$ can cause the outcome of the event to appear unstable and difficult to predict.
-
-\begin{figure}[h!]
-\vspace{-2em}
-\centering
-\hspace*{-3.5em} 	\includegraphics{Figures/LegendExamplePlot} % requires the graphicx package
-\vspace{-3.5em}
-
-\hspace{-0.5em}
-\subfigure[\textit{Will the expansion of the European bailout fund be ratified by all 17 Eurozone nations before 1 November 2011?}]{
-\hspace*{-1em}  \includegraphics[width= 0.49\textwidth]{Figures/ExamplePlot1b}
-\label{Example1b}
-} \hspace{0.5em}
-\subfigure[\textit{Will the Nikkei 225 index finish trading at or above 9,500 on 30 September 2011?}]{
-\hspace*{-1em}  \includegraphics[width= 0.49\textwidth]{Figures/ExamplePlot2b}
-\label{Example2b}
-}
-
-\caption[Optional caption for list of figures]{Scatterplots of the probability forecasts given for two questions in our dataset.
-The shadings represents the self-reported expertise of the expert who provided the probability forecast.
-The solid line gives the posterior mean of the calibrated crowd belief as estimated by $\text{SAC}_{\text{LOG}}$.
-The surrounding dashed lines connect the point-wise 95\% posterior intervals.}
-\label{ExamplePlotsFinal}
-\end{figure}
-
-As a final illustration of our model, we return to the two example questions introduced in Section \ref{data}.
-Figure \ref{ExamplePlotsFinal} is a copy of Figure \ref{ExamplePlots} with the addition of a solid line surrounded by a dashed band.
-The solid line represents the posterior mean of the calibrated crowd belief as estimated by $\text{SAC}_{\text{LOG}}$.
-The dashed lines connect the point-wise 95\% posterior intervals across different time points.
-Given that $\hat{\sigma}_k^2 = 2.43$  and $\hat{\sigma}_k^2 = 1.77$ for the questions depicted in Figures \ref{Example1b} and \ref{Example2b}, respectively, the first question provokes more disagreement among the experts than the second one.
-Intuitively this makes sense because the target event in Figure \ref{Example1b}  is determined by several conditions that may change radically from one day to the next while the target event in Figure \ref{Example2b} is determined by a relatively steady stock market index.
-Therefore it is not surprising to find that in Figure \ref{Example1b} $\hat{\tau}_k^2 = 0.269$, which is much higher than $\hat{\tau}_k^2 = 0.039$ in Figure \ref{Example2b}.
-We may conclude that the first question is inherently more difficult than the second one.
-
-\section{Discussion}
-This paper began with an introduction of a rather unorthodox but nonetheless realistic time-series setting where probability forecasts are made very infrequently by a heterogeneous group of experts.
-The resulting data is too sparse to be modeled well with standard time-series methods.
-In response to this lack of appropriate modeling procedures, our work introduces an interpretable time-series model that incorporates self-reported expertise to capture a sharp and well-calibrated estimate of the crowd belief.
-The model estimation is performed in two steps: The first step, known as the \textit{sampling step}, samples constrained versions of the model parameters via Gibbs sampling.
-The sampling is done from standard distributions with fast convergence to the target distribution (see Appendix \ref{appendix} for technical details).
-The second step, known as the \textit{calibration step}, uses a one-dimensional optimization procedure to transform the constrained parameter values to their unconstrained counterparts.
-To the best of our knowledge, this procedure extends the forecasting literature into rather unexplored areas of probability aggregation.
-
-\subsection{Summary of Findings}
-The model was applied to an unusually large dataset on expert probability forecasts.
-The estimated crowd belief was found to be sharp and well-calibrated under both in- and out-of-sample settings.
-This has direct implications  on predictive power and model interpretability.
-First, the model was shown to outperform other probability aggregators in terms of forecasting ability.
-In particular, the model was deemed a well-balanced compromise that avoids over-fitting without being overly conservative.
-Second, the crowd belief was used as the no-bias reference point to study the bias among groups of experts with different levels of self-reported expertise.
-All the groups were found to be under-confident.
-The under-confidence, however, decreased as the level of self-reported expertise increased.
-This result is about groups of experts and hence does not conflict with the well-known result of the individuals being over-confident (see, e.g.,  \citet{lichtenstein1977calibration, morgan1992uncertainty, bier2004implications}).
-Besides making predictions or studying group-level bias, the model can be used to generate estimates of many problem-specific parameters.
-These quantities have clear interpretations and can be combined in many interesting ways to explore a range of hypotheses about different types of questions and expert behavior.
-
-\subsection{Limitations and Directions for Future Research}
-Our model preserves parsimony while addressing the main challenges in modeling sparse probability forecasting data.
-Therefore it can be viewed as a basis for many future extensions.
-To give some ideas, recall that most of the model parameters were assumed constant over time.
-It is intuitively reasonable, however, that these parameters behave differently during different time intervals of the question.
-For instance, the level of disagreement (represented by $\sigma^2_k$ in our model) among the experts can be expected to decrease towards the final time point when the question  resolves.
-This hypothesis could be explored by letting $\sigma^2_{t,k}$ evolve dynamically as a function of the previous term $\sigma^2_{t-1,k}$ and random noise.
-Furthermore, this parameter along with the bias term can be explored at an individual level if the experts updated their forecasts very frequently.
-Estimates of $\sigma_{i,k}^2$ and $b_{i,k}$ could then be used to separate the accurate forecasters from the inaccurate ones.
-
-If the experts were asked to provide personal information during the data collection process, it may be of interest to study the forecasting behavior of different kinds of experts.
-For instance, this paper modeled the bias separately within each expertise group.
-This is by no means restricted to the study of bias or its relation to self-reported expertise.
-Different parameter dependencies could be constructed based on many other expert characteristic, such as gender, education, or specialty, to produce a range of novel insight on the forecasting behavior across different groups of experts.
-It would also be useful to know how expert characteristics interact with question types, such as economic, domestic, or international.
-This way the researcher can gain insight about the kind of experts who generally perform well on certain types of questions.
-The results would be of interest to the decision-maker who could use the information as a basis for consulting only a high-performing subset of the available experts.
-
-%If the experts updated their probability forecasts very frequently, it is possible to use the model to estimate expert-specific parameters such as $\sigma_{k,i}^2$ or $b_{i,k}$.
-Together these estimates can be used to separate the accurate from the inaccurate forecasters.
-The estimates can also provide insight on the responsiveness and
-
-Given that decision-makers can have different preferences on calibration and sharpness, it may also be of interest to study hidden processes with other goals besides maximizing sharpness subject to calibration.
-For instance, a government official may want to sacrifice some calibration for extra sharpness.
-This can be easily achieved by changing the optimization criterion in Equation (\ref{OSE}) to an appropriate function that meets the new goal.
-
-Other future directions could aim to remove some of the obvious limitations of our model.
-For instance, recall that the random components are assumed to follow a normal distribution.
-This is a strong assumption that may not always be justified.
-Logit-probabilities, however, have been modeled with the normal distribution before (see, e.g., \citet{Erev1994}).
-Furthermore, the normal distribution is a rather standard assumption in psychological models (see, e.g.,  signal-detection theory  in \citet{tanner1954decision}).
-A second limitation resides in the assumption that both the observed and hidden processes are expected to grow linearly.
-This assumption could be relaxed, for instance, by adding higher order terms to the model.
-A more complex model, however, is likely to sacrifice interpretability.
-Given that our model can detect very intricate patterns in the crowd belief  (see Figure \ref{ExamplePlotsFinal}), compromising interpretability for the sake of facilitating non-linear growth is hardly necessary.
-
-\section{Acknowledgements}
-\noindent
-This research was supported by a research contract to the University
-of Pennsylvania and the University of California from the Intelligence Advanced Research Projects Activity
-(IARPA) via the Department of Interior National Business Center
-contract number D11PC20061. The U.S. Government is authorized to reproduce and
-distribute reprints for Government purposes notwithstanding any
-copyright annotation thereon. Disclaimer: The views and conclusions
-expressed herein are those of the authors and should not be
-interpreted as necessarily representing the official policies or
-endorsements, either expressed or implied, of IARPA, DoI/NBC, or the
-U.S. Government.
-
-We deeply appreciate the project management
-skills and work of Terry Murray and David Wayrynen, which went far
-beyond the call-of-duty on this project.
-
-\appendix
 \section{Technical Details of the Sampling Step}
@@
 \end{align*}
-where the initial values, $X_{0|0,k}$ and $P_{0|0, k}$, are equal to $0$ and $1$, respectively. The update step is 
-\begin{align*}
-e_{t,k} &= Y_{i,t,k} - b_{i,k} X_{t | t-1, k} \\
-S_{t,k} &=  \sigma_k^2 + b_{i,k}^2 P_{t|t-1, k}\\
-K_{t,k} &=  P_{t|t-1, k} b_{i,k} S_{t,k}^{-1} \\
-X_{t|t, k} &= X_{t|t-1, k} + K_{t,k} e_{t,k} \\
-P_{t|t,k} &= (1 - K_{t,k} b_{i,k}) P_{t|t-1,k},
-\end{align*}
-where $b_{i,k}$ is the corresponding bias term for the $i$th expert in the $k$th question.
-The update step is repeated sequentially for each observation $Y_{i,t,k}$ given at time $t$.
+where the initial values, $X_{0|0,k}$ and $P_{0|0, k}$, are equal to $0$ and $1$, respectively.
+
+\begin{algorithm}
+\caption{The update step of the FFBS algorithm.
+$N_{t,k}$ denotes the number of forecasts made at time $t$ for question $k$.
+The subindex  $j(i)$ denotes the $i$th expert's self-assessed expertise group.}
+\label{update}
+\begin{algorithmic}
+\For{$i = 1, 2, \dots, N_{t,k}$}  
+\State $e_{i, t,k} = Y_{i,t,k} - b_{j(i)} X_{t | t-1, k}$
+\State $S_{i,t,k} =  \sigma_k^2 + b_{j(i)}^2 P_{t|t-1, k}$
+\State $K_{i,t,k} =  P_{t|t-1, k} b_{j(i)} S_{i,t,k}^{-1}$
+\State $X_{t|t, k} = X_{t|t-1, k} + K_{i, t,k} e_{i,t,k}$
+\State $P_{t|t,k} = (1 - K_{i, t,k} b_{j(i)}) P_{t|t-1,k}$
+\If {$i \neq N_{t,k}$}
+\State $X_{t|t-1, k} = X_{t|t, k}$
+\State $P_{t|t-1, k} = P_{t|t, k}$
+\EndIf
+%\If {$i\geq maxval$}
+%    \State $i\gets 0$
+%\Else
+%    \If {$i+k\leq maxval$}
+%        \State $i\gets i+k$
+%    \EndIf
+%\EndIf
+\EndFor
+\end{algorithmic}
+\end{algorithm}
+%\textbf{repeat} for $i = 1, 2, \dots, N_{t,k}$:
+%\begin{align*}
+%e_{i, t,k} &= Y_{i,t,k} - b_{j(i)} X_{t | t-1, k} \\
+%S_{i,t,k} &=  \sigma_k^2 + b_{j(i)}^2 P_{t|t-1, k}\\
+%K_{i,t,k} &=  P_{t|t-1, k} b_{j(i)} S_{i,t,k}^{-1} \\
+%X_{t|t, k} &= X_{t|t-1, k} + K_{t,k} e_{i,t,k} \\
+%P_{t|t,k} &= (1 - K_{t,k} b_{i,k}) P_{t|t-1,k}\\
+%\end{align*}
+The update step is given by Algorithm \ref{update}.
+The update is repeated sequentially for each observation $Y_{i,t,k}$ given at time $t$.
 For each such repetition, the previous posterior values, $X_{t|t, k}$ and $P_{t|t, k}$, are considered as the new prior values, $X_{t|t-1, k}$ and $P_{t|t-1, k}$.
